### 2018-05-15 paper/paper.tex
First commit

% !TEX program = pdflatex
\input{preamble/preamble.tex}
\input{preamble/preamble_math.tex}
\input{preamble/preamble_acronyms.tex}

\DeclareMathOperator*{\tr}{tr}

\newcommand{\UN}[1]{\ensuremath{\left|#1\right|_\infty}}
\newcommand{\EN}[1]{\ensuremath{\left|#1\right|}}
\newcommand{\kldiv}[2]{\ensuremath{D\left(#1||#2\right)}}

\usepackage{cancel}

% \linenumbers

\title{Markov Link Method for calibrating without joint measurement, including the case of destructive measurements}
\author{Jackson Loper}

\usepackage{amsthm}
\newtheorem*{thm}{Theorem}
\newtheorem{lem}{Lemma}
\newtheorem{conj}{Conjecture}

\theoremstyle{definition}
\newtheorem{example}{Example}

\begin{document}
\maketitle

\begin{abstract}
A proliferation of new experimental tools has left a serious gap: calibration. Two thermometers can be calibrated against each other by simply measuring the same bodies of water with both thermometers, but the problem is much harder for many modern tools.  One common problem is that we do not have measurements from the same ``body of water'' for both tools.  We propose the Markov Link Method (MLM) as a way to overcome this difficulty.   This method produces consistent estimators that tightly bound the calibration, i.e. the conditional distribution of one tool's measurement given another tool's measurement. It achieves this without any measurement data from both tools applied to the same ``bodies of water.'' Moreover, MLM makes zero assumptions about what calibrations we might expect to see, instead applying a subpopulation-based conditional independence assumption.   We evaluate MLM on a pair of single-cell RNA techniques, obtaining precise calibrations between the tools as well as more accurate models for each tool separately.
\end{abstract}

The modern setting is rife with experimental measurement tools, and it can be very frustrating to understand how the output of these tools relate to one another.  This problem is known as ``calibration'' or ``zeroing'' \cite{bipm2008international}.  A calibration tells us what readings we should expect from one tool, given the reading we obtained from another tool.   Calibration additionally must give uncertainty bounds for how much we can trust those expectations.  Calibration between measurement tools allows us to combine experimental results from different labs and different methodologies into larger scientific theories.

Formally, a calibration is simply a conditional distribution.  We will denote it by $q^*(y|x)$.  As input, this conditional distribution takes the measurement results from one tool on a particular specimen, $x$.  As output, it yields the probability of obtaining result $y$ from a second tool applied to measure the same specimen.  One way to learn the calibration is to measure the same specimens under both tools.  We call this ``joint measurement.''  Unfortunately, calibrations are often required even when joint measurement is unavailable.  For example, if the measurement tool alters the specimen being measured, joint measurement is simply impossible.  In other cases, it may be expensive or impractical.

We here propose the Markov Link Method (MLM) to estimate calibrations between tools.  The MLM can be trained without any joint measurement.  The key idea is to use multiple subpopulations of specimens.  If each subpopulation captures a different slice of the overall population, we can obtain tight bounds on the true calibration.  This is true even if each subpopulation is highly heterogeneous.  By integrating information from all the subpopulations we can make rigorous deductions about what the calibration might be.  MLM also gives suggestions about which further subpopulations might be helpful to study in order to further refine our knowledge of the true calibration.  The method can also be naturally extended to calibration distributions among many tools.

\section{The Markov Link Method assumption}

To make these ideas rigorous, let us develop a little bit of notation.  Let us say we are considering the members of a large population.  For example, each ``specimen`` in the large population might be a human cell or a piece of metal which needs to be tested.  We will assume there are three basic properties of interest for each specimen $i$:

\begin{enumerate}
\item $\ell_i$, the subpopulation or side information.  We assume that the overall population can be split into subpopulations of interest.  For example, we could define subpopulations by looking at cells in different parts of the body, or cells with different sizes.  
\item $X_i$, the result of measuring a specimen with tool I.  For example, perhaps tool I takes a picture of the cell with a destructive electron microscopy method.
\item $Y_i$, the result of measuinrg a specimen with tool II.  For example, perhaps tool II measures the RNA expression of the cell with a destructive sequencing method.
\end{enumerate}

We here consider the case that ``joint measurement'' is impossible or impractical.  In terms of the notation above, that means that for any given specimen we can either observe $\ell_i, X_i$ or $\ell_i,Y_i$.  We can never observe $\ell_i,X_i,Y_i$ for any specimen $i$.  Despite this obstacle, we would like to estimate the conditional distribution $q^*(y|x)\triangleq \mathbb{P}(Y_i=y|X_i=x)$.  Under suitable assumptions, we will see that such estimation is indeed possible using the side information $\ell$.  However, this estimation suffers from so-called \emph{identifiability problems}, which we will describe in detail in the next section.  Fortunately, this identifiability error can be quantified, allowing us to determine exactly what our data tells us about the true calibration.  If the identifiability error is large, the Markov Link method suggests future experiments based on new subpopulations which will hone down on the identifiability problem.

\begin{figure}
\hfill{}\includegraphics[width=.7\textwidth]{../images/bigpicture}\hfill{}
\caption{We consider the case that for each specimen there are two ways we might measure it.  We divide the specimens into two groups.  The specimens on the left are measured with technique I and the specimens on the right are measured with technique II.  In the end, we are left wondering: what would have happened if we had measured the specimens on the left with technique II?  Or what if we had measured the specimens on the right with technique I?  The Markov Link Method gives a way to answer this problem using a subpopultaion analysis.  For each specimen, we take $\ell$ to indicate the subpopulation (indicated here by the color of the neuron, i.e.\ yellow or purple), $X$ to indicate the results we would obtain from that specimen if we measured it with method I, and $Y$ to indicate the results we would obtain from that specimen if we meausured it with method II. \label{fig:bigpicture}}
\end{figure}

The key assumption of the Markov Link Method is a conditional independence assumption: that $\mathbb{P}(Y|X,\ell_i)$ is independent of $\ell$, i.e. $\mathbb{P}(Y=y|X=x,\ell)=q^*(y|x)$ for every value of $\ell$.  Intuitively, this signifies that the manner in which $X$ and $Y$ are related is the same for each subpopulation.  If this assumption is not met, then the method presented here is not applicable.  The validity of these assumptions for a given situation should be closely contemplated.

Let us consider a few real-world examples where this assumption may apply.

\begin{itemize}
    \item Quality control for manufacturing.   The surest way to test the reliability of a part is to construct a machine that pushes the part until it breaks.  However, how can we test the reliability of the machine that performs the test?  In each test run there will be some variability induced by the machine itself, which induces a measurement error  In practice, some kind of assumptions about part homogeneity are used to approximate this error (cf.\ \cite{de2005gauge}).  However, if we have two testing machines we can use the MLM to obtain a calibration between the machines, even though we can never test the same part with both machines.  This enables us to bound the overall measurement error.  In this case, $\ell$ might indicate the type of a part being tested, $X$ would indicate the reliability of a part as measured by one machine, and $Y$ would indicate the reliability of a part as measured by another machine.  If the error in machine $Y$ is not correlated to the part type $\ell$, then the MLM assumption certainly holds.  Even if the error is correlated, the MLM assumption may still hold.  For example, imagine that the $Y$ error is correlated with the absolute reliability of the part, this may pose no problem because that reliability is measured by $X$.  As long as the correlations are the result of something which is measured in $X$, the conditional independence assumption will hold.

    \item Combining knowledge across experimental modalities: morphology and transcriptomics.  There are different ways to think about the different types of cells in an organism.  A traditional approach is to classify cells based on what they look like (cf.\ \cite{steinman1973identification,bloomfield1982physiological}).  A more modern approach is to assay the cell's transcriptome (cf.\ \citep{tasic2017shared}).  Unfortunately, modern high-resolution cell photography and single-cell sequencing technologies are both destructive.  As a result, we can't get both kinds of data for the same specimens.  For cells native to regions full of diverse cell-types, finding a correspondence is a real problem.  The result is two completely independent classifications of cells, one for each way of looking at the cell.  MLM allows us to estimate the relationship between those two classification systems, yielding a wholistic understanding of the different types of cells.  In this case, $\ell$ might indicate some side information such as where in the body the cell was found, $X$ would indicate the classification of the cell according to its transcriptomics, and $Y$ would indicate the classification of a cell according to its morphology.  We expect that cell morphology is largely a function of cell transcriptomics.  Thus, as long as the $X$ measurement is sufficiently detailed, we expect that any correlations between $Y$ and $\ell$ would be explained by $X$.  That is, the MLM assumption holds.

    \item Cancer treatment efficacy prediction.  Starting from in-vivo human cancers, many cell-lines have been cultured over the years.  These cell cultures live indefinitely on plates.  Many experiments have been performed to see how these cancer cells respond to treatment.  However, if a treatment work on a particular cultured cell-line, what can we say about which kinds of cancer might respond well to that treatment?  Coarse side-information such as original cancer location is often available for both in-vivo and cultured cells, but this is often a surprisingly weak signal.  Cell transcriptomes provides much more specific information about the cancer, and thus, in theory, what treatments might be appropriate (cf.\ \cite{cieslik2018cancer}).  However, we know that cultured cell-lines look quite different from in-vivo cells (cf. \cite{imamura2015comparison,haibe2013inconsistency}).  Moreover, we have very little joint measurement between the human cells which originally gave rise to a cell culture and the cells that survived to beceome the cell line.  The Markov Link method can leverage the common side-information together with separate transcriptome information to produce a fine-grained correspondence between in-vivo and cultured cells.  This correspondence can used to propose new treatments for cancers.  Here $\ell$ might indicate cancer location, $X$ might indicate transcriptomic expression of cultured cells, and $Y$ might indicate transcriptomic expression of in-vivo cells.  As the transcriptomic expression is much more informative than the cancer location, it is plausible that $X$ might be sufficient to explain any correlations between $\ell$ and $Y$.  Thus the MLM assumption may hold.

    \item Text/image correspondence.  Automatic image captioning is an ongoing effort in machine learning (cf.\ \cite{srivastava2018survey}).  There are three types of data available to help develop such algorithms: text-only data, image-only data, and paired-text-and-image data.  Obviously the last kind is the most useful for automatic image captioning, but there is much less of it.  The Markov Link Method suggests one way to use the more plentiful text-only and image-only data.  We can first apply classic machine learning techniques to get coarse labels for both kinds of data.  Using this side-information to identify subpopulations, the MLM can then deduce a fine-grained correspondence between text and images by combining information from across all the subpopulations.  Here $\ell$ would indicate coarse labels such as ``cat'' or ``street scene.''  These labels could be derived from either images or text and can be trained in a supervised fashion.  Then, $X$ would indicate the image and $Y$ would indicate a caption.  Since the caption should be determined by the picture $X$, the MLM assumption may hold.

    \item Replication crisis and lab effects.  Replicating a published study is not always an easy thing to do.  This difficulty is commonly attributed to selective publication bias, bad design, poor description of methods, and even outright fraud \cite{baker2016reproducibility}.  A calibration would allow us to understand this problem in detail.  If two labs perform identical experiments and get different data, that does not mean we need to throw out both datasets.  Instead, we can use MLM to calibrate the tools.  Once the tools are properly calibrated, we can combine both datasets.  Unlike other tools to deal with lab or batch effects (e.g. \cite{crow2018characterizing,johnson2007adjusting}), MLM makes zero assumptions about what calibrations we might expect.  In this case, $\ell$ would indicate subpopulations which both labs could access.  For example, we can take several batches of mice; for each batch we can send half to one lab and half to the other lab.  $X$ will the indicate the full results from each specimen examined in one lab and $Y$ coarser information from specimens examined in the other.  If the $X$ data is sufficiently detailed, the MLM assumption may hold.  
\end{itemize}

The main contributions of this paper are summarized here:

\begin{itemize}
\item Posing of the Markov Link Method assumption as an assumption that may make it possible to perform calibrations without joint measurement.
\item An analysis of some identifiability problems posed by the Markov Link Method assumption.  Although it turns out precise identifiability is often impossible, we formulate an asymptotic theory for bounding the magnitude of the this problem.  In particular, although we often cannot hope to exactly determine the calibration $q^*(y|x)\triangleq\mathbb{P}(Y=y|X=x)$, we can estimate a so-called polytope $\hat \Theta$ which captures our uncertainty about $q^*$ due to identifiability problems.
\item The result of the identifiability analysis is a polytope $\hat \Theta$.  We investigate various ways of measuring the extent of this polytope.  We see how it is possible to understand what this ``extent'' means in terms of the identifiability problem.  When the extent is small, we see that the identifiability problem is actually quite minor.
\item We apply the MLM to two single-cell transcriptomic datasets.  The result uncovers an important and unexpected property of one of the experimental methods.  
\end{itemize}

Our identifiability analysis stands on the shoulders of a long history of relating probabilistic assumptions to probabilistic inequalities on unidentifiable parameters.  Indeed, the core idea of this work is to take an assumption (the Markov Link assumption) and use it to place inequality bounds on an unidentifiable quantity (namely $q^*$).  When inequalities are fairly tight on all sides, we see that much can be learned even when what we want isn't identifiable.  Much of the prior literature in this kind of direction comes from research into causality.  For example, in \cite{bonet2001instrumentality} Bonet uses polytopes not unlike the ones seen here to explore whether a variable can be used as an instrument.  The famous Clauser-Horne-Shimony-Holt inequality was designed to help answer causality questions in quantum physics, but it also sheds light on what distributions are consistent with certain assumptions \cite{clauser1969proposed}.  Indeed the physics literature has contributed many key inequalities (cf. \cite{chaves2014inferring}, \cite{kela2017semidefinite}, and the references therein).  Perhaps the closest work to this one would be \cite{makarov1982estimates}, which used two marginal distributions to get bounds on a property of the joint distribution (namely the distribution of the sum).  We advance this approach to a more general-purpose technique, both by using many subpopulations to refine our estimates and by considering the entire space of possible joint distributions instead of simply a particular aspect of the joint.

The remainder of this work proceeds as follows.  First we give a careful consideration to the Markov Link Method assumption and how it allows us to use side information to estimate the calibration.  We also see when this side information will and will not be helpful.  Next, we outline the Markov Link Method and describe a new asymptotic consistency result concerning the identifiability issues.   Finally, we consider a real dataset.  We perform the method and then attempt to criticize the result three ways: through concerns about not having enough data, through concerns about modeling assumptions, and through concerns about identifiability.  The method largely prevails against these criticisms, and reveals an unexpected feature of the two techniques.

\section{Precise problem formulation and discussion of the identifiability problem}

Let us now be completely rigorous.  We shall make three assumptions:

\begin{enumerate}
\item For each specimen $i$, the distribution of $X_i,Y_i|\ell_i$ may be written 
\[
\mathbb{P}(X_i=x,Y_i=y|\ell_i) = p^*(x|\ell_i)q^*(y|x)
\]
This is the central assumption we have discussed at length above.  For convenience we will also introduce the notation
\[
\mathbb{P}(Y=y|\ell) = h^*(y|\ell)=\sum_x p^*(x|\ell) q^*(y|x)
\]
\item Joint measurement is unavailable.  In particular, we will assume we have $n+m$ individual specimens.  Of these, we have observed $\ell_i,X_i$ for $i\in\{1\cdots n\}$ and $\ell_i,Y_i$ for $i\in n+1\cdots n+m$.  
\item $X$ and $Y$ are discrete random variables with finite support (the general concepts here will apply more generally, but we leave it for future work).  If the data is not discrete, we can always make it so by dividing it into suitable bins.  In this simple case, we may summarize all of this information with two matrices:
\[
N^X_{\ell x} = |\{i\leq n:\ \ell_i=\ell,X_i=x\}|  \qquad N^Y_{\ell y} = |\{i>n:\ \ell_i=\ell,Y_i=x\}|
\]
Thus $N^X,N^Y$ are matrices counting the number of each kind of observation for method I and method II respectively. 
\end{enumerate}

Our goal will be to use $N^X,N^Y$ to estimate $q^*$, the calibration.  However, \emph{the data from the matrices $N^X,N^Y$ only enable us to estimate $p^*,h^*$}.  They do not enable us to directly estimate $q^*$.  Therefore, we define
\[
\Theta(p,h) \triangleq  \left\{q:\ \sum_x p(x|\ell)q(y|x) = h(y|\ell)\ \forall \ell,y,\ q(y|x)\geq 0\ \forall x,y, \sum_y q_{xy}=1\ \forall x\right\}
\]
as the set of values of the calibration $q$ which are consistent with a given value of $p$ (the conditional distribution of $X|\ell$) and $h$ (the conditional distribution of $Y|\ell$).  Any effort to estimate $q^*$ must therefore overcome two fundamentally different challenges:

\begin{description}
\item[Not-enough-data problems] We don't have infinite data, so we can't hope to exactly determine $p^*,h^*$.
\item[Identifiability problems] Even if we knew $p^*,h^*$ exactly, it is often impossible to know the value of $q^*$.  We can only ever know that it lies somewhere in $\Theta(p^*,h^*)$.  
\end{description}

Let us now take a moment to understand the set $\Theta$.  This set tells us how the things we can estimate (i.e.\ $p^*,h^*$) inform us about what we want to know (i.e.\ $q^*$).  To understand a bit more concretely how this works, it may be helpful to think of $p^*,h^*,q^*$ as matrices.  From this point of view, one aspect of the definition of $\Theta$ can be written more concisely as a matrix equality constraint on $q^*$: if $q \in \Theta(p^*,h^*)$, then $p^*q=h^*$.  The consequences of this equation depend upon whether $p^*$ has a left-pseudoinverse.
\begin{itemize}
    \item If $p^*$ has a left-pseudoinverse then this equation allows us to uniquely determine what we want ($q^*$) in terms of what we know ($p^*,h^*$).  In general, this will happen when the number of subpopulations outnumbers the number of different states that $X$ can take on.  In this case, the problem should be straightforward to solve.   
    \item If $p^*$ does not have a left-pseudoinverse, then there is an identifiability issue: we can never hope to exactly determine $q^*$.  
\end{itemize}

In this paper we will focus almost entirely on the second case.  In this second case there is a genuine identifiability issue;  it is \emph{impossible} to ever determine the true value of $q^*$, regardless of how much data we have.  It therefore becomes of paramount importance to understand the magnitude of this identifiability problem.  In particular, even though we cannot know $q^*$ exactly, can we place bounds on what $q^*$ could be?

To gain intuition, let us consider a few examples.\vspace{.1in}

\begin{example}
Let $\ell \in \{1,2\}$, $X\in \{1,2,3\}$, $Y\in \{1,2\}$, and 
\[
p^{*}=\left(\begin{array}{ccc}
40\% & 50\% & 10\% \\
10\% & 10\% & 80\%
\end{array}\right)\qquad h^{*}=\left(\begin{array}{cc}
20\% & 80\% \\
40\% & 60\%
\end{array}\right)
\]
That is, for example, $p^*_{1,2}=\mathbb{P}(X=2|\ell=1)=50\%$ and $h^*_{2,1}\mathbb{P}(Y=1|\ell=2)=40\%$.  Let us look at a single equation in the system $p^* q^* = h^*$ entailed by the MLM assumption.  For example, $h^*_{2,1}$, we get
\[
0.4 = h^*_{2,1} = \sum_x p^*_{2,x} q^*_{x,1} = .1 q^*_{1,1} + .1 q^*_{2,1} + .8 q^*_{3,1} 
\]
This immediately tells us something about $q^*$.  For example we now know that $q^*_{1,1} = 10(.4 - .1 q^*_{2,1} - 0.8 q^*_{3,1})$.  But it actually tells us much more than that.  Observe that $0\leq q^*_{xy}\leq 1$ for every $x,y$; in other words, probabilities must be positive by definition.  Thus, in particular,
\[
0.4 = .1 q^*_{1,1} + .1 q^*_{2,1} + .8 q^*_{3,1} \leq .2 + .8 q^*_{3,1}
\]
It follows immediately that $q^*_{3,1}\geq .25$.  Each equation in the MLM system yields insights of this kind; together these insights form the constraints that define $\Theta$.  The more different subpopulations you have, the more equations of this kind you will have.  
\end{example}

\vspace{.1in}

\begin{example}
Let $\ell \in \{1\}$, $X\in \{1,2,3,\cdots 11\}$, $Y\in \{1,2\}$, and 
\[
p^{*}=\left(\begin{array}{ccccc}
90\% & 1\% & 1\% & \cdots & 1\% \\
\end{array}\right)\qquad h^{*}=\left(\begin{array}{cc}
95\% & 5\% \\
\end{array}\right)
\]
Here we have only \emph{one} population.  Still we are able to say something quite significant; using similar reasoning to that shown above we see that $q^*_{1,1}\geq 94.\bar4\%$.  
\end{example}

\vspace{.1in}

\begin{figure}
\hfill{}\includegraphics[width=.7\textwidth]{../images/polytope}\hfill{}
\caption{We consider the simple case $\ell=1$, $X\in \{1,2\}$, $Y\in \{1,2\}$.  Then the space of values of $q$ which are valid probability distributions can be understood as a box.  The space of values of $q$ which satisfy the matrix equation $p^* q = h^*$ can be understood as a line that passes through that box.  The set of possible values of $q$ which are consistent with the matrix equation and are also valid probability distributions forms the set $\Theta$, a line segment.  We know that the true value of $q^*$ must lie somewhere inside this $\Theta$.  Depending upon the exact values of $p^*,h^*$, this line segment may be larger or smaller.  For example, if the line crosses close to a corner of the box, then $\Theta$ may be restricted to a very small region near that corner.  \label{fig:polytopeexample}}
\end{figure}

\begin{example}
Finally, let us consider one class of examples in which the entire situation can be visualized: $\ell=1$, $X\in \{1,2\}$, $Y\in \{1,2\}$.   Let $\theta^*_1 = q_^*{1,1}$ and $\theta^*_2 = q_^*{2,1}$.  Notice that in this simple case $\mathbb{P}(Y=2|\ell) = 1- \mathbb{P}(Y=1|\ell)$, so the two values $\theta_^*1,\theta_2^*$ completely define $q^*$.  The matrix equation $p^* q^* = h^*$ corresponds to a linear constraint on $\theta^*$.  On the other hand, the fact that $q^*$ must be a valid probability distribution tells us that $\theta_1^*,\theta_2^*$ must lie inside the box $[0,1]\times[0,1]$.   Combining the matrix equation with the constraints of the box, we learn that $q^*$ lies inside a certain bounded, one-dimensional space.  This space is precisely $\Theta$.  The length of this line segment depends on the exact values of $p^*,h^*$.  In some cases it is quite large, i.e.\ it will be difficult to know much about the true value of $q^*$.  However, in other cases the positivity restrictions force $\Theta$ to be a very small region indeed.  See Figure \ref{fig:polytopeexample} to visualize this.  In general, this $\Theta$ is much higher-dimensional and harder to visualize, but in this simple case we can see it clearly.
\end{example}

\section{The Markov Link Method}

Let $N^X,N^Y$ denote matrices describing the original observed data. The purpose of the Markov Link Method is to try to use these matrices to estimate $q^*(y|x)=\mathbb{P}(Y=y|X=x)$, the calibration between $X$ and $Y$.  The Markov Link Method proceeds in two main steps:

\begin{description}
\item[Estimation]  Estimate $p^*$ with robust pseudocount estimator:
\[
\hat p(x|\ell) \triangleq \frac{1+N^X_{\ell x}}{\sum_{x'} N^X_{\ell x'}}
\]
and then calculate a possible guess for $q^*$, namely
\[
\hat q = \argmax_q \sum_{\ell,y}N_{\ell y}\log\left(\sum_{x}\hat p(x|\ell)q(y|x)\right) + \kappa \sum_{xy} \log q(y|x)
\]
The entropic regularizer $\kappa$ makes it possible to obtain a unique estimate for $q^*$, despite the identifiability problem discussed above.   In practice, we took $\kappa$ to be any small constant less than 1.  The optimization problem which defines $\hat q$ is convex and easily solved.

Note that although this gives us a unique estimate for $q^*$, the identifiability problems have not magically vanished.  There is good reason to suppose that $\hat q$ is near to the set $\Theta(p^*,h^*)$, but if the set $\Theta$ is very large then it may be that $\hat q$ is very different from the true $q^*$.   Therefore, there is a second step we take: a criticism step.  This step should be taken for any model, but one which is particularly important in this case.  
\item[Criticism]  Level criticisms at the learned model to see whether $\hat q$ can be trusted.  
\begin{enumerate}
\item Use bootstrap to determine if we have enough data.
\item Use a held-out log-likelihood method to make basic sanity checks as to whether the MLM assumption fits with the data.
\item Measure the extent of $\Theta(\hat p,\hat p \hat q)$ to see whether we have significant identifiability issues.
\end{enumerate}
\end{description}

The very final point merits a little bit more discussion.  In the previous section, we saw that the identifiability issues could be summed up as follows: we can't generally tell what $q^*$ is, only that it lies somewhere inside the set $\Theta(p^*,h^*)$.  The extent of the identifiability problems can therefore be understood in terms of the extent of this set.  However, in the method above, we look at the extent of a different set: $\Theta(\hat p,\hat p \hat q)$.  This set is based on mere estimates: $p^* \approx \hat p$ and $h^* \approx \hat p \hat q$.  It is not immediately obvious that the size of this \emph{estimated} set would give us an indication of the size of the \emph{true} set of interest, $\Theta(p^*,h^*)$.  Fortunately, it is straightforward to show that under mild conditions this is indeed the case.  This is the content of the main analytical theorem of this paper, for which a precise statement may be found in Appendix \ref{sec:proof}.  It is in part due to the structure of this proof that we choose to estimate $h^*$ with $\hat p \hat q$ instead of a more traditional estimator.

\section{Empirical results}

\begin{figure}
\hfill{}
\includegraphics[width=.5\textheight]{../images/plxy}
\hfill{}

\caption{Input: two tables.  The Allen Institute had access to various cre/lox cell selection techniques.  Each technique pulls out a different group of cells.  Once the cells were selected, they were then either subjected to technique I technique II (`patch').  Technique I gives a very complete analysis of the gene expression of the cell.  Technique II gives a less complete analysis, but yields additional electrophysiological data that may be of interest.  For each subpopulation and each technique, data was gathered to estimate the distribution of different cell-types.  These are shown above.  Each column corresponds to a different cell-type.  Each row corresponds to a different subpopulation.  Each entry gives the proportion of a particular cell-type found within a particular population, according to a particular technique.  If the two methods were perfectly calibrated, we would expect the two tables to look identical.  \label{fig:alleninput}}
\end{figure}

\subsection{Background}

Our motivation for this problem arose from looking at Allen Institute cell-type assignment of cells, performed using two different experimental techniques (also called experimental ``modalities'').  Each modality would take a cell and determine what ``type'' of cell it was.  However, as part of that process it would destroy the cell.

Best efforts were made to use biological intuition to calibrate the methods.  For example, if a cell was designated as ``Lamp5 Egln3\_1'' celltype using one method, the hope was that it would also be given the same designation if it was processed using the other method.  The two methods were designed by scientists to achieve this goal.  However, each method has its own biases and errors, and it was not obvious whether this effort was successful.  In particular, it seemed clear that in some cases cells labelled one way with one method would get labelled another way with another method, but it was not clear how often this occurred.

Fortunately, there was a kind of information that seemed like it might help determine whether the two methods were properly calibrated: sub-populations.  Using a cre/lox system (cf.\ \citep{tasic2017shared}) they were able to pick out specific, overlapping subpopulations of neurons.  Each subpopulation was expected to contain different proportions of the different cell-types.  For each subpopulation and each method, many specimens were sampled and their cell-types determined.  If the methods were perfectly calibrated, we would expect that both methods would yield the same distribution of cell-types in each subpopulation.

Towards this effort, data was collected for each subpopulation and each method.  The result of this process was two tables, shown in Figure \ref{fig:alleninput}.  Perhaps not surprisingly, it was found that the distribution of cell-types appeared different under the two modalities.  In fact, in one of the methods some of the cells were designated as ``unknown,'' so even the set of cell-types was not the same between the two groups.  Clearly the methods were not perfectly calibrated -- but how big of a problem was it?  It is not obvious to know just by staring at Figure \ref{fig:alleninput}.  A more quantitative method was needed.

\subsection{Estimation}

The first step of the Markov Link Method was simply to estimate the calibration $q^*$.  Taking $\kappa=.1$, we obtained a calibration found in Figure \ref{fig:allenout}.  While this result seemed to generally suggest the methods were well-calibrated, there are some striking divergences.  For example, this calibration suggests a that celltypes designated as celltype ``10'' in one method are actually quite likely to be designated as celltype ``7'' in the other technique.  Both celltypes are associated with similar neuronal types with high expression of somatostatin.  However, type ``10'' is associated with genes such as `C1ql3,' `Chodl,'  and `Nts,' whereas type ``7'' is associated with `Etv1' and `Myh8.'

Notice that there are several fundamentally different reasons we could see these kinds of miscalibrations:

\begin{itemize}
    \item The methods are actually miscalibrated.
    \item The subpopulations were unevenly sampled.  It may be that certain techniques tend to cause cell-death in certain cell-types more than others.  These cells are then discarded from the data.  If this process is different between the two techniques, it could have an important impact on the calibration estimated here.
    \item The calibration was poorly estimated.
\end{itemize}

It is to the third point we now turn.

\subsection{Criticism}

TODO:
\begin{enumerate}
\item Nonparametric bootstrap (avg distance 0.66, Figure \ref{fig:nonparametricbootstrap})
\item Parametric bootstrap (avg distance .74, Figure \ref{fig:parametricbootstrap})
\item Held out likelihood (2.21 nats vs our method 2.14 nats)
\item RUX samples (avg distance .16, Figure \ref{fig:ruxsamples})
\item Uniform samples (avg distance .02, Figure \ref{fig:uniformsamples})
\end{enumerate}

\begin{figure}
\includegraphics[width=0.8\textwidth]{../images/nonparametricbootstrap}
\caption{Nonparametric bootstrap.\label{fig:nonparametricbootstrap}}
\end{figure}

\begin{figure}
\includegraphics[width=0.8\textwidth]{../images/parametricbootstrap}
\caption{Parametric bootstrap.\label{fig:parametricbootstrap}}
\end{figure}

\begin{figure}
\includegraphics[width=0.8\textwidth]{../images/unif}
\caption{RUX samples.\label{fig:ruxsamples}}
\end{figure}

\begin{figure}
\includegraphics[width=0.8\textwidth]{../images/rux}
\caption{Uniform samples.\label{fig:uniformsamples}}
\end{figure}

\section{Conclusions}

When joint measurement is impossible, it can be difficult to calibrate two methods against each other or understand how they may be related.  Here we show that a simple Markov assumption can make it possible to actually learn quite a lot.  Although the exact relationship may not be identifiable, a polytope of possible relationships can be identified, and this polytope may in fact be quite small indeed.  By exploring this polytope, we can understand what we know -- and what we don't know -- about the relationship between measurements.

The Markov assumption is of course not the only one that we could have used, and may not be valid in every case.  For example, it has been speculated that some cell types tend to die more often in one experimental modality than another, and these cells are not part of the data.  This violates our assumptions.  However, assuming this death rate can be roughly measured, it can be adjusted for, yielding a different but equally meaningful assumption about the data.  Moreover, if this method yields bizarre results, it may give useful clues as to exactly how cell death may happen differently in the two modalities.

Once we accept that what we're interested in may not be fully identifiable, any of a wide variety of assumptions can help us obtain practical bounds.  Although we may not be able to learn exactly what we want, we can learn a set of possibilities.  By probing this set carefully with uniform samplers and extremal tests, we can learn what the data actually has to say and what experiments we need to do to learn more.

\bibliographystyle{unsrt}
\bibliography{refs}

\appendix

\section{Dikin sampler}

\label{sec:dikin}

Consider a convex polytope $T=\{x:\ Ax\leq b\}$.  We have implemented a method for sampling from this polytope, based on the paper \citep{kannan2012random}.  This method makes use of the Dikin ellipsoids, $E(x)$.  For any $x$, these are defined by

\begin{itemize}
    \item Computing the distance from $x$ to each facet of the polytope, i.e. $d_i = b_i - \sum_j A_{ij} X_j$.
    \item Constructing $\tilde A$ as $\tilde A_{ij} = A_{ij} / d_i$.
    \item Define $E(x) = \{y:\ |\tilde A (X-y)|\leq 1\}$.  
\end{itemize}

We can use these ellpsoids to efficiently sample the polytope $T$.    At each step, we have some point $X\in T$, and we would to use this point to obtain a new sample $Y$, such that by iterating this process we asymptotically obtain samples which are uniform in $T$.  Here is how we use $X$ to get $Y$:

\begin{algorithm}[H]
 \KwData{A point $X\in T$}
 \KwResult{A point $Y\in T$}
 \vspace{.1in}
  Sample a proposal $\tilde Y$, uniformly from $E(X)$\;
    \eIf{$X \in E(\tilde Y)$}{
        Sample $U \sim \mathrm{Uniform}[0,1]$\;
        \eIf{$U \leq \mathrm{Vol}(E(X))/\mathrm{Vol}(E(\tilde Y)) \leq 1$}{
            Let $Y \gets \tilde Y$\;
        }{
            Let $Y \gets X$\;
        }
    }{
        Let $Y\gets X$\;
    }
 \caption{Dikin sampler step}
\end{algorithm}

It is easy to show that the stationary distribution of the Markov chain found by iterating these Dikin sampler steps is indeed uniform on $T$.  To ensure an numerically robust method in the face of high-dimensional and nearly degenerate matrices, we take the following approach to robustly sampling from the ellipsoid:

\begin{algorithm}[H]
 \KwData{An $n\times m$ matrix $\tilde A$}
 \KwResult{A point $X$ sampled uniformly from $\{x:\ |Ax|\leq 1\}$}
 \vspace{.1in}
  Sample $Z$ as an $n$-dimensional normal variables vector\;
  Let $X$ denote the solution to the least squares problem $\min_x|\tilde Ax - Z|$\;
  Normalize $X$ by $X \gets X/|\tilde AX|$\;
  Sample $U \sim \mathrm{Uniform}[0,1]$\;
  Scale $X$ by $X \gets X \times U^{1/m}$\;
 \caption{Ellipsoid sampler}
\end{algorithm}

\section{Proof of the theorem}

\label{sec:proof}

For the benefit of the reader, we here repeat the statement of our theorem in more explicit terms.

\begin{itemize}
\item Let $\UN{\cdot}$ denote the uniform norm (i.e.\ the maximum absolute value) and $\EN{\cdot}$ denote the Euclidean norm (i.e.\ the square root of the sum of the squares).  In the case of matrices, this Euclidean norm goes by the name of the Frobenius norm.  Recall that in this norm matrices satisfy a Cauchy-Schwarz like equality, $\EN{pq}\leq\EN{p}\EN{q}$.  Also recall that $\UN{a} \leq EN{a} \leq \sqrt{n}\UN{a}$ where $n$ is the number of entries in $a$.
\item Let $T_{a,b}$ denote the transition matrix polytope, i.e. the set of $a\times b$ matrices whose rows sum to 1 and whose entries are all positive.
\item Let $|\Omega_\ell|,|\Omega_X|,|\Omega_Y| \in \mathbb{N}$.
\item Let $p^*\in T_{|\Omega_\ell|,|\Omega_X|}$.
\item Let $q^*\in T_{|\Omega_X|,|\Omega_Y|}$.
\item We require the matrix $q^*$ has strictly positive entries, $q^*_{xy}\geq c>0$.
\item We require that the rows of $p^*$ are linearly independent.
\item Let $\hat p$ denote an empirical transition matrix drawn by obtaining $N_{X,\ell}$ samples for each row of $p^*$, i.e. we have samples $(\ell_1,X_1) \cdots (\ell_1,X_n)$ such that $\mathbb{P}(X_i=x) = p^*_{\ell_i,x}$, $N_{X,\ell}=\sum_{i=1}^n \mathbb{I}_{\ell_i=\ell}$, and $\hat p_{\ell x} = \sum_{i=1}^n \mathbb{I}_{X_i=x,\ell_i=\ell} / N_{X,\ell}$. 
\item Let $\hat h$ denote an empirical transition matrix drawn by obtainin $N_{Y,\ell}$ samples for each row of $h^*=p^*q^*$.  
\end{itemize}

Now fix any $\kappa>0$.  Let
\begin{align}
\hat q = \argmax_q \left(\sum_\ell N_{Y,\ell}\sum_{y}\hat h(y|\ell)\log\left(\sum_{x}\hat p(x|\ell)q(y|x)\right) + \kappa \sum_{xy} \log q(y|x)\right)\label{eq:mindef}
\end{align}
and $\hat \Theta=\{q:\ \hat p\hat q=\hat pq\} \cap T_{|\Omega_X|,|\Omega_Y|}$.  
\vspace{.1in}
\begin{thm}
If $N_{X,\ell},N_{Y,\ell}\rightarrow \infty$ in such a way that $N_{Y,\ell'}/\sum_{\ell}N_{Y,\ell} \geq \rho>0$ for each $\ell'$, then $\inf_{q\in \hat\Theta} \UN{q^*-q}\rightarrow 0$ in probability.
\end{thm}

\begin{proof}
It is well-known that $\hat p\rightarrow p^*$ in probability (in both the uniform or the Euclidean norm, which are of course equivalent in this case).  It is easy to see that the same goes for $\hat p\hat q\rightarrow h^*$ (see Lemma \ref{lem:easyconsistency}).  Thus, intuitively, the difficulty is this: by allowing ourselves to ensure $\UN{\hat p - p^*},\EN{\hat p - p^*},\UN{\hat p\hat q - h^*},\EN{\hat p\hat q - h^*}$ sufficiently small, can we find some $\tilde q\in \hat \Theta$ so that $\UN{\tilde q - q^*}$ is arbitrarily small?  It turns out we can.

Recall that $c>0$ is the smallest value of $q^*_{xy}$.  Fix any $\epsilon<c, p^*,q^*$.  Let the right inverse of a matrix be defined by $a^\dagger \triangleq a^T (aa^T)^{-1}$.  Note that since $p^*$ has linearly independent rows, this is well-defined and continuous in a small neighborhood around $p^*$.  Let $M=\EN{(p^*)^\dagger}$.  Find $\delta$ small enough so that if $\UN{p-p^*}<\delta$ then $\EN{p^\dagger}<2M$.  Taking a further smaller $\delta$ if necessary, ensure that if $\UN{p^*-p}<\delta$ then $\EN{p^*-p}$ is less than $\epsilon / 4M\sqrt{|\Omega_X||\Omega_Y|}$.  
Now fix any $\hat p,\hat q$ with $\UN{\hat p-p^*}<\delta$ and $\EN{\hat p \hat q - p^*q^*}<\epsilon/4M$.  Take
\[
\tilde q = q^* + \hat p^\dagger \hat p (\hat q - q^*)
\] 
Then we make the following observations:
\begin{itemize}
  \item Let us compute $\EN{\tilde q -q^*}$. We have
   \begin{align*}
   \EN{\tilde q -q^*} &= \EN{\hat p^\dagger \hat p (\hat q - q^*)} \leq 2M \EN{\hat p \hat q - \hat p q^*} \\
        &\leq 2M \EN{\hat p \hat q - p^* q^*} + 2M\EN{(p^* - \hat p)q^*}\\
        &\leq 2M \frac{\epsilon}{4M} + \frac{2M\epsilon}{4M\sqrt{|\Omega_X||\Omega_Y|}}\sqrt{|\Omega_X||\Omega_Y|}\UN{q^*} \leq \epsilon 
   \end{align*}
  \item $\hat p \tilde q = \hat p q^* + \hat p \hat q - \hat p q^* = \hat p \hat q$
  \item The rows of $\tilde q$ sum to 1.  This is easy to see, because the rows of $q^*$ sum to 1 and the rows of $\hat q$ sum to 1, and so $\tilde q \mathbf{1} = q^* \mathbf{1} + \hat p^\dagger \hat p (\hat q - q^*) \mathbf{1} = \mathbf{1} + 0$ as desired.
  \item The entries of $\tilde q$ are positive.   Indeed, the the smallest value of $q^*$ is $c$, and we have already argued that $\UN{\tilde q - q^*}\leq \epsilon$.  Thus the smallest value of $\tilde q$ is at least $c-\epsilon$, and we have required $\epsilon<c$.  
\end{itemize}
Thus $\UN{\tilde q - q^*}<\epsilon$ and $\tilde q \in \hat \Theta$.

In conclusion, we see that by taking $\hat p$ sufficiently close to $p^*$ and $\hat p \hat q$ sufficiently close to $p^*q^*$, we can ensure that the set $\hat \Theta$ contains a close which is arbitrarily close to the true $q^*$.  Since $\hat p$ and $\hat p \hat q$ are themselves consistent estimators, this completes the proof.
\end{proof}

\begin{lem}
\label{lem:easyconsistency} If $N_{X,\ell},N_{Y,\ell}\rightarrow \infty$ in such a way that $N_{Y,\ell'}/\sum_{\ell}N_{Y,\ell} \geq \rho>0$ for each $\ell'$, then $\UN{p^*q^*-\hat p\hat q},\EN{p^*q^*-\hat p\hat q}\rightarrow 0$ in probability.
\end{lem}
\begin{proof}
Our first task is to make a short study of the continuity of KL divergences on categorical distributions when the probabilities are bounded away from zero.  Recall that we have insisted $q^*_{xy}\geq c>0$ for every $x,y$ -- and this also means $(pq^*)_{\ell y}\geq c$ for every $\ell,y$, since each row of $p$ is itself a probability distribution.  Moreover, observe that the KL-divergence on $|\Omega_Y|$-dimensional distributions, $\kldiv{\hat r}{\tilde r} \triangleq \sum_{y} \hat r_y \log \hat r_y / \tilde r_y$, is \emph{uniformly} continuous on the space of such distributions whose minimum probability is greater than any fixed positive constant.   It follows that the map $h,p,q \mapsto D(h_\ell||(pq)_\ell)$ is also uniformly continuous on a space where $h$ and $q$ are strictly greater than some fixed positive constant.

With this in hand, the remainder of the proof follows naturally, using the well-known results that empirical distributions are consistent, i.e.\ $\hat p \rightarrow p^*$ and $\hat h \rightarrow p^*q^*$ in probability.

Fix any $\epsilon,\pi$.  Let $\delta$ the modulus of continuity in the norm $\UN{\cdot}$ at level $\epsilon\rho$ for the map $h,p,q \mapsto \kldiv{h_\ell}{(pq)_\ell}$ restricted to the domain where $h,q>c/2$.  Select $N$ large enough so that $\frac{1}{N_{Y,\ell}} \kappa |\Omega_X||\Omega_Y| \log \frac{1}{c} < \epsilon$ for each $\ell$ and so that with probability at least $\pi$ we have that $\hat h,\hat p$ so that $\UN{\hat h-p^*q^*},\UN{\hat p-p^*}\leq \delta,\UN{\hat h - p^*q^*}<c/2$. Then, with probability $\pi$, we must have
%
\begin{align*}
|\kldiv{\hat h_\ell}{(\hat p \hat q)_\ell} - \kldiv{h^*_\ell}{(\hat p \hat q)_\ell}| \leq \rho \epsilon \\
\kldiv{\hat h_\ell}{(\hat p q^*)_\ell} = |\kldiv{\hat h_\ell}{(\hat p q^*)_\ell} - \kldiv{(p^*q^*)_\ell}{(p^* q^*)_\ell}| \leq \rho\epsilon 
\end{align*}
%
Now, since $\hat q$ is defined as the maximizer of a certain quantity (Equation \ref{eq:mindef}), we may be sure that it is greater than the same quantity evaluated at $q=q^*$.  That is,
%
\begin{align*}
0 \leq& \sum_\ell N_{Y,\ell}\sum_{y}\hat h(y|\ell)\log\frac{\sum_{x}\hat p(x|\ell)\hat q(y|x)}{\sum_{x}\hat p(x|\ell)q^*(y|x)} + \kappa \sum_{xy} \log \frac{\hat q(y|x)}{q^*(y|x)}\\
 =& \sum_\ell N_{Y,\ell}\left(\kldiv{\hat h_\ell}{(\hat pq^*)_\ell} - \kldiv{\hat h_\ell}{(\hat p\hat q)_\ell}\right) + \kappa \sum_{xy} \log \frac{\hat q(y|x)}{q^*(y|x))}
\end{align*}
%
Applying our continuity results, it follows that
%
\begin{gather*}
 \sum_\ell N_{Y,\ell}\left(\kldiv{\hat h^*_\ell}{(\hat p\hat q)_\ell}\right)  \leq 2\left(\sum_\ell N_{Y,\ell}\right)\epsilon  + \kappa |\Omega_X||\Omega_Y| \log \frac{1}{c} 
\end{gather*}
%
Note that the left-hand summands are all positive.  So, in particular, it follows that for each $\ell'$, applying the uniformity condition $\rho$, we have
%
\begin{gather*}
\kldiv{\hat h^*_{\ell'}}{(\hat p\hat q)_{\ell'}}  \leq 2\epsilon  + \frac{1}{N_{Y,\ell'}} \kappa |\Omega_X||\Omega_Y| \log \frac{1}{c} \leq 3\epsilon
\end{gather*}

That is, we have shown convergence of probability in the KL sense: for any $\epsilon,\pi$ we can find $N$ high enough so that $\kldiv{\hat h^*_{\ell'}}{(\hat p\hat q)_{\ell'}}<\epsilon$ with at least probability $\pi$.  This, in turn yields convergence in probability in the Euclidean or uniform metrics by Pinsker's inequality. 
\end{proof}

\end{document}

### 2018-05-15 markovlink/__init__.py

@@ -1,0 +1,10 @@
+from . import estimation
+from . import exploration
+from . import criticize
+
+train_q = estimation.train_q
+find_extremal = exploration.find_extremal
+sample_prux = exploration.sample_prux
+sample_rux = exploration.sample_rux
+sample_uniform = exploration.sample_uniform
+train_test_split = criticize.train_test_split

### 2018-05-15 markovlink/criticize.py

@@ -1,0 +1,103 @@
+import numpy as np
+import numpy.random as npr
+
+
+def project_to_polytope(p,qdef,newq):
+    '''
+    Given 
+    - polytope defined by p @ q = p @ qdef and q>=0
+    - a point newq
+
+    Output
+    - the point in the polytope which is closest to newq
+
+    '''
+
+    try:
+        import cvxopt
+
+        raveln= np.prod(newq.shape)
+
+        # equality constraint is that A q = b
+        A,b = formulate_polytope(p,qdef)
+
+        # inequality constraint is that 
+        # (-I)q<=0
+        G=-np.eye(raveln)
+        h=np.zeros(raveln)
+
+        # thing to minimize is
+        #   .5 |q - newq|^2
+        # = .5 |q|^2 - <newq,q> + ...
+        # = .5 q^T I q + (-newq^T) q + ...
+        P = np.eye(raveln)
+        d = -newq.ravel()
+
+        '''
+        min   |q - qxy|^2
+        subj  p q = p qdef
+
+        Has become
+        
+        min   .5 q^T P q + d^T q
+        subj  Aq = b
+              Gq <= h
+
+        '''
+        
+        mtx=lambda x: cvxopt.matrix(x)
+        
+        rez=cvxopt.solvers.qp(mtx(P),mtx(d),G=mtx(G),h=mtx(h),A=mtx(A),b=mtx(b))
+        
+        if rez['status']=='optimal':
+            return np.array(rez['x']).reshape(newq.shape)
+        else:
+            raise Exception("Polytope projection failed!")
+    except ModuleNotFoundError:
+        raise("You don't seem to have CVXOPT.  We depend on CVXOPT for polytope projection :(.  Try conda install cvxopt?")
+    
+
+def train_test_split(Nly,m):
+    '''
+    Input:
+    - an array of positive integers, Nly
+
+    Output:
+    - an array train,test
+
+    For each row, we hold out m of the counts from Nly
+    '''
+
+    assert (np.sum(Nly,axis=1)>=m).all(),"Not all rows of Nly have %d total count; can't hold out that much"%m
+
+    Nly=np.require(Nly)
+    Nshp=Nly.shape
+
+    # sample (l,Y) events, uniformly without replacement
+    # from the (l,Y) events indicated by Nly
+
+    # we iteratively select a random point from Nly
[…]

### 2018-05-15 markovlink/estimation.py

@@ -1,0 +1,60 @@
+import numpy as np
+
+import sys
+
+def emupdate(p,q,kappa,Nly):
+    d1,d2=p.shape
+    d2,d3=q.shape
+    pi = p.reshape((d1,d2,1)) * q.reshape((1,d2,d3))
+    pi=pi/np.sum(pi,axis=1,keepdims=True)
+    
+    L=pi*Nly.reshape((d1,1,d3))
+    L=np.sum(L,axis=0) + kappa
+    
+    L = L / np.sum(L,axis=1,keepdims=True)
+    
+    return L
+
+
+def check_transition_matrix(a):
+    return (np.abs(np.sum(a,axis=1) - 1).max()<1e-9) and (np.min(a)>-1e-9)
+
+def train_q(p,Nly,iterations=5000,kappa=.0001,verbose=False):
+    '''
+    Uses EM algorithm to find the qhat which optimizes of the markov link method objective.
+
+    - p -- estimate empirical distribution of X|l (each row sums to 1.0)
+    - Nly -- Nly[l,y] is the number of times samples from subpopulation l yielded Y=y 
+        (each row sums to the number of samples for subpopulation l under technique II)
+    - kappa -- regularization 
+    - verbose -- if true, prints out updates as it trains to indicate progress
+    '''
+
+    Nly = np.require(Nly,dtype=np.float)
+    p=np.require(p,dtype=np.float)
+
+    assert len(Nly.shape)==2,"Nly should be a matrix"
+    assert len(p.shape)==2,"p should be a matrix"
+    d1,d2=Nly.shape
+    d1p,d3=Nly.shape
+    assert d1==d1p,'The first dimensions of Nly and p should be the same'
+    assert np.abs(np.sum(p,axis=1)-1).max()<1e-9,"The rows of p should sum to 1 and all entries should be positive"
+
+    # initialize qxy
+    qxy = np.ones((p.shape[1],Nly.shape[1])) / Nly.shape[1]
+
+    # run emupdate repeatedly
+    if verbose:
+        # ersatz tqdm replacement (didn't want to include tqdm as a requirement)
+
+        update_interval=1+iterations//20
+        for i in range(iterations): 
+            if i%update_interval==0:
+                sys.stdout.write('(%d/%d)'%(i,iterations))
+                sys.stdout.flush()
+            qxy=emupdate(p,qxy,kappa,Nly)
+    else:
+        for i in range(iterations):
+            qxy=emupdate(p,qxy,kappa,Nly)
+
+    return qxy

### 2018-05-15 markovlink/exploration.py

@@ -1,0 +1,246 @@
+import scipy as sp
+import scipy.optimize
+import numpy as np
+import numpy.random as npr
+import sys
+
+from . import estimation
+
+def sample_prux(plx,qxy,n):
+    '''
+    Consider the polytope of transition matrices q such that plx @ q = plx @ qxy.
+
+    Choose a direction uniformly at random, and find the pair of vertices in that polytope
+    which are most extremal along that axis.  Calculate the distance between those vertices
+    projected along the direction, as well as their distance in absolute terms.  
+
+    Repeat n times.
+
+    Returns a list of the form
+
+    [
+        (direction,samp1,samp2,distance_along_axis,absolute_distance),
+        (direction,samp1,samp2,distance_along_axis,absolute_distance),
+        (direction,samp1,samp2,distance_along_axis,absolute_distance),
+    ]
+    '''
+
+    lst=[]
+    for i in range(n):
+        direc=npr.randn(*qxy.shape)
+        direc=direc/np.linalg.norm(direc)
+        samp1=find_extremal(plx,qxy,direc)
+        samp2=find_extremal(plx,qxy,direc2)
+        lst.append((direc,samp1,samp2,np.sum(direc*(samp1-samp2)),np.linalg.norm(samp1-samp2)))
+    return lst
+
+def sample_rux(plx,qxy,n,verbose=True,update_interval=1):
+    '''
+    Consider the polytope of transition matrices q such that plx @ q = plx @ qxy.
+
+    Choose a direction uniformly at random, and find vertex in that polytope
+    which is most extremal along that axis.  
+
+    Repeat n times.
+
+    Returns a list of samples.
+    '''
+
+    lst=[]
+    for i in range(n):
+        if verbose and i%update_interval==0:  # ersatz tqdm replacement (didn't want to include tqdm as a requirement)
+            sys.stdout.write('(%d/%d)'%(i,n))
+            sys.stdout.flush()
+        direc=npr.randn(*qxy.shape)
+        samp1=find_extremal(plx,qxy,direc)
+        lst.append(samp1)
+    return np.array(lst)
+
+def find_extremal(plx,qxy,c):
+    '''
+    Consider the polytope of transition matrices q such that plx @ q = plx @ qxy.
+    Find the most extremal vertex of this polytope in the direction of c.
[…]

### 2018-05-15 paper/paper.tex
added qhat figure

@@
 
+\begin{figure}
+\includegraphics[width=0.8\textwidth]{../images/qhat}
+\caption{The $\hat q$ estimated by the MLM method with $\kappa=.1$.
+Here we find that the calibration is largely diagonal.
+That is, for the most part, if a cell-type was designated as a particular type according to one method, we expect it to be designated by the same type in the other method.
+However, there are some striking deviations from diagonality.
+Note that the number of cell-types is not exactly the same between the two methods, but this provides no barrier to implementin the MLM method.
+The final row and column are in some sense ``miscellaneous'' and can be disregarded to some extent.\label{fig:allenout}}
+\end{figure}
+
 The first step of the Markov Link Method was simply to estimate the calibration $q^*$.

### 2018-05-21 paper/paper.tex
Used full nxx types for the analysis

--- 2018-05-15 paper/paper.tex
@@
 
+\newcommand{\figgygr}[2]{\includegraphics[width=#1\textwidth]{#2}}
+% \newcommand{\figgygr}[2]{picture here}
+\newcommand{\figgy}[1]{\begin{figure}\fbox{\begin{minipage}{\textwidth}#1\end{minipage}}\end{figure}}
+
 \usepackage{cancel}
@@
 \title{Markov Link Method for calibrating without joint measurement, including the case of destructive measurements}
-\author{Jackson Loper}
+\author{Jackson Loper, Osnat Penn, Trygve Bakken, David Blei, Liam Paninski}
 
@@
 \newtheorem{example}{Example}
+\newtheorem{definition}{Definition}
 
@@
 It achieves this without any measurement data from both tools applied to the same ``bodies of water.'' Moreover, MLM makes zero assumptions about what calibrations we might expect to see, instead applying a subpopulation-based conditional independence assumption.
-We evaluate MLM on a pair of single-cell RNA techniques, obtaining precise calibrations between the tools as well as more accurate models for each tool separately.
+We evaluate MLM on a pair of single-cell RNA techniques, obtaining a calibration between the tools.
 \end{abstract}
@@
 The modern setting is rife with experimental measurement tools, and it can be very frustrating to understand how the output of these tools relate to one another.
-This problem is known as ``calibration'' or ``zeroing'' \cite{bipm2008international}.
-A calibration tells us what readings we should expect from one tool, given the reading we obtained from another tool.
-Calibration additionally must give uncertainty bounds for how much we can trust those expectations.
+This problem is known as ``calibration'' or ``zeroing.''  A calibration tells us what readings we should expect from one tool, given the reading we obtained from another tool.
+Calibration additionally must give uncertainty bounds for how much we can trust those expectations  \cite{bipm2008international}.
 Calibration between measurement tools allows us to combine experimental results from different labs and different methodologies into larger scientific theories.
@@
 We will denote it by $q^*(y|x)$.
-As input, this conditional distribution takes the measurement results from one tool on a particular specimen, $x$.
+As input, this conditional distribution takes the measurement result $x$ obtained from one tool on a particular specimen.
 As output, it yields the probability of obtaining result $y$ from a second tool applied to measure the same specimen.
@@
 We call this ``joint measurement.''  Unfortunately, calibrations are often required even when joint measurement is unavailable.
-For example, if the measurement tool alters the specimen being measured, joint measurement is simply impossible.
+For example, if the measurement tool significantly alters the specimen being measured, joint measurement is simply impossible.
 In other cases, it may be expensive or impractical.
@@
 
+\section{Relation to prior work}
+
+Our main goal is to achieve calibration without joint measurement.
+The main obstacle is what is known in the statistics world as an `identifiability problem.'  Due to this problem, we simply \emph{cannot} directly estimate the calibration that we are interested in.
+The calibration is simply not ``identifiable.''  Nonetheless, all is not lost.
+Our main contribution is to show that one can bound the extent of this identifiability problem.
+In short, the fundamental problem cannot generally be vanquished, but we can put it in its place.
+
+Our identifiability analysis stands on the shoulders of a long history of turning probabilistic assumptions into bounds on unidentifiable parameters.
+The core idea of the MLM is to take an assumption (the `Markov Link assumption,' which we will define shortly) and use it to place bounds on an unidentifiable quantity (namely the calibration $q^*$).
+When these bounds are fairly tight on all sides, we see that much can be learned even though what we want isn't identifiable.
+Much of the prior literature in this kind of direction comes from research into causality.
+For example, in \cite{bonet2001instrumentality} Bonet uses polytopes not unlike the ones seen here to explore whether a variable can be used as an instrument.
+The famous Clauser-Horne-Shimony-Holt inequality was designed to help answer causality questions in quantum physics, but it also sheds light on what distributions are consistent with certain assumptions \cite{clauser1969proposed}.
+Indeed the physics literature has contributed many key inequalities (cf.
+\cite{chaves2014inferring}, \cite{kela2017semidefinite}, and the references therein).
+Perhaps the closest work to this one would be \cite{makarov1982estimates}, which used two marginal distributions to get bounds on a property of the joint distribution (namely the distribution of the sum).
+We advance this approach to a more general-purpose technique, both by using many subpopulations to refine our estimates and by considering the entire space of possible joint distributions instead of simply a particular aspect of the joint.
+
+It is perhaps worth mentioning that the actual theorem presented here has probably been derived before (although we did not find it in our review of the literature).
+Other results in the same flavor as the one presented here might also find a practical use for modern problems.
+There is a treasure-trove of ideas in the causality literature; this wealth has not yet been brought fully to bear on the challenging and important problems of calibration for modern experimental modalities.
+Our primary intention with this article is to bring attention to the practical utility of these kinds of result.
+
 \section{The Markov Link Method assumption}
 
-To make these ideas rigorous, let us develop a little bit of notation.
-Let us say we are considering the members of a large population.
-For example, each ``specimen`` in the large population might be a human cell or a piece of metal which needs to be tested.
+To make our ideas rigorous, let us develop a little bit of notation.
+Let us say we are considering the specimens of a large population.
+For example, each specimen in the large population might be a human cell.
+Each specimen could also be be a piece of metal which needs to be tested.
 We will assume there are three basic properties of interest for each specimen $i$:
@@
 \begin{enumerate}
-\item $\ell_i$, the subpopulation or side information.
+\item $\ell_i$, the subpopulation.
 We assume that the overall population can be split into subpopulations of interest.
-For example, we could define subpopulations by looking at cells in different parts of the body, or cells with different sizes.
+For example, we could define subpopulations by looking at cells in different parts of the body or cells with different sizes.
 \item $X_i$, the result of measuring a specimen with tool I.  For example, perhaps tool I takes a picture of the cell with a destructive electron microscopy method.
@@
 Despite this obstacle, we would like to estimate the conditional distribution $q^*(y|x)\triangleq \mathbb{P}(Y_i=y|X_i=x)$.
-Under suitable assumptions, we will see that such estimation is indeed possible using the side information $\ell$.
-However, this estimation suffers from so-called \emph{identifiability problems}, which we will describe in detail in the next section.
-Fortunately, this identifiability error can be quantified, allowing us to determine exactly what our data tells us about the true calibration.
-If the identifiability error is large, the Markov Link method suggests future experiments based on new subpopulations which will hone down on the identifiability problem.
+To make this estimation at all possible, we need some kind of assumption.
 
-\begin{figure}
-\hfill{}\includegraphics[width=.7\textwidth]{../images/bigpicture}\hfill{}
+\figgy{
+\hfill{}\figgygr{.7}{../images/bigpicture}\hfill{}
 \caption{We consider the case that for each specimen there are two ways we might measure it.
@@
 Or what if we had measured the specimens on the right with technique I?
-The Markov Link Method gives a way to answer this problem using a subpopultaion analysis.
-For each specimen, we take $\ell$ to indicate the subpopulation (indicated here by the color of the neuron, i.e.\ yellow or purple), $X$ to indicate the results we would obtain from that specimen if we measured it with method I, and $Y$ to indicate the results we would obtain from that specimen if we meausured it with method II.
+The Markov Link Method gives a way to answer this problem using a subpopulation analysis.
 \label{fig:bigpicture}}
-\end{figure}
+}
 
-The key assumption of the Markov Link Method is a conditional independence assumption: that $\mathbb{P}(Y|X,\ell_i)$ is independent of $\ell$, i.e.
-$\mathbb{P}(Y=y|X=x,\ell)=q^*(y|x)$ for every value of $\ell$.
-Intuitively, this signifies that the manner in which $X$ and $Y$ are related is the same for each subpopulation.
+The key assumption of the Markov Link Method is a conditional independence: that $\mathbb{P}(Y|X,\ell)$ is independent of $\ell$, i.e.
+
+\begin{center}
+\fbox{\parbox[c]{4in}{
+\vspace{.1in}
+\begin{center}
+\textbf{The Markov Link Method Assumption}\\$\mathbb{P}(Y=y|X=x,\ell)=q^*(y|x)$ for every value of $\ell$.
+\end{center}
+\vspace{.1in}
+}}
+\end{center}
+
+Intuitively, this signifies that the manner in which $X$ predicts $Y$ would be the same for each subpopulation.
 If this assumption is not met, then the method presented here is not applicable.
@@
 However, how can we test the reliability of the machine that performs the test?
-In each test run there will be some variability induced by the machine itself, which induces a measurement error  In practice, some kind of assumptions about part homogeneity are used to approximate this error (cf.\ \cite{de2005gauge}).
+In each test run there will be some variability induced by the machine itself, which induces a measurement error.
+In practice, some kind of assumptions about part homogeneity are used to approximate this error (cf.\ \cite{de2005gauge}).
 However, if we have two testing machines we can use the MLM to obtain a calibration between the machines, even though we can never test the same part with both machines.
@@
 Even if the error is correlated, the MLM assumption may still hold.
-For example, imagine that the $Y$ error is correlated with the absolute reliability of the part, this may pose no problem because that reliability is measured by $X$.
-As long as the correlations are the result of something which is measured in $X$, the conditional independence assumption will hold.
+For example, imagine that the $Y$ error is correlated with the absolute reliability of the part, this may pose no problem if that reliability is adequately measured by $X$.
 
@@
 Unfortunately, modern high-resolution cell photography and single-cell sequencing technologies are both destructive.
-As a result, we can't get both kinds of data for the same specimens.
-For cells native to regions full of diverse cell-types, finding a correspondence is a real problem.
+As a result, we can't always get both kinds of data for the same specimens.
+For cells native to regions full of diverse cell-types, it is thus quite hard to grasp the correspondence between these different kinds of classification systems.
 The result is two completely independent classifications of cells, one for each way of looking at the cell.
 MLM allows us to estimate the relationship between those two classification systems, yielding a wholistic understanding of the different types of cells.
-In this case, $\ell$ might indicate some side information such as where in the body the cell was found, $X$ would indicate the classification of the cell according to its transcriptomics, and $Y$ would indicate the classification of a cell according to its morphology.
+In this case, $\ell$ might indicate some side information such as where in the body the cell was found, $X$ would indicate a detailed classification of the cell according to its transcriptomics, and $Y$ would indicate a coarser classification of a cell according to its morphology.
 We expect that cell morphology is largely a function of cell transcriptomics.
@@
 Many experiments have been performed to see how these cancer cells respond to treatment.
-However, if a treatment work on a particular cultured cell-line, what can we say about which kinds of cancer might respond well to that treatment?
+However, if a treatment works on a particular cultured cell-line, what can we say about whether a treatment will work on an actual in-vivo cancer inside a patient?
 Coarse side-information such as original cancer location is often available for both in-vivo and cultured cells, but this is often a surprisingly weak signal.
@@
 \cite{imamura2015comparison,haibe2013inconsistency}).
-Moreover, we have very little joint measurement between the human cells which originally gave rise to a cell culture and the cells that survived to beceome the cell line.
-The Markov Link method can leverage the common side-information together with separate transcriptome information to produce a fine-grained correspondence between in-vivo and cultured cells.
-This correspondence can used to propose new treatments for cancers.
+These cell cultures are subject to quite different pressures, due to the fact that they survive on a plate instead of inside a human being.
+The Markov Link method can leverage the common side-information together with separate transcriptome information to understand the correspondence between in-vivo and cultured cells.
+If a particular drug is effective on a particular cultured cell-line, we can then look at the corresponding in-vivo transcriptomic profile.
+If we find human cancers that match this profile, they might be good candidates for further research using this particular drug.
 Here $\ell$ might indicate cancer location, $X$ might indicate transcriptomic expression of cultured cells, and $Y$ might indicate transcriptomic expression of in-vivo cells.
@@
 Here $\ell$ would indicate coarse labels such as ``cat'' or ``street scene.''  These labels could be derived from either images or text and can be trained in a supervised fashion.
-Then, $X$ would indicate the image and $Y$ would indicate a caption.
-Since the caption should be determined by the picture $X$, the MLM assumption may hold.
+$X$ would indicate the image and $Y$ would indicate a caption.
+If caption is largely determined by the picture $X$, the MLM assumption may hold.
 
@@
 
-The main contributions of this paper are summarized here:
+The rest of the paper proceeds as follows:
 
 \begin{itemize}
-\item Posing of the Markov Link Method assumption as an assumption that may make it possible to perform calibrations without joint measurement.
-\item An analysis of some identifiability problems posed by the Markov Link Method assumption.
-Although it turns out precise identifiability is often impossible, we formulate an asymptotic theory for bounding the magnitude of the this problem.
-In particular, although we often cannot hope to exactly determine the calibration $q^*(y|x)\triangleq\mathbb{P}(Y=y|X=x)$, we can estimate a so-called polytope $\hat \Theta$ which captures our uncertainty about $q^*$ due to identifiability problems.
-\item The result of the identifiability analysis is a polytope $\hat \Theta$.
-We investigate various ways of measuring the extent of this polytope.
-We see how it is possible to understand what this ``extent'' means in terms of the identifiability problem.
-When the extent is small, we see that the identifiability problem is actually quite minor.
-\item We apply the MLM to two single-cell transcriptomic datasets.  The result uncovers an important and unexpected property of one of the experimental methods.  
+\item We describe the identifiability problem and describe how our analysis puts this problem in our place.
+We give several examples attempting to give intuition as to why the identifiability problem may not be as severe as it first appears.
+\item Applying these ideas, we develop the Markov Link Method for calibration without joint measurements.
+\item We apply the MLM to two single-cell transcriptomic methods.
+The result uncovers some ways in which the methods appear to be well-calibrated and other areas where the calibration may not be as good.
 \end{itemize}
 
-Our identifiability analysis stands on the shoulders of a long history of relating probabilistic assumptions to probabilistic inequalities on unidentifiable parameters.
-Indeed, the core idea of this work is to take an assumption (the Markov Link assumption) and use it to place inequality bounds on an unidentifiable quantity (namely $q^*$).
-When inequalities are fairly tight on all sides, we see that much can be learned even when what we want isn't identifiable.
-Much of the prior literature in this kind of direction comes from research into causality.
-For example, in \cite{bonet2001instrumentality} Bonet uses polytopes not unlike the ones seen here to explore whether a variable can be used as an instrument.
-The famous Clauser-Horne-Shimony-Holt inequality was designed to help answer causality questions in quantum physics, but it also sheds light on what distributions are consistent with certain assumptions \cite{clauser1969proposed}.
-Indeed the physics literature has contributed many key inequalities (cf.
-\cite{chaves2014inferring}, \cite{kela2017semidefinite}, and the references therein).
-Perhaps the closest work to this one would be \cite{makarov1982estimates}, which used two marginal distributions to get bounds on a property of the joint distribution (namely the distribution of the sum).
-We advance this approach to a more general-purpose technique, both by using many subpopulations to refine our estimates and by considering the entire space of possible joint distributions instead of simply a particular aspect of the joint.
+\section{The identifiability problem}
 
-The remainder of this work proceeds as follows.
-First we give a careful consideration to the Markov Link Method assumption and how it allows us to use side information to estimate the calibration.
-We also see when this side information will and will not be helpful.
-Next, we outline the Markov Link Method and describe a new asymptotic consistency result concerning the identifiability issues.
-Finally, we consider a real dataset.
-We perform the method and then attempt to criticize the result three ways: through concerns about not having enough data, through concerns about modeling assumptions, and through concerns about identifiability.
-The method largely prevails against these criticisms, and reveals an unexpected feature of the two techniques.
-
-\section{Precise problem formulation and discussion of the identifiability problem}
-
-Let us now be completely rigorous.  We shall make three assumptions:
+Let us make three assumptions:
 
@@
 Of these, we have observed $\ell_i,X_i$ for $i\in\{1\cdots n\}$ and $\ell_i,Y_i$ for $i\in n+1\cdots n+m$.
-\item $X$ and $Y$ are discrete random variables with finite support (the general concepts here will apply more generally, but we leave it for future work).
-If the data is not discrete, we can always make it so by dividing it into suitable bins.
-In this simple case, we may summarize all of this information with two matrices:
-\[
+\item $X$ and $Y$ are discrete random variables with finite support (the general concepts here will apply more generally, but we leave it for future work; if the data is not discrete, we can always make it so by dividing it into suitable bins).
+In this simple case, we may summarize all of our observed data in two matrices:
+\begin{align}
 N^X_{\ell x} = |\{i\leq n:\ \ell_i=\ell,X_i=x\}|  \qquad N^Y_{\ell y} = |\{i>n:\ \ell_i=\ell,Y_i=x\}|
-\]
+\label{eq:twomatrices}
+\end{align}
 Thus $N^X,N^Y$ are matrices counting the number of each kind of observation for method I and method II respectively. 
@@
 
-\begin{description}
-\item[Not-enough-data problems] We don't have infinite data, so we can't hope to exactly determine $p^*,h^*$.
-\item[Identifiability problems] Even if we knew $p^*,h^*$ exactly, it is often impossible to know the value of $q^*$.  We can only ever know that it lies somewhere in $\Theta(p^*,h^*)$.  
-\end{description}
+\begin{enumerate}
+\item Not-enough-data problems. We don't have infinite data, so we can't know the exact values of $p^*,h^*$.
+\item Identifiability problems. Even if we knew $p^*,h^*$ exactly, it is often impossible to know the value of $q^*$.  We can only ever know that it lies somewhere in $\Theta(p^*,h^*)$.  
+\end{enumerate}
+
+This point of view suggests a practical procedure for estimating $q^*$:
+
+\begin{enumerate}
+    \item Use traditional methods to develop estimators $\hat h,\hat p$ so that with more and more data we can guarantee that $\hat h\approx h^*$ and $\hat p\approx p^*$ with high probability.
+In particular, we will use a robust pseudocount estimator for $p^*$ and use the MLM assumption to get a high-quality estimator for $h^*$ that leverages information from $N^X$ and $N^Y$.
+    \item Then we may suppose that $q^*$ is probably somewhere \emph{near} to $\Theta(\hat p,\hat h)$.  
+\end{enumerate}
+
+It is not completely obvious that this will work.
+Certainly $\hat p\approx p^*$ and $\hat h\approx h^*$.
+Certainly $q^*\in \Theta(p^*,h^*)$.
+But does that guarantee that $q^*$ is nearby to the set $\Theta(\hat p,\hat h)$?
+Something could go quite wrong in this line of reasoning.
+Fortunately, subject to mild assumptions, one can show that nothing goes wrong.
+This is the content of our main technical contribution.
+It can be found in Appendix \ref{sec:proof}.
 
 Let us now take a moment to understand the set $\Theta$.
-This set tells us how the things we can estimate (i.e.\ $p^*,h^*$) inform us about what we want to know (i.e.\ $q^*$).
+This set encodes exactly how it is that the Markov Link assumption enables us to understand something about the calibration $q^*$.
+That is, this set tells us how the things we can estimate (i.e.\ $p^*,h^*$) inform us about what we want to know (i.e.\ $q^*$).
 To understand a bit more concretely how this works, it may be helpful to think of $p^*,h^*,q^*$ as matrices.
 From this point of view, one aspect of the definition of $\Theta$ can be written more concisely as a matrix equality constraint on $q^*$: if $q \in \Theta(p^*,h^*)$, then $p^*q=h^*$.
-The consequences of this equation depend upon whether $p^*$ has a left-pseudoinverse.
+This equation is in fact a literal mathematical rendering of the Markov Link assumption of conditional independence.
+The consequences of this assumption and its corresponding equation depend upon whether $p^*$ has a left-pseudoinverse.
 \begin{itemize}
@@
 In this paper we will focus almost entirely on the second case.
-In this second case there is a genuine identifiability issue;  it is \emph{impossible} to ever determine the true value of $q^*$, regardless of how much data we have.
-It therefore becomes of paramount importance to understand the magnitude of this identifiability problem.
-In particular, even though we cannot know $q^*$ exactly, can we place bounds on what $q^*$ could be?
+In this second case there is a genuine identifiability issue;  $\Theta(p^*,h^*)$ is nontrivial and so it is \emph{impossible} to ever determine the true value of $q^*$, regardless of how much data we have.
+All we have is the equality and inequality bounds that arise from our knowledge that $q^* \in \Theta(p^*,h^*)$.
 
-To gain intuition, let us consider a few examples.\vspace{.1in}
+To gain intuition about the equality and inequality bounds that define the set $\Theta$, let us consider a few examples.\vspace{.1in}
 
@@
 \]
-This immediately tells us something about $q^*$.
-For example we now know that $q^*_{1,1} = 10(.4 - .1 q^*_{2,1} - 0.8 q^*_{3,1})$.
-But it actually tells us much more than that.
-Observe that $0\leq q^*_{xy}\leq 1$ for every $x,y$; in other words, probabilities must be positive by definition.
-Thus, in particular,
+We thereby obtain an important constraint on what the calibration $q^*$ might be.
+
+This equation also takes on additional significance in light of the inequality bounds we have: $0\leq q^*_{xy}\leq 1$ for every $x,y$.
+In other words, probabilities must be between 0 and 1, by definition.
+This may seem trivial, but it has important consequences in light of the constraints imposed by the equation $p^*q=h^*$.
+In particular, taking another look at that same equation, we see that
 \[
@@
 
-\begin{figure}
-\hfill{}\includegraphics[width=.7\textwidth]{../images/polytope}\hfill{}
+\figgy{
+\hfill{}\figgygr{.7}{../images/polytope}\hfill{}
 \caption{We consider the simple case $\ell=1$, $X\in \{1,2\}$, $Y\in \{1,2\}$.
-Then the space of values of $q$ which are valid probability distributions can be understood as a box.
+In this case the space of values of the calibration $q$ which are valid probability distributions can be understood as a box.
 The space of values of $q$ which satisfy the matrix equation $p^* q = h^*$ can be understood as a line that passes through that box.
-The set of possible values of $q$ which are consistent with the matrix equation and are also valid probability distributions forms the set $\Theta$, a line segment.
+The set of possible values of $q$ which are consistent with the matrix equation and are also valid probability distributions forms the set $\Theta$, which in this case can be identified with a line segment.
 We know that the true value of $q^*$ must lie somewhere inside this $\Theta$.
-Depending upon the exact values of $p^*,h^*$, this line segment may be larger or smaller.
+Depending upon the exact values of $p^*,h^*$, the corresponding line segment may be larger or smaller.
 For example, if the line crosses close to a corner of the box, then $\Theta$ may be restricted to a very small region near that corner.
 \label{fig:polytopeexample}}
-\end{figure}
+}
 
@@
 Finally, let us consider one class of examples in which the entire situation can be visualized: $\ell=1$, $X\in \{1,2\}$, $Y\in \{1,2\}$.
-Let $\theta^*_1 = q_^*{1,1}$ and $\theta^*_2 = q_^*{2,1}$.
-Notice that in this simple case $\mathbb{P}(Y=2|\ell) = 1- \mathbb{P}(Y=1|\ell)$, so the two values $\theta_^*1,\theta_2^*$ completely define $q^*$.
-The matrix equation $p^* q^* = h^*$ corresponds to a linear constraint on $\theta^*$.
-On the other hand, the fact that $q^*$ must be a valid probability distribution tells us that $\theta_1^*,\theta_2^*$ must lie inside the box $[0,1]\times[0,1]$.
+Notice that $q$ is completely defined by $\theta_1=\mathbb{P}(Y=1|x=1)$ and $\theta_2=\mathbb{P}(Y=1|x=2)$.
+The matrix equation $p^* q^* = h^*$ corresponds to a linear constraint on $\theta_1,\theta_2$, namely $p^*_{11}\theta_1 + p^*_{12}\theta_2 = h^*_{11}$.
+On the other hand, the fact that $q^*$ must be a valid probability distribution tells us that $\theta_1,\theta_2$ must lie inside the box $[0,1]\times[0,1]$.
 Combining the matrix equation with the constraints of the box, we learn that $q^*$ lies inside a certain bounded, one-dimensional space.
 This space is precisely $\Theta$.
-The length of this line segment depends on the exact values of $p^*,h^*$.
+It can be identified with a line segment, and the depends on the exact values of $p^*,h^*$.
 In some cases it is quite large, i.e.\ it will be difficult to know much about the true value of $q^*$.
@@
 See Figure \ref{fig:polytopeexample} to visualize this.
-In general, this $\Theta$ is much higher-dimensional and harder to visualize, but in this simple case we can see it clearly.
+In real-world cases this $\Theta$ is much higher-dimensional and harder to visualize, but in this simple case we can see it clearly.
 \end{example}
@@
 
-Let $N^X,N^Y$ denote matrices describing the original observed data.
+Let us now exactly describe our approach for calibrating in the absence of joint measurement.
+As in Equation (\ref{eq:twomatrices}) above, we assume that $X$ and $Y$ are discrete random variables and $N^X,N^Y$ are the matrices describing the  observed data.
 The purpose of the Markov Link Method is to try to use these matrices to estimate $q^*(y|x)=\mathbb{P}(Y=y|X=x)$, the calibration between $X$ and $Y$.
-The Markov Link Method proceeds in two main steps:
+The Markov Link Method has three main components:
 
 \begin{description}
-\item[Estimation]  Estimate $p^*$ with robust pseudocount estimator:
-\[
-\hat p(x|\ell) \triangleq \frac{1+N^X_{\ell x}}{\sum_{x'} N^X_{\ell x'}}
-\]
+\item[Estimation]  Estimate $p^*$ with robust pseudocount estimator (cf.\ \cite{flach2012machine}):
+\begin{align}
+\hat p(x|\ell) \triangleq \frac{1+N^X_{\ell x}}{\sum_{x'} N^X_{\ell x'}} \label{eq:pseudocountp}
+\end{align}
 and then calculate a possible guess for $q^*$, namely
-\[
+\begin{align}
 \hat q = \argmax_q \sum_{\ell,y}N_{\ell y}\log\left(\sum_{x}\hat p(x|\ell)q(y|x)\right) + \kappa \sum_{xy} \log q(y|x)
-\]
-The entropic regularizer $\kappa$ makes it possible to obtain a unique estimate for $q^*$, despite the identifiability problem discussed above.
-In practice, we took $\kappa$ to be any small constant less than 1.
-The optimization problem which defines $\hat q$ is convex and easily solved.
+\label{eq:mlmq}
+\end{align}
+The Dirichlet prior regularizer $\kappa$ ensures this optimization problem has a unique solution; indeed, one may easily verify that the objective is strictly convex.
+In practice, we took $\kappa$ to be any small constant.
+We use a majorization-minimization approach to solve this optimization problem.
+That is, we first take an initial guess for $\hat q$.
+We then define $\tilde \pi_{\ell,x,y} \gets \hat p(x|\ell)q(y|x) / \sum_{x'} \hat p(x'|\ell)q(y|x')$.
+We then update $\hat q$ by  $\hat q \gets \argmax_q \sum_{\ell,y,x}N_{\ell y}\tilde \pi_{\ell,x,y}\log\left(\hat p(x|\ell)q(y|x)\right) + \kappa \sum_{xy} \log q(y|x)$, which can be solved in closed form.
+We then iterate this procedure, redefining $\tilde \pi$ and then redefining $\hat q$ in each iteration.
+It is well-known that since the overall problem is convex we can guarantee that this procedure converges to the global optima (cf.\ \cite{wu1983convergence} for a detailed exposition).
+In practice we found 5000 iterations to be sufficient for convergence.
 
-Note that although this gives us a unique estimate for $q^*$, the identifiability problems have not magically vanished.
-There is good reason to suppose that $\hat q$ is near to the set $\Theta(p^*,h^*)$, but if the set $\Theta$ is very large then it may be that $\hat q$ is very different from the true $q^*$.
-Therefore, there is a second step we take: a criticism step.
-This step should be taken for any model, but one which is particularly important in this case.
-\item[Criticism]  Level criticisms at the learned model to see whether $\hat q$ can be trusted.  
-\begin{enumerate}
-\item Use bootstrap to determine if we have enough data.
-\item Use a held-out log-likelihood method to make basic sanity checks as to whether the MLM assumption fits with the data.
-\item Measure the extent of $\Theta(\hat p,\hat p \hat q)$ to see whether we have significant identifiability issues.
-\end{enumerate}
+This yields a high-quality, consistent estimator for $h^*(y|\ell)\triangleq \mathbb{P}(Y=y|\ell)$, namely $\sum_x \hat p(x|\ell) \hat q(y|x)$ (cf.\ the lemma in Appendix \ref{sec:proof} for a simple consistency proof).
+It also gives us a unique estimate for $q^*$, although this estimate may suffer from errors due to incorrect assumptions, lack of data, or identifiability.
+Therefore, a vital part of the MLM is a criticism step.
+\item[Model criticism]  The MLM assumption cannot be checked directly.
+However, we can check certain aspects of it.
+In particular, the MLM assumption suggests a natural estimator for the distribution .
+On the other hand, a more traditional approach would etimate $h^*$ from $N^Y$ using a simple pseudocount method.
+We can compare both estimators for $h^*$ using held-out log-likelihood of $Y$ data.
+If the MLM estimate is considerably worse than the pseudocount method, it may indicate that the MLM assumption doesn't fit with the data.
+\item[Uncertainty characterization] Measure our uncertainy about $q^*$ using \emph{Bootstrapped Rotationally Uniform eXtremal Paired (BRUXP)} samples and \emph{Rotationally Uniform eXtremal Paired (BRUXP)}.
+The variability in these samples capture error both due to insufficient data and due to identifiability issues.
 \end{description}
 
-The very final point merits a little bit more discussion.
-In the previous section, we saw that the identifiability issues could be summed up as follows: we can't generally tell what $q^*$ is, only that it lies somewhere inside the set $\Theta(p^*,h^*)$.
-The extent of the identifiability problems can therefore be understood in terms of the extent of this set.
-However, in the method above, we look at the extent of a different set: $\Theta(\hat p,\hat p \hat q)$.
-This set is based on mere estimates: $p^* \approx \hat p$ and $h^* \approx \hat p \hat q$.
-It is not immediately obvious that the size of this \emph{estimated} set would give us an indication of the size of the \emph{true} set of interest, $\Theta(p^*,h^*)$.
-Fortunately, it is straightforward to show that under mild conditions this is indeed the case.
-This is the content of the main analytical theorem of this paper, for which a precise statement may be found in Appendix \ref{sec:proof}.
-It is in part due to the structure of this proof that we choose to estimate $h^*$ with $\hat p \hat q$ instead of a more traditional estimator.
+We now describe this very last step in more detail:
+
+\vspace{.2in}
+\begin{definition}
+Let $d$ denote a random direction in calibration space.
+That is, let $d$ denote a matrix of the same dimensions as the calibration $q$, each entry of which is sampled from a random normal distribution.
+Sample with replacement to produce a surrogate dataset $N^{X,(1)},N^{Y,(1)}$, estimate $\hat p^{(1)},\hat q^{(1)}$ using the surrogate dataset and the method outlined above, and define $\hat \Theta^{(1)} \triangleq \Theta(\hat p^{(1)},\hat p^{(1)}\hat q^{(1)})$.
+Repeat this process to obtain $\hat \Theta^{(2)} \triangleq \Theta(\hat p^{(2)},\hat p^{(2)}\hat q^{(2)})$.
+Now, let us take $\tilde q^{(1)}$ to be the most extremal vertex of $\hat \Theta^{(1)}$ in the direction $d$ and $\tilde q^{(2)}$ to be the most extremal vertex of $\hat \Theta^{(2)}$ in the direction $-d$.
+The tuple $d,\tilde q^{(1)},\tilde q^{(2)}$ is said to be a \textbf{Bootstrapped Rotationally Uniform eXtremal Paired (BRUXP)} sample.
+If we adopt this same procedure without the resampling-with-replacement aspect, we will refer to it as \textbf{Rotationally Uniform eXtremal Paired (RUXP)}.
+\end{definition}
+
+For each BRUXP sample, the pair $\tilde q^{(1)},\tilde q^{(2)}$ gives for the range of what $q^*$ might be along the direction $d$.
+The variability amongst many such samples indicates our uncertainty about the true $\hat q$.
+
+We cannot make general claims about the bootstrap in this case, due to the usual difficulties with bootstrap and extremal statistics (cf.\ \cite{andrews2000inconsistency}).
+However, we do at least know that this procedure correctly handles the identifiability issue in the asymptotic limit.
+In particular, subject to mild reguarity conditions, we can show that $\inf_{q\in \Theta(\hat p, \hat p \hat q)} |q-q^*|\rightarrow 0$ as we gather sufficient data.
+That is, with sufficient data BRUXP samples are guaranteed to represent the uncertainty we face from identifiability concerns.
+An exact statement may be found in Appendix \ref{sec:proof}.
+
+The the variation amongst the BRUXP samples is necessarily due to a combination of effects from insufficient data and effects from the identifiability issues.
+To separate issues, we recommend that one also look at Rotationally Uniform eXtremal Paired (RUXP) samples (i.e.\ BRUXP samples taken without using the sampling-with-replacement procedure).
+The variability among these samples corresponds primarily to uncertainty due to identifiability concerns, and ignores variability due to insufficient data.
+
+Code for this procedure is available at https://github.com/jacksonloper/markov-link-method.
+All calculations in this paper can be replicated by going through a tutorial-style ipython notebook that can be found at the same address.
 
@@
 
-\begin{figure}
-\hfill{}
-\includegraphics[width=.5\textheight]{../images/plxy}
-\hfill{}
-
+\figgy{
+\begin{minipage}[c]{0.15\textwidth}
+\figgygr{.15}{../images/Nlx}
+\end{minipage}\hfill
+\begin{minipage}[c]{0.15\textwidth}
+\figgygr{.15}{../images/Nly}
+\end{minipage}\hfill
+\begin{minipage}[c]{0.3\textwidth}
 \caption{Input: two tables.
@@
 Each technique pulls out a different group of cells.
-Once the cells were selected, they were then either subjected to technique I technique II (`patch').
+Once the cells were selected, they were then either subjected to technique I or technique II.
 Technique I gives a very complete analysis of the gene expression of the cell.
@@
 These are shown above.
-Each column corresponds to a different cell-type.
-Each row corresponds to a different subpopulation.
-Each entry gives the proportion of a particular cell-type found within a particular population, according to a particular technique.
-If the two methods were perfectly calibrated, we would expect the two tables to look identical.
+Each row corresponds to a different cell-type.
+Each column corresponds to a different subpopulation.
+Each entry gives the number of a particular cell-type found within a particular population, according to a particular technique.
 \label{fig:alleninput}}
-\end{figure}
+\end{minipage}\hfill
+}
 
@@
 Best efforts were made to use biological intuition to calibrate the methods.
-For example, if a cell was designated as ``Lamp5 Egln3\_1'' celltype using one method, the hope was that it would also be given the same designation if it was processed using the other method.
-The two methods were designed by scientists to achieve this goal.
+For example, both methods have a notion of a ``Lamp5 Egln3\_1'' cell-type.
+If a cell was designated as ``Lamp5 Egln3\_1'' celltype using one method, the hope was that it would also be given the same designation if it was processed using the other method.
+The two methods were designed to achieve this goal.
 However, each method has its own biases and errors, and it was not obvious whether this effort was successful.
@@
 Perhaps not surprisingly, it was found that the distribution of cell-types appeared different under the two modalities.
-In fact, in one of the methods some of the cells were designated as ``unknown,'' so even the set of cell-types was not the same between the two groups.
+Indeed, it was found necessary in method II to designate some cells with certain ``nxx'' types that indicated different kinds of uncertainty about the cell-type.
+Thus even the set of cell-types was not the same between the two groups.
 Clearly the methods were not perfectly calibrated -- but how big of a problem was it?
@@
 
-The first step of the Markov Link Method was simply to estimate the calibration $q^*$.
-Taking $\kappa=.1$, we obtained a calibration found in Figure \ref{fig:allenout}.
-While this result seemed to generally suggest the methods were well-calibrated, there are some striking divergences.
-For example, this calibration suggests a that celltypes designated as celltype ``10'' in one method are actually quite likely to be designated as celltype ``7'' in the other technique.
-Both celltypes are associated with similar neuronal types with high expression of somatostatin.
-However, type ``10'' is associated with genes such as `C1ql3,' `Chodl,'  and `Nts,' whereas type ``7'' is associated with `Etv1' and `Myh8.'
+\figgy{
+\figgygr{0.8}{../images/qhat}
+\caption{The $\hat q$ estimated by the MLM method with $\kappa=.01$.
+Notice the mass along a somewhat diagonally-oriented slice of the calibration.
+This corresponds to cell-types which appear to be designated with the same cell-type by both methods.
+The large region on the right corresponds to the ``nxx'' types in Tool II; these indicate different forms of uncertainty about the type of the cell.\label{fig:allenout}}
+}
 
-Notice that there are several fundamentally different reasons we could see these kinds of miscalibrations:
+The first step of the Markov Link Method is to estimate the calibration with the estimator $\hat q$ according to equations (\ref{eq:pseudocountp}) and (\ref{eq:mlmq}).
+Taking $\kappa=.01$, we obtained a calibration found in Figure \ref{fig:allenout}.
+From this, it appears that there are many cells which would be identified with the same cell type in either method.
+For example, according to $\hat q$, if a cell is classified as type ``VISp L5 Endou'' by tool I, it appears there is a 91\% chance it will be classified the same way by tool II.
+However, there are other cell-types which seem to have more ambiguity.
+Lest we make too many conclusions without considering the uncertainty in our estimates of $q^*$, let us proceed to the criticism phase of the MLM.
+
+\subsection{Model criticism}
+
+The Markov Link Method assumption suggests a natural estimator for the distribution of tool II results ($Y$) under each subpopulation ($\ell$).
+We have denoted this distribution $h^*$, and the MLM estimates this distribution with $h^*(y|\ell)\approx \sum_x \hat p(x|\ell) \hat q(y|\ell)$.
+A more traditional method would be to estimate $h^*$ using a pseudocount method (cf.\ \cite{flach2012machine}).
+We can compare these methods using held-out log likelihood.
+That is, we train both methods using a subset of our data.
+Then we compute the negative log likelihood of the held out data under both models.
+This indicates how concisely the model can represent the held out data.
+If the MLM method is much worse, it may be a sign that the MLM assumption is poor.
+Smaller values are better:
 
 \begin{itemize}
-    \item The methods are actually miscalibrated.
-    \item The subpopulations were unevenly sampled.
-It may be that certain techniques tend to cause cell-death in certain cell-types more than others.
-These cells are then discarded from the data.
-If this process is different between the two techniques, it could have an important impact on the calibration estimated here.
-    \item The calibration was poorly estimated.
+    \item Traditional estimator: 3.96 nats per cell
+    \item MLM estimator: 3.70 nats per cell
 \end{itemize}
 
-It is to the third point we now turn.
+This is encouraging.
+Our model is as good at predicting held-out data as a more traditional pseudocount estimator (which makes no assumptions on the conditional independence structure of the data).
+In fact, the MLM estimate is able to outperform the traditional estimate, presumably because it can leverage the data in $N^X$ to help us learn about the distributions on $Y$.
+At the very least, it is not a clear sign that the MLM assumption is inappropriate for this data.
 
-\subsection{Criticism}
+\subsection{Uncertainty characterization}
 
-TODO:
-\begin{enumerate}
-\item Nonparametric bootstrap (avg distance 0.66, Figure \ref{fig:nonparametricbootstrap})
-\item Parametric bootstrap (avg distance .74, Figure \ref{fig:parametricbootstrap})
-\item Held out likelihood (2.21 nats vs our method 2.14 nats)
-\item RUX samples (avg distance .16, Figure \ref{fig:ruxsamples})
-\item Uniform samples (avg distance .02, Figure \ref{fig:uniformsamples})
-\end{enumerate}
+We obtained BRUXP samples and RUXP samples for this dataset.  These can be used to understand our uncertainty about the calibration $q^*$ in an number of ways.
 
-\begin{figure}
-\includegraphics[width=0.8\textwidth]{../images/nonparametricbootstrap}
-\caption{Nonparametric bootstrap.\label{fig:nonparametricbootstrap}}
-\end{figure}
+\figgy{
+\figgygr{0.8}{../images/ruxquant}
+\caption{Here we plot a histogram of $|\tilde q^{(1)}_i- \tilde q^{(2)}_i|$ among fifty BRUXP samples $\{d_i,\tilde q^{(1)}_i,\tilde q^{(2)}_i\}_{i\in 1\cdots 50}$.
+Each distance estimates the extent of our uncertainty about the calibration $q$ along the direction $d$.
+We can compare this with the extent of the uncertainty we had before we applied our method, i.e.
+the diameter of the set of probability matrices with these dimensions.
+In this case the diameter of that set is exactly 14.
+In this sense, the MLM has enabled us to roughly half our uncertainty about the calibration.
+\label{fig:ruxquant}}
+}
 
-\begin{figure}
-\includegraphics[width=0.8\textwidth]{../images/parametricbootstrap}
-\caption{Parametric bootstrap.\label{fig:parametricbootstrap}}
-\end{figure}
+The distances between the paired vector in a BRUXP sample give us one kind of insight into the extent of our uncertainty.
+Recall that for each direction $d$ the BRUXP sample comprises the extremal vertex in the direction $d$ as well as the extremal vertex in the direction $-d$.
+The distance between these two vertices gives a kind of measure for the extent of our uncertainty.
+We plot a histogram of these distances in Figure \ref{fig:ruxquant}.
+This figure suggest that this extent is around 7 (although we caution that general results on polytope diameter estimation are not encouraging, cf.
+\citep{brieden1998approximation}).
+By contrast, the diameter of the entire space of valid probability matrices is $14$.
+In this sense, the Markov Link Method enables us to half our uncertainty about the calibration.
 
-\begin{figure}
-\includegraphics[width=0.8\textwidth]{../images/unif}
-\caption{RUX samples.\label{fig:ruxsamples}}
-\end{figure}
+\figgy{
+\figgygr{0.6}{../images/VipRspo4}
+\caption{Each column represents a possible manner in which method I and method II are calibrated for cells which are designated as type `Vip Rspo4' by method I.
+Together, the columns form a notion of our uncertainty about this aspect of the calibration.
+This plot shows that cells designated as `Vip Rspo4' by method I may be designated by method II as taking on various `Vip'-related celltypes and or some of the `nxx' celltypes.
+We cannot be sure which of these columns more faithfully represents the true calibration between the methods.
+\label{fig:ruxqual}}
+}
 
-\begin{figure}
-\includegraphics[width=0.8\textwidth]{../images/rux}
-\caption{Uniform samples.\label{fig:uniformsamples}}
-\end{figure}
+Qualitatively, we can inspect the BRUXP samples by focusing on a single value of $X$ at a time.
+For example, in Figure \ref{fig:ruxqual} we focus on the distribution of $Y$ given that method I measures a specimen as having the type `Vip Rspo4.'   In this figure we look at various BRUXP samples, lining up all the pairs into a single heatmap.
+We see that cells designated as type `Vip Rspo4' by method I may be associated with one of the `Vip'-related celltypes in method II or an `n53' or `n57' type (recall that these types indicate a group of possible cell types, of which `Vip Rspo4' is one).
+For example, it may be that method II would identify such cells as `Vip Htr1f' cells.
+This is potentially concerning.
+On the other hand, the plot above suggests that these cells are \emph{not} ever identified as `Vip Lmol\_1' cells.
+That is, method II is able to accurately discern between cells designated as `Vip Rspo4' by method I and cells designated as `Vip Lmol\_1' by method I.
+This is encouraging.
+To further resolve the various remaining ambiguities in the calibration, we would need to perform additional experiments involving a subpopulation that might be likely to include `Vip Rspo4' cells but no `Vip Htr1f' cells.
+
+Let us next look at the BRUXP samples for cells designated by tool I as having the `VISp L5 Endou' type; these samples are fairly unequivocal.
+Figure \ref{fig:ruxqual2} gives strong support for the notion that the methods are very well-calibrated with respect to this cell-type.
+In all of the BRUXP samples we see that a cell designated as `VISp L5 Endou' by method I would be likely to be designated with the same type with method II.
+
+\figgy{
+\figgygr{0.6}{../images/Endou}
+\caption{Each column represents a possible manner in which the methods are calibrated for cells designated by method I as being type 'VISp L5 Endou.'  Notice that in every column there is strong signal at the corresponding type in method II.
+This suggests a good calibration.
+\label{fig:ruxqual2}}
+}
+
+In the interests of trying to disentangle our uncertainty due to insufficient data from our uncertainty due to identifiability, we repeat the histogram-of-distances analysis on RUXP samples.
+Recall that RUXP samples are just like BRUXP samples, only we do not perform the resampling with replacement.
+This RUXP histogram gives us a lower bound on the diameter of the set $\Theta(\hat p,\hat q)$.
+We expect this diameter to be smaller than the extent we measured in the first analysis: the RUXP analysis only captures uncertainty due to identifiability whereas the BRUXP analysis also captures our uncertainty due to insufficient data.
+This is indeed what we find.
+Figure \ref{fig:ruxquant2} shows the result of the histogram-of-distances analysis on the RUXP samples.
+We see distances are generally closer to 6 and never as high as 7.
+Recall that 7 was the typical distance found in the original BRUXP analysis.
+
+\figgy{
+\figgygr{0.6}{../images/ruxquant2}
+\caption{Here we plot a histogram of $|\tilde q^{(1)}_i- \tilde q^{(2)}_i|$ among fifty RUXP samples $\{d_i,\tilde q^{(1)}_i,\tilde q^{(2)}_i\}_{i\in 1\cdots 50}$.
+Each distance estimates the extent of our uncertainty about the calibration $q$ along the direction $d$ that is due to identifiability issues alone.
+These distances do not measure the extent of our uncertainty due to insufficient data.
+We can compare this with the kinds of distances we saw in Figure \label{fig:ruxquant2}, which were generally larger because they measured uncertainty due to both insufficient data and identifiability issues.
+\label{fig:ruxquant2}}
+}
+
+With more samples, we expect that we could get our original BRUXP analysis in Figure \ref{fig:ruxquant} to look more like the RUXP analysis in Figure \ref{fig:ruxquant2}.
+However, it appears that more samples might not be sufficient to bring the diameter of our uncertainty below 6.
+For that, we would need different \emph{kinds} of experiments, using different subpopulations.
+For example, as we saw above, we could reduce some of the identifiability concerns with experiments using a new subpopulation that included `Vip Rspo4' cells but not `Vip Htr1f.'
+
+Finally, in this special case there is an additional analysis that is appropriate.
+The team producing method I and method II designed these methods to have corresponding types.
+The hope was that a cell designated as `Vip Rspo4' in method I would also be designated as `Vip Rspo4' in method II.
+As we have seen above, we cannot be exactly sure what the calibration is.
+Is it perhaps possible that the calibration is nearly perfect?
+We can place a kind of upper bound on the precision of the calibration by looking for the \emph{best possible} calibration that is consistent with the data we have observed.
+In particular, let $d$ denote the direction of ``perfect calibration'' in which method I and method II agree perfectly.
+We can find the value of $q \in \Theta(\hat p,\hat p \hat q)$ which is most extreme in this direction.
+
+\figgy{
+\figgygr{.8}{../images/bestq}
+\caption{The value of $q$ that yields the most precise correspondence between the two cell types while remaining consistent with the observed data.
+This gives a kind of upper bound on how well the methods may be calibrated.
+As we have seen throughout, there is an extent to which we simply cannot be sure what the calibration is.
+Within all this uncertainty, this figure shows something like the rosiest possible picture of how well the methods might be calibrated.
+The strong diagonal components in this calibration show a good correspondence, i.e.
+cells designated as a particular type under method I are likely to be associated with the same type in method II.
+We caution that there is no reason to suppose that this \emph{is} the true calibration, but this figure shows that the data does not rule out that possibility.\label{fig:bestq}}
+}
+
+We show this `best possible $q$' in Figure \ref{fig:bestq}.
+The result is very encouraging: this best $q$ suggests a strong calibration between the two methods.
+We emphasize that our method does not prove this calibration by any means; as we have seen, the data only allow us to know that the calibration up to a considerable degree of uncertainty.
+However, we can conclude that the observed data does not \emph{rule out} the possibility that the methods are actually quite well calibrated on the whole.
+Of course, even for this best $q$ we do see certain miscalibrations that might be worth further investigation.
+For example, it appears that cells designated as `Vip Serpinf1' by method I might be designated as type `Lamp5 Slc35d3' by method II.
 
@@
 Here we show that a simple Markov assumption can make it possible to actually learn quite a lot.
-Although the exact relationship may not be identifiable, a polytope of possible relationships can be identified, and this polytope may in fact be quite small indeed.
-By exploring this polytope, we can understand what we know -- and what we don't know -- about the relationship between measurements.
+Although the exact relationship may not be identifiable, we can rigorously bound our uncertainty.
+We have proposed the Markov Link Method as procedure to estimate the calibration and understand our uncertainty regarding that estimate.
+We investigated a real-world calibration problem; the MLM gave bounds on the accuracy of the calibration and also the suggested directions for future experiments to further refine our uncertainty about this calibration.
+Code is published at https://github.com/jacksonloper/markov-link-method, including a tutorial-style ipython notebook detailing every calculation used in this paper.
 
-The Markov assumption is of course not the only one that we could have used, and may not be valid in every case.
-For example, it has been speculated that some cell types tend to die more often in one experimental modality than another, and these cells are not part of the data.
-This violates our assumptions.
+The Markov assumption is of course not the only one that we could have used, and may not be valid in every case; future work may be to investigate others.
+For example, it has been speculated that some cell types tend to die more often in one experimental modality than another, and these cells are simply excised from the data without comment.
+This would violate our assumptions.
 However, assuming this death rate can be roughly measured, it can be adjusted for, yielding a different but equally meaningful assumption about the data.
-Moreover, if this method yields bizarre results, it may give useful clues as to exactly how cell death may happen differently in the two modalities.
+Indeed, if the MLM method gives insensible results, this could actually serve as a useful clue that this disproportionate cell death is happening.
+
+Another potential direction for future work is suggested by the `nxx' types found in method II of our empirical investigation.
+These types are meant to indicate some uncertainty about the exact cell type.
+Each one is associated with a family of cell types.
+Our analysis did not need to rely on this information to proceed.
+However, making better use of it could considerably improve our results.
+Developing a rigorous framework to incorporate these kinds of structured data would be a worthy endeavor.
 
@@
 Although we may not be able to learn exactly what we want, we can learn a set of possibilities.
-By probing this set carefully with uniform samplers and extremal tests, we can learn what the data actually has to say and what experiments we need to do to learn more.
+By probing this set carefully, we can learn what the data actually has to say and what experiments we need to do to learn more.
 
@@
 \appendix
-
-\section{Dikin sampler}
-
-\label{sec:dikin}
-
-Consider a convex polytope $T=\{x:\ Ax\leq b\}$.
-We have implemented a method for sampling from this polytope, based on the paper \citep{kannan2012random}.
-This method makes use of the Dikin ellipsoids, $E(x)$.
-For any $x$, these are defined by
-
-\begin{itemize}
-    \item Computing the distance from $x$ to each facet of the polytope, i.e. $d_i = b_i - \sum_j A_{ij} X_j$.
-    \item Constructing $\tilde A$ as $\tilde A_{ij} = A_{ij} / d_i$.
-    \item Define $E(x) = \{y:\ |\tilde A (X-y)|\leq 1\}$.  
-\end{itemize}
-
-We can use these ellpsoids to efficiently sample the polytope $T$.
-At each step, we have some point $X\in T$, and we would to use this point to obtain a new sample $Y$, such that by iterating this process we asymptotically obtain samples which are uniform in $T$.
-Here is how we use $X$ to get $Y$:
-
-\begin{algorithm}[H]
- \KwData{A point $X\in T$}
- \KwResult{A point $Y\in T$}
- \vspace{.1in}
-  Sample a proposal $\tilde Y$, uniformly from $E(X)$\;
-    \eIf{$X \in E(\tilde Y)$}{
-        Sample $U \sim \mathrm{Uniform}[0,1]$\;
-        \eIf{$U \leq \mathrm{Vol}(E(X))/\mathrm{Vol}(E(\tilde Y)) \leq 1$}{
-            Let $Y \gets \tilde Y$\;
-        }{
-            Let $Y \gets X$\;
-        }
-    }{
-        Let $Y\gets X$\;
-    }
- \caption{Dikin sampler step}
-\end{algorithm}
-
-It is easy to show that the stationary distribution of the Markov chain found by iterating these Dikin sampler steps is indeed uniform on $T$.
-To ensure an numerically robust method in the face of high-dimensional and nearly degenerate matrices, we take the following approach to robustly sampling from the ellipsoid:
-
-\begin{algorithm}[H]
- \KwData{An $n\times m$ matrix $\tilde A$}
- \KwResult{A point $X$ sampled uniformly from $\{x:\ |Ax|\leq 1\}$}
- \vspace{.1in}
-  Sample $Z$ as an $n$-dimensional normal variables vector\;
-  Let $X$ denote the solution to the least squares problem $\min_x|\tilde Ax - Z|$\;
-  Normalize $X$ by $X \gets X/|\tilde AX|$\;
-  Sample $U \sim \mathrm{Uniform}[0,1]$\;
-  Scale $X$ by $X \gets X \times U^{1/m}$\;
- \caption{Ellipsoid sampler}
-\end{algorithm}
 
@@
 It is well-known that $\hat p\rightarrow p^*$ in probability (in both the uniform or the Euclidean norm, which are of course equivalent in this case).
-It is easy to see that the same goes for $\hat p\hat q\rightarrow h^*$ (see Lemma \ref{lem:easyconsistency}).
-Thus, intuitively, the difficulty is this: by allowing ourselves to ensure $\UN{\hat p - p^*},\EN{\hat p - p^*},\UN{\hat p\hat q - h^*},\EN{\hat p\hat q - h^*}$ sufficiently small, can we find some $\tilde q\in \hat \Theta$ so that $\UN{\tilde q - q^*}$ is arbitrarily small?
+It is easy (albeit technical) to see that the same goes for $\hat p\hat q\rightarrow h^*$ (see Lemma \ref{lem:easyconsistency}).
+Thus, intuitively, the difficulty is this: by ensuring $\UN{\hat p - p^*},\EN{\hat p - p^*},\UN{\hat p\hat q - h^*},\EN{\hat p\hat q - h^*}$ sufficiently small, can we find some $\tilde q\in \hat \Theta$ so that $\UN{\tilde q - q^*}$ is arbitrarily small?
 It turns out we can.

### 2018-05-21 markovlink/__init__.py

@@ -1,10 +1,7 @@
 from . import estimation
 from . import exploration
-from . import criticize
 
 train_q = estimation.train_q
-find_extremal = exploration.find_extremal
-sample_prux = exploration.sample_prux
-sample_rux = exploration.sample_rux
-sample_uniform = exploration.sample_uniform
-train_test_split = criticize.train_test_split
+sample_bootstrap_rux = exploration.sample_bootstrap_rux
+train_test_split = exploration.train_test_split
+get_rux_tildetheta=exploration.get_rux_tildetheta

### 2018-05-21 markovlink/criticize.py

@@ -1,103 +1,0 @@
-import numpy as np
-import numpy.random as npr
-
-
-def project_to_polytope(p,qdef,newq):
-    '''
-    Given 
-    - polytope defined by p @ q = p @ qdef and q>=0
-    - a point newq
-
-    Output
-    - the point in the polytope which is closest to newq
-
-    '''
-
-    try:
-        import cvxopt
-
-        raveln= np.prod(newq.shape)
-
-        # equality constraint is that A q = b
-        A,b = formulate_polytope(p,qdef)
-
-        # inequality constraint is that 
-        # (-I)q<=0
-        G=-np.eye(raveln)
-        h=np.zeros(raveln)
-
-        # thing to minimize is
-        #   .5 |q - newq|^2
-        # = .5 |q|^2 - <newq,q> + ...
-        # = .5 q^T I q + (-newq^T) q + ...
-        P = np.eye(raveln)
-        d = -newq.ravel()
-
-        '''
-        min   |q - qxy|^2
-        subj  p q = p qdef
-
-        Has become
-        
-        min   .5 q^T P q + d^T q
-        subj  Aq = b
-              Gq <= h
-
-        '''
-        
-        mtx=lambda x: cvxopt.matrix(x)
-        
-        rez=cvxopt.solvers.qp(mtx(P),mtx(d),G=mtx(G),h=mtx(h),A=mtx(A),b=mtx(b))
-        
-        if rez['status']=='optimal':
-            return np.array(rez['x']).reshape(newq.shape)
-        else:
-            raise Exception("Polytope projection failed!")
-    except ModuleNotFoundError:
-        raise("You don't seem to have CVXOPT.  We depend on CVXOPT for polytope projection :(.  Try conda install cvxopt?")
-    
-
-def train_test_split(Nly,m):
-    '''
-    Input:
-    - an array of positive integers, Nly
-
-    Output:
-    - an array train,test
-
-    For each row, we hold out m of the counts from Nly
-    '''
-
-    assert (np.sum(Nly,axis=1)>=m).all(),"Not all rows of Nly have %d total count; can't hold out that much"%m
-
-    Nly=np.require(Nly)
-    Nshp=Nly.shape
-
-    # sample (l,Y) events, uniformly without replacement
-    # from the (l,Y) events indicated by Nly
-
-    # we iteratively select a random point from Nly
[…]

### 2018-05-21 markovlink/exploration.py

@@ -6,57 +6,46 @@ import sys
 
 from . import estimation
 
-def sample_prux(plx,qxy,n):
-    '''
-    Consider the polytope of transition matrices q such that plx @ q = plx @ qxy.
+def get_rux_tildetheta(d):
+    q1=np.zeros(d.shape)
+    q1[np.r_[0:d.shape[0]],np.argmax(d,axis=1)]=1
+    
+    q2=np.zeros(d.shape)
+    q2[np.r_[0:d.shape[0]],np.argmax(-d,axis=1)]=1
 
-    Choose a direction uniformly at random, and find the pair of vertices in that polytope
-    which are most extremal along that axis.  Calculate the distance between those vertices
-    projected along the direction, as well as their distance in absolute terms.  
+    return d,q1,q2
 
-    Repeat n times.
+def sample_bootstrap_rux(Nlx,Nly,verbose=True,kappa=.01,use_cvxopt=False,direc=None,bootstrap=True):
 
-    Returns a list of the form
+    if direc is None:
+        direc=npr.randn(Nlx.shape[1],Nly.shape[1])
 
-    [
-        (direction,samp1,samp2,distance_along_axis,absolute_distance),
-        (direction,samp1,samp2,distance_along_axis,absolute_distance),
-        (direction,samp1,samp2,distance_along_axis,absolute_distance),
-    ]
-    '''
+    # do a resample, get extremal
+    if bootstrap:
+        Nlx2=resample(Nlx)
+        Nly2=resample(Nly)
+    else:
+        Nlx2=Nlx
+        Nly2=Nly
+    plx = (Nlx2+1)/np.sum(Nlx2+1,axis=1,keepdims=True)
+    qxy=estimation.train_q(plx,Nly2,kappa=kappa)
+    OPT1=find_extremal(plx,qxy,direc,use_cvxopt=use_cvxopt)
 
-    lst=[]
-    for i in range(n):
-        direc=npr.randn(*qxy.shape)
-        direc=direc/np.linalg.norm(direc)
-        samp1=find_extremal(plx,qxy,direc)
-        samp2=find_extremal(plx,qxy,direc2)
-        lst.append((direc,samp1,samp2,np.sum(direc*(samp1-samp2)),np.linalg.norm(samp1-samp2)))
-    return lst
+    # do another resample, get another extremal
+    if bootstrap:
+        Nlx2=resample(Nlx)
[…]

### 2018-07-02 paper/paper.tex
Fixed picture formatting
Latest and greatest

@@
 \newcommand{\UN}[1]{\ensuremath{\left|#1\right|_\infty}}
+\newcommand{\TV}[1]{\ensuremath{\left\Vert #1\right\Vert_{TV}}}
 \newcommand{\EN}[1]{\ensuremath{\left|#1\right|}}
@@
 \usepackage{amsthm}
-\newtheorem*{thm}{Theorem}
+\newtheorem{thm}{Theorem}
 \newtheorem{lem}{Lemma}
@@
 the conditional distribution of one tool's measurement given another tool's measurement.
-It achieves this without any measurement data from both tools applied to the same ``bodies of water.'' Moreover, MLM makes zero assumptions about what calibrations we might expect to see, instead applying a subpopulation-based conditional independence assumption.
+It achieves this without any measurement data from both tools applied to the same ``bodies of water.'' Moreover, MLM can be applied even if we cannot make any assumptions about what calibrations we might expect to see; instead, it uses a subpopulation-based conditional independence assumption.
 We evaluate MLM on a pair of single-cell RNA techniques, obtaining a calibration between the tools.
@@
 We will denote it by $q^*(y|x)$.
-As input, this conditional distribution takes the measurement result $x$ obtained from one tool on a particular specimen.
-As output, it yields the probability of obtaining result $y$ from a second tool applied to measure the same specimen.
-One way to learn the calibration is to measure the same specimens under both tools.
+Let us say we obtain measurement result $x$ from one tool on a particular specimen.
+The number $q(y|x)$ indicates the probability of obtaining result $y$ from a second tool applied to measure the same specimen.
+One way to learn the calibration is to measure the same specimens with both tools.
 We call this ``joint measurement.''  Unfortunately, calibrations are often required even when joint measurement is unavailable.
@@
 We here propose the Markov Link Method (MLM) to estimate calibrations between tools.
-The MLM can be trained without any joint measurement.
+The MLM can be applied without any joint measurement.
 The key idea is to use multiple subpopulations of specimens.
@@
 MLM also gives suggestions about which further subpopulations might be helpful to study in order to further refine our knowledge of the true calibration.
-The method can also be naturally extended to calibration distributions among many tools.
 
-\section{Relation to prior work}
-
-Our main goal is to achieve calibration without joint measurement.
-The main obstacle is what is known in the statistics world as an `identifiability problem.'  Due to this problem, we simply \emph{cannot} directly estimate the calibration that we are interested in.
-The calibration is simply not ``identifiable.''  Nonetheless, all is not lost.
-Our main contribution is to show that one can bound the extent of this identifiability problem.
-In short, the fundamental problem cannot generally be vanquished, but we can put it in its place.
-
-Our identifiability analysis stands on the shoulders of a long history of turning probabilistic assumptions into bounds on unidentifiable parameters.
-The core idea of the MLM is to take an assumption (the `Markov Link assumption,' which we will define shortly) and use it to place bounds on an unidentifiable quantity (namely the calibration $q^*$).
-When these bounds are fairly tight on all sides, we see that much can be learned even though what we want isn't identifiable.
-Much of the prior literature in this kind of direction comes from research into causality.
-For example, in \cite{bonet2001instrumentality} Bonet uses polytopes not unlike the ones seen here to explore whether a variable can be used as an instrument.
-The famous Clauser-Horne-Shimony-Holt inequality was designed to help answer causality questions in quantum physics, but it also sheds light on what distributions are consistent with certain assumptions \cite{clauser1969proposed}.
-Indeed the physics literature has contributed many key inequalities (cf.
-\cite{chaves2014inferring}, \cite{kela2017semidefinite}, and the references therein).
-Perhaps the closest work to this one would be \cite{makarov1982estimates}, which used two marginal distributions to get bounds on a property of the joint distribution (namely the distribution of the sum).
-We advance this approach to a more general-purpose technique, both by using many subpopulations to refine our estimates and by considering the entire space of possible joint distributions instead of simply a particular aspect of the joint.
-
-It is perhaps worth mentioning that the actual theorem presented here has probably been derived before (although we did not find it in our review of the literature).
-Other results in the same flavor as the one presented here might also find a practical use for modern problems.
-There is a treasure-trove of ideas in the causality literature; this wealth has not yet been brought fully to bear on the challenging and important problems of calibration for modern experimental modalities.
-Our primary intention with this article is to bring attention to the practical utility of these kinds of result.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{The Markov Link Method assumption}
 
-To make our ideas rigorous, let us develop a little bit of notation.
+The MLM is only applicable if a certain assumption is met.
+In this section we will articulate this assumption as clearly as possible.
 Let us say we are considering the specimens of a large population.
@@
 \item $\ell_i$, the subpopulation.
-We assume that the overall population can be split into subpopulations of interest.
+We assume that we can identify a variety of these subpopulations.
+They are not required to be disjoint subpopulations; it is only necessary that each subpopulation captures a different slice of the overall population.
 For example, we could define subpopulations by looking at cells in different parts of the body or cells with different sizes.
 \item $X_i$, the result of measuring a specimen with tool I.  For example, perhaps tool I takes a picture of the cell with a destructive electron microscopy method.
-\item $Y_i$, the result of measuinrg a specimen with tool II.  For example, perhaps tool II measures the RNA expression of the cell with a destructive sequencing method.
+\item $Y_i$, the result of measuring a specimen with tool II.  For example, perhaps tool II measures the RNA expression of the cell with a destructive sequencing method.
 \end{enumerate}
@@
 We can never observe $\ell_i,X_i,Y_i$ for any specimen $i$.
-Despite this obstacle, we would like to estimate the conditional distribution $q^*(y|x)\triangleq \mathbb{P}(Y_i=y|X_i=x)$.
-To make this estimation at all possible, we need some kind of assumption.
+We would like to estimate as much as possible about the distribution of $X,Y|\ell$.
+In this paper, we will consider the case that this estimation is facilitated by an additional assumption:
 
@@
 
-The key assumption of the Markov Link Method is a conditional independence: that $\mathbb{P}(Y|X,\ell)$ is independent of $\ell$, i.e.
-
 \begin{center}
@@
 Intuitively, this signifies that the manner in which $X$ predicts $Y$ would be the same for each subpopulation.
+In the language of statistics, the assumption is that the variable $X$ is `sufficient' for $\ell$.
 If this assumption is not met, then the method presented here is not applicable.
-The validity of these assumptions for a given situation should be closely contemplated.
+If it \emph{is} met, it opens the door for new methods to estimate three important objects:
 
-Let us consider a few real-world examples where this assumption may apply.
+\begin{itemize} 
+    \item The calibration $q^*(y|x)$ 
+    \item The distribution of tool-I measurement values found in each subpopulation, $p^*(x|\ell)$ 
+    \item The distribution of tool-II measurement values found in each subpopulation,  $h^*(y|\ell)$.  
+\end{itemize}
+
+Under the MLM assumption, it is straightforward to see that these three objects are all bound together by the formula $h^*(y|\ell) = \sum_x p^*(x|\ell)q^*(y|x)$ for each $\ell,y$.
+Using this equation, we will see that it is possible to estimate properties of the calibration $q^*$ even without access to joint measurement.
+It also makes it possible to perform more efficient estimation of $h^*$.
+Theoretically we could also get more efficient estimation of $p^*$, but we did not find plausible cases where this actually occured.
+
+\figgy{
+\hfill{}\figgygr{.7}{../images/plate}\hfill{}
+\caption{Another way to articulate the Markov Link Method is to look at the graphical model associated with the MLM assumption.
+This reveals that we have two kinds of data.
+In one case (above), we can observe $\ell,X$, but $Y$ is hidden.
+For other kinds of data (below) we can observe $\ell,Y$, but $X$ is hidden.
+All of this data is joined by the unknown parameters $p,q$.
+Here $p$ indicates the distribution of tool-I measurement values from each subpopulation, and $q$ indicates the calibration.
+Our task is to use these different kinds of data to determine the parameters $p,q$ as best as we can.
+Once these parameters are known, notice that distribution of tool-II measurement values for each subpopulation, which we denote $h$, can be determined immediately using the Markov Link Assumption: $h^*(y|\ell) = \sum_x p^*(x|\ell)q^*(y|x)$.
+\label{fig:platediag}}
+}
+
+\section{Examples where the Markov Link Method may apply}
+
+The validity of the Markov Link Method assumption for a given situation should be closely contemplated.  Let us consider a few real-world examples where this assumption may apply.
 
@@
     \item Quality control for manufacturing.
-The surest way to test the reliability of a part is to construct a machine that pushes the part until it breaks.
+One way to test the reliability of a part is to construct a machine that pushes the part until it breaks.
 However, how can we test the reliability of the machine that performs the test?
@@
 That is, the MLM assumption holds.
+
+    \item Radiometric calibration.
+Different cameras measure light differently.
+For example, each camera has different lens distortions.
+Different cameras also have different ways of transforming photon counts into digital information.
+Fortunately, joint measurement is generally possible with cameras; simply take a picture of the same object with both cameras.
+Unfortunately, an adequate amount of joint measurement is sometimes hard to come by.
+For example, with expensive astronomy-grade settings, it can be difficult to balance the need for calibration with the total amount of the sky one wants to cover (cf.\ \cite{padmanabhan2008improved}).
+For example, instead of requiring different cameras to take pictures of exactly the same portion of sky at exactly the same time, the subpopulations $\ell$ could represent portions of the sky.
+Various conditions may cause these portions of the sky to appear differently over time, but if we assume this variability is independent of the calibration itself, the MLM assumption may apply.
 
@@
 This difficulty is commonly attributed to selective publication bias, bad design, poor description of methods, and even outright fraud \cite{baker2016reproducibility}.
-A calibration would allow us to understand this problem in detail.
+However, some of the problem may simply be a matter of calibration.
 If two labs perform identical experiments and get different data, that does not mean we need to throw out both datasets.
-Instead, we can use MLM to calibrate the tools.
-Once the tools are properly calibrated, we can combine both datasets.
+Instead, we can use MLM to calibrate the processes used by each lab.
+Once the labs are properly calibrated, we can meaningfully combine both datasets.
 Unlike other tools to deal with lab or batch effects (e.g.
@@
 
-The rest of the paper proceeds as follows:
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{The Markov Link Method: input and output}
+
+The Markov Link Method is a procedure from estimating the distribution $X,Y|\ell$ from data, under the Markov Link Method assumption.
+Details on the exact algorithm and convergence results are given in Section \ref{sec:mlm}.
+For now, let us content ourselves with looking at the input and the output of the MLM:
+
+\begin{description}
+    \item[Input] We assume we are given $n$ samples using tool I and $m$ samples using tool II.
+    \begin{itemize}
+        \item Samples $(\ell_1,X_1)\cdots (\ell_n,X_n)$ sampled such that $\mathbb{P}(X_i=x|\ell_i=\ell)= p^*(x|\ell)$
+        \item Samples $(\ell_{n+1},Y_{n+1})\cdots (\ell_{n+m},Y_{n+m})$ sampled such that $\mathbb{P}(Y_i=x|\ell_i=\ell)= h^*(x|\ell)$
+    \end{itemize}
+    We assume that $\ell,X,Y$ are discrete random variables with finite support.
+We also assume that the MLM assumption holds, i.e.\ $h^*(x|\ell)=\sum_y p^*(x|\ell)q^*(x|y)$ for some unknown $q^*$.
+It would be interesting to generalize the method to non-discrete random variables, but we leave this for future work.
+Note that if the random variables are low-dimensional and continuous, the MLM can still be applied by discretizing the continuous space into a finite number of bins.
+    \item[Output] Our goal is to estimate $p^*,q^*,h^*$ from the samples.  In this effort, we return:
+    \begin{itemize}
+        \item Point estimates, $\hat p,\hat q,\hat h$.  
+        \item Quantitative estimates of the variability of these point estimates due to insufficient data, $\Delta_p,\Delta_q,\Delta_h$.  
+        \item An quantitative estimate of the variability of $\hat q$ due to so-called `identifiability' issues, $\bar d$.  
+        \item Variability visualizations indicating a set of ``plausible'' values for each row of $q$; this gives a qualitative counterpart to the quantitative measures of variability.
+    \end{itemize}
+    Precise definitions and mathematical results are given in section \ref{sec:mlm}.
+\end{description}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Simulation results}
+
+Before we show the results of the MLM on real data, let us look at some simulation results to see what the MLM can and cannot achieve.  In particular, we will:
 
 \begin{itemize}
-\item We describe the identifiability problem and describe how our analysis puts this problem in our place.
-We give several examples attempting to give intuition as to why the identifiability problem may not be as severe as it first appears.
-\item Applying these ideas, we develop the Markov Link Method for calibration without joint measurements.
-\item We apply the MLM to two single-cell transcriptomic methods.
-The result uncovers some ways in which the methods appear to be well-calibrated and other areas where the calibration may not be as good.
+    \item Pick a set of possible parameter values for $p^*,q^*,h^*$.
+    \item Generate data according to these distributions, namely $n$ samples according to $p^*$ and $m$ samples according to $h^*$.  
+    \item Pretend to forget the true values of $p^*,q^*,h^*$.
+    \item Apply the Markov Link Method to the generated samples.
+    \item Compare the results of the MLM with the original true values.
 \end{itemize}
 
-\section{The identifiability problem}
-
-Let us make three assumptions:
+Our purpose is to see how well the method performs if all of its assumptions are met. We will measure performance two ways:
 
 \begin{enumerate}
-\item For each specimen $i$, the distribution of $X_i,Y_i|\ell_i$ may be written 
-\[
-\mathbb{P}(X_i=x,Y_i=y|\ell_i) = p^*(x|\ell_i)q^*(y|x)
-\]
-This is the central assumption we have discussed at length above.  For convenience we will also introduce the notation
-\[
-\mathbb{P}(Y=y|\ell) = h^*(y|\ell)=\sum_x p^*(x|\ell) q^*(y|x)
-\]
-\item Joint measurement is unavailable.
-In particular, we will assume we have $n+m$ individual specimens.
-Of these, we have observed $\ell_i,X_i$ for $i\in\{1\cdots n\}$ and $\ell_i,Y_i$ for $i\in n+1\cdots n+m$.
-\item $X$ and $Y$ are discrete random variables with finite support (the general concepts here will apply more generally, but we leave it for future work; if the data is not discrete, we can always make it so by dividing it into suitable bins).
-In this simple case, we may summarize all of our observed data in two matrices:
-\begin{align}
-N^X_{\ell x} = |\{i\leq n:\ \ell_i=\ell,X_i=x\}|  \qquad N^Y_{\ell y} = |\{i>n:\ \ell_i=\ell,Y_i=x\}|
-\label{eq:twomatrices}
-\end{align}
-Thus $N^X,N^Y$ are matrices counting the number of each kind of observation for method I and method II respectively. 
+    \item First, we will measure the error of the estimators $\hat p,\hat q,\hat h$, i.e.\ how much they differ from the true generating distributions $p^*,q^*,h^*$.
+For $h$ we will compare this measurement error against a standard pseudocount estimator (we will not do the same for $p$, because the MLM estimator for $p$ \emph{is} a pseudocount estimator).
+We refer the reader to Section \ref{sec:mlm} for details on this estimator.
+    \item Second, we will look at how well the MLM has estimated its uncertainty about the estimates.
+We know that there will always be some amount of error.
+For this reason, the MLM offers two tools to estimate the amount of error that may be present.
+In this section we will particularly focus on the quantities $\Delta_p,\Delta_q,\Delta_h$ (which approximately indicate variability due to insufficient data) and $\bar d$ (which approximately indicates variability due to so-called identifiability concerns).
+For a precise explanation of these two sources of variability, we refer the reader to Section \ref{sec:mlm}.
 \end{enumerate}
 
-Our goal will be to use $N^X,N^Y$ to estimate $q^*$, the calibration.
-However, \emph{the data from the matrices $N^X,N^Y$ only enable us to estimate $p^*,h^*$}.
-They do not enable us to directly estimate $q^*$.
-Therefore, we define
-\[
-\Theta(p,h) \triangleq  \left\{q:\ \sum_x p(x|\ell)q(y|x) = h(y|\ell)\ \forall \ell,y,\ q(y|x)\geq 0\ \forall x,y, \sum_y q_{xy}=1\ \forall x\right\}
-\]
-as the set of values of the calibration $q$ which are consistent with a given value of $p$ (the conditional distribution of $X|\ell$) and $h$ (the conditional distribution of $Y|\ell$).
-Any effort to estimate $q^*$ must therefore overcome two fundamentally different challenges:
+For our first simulation, we will look at a case in which the MLM performs well.
+We consider twenty different subpopulations (which we will label $\{1,2,3,\cdots 20\}$).
+We will assume that tool I always returns an integer measurement in $\{1,2,3,4,5,6\}$ and tool II returns a measurement in $\{1,2,3,4,5,6\}$.
+We will assume we have one hundred samples for each subpopulation for each method.
+Some results of this simulation can be seen in Figure \ref{fig:simulation1gt}.
+This figure shows that the MLM gives us an edge in estimating $h$, performing better than a traditional pseudocount estimator.
+It also allows us to estimate $q$ with decent accuracy (which is only possible because of the Markov Link Assumption).
+The figure also shows that the $\Delta$ variability measures seem to accurately predict the estimation error.
+The $\bar d$ for this simulation was identically zero, because there were no identifiability issues in this case; this is related to the fact that there were more subpopulations than there were possible values that tool I could return.
 
-\begin{enumerate}
-\item Not-enough-data problems. We don't have infinite data, so we can't know the exact values of $p^*,h^*$.
-\item Identifiability problems. Even if we knew $p^*,h^*$ exactly, it is often impossible to know the value of $q^*$.  We can only ever know that it lies somewhere in $\Theta(p^*,h^*)$.  
-\end{enumerate}
-
-This point of view suggests a practical procedure for estimating $q^*$:
-
-\begin{enumerate}
-    \item Use traditional methods to develop estimators $\hat h,\hat p$ so that with more and more data we can guarantee that $\hat h\approx h^*$ and $\hat p\approx p^*$ with high probability.
-In particular, we will use a robust pseudocount estimator for $p^*$ and use the MLM assumption to get a high-quality estimator for $h^*$ that leverages information from $N^X$ and $N^Y$.
-    \item Then we may suppose that $q^*$ is probably somewhere \emph{near} to $\Theta(\hat p,\hat h)$.  
-\end{enumerate}
-
-It is not completely obvious that this will work.
-Certainly $\hat p\approx p^*$ and $\hat h\approx h^*$.
-Certainly $q^*\in \Theta(p^*,h^*)$.
-But does that guarantee that $q^*$ is nearby to the set $\Theta(\hat p,\hat h)$?
-Something could go quite wrong in this line of reasoning.
-Fortunately, subject to mild assumptions, one can show that nothing goes wrong.
-This is the content of our main technical contribution.
-It can be found in Appendix \ref{sec:proof}.
-
-Let us now take a moment to understand the set $\Theta$.
-This set encodes exactly how it is that the Markov Link assumption enables us to understand something about the calibration $q^*$.
-That is, this set tells us how the things we can estimate (i.e.\ $p^*,h^*$) inform us about what we want to know (i.e.\ $q^*$).
-To understand a bit more concretely how this works, it may be helpful to think of $p^*,h^*,q^*$ as matrices.
-From this point of view, one aspect of the definition of $\Theta$ can be written more concisely as a matrix equality constraint on $q^*$: if $q \in \Theta(p^*,h^*)$, then $p^*q=h^*$.
-This equation is in fact a literal mathematical rendering of the Markov Link assumption of conditional independence.
-The consequences of this assumption and its corresponding equation depend upon whether $p^*$ has a left-pseudoinverse.
-\begin{itemize}
-    \item If $p^*$ has a left-pseudoinverse then this equation allows us to uniquely determine what we want ($q^*$) in terms of what we know ($p^*,h^*$).
-In general, this will happen when the number of subpopulations outnumbers the number of different states that $X$ can take on.
-In this case, the problem should be straightforward to solve.
-    \item If $p^*$ does not have a left-pseudoinverse, then there is an identifiability issue: we can never hope to exactly determine $q^*$.  
-\end{itemize}
-
-In this paper we will focus almost entirely on the second case.
-In this second case there is a genuine identifiability issue;  $\Theta(p^*,h^*)$ is nontrivial and so it is \emph{impossible} to ever determine the true value of $q^*$, regardless of how much data we have.
-All we have is the equality and inequality bounds that arise from our knowledge that $q^* \in \Theta(p^*,h^*)$.
-
-To gain intuition about the equality and inequality bounds that define the set $\Theta$, let us consider a few examples.\vspace{.1in}
-
-\begin{example}
-Let $\ell \in \{1,2\}$, $X\in \{1,2,3\}$, $Y\in \{1,2\}$, and 
-\[
-p^{*}=\left(\begin{array}{ccc}
-40\% & 50\% & 10\% \\
-10\% & 10\% & 80\%
-\end{array}\right)\qquad h^{*}=\left(\begin{array}{cc}
-20\% & 80\% \\
-40\% & 60\%
-\end{array}\right)
-\]
-That is, for example, $p^*_{1,2}=\mathbb{P}(X=2|\ell=1)=50\%$ and $h^*_{2,1}\mathbb{P}(Y=1|\ell=2)=40\%$.
-Let us look at a single equation in the system $p^* q^* = h^*$ entailed by the MLM assumption.
-For example, $h^*_{2,1}$, we get
-\[
-0.4 = h^*_{2,1} = \sum_x p^*_{2,x} q^*_{x,1} = .1 q^*_{1,1} + .1 q^*_{2,1} + .8 q^*_{3,1} 
-\]
-We thereby obtain an important constraint on what the calibration $q^*$ might be.
-
-This equation also takes on additional significance in light of the inequality bounds we have: $0\leq q^*_{xy}\leq 1$ for every $x,y$.
-In other words, probabilities must be between 0 and 1, by definition.
-This may seem trivial, but it has important consequences in light of the constraints imposed by the equation $p^*q=h^*$.
-In particular, taking another look at that same equation, we see that
-\[
-0.4 = .1 q^*_{1,1} + .1 q^*_{2,1} + .8 q^*_{3,1} \leq .2 + .8 q^*_{3,1}
-\]
-It follows immediately that $q^*_{3,1}\geq .25$.
-Each equation in the MLM system yields insights of this kind; together these insights form the constraints that define $\Theta$.
-The more different subpopulations you have, the more equations of this kind you will have.
-\end{example}
+\figgy{
+\hfill{}\figgygr{.7}{../images/simulation1Groundtruth}\hfill{}
 
@@
 
-\begin{example}
-Let $\ell \in \{1\}$, $X\in \{1,2,3,\cdots 11\}$, $Y\in \{1,2\}$, and 
-\[
-p^{*}=\left(\begin{array}{ccccc}
-90\% & 1\% & 1\% & \cdots & 1\% \\
-\end{array}\right)\qquad h^{*}=\left(\begin{array}{cc}
-95\% & 5\% \\
-\end{array}\right)
-\]
-Here we have only \emph{one} population.  Still we are able to say something quite significant; using similar reasoning to that shown above we see that $q^*_{1,1}\geq 94.\bar4\%$.  
-\end{example}
+\hfill{}\begin{tabular}{r|p{.3in}p{.7in}p{1.0in}}
+Variable & MLM error & Pseudocount error & $\Delta$ \\
+\hline
+$p$ & 1.5 & N/A & 1.6 \\
+$q$ & 0.6 & N/A & 0.6 \\
+$h$ & 0.9 & 1.4 & 1.0 \\ 
+\end{tabular}\hfill{}
 
-\vspace{.1in}
+\caption{For our first simulation, we imagine that the true values of $p^*,q^*,h^*$ are indicated by the first row of heatmaps above.
+The color of each entry indicates a probability between 0 and 1, with 0 being represented by black and 1 being represented by white.
+For example, the entry at the third row and the fifth column of the $q$ plot indicates $\mathbb{P}(Y=5|X=3)=q^*(5|3)$.
+The second row of heatmaps indicates the estimates for these quantities returned by the MLM.
+We can also look at this more quantitatively, as shown in the table.
+For each variable, $p,q,h$, we can compute the total variation distance between our estimates and the truth.
+For $h$ we can also compare the error we get when using a more standard pseudocount estimator.
+Finally, in the right-most column of the table, we show the $\Delta$ variability measures for each variable.
+These are intended to roughly indicate the level of error we would expect due to insufficient data.
+They are caclulated using a bootstrap method.
+We can see that the MLM variability measured do seem to give a good indication of the actual error level.
+\label{fig:simulation1gt}}
+}
+
+For our second simulation, we will look at a case where the MLM does a good job of estimating $p,h$, but does not succeed in accurately estimating $q$.
+We use the same data as in the previous simulation, but we will merge groups of subpopulations into four larger subpopulations.
+This will make it easier to estimate $h(y|\ell),p(x|\ell)$, since we have fewer subpopulations and for each one we will have five times as much data.
+The results are shown in Figure \ref{fig:simulation2gt}.
+As we would expect, estimation error of $p,h$ improves.
+However, estimation of $q$ does not improve at all.
+In fact, it is considerably worse.
+To understand this, we run a further simulation with the same true distributions but a virtually infinite number of samples.
+Even in this case, we see that accurate estimation of $q$ is still not achieved; we end up with a final error of $0.79$.
+This significant error is due to a fundamental identifiability issue which is discussed in detail in Section \ref{sec:mlm}.
+Accurate estimation of $q$ is simply impossible in this case, regardless of the number of samples.
+This is related to the fact that the number of subpopulations is smaller than the number of different values that tool I can report.
+What is worse, the $\Delta_q$ reports only $.01$ for the predicted error of $\hat q$.
+It is for this reason we must also look at $\bar d$, which approximates the error due to identifiability.
+For this simulation we calculate that $\bar d=3.5$, which captures the fact that our estimator $\hat q$ may suffer from nontrivial identifiability errors.
+Indeed, it appears to substantially overestimate the total error.
+This problem is somewhat unavoidable without additional assumptions.
+A precise statement of the theoretical guarantees regarding the ability of the MLM to estimate its own error due to identifiability issues can be found in Section \ref{sec:mlm}.
+I
 
 \figgy{
-\hfill{}\figgygr{.7}{../images/polytope}\hfill{}
-\caption{We consider the simple case $\ell=1$, $X\in \{1,2\}$, $Y\in \{1,2\}$.
-In this case the space of values of the calibration $q$ which are valid probability distributions can be understood as a box.
-The space of values of $q$ which satisfy the matrix equation $p^* q = h^*$ can be understood as a line that passes through that box.
-The set of possible values of $q$ which are consistent with the matrix equation and are also valid probability distributions forms the set $\Theta$, which in this case can be identified with a line segment.
-We know that the true value of $q^*$ must lie somewhere inside this $\Theta$.
-Depending upon the exact values of $p^*,h^*$, the corresponding line segment may be larger or smaller.
-For example, if the line crosses close to a corner of the box, then $\Theta$ may be restricted to a very small region near that corner.
-\label{fig:polytopeexample}}
+$500$ samples per population:
+\hfill{}\begin{tabular}{r|p{.7in}p{1.0in}}
+Variable & MLM error & $\Delta$ \\
+\hline
+$p$ & 0.12 & 0.16 \\
+$q$ & 1.33 & 1.19 \\
+$h$ & 0.17 & 0.15 \\ 
+\end{tabular}\hfill{}
+
+\vspace{.2in}
+
+$10^{10}$ samples per population:
+\hfill{}\begin{tabular}{r|p{.7in}p{1.0in}}
+Variable & MLM error & $\Delta$ \\
+\hline
+$p$ & $10^{-7}$ & $10^{-3}$ \\
+$q$ & $0.79$ & $0.01$ \\
+$h$ & $10^{-3}$ & $10^{-3}$ \\
+\end{tabular}\hfill{}
+
+\caption{Our second simulation is similar to our first.
+However, instead of twenty subpopulations, we will assume we only have five subpopulations with four times the amount of data for each subpopulation.
+This makes estimating $p,h$ easier, but it doesn't make estimating $q$ any easier.
+\label{fig:simulation2gt}  Even when we give ourselves a virtually unlimited number of samples, perfect estimation of $q$ is just not possible.
+The $\Delta$ measures of the MLM do \emph{not} accurately reflect this fact.
+It is for this reason that the MLM additionally provide the quantity $\bar d$.
+For this simulation $\bar d=3.5$, which allows us to know that there are identifiability concerns, and also to bound the magnitude of those concerns.}
 }
 
-\begin{example}
-Finally, let us consider one class of examples in which the entire situation can be visualized: $\ell=1$, $X\in \{1,2\}$, $Y\in \{1,2\}$.
-Notice that $q$ is completely defined by $\theta_1=\mathbb{P}(Y=1|x=1)$ and $\theta_2=\mathbb{P}(Y=1|x=2)$.
-The matrix equation $p^* q^* = h^*$ corresponds to a linear constraint on $\theta_1,\theta_2$, namely $p^*_{11}\theta_1 + p^*_{12}\theta_2 = h^*_{11}$.
-On the other hand, the fact that $q^*$ must be a valid probability distribution tells us that $\theta_1,\theta_2$ must lie inside the box $[0,1]\times[0,1]$.
-Combining the matrix equation with the constraints of the box, we learn that $q^*$ lies inside a certain bounded, one-dimensional space.
-This space is precisely $\Theta$.
-It can be identified with a line segment, and the depends on the exact values of $p^*,h^*$.
-In some cases it is quite large, i.e.\ it will be difficult to know much about the true value of $q^*$.
-However, in other cases the positivity restrictions force $\Theta$ to be a very small region indeed.
-See Figure \ref{fig:polytopeexample} to visualize this.
-In real-world cases this $\Theta$ is much higher-dimensional and harder to visualize, but in this simple case we can see it clearly.
-\end{example}
-
-\section{The Markov Link Method}
-
-Let us now exactly describe our approach for calibrating in the absence of joint measurement.
-As in Equation (\ref{eq:twomatrices}) above, we assume that $X$ and $Y$ are discrete random variables and $N^X,N^Y$ are the matrices describing the  observed data.
-The purpose of the Markov Link Method is to try to use these matrices to estimate $q^*(y|x)=\mathbb{P}(Y=y|X=x)$, the calibration between $X$ and $Y$.
-The Markov Link Method has three main components:
-
-\begin{description}
-\item[Estimation]  Estimate $p^*$ with robust pseudocount estimator (cf.\ \cite{flach2012machine}):
-\begin{align}
-\hat p(x|\ell) \triangleq \frac{1+N^X_{\ell x}}{\sum_{x'} N^X_{\ell x'}} \label{eq:pseudocountp}
-\end{align}
-and then calculate a possible guess for $q^*$, namely
-\begin{align}
-\hat q = \argmax_q \sum_{\ell,y}N_{\ell y}\log\left(\sum_{x}\hat p(x|\ell)q(y|x)\right) + \kappa \sum_{xy} \log q(y|x)
-\label{eq:mlmq}
-\end{align}
-The Dirichlet prior regularizer $\kappa$ ensures this optimization problem has a unique solution; indeed, one may easily verify that the objective is strictly convex.
-In practice, we took $\kappa$ to be any small constant.
-We use a majorization-minimization approach to solve this optimization problem.
-That is, we first take an initial guess for $\hat q$.
-We then define $\tilde \pi_{\ell,x,y} \gets \hat p(x|\ell)q(y|x) / \sum_{x'} \hat p(x'|\ell)q(y|x')$.
-We then update $\hat q$ by  $\hat q \gets \argmax_q \sum_{\ell,y,x}N_{\ell y}\tilde \pi_{\ell,x,y}\log\left(\hat p(x|\ell)q(y|x)\right) + \kappa \sum_{xy} \log q(y|x)$, which can be solved in closed form.
-We then iterate this procedure, redefining $\tilde \pi$ and then redefining $\hat q$ in each iteration.
-It is well-known that since the overall problem is convex we can guarantee that this procedure converges to the global optima (cf.\ \cite{wu1983convergence} for a detailed exposition).
-In practice we found 5000 iterations to be sufficient for convergence.
-
-This yields a high-quality, consistent estimator for $h^*(y|\ell)\triangleq \mathbb{P}(Y=y|\ell)$, namely $\sum_x \hat p(x|\ell) \hat q(y|x)$ (cf.\ the lemma in Appendix \ref{sec:proof} for a simple consistency proof).
-It also gives us a unique estimate for $q^*$, although this estimate may suffer from errors due to incorrect assumptions, lack of data, or identifiability.
-Therefore, a vital part of the MLM is a criticism step.
-\item[Model criticism]  The MLM assumption cannot be checked directly.
-However, we can check certain aspects of it.
-In particular, the MLM assumption suggests a natural estimator for the distribution .
-On the other hand, a more traditional approach would etimate $h^*$ from $N^Y$ using a simple pseudocount method.
-We can compare both estimators for $h^*$ using held-out log-likelihood of $Y$ data.
-If the MLM estimate is considerably worse than the pseudocount method, it may indicate that the MLM assumption doesn't fit with the data.
-\item[Uncertainty characterization] Measure our uncertainy about $q^*$ using \emph{Bootstrapped Rotationally Uniform eXtremal Paired (BRUXP)} samples and \emph{Rotationally Uniform eXtremal Paired (BRUXP)}.
-The variability in these samples capture error both due to insufficient data and due to identifiability issues.
-\end{description}
-
-We now describe this very last step in more detail:
-
-\vspace{.2in}
-\begin{definition}
-Let $d$ denote a random direction in calibration space.
-That is, let $d$ denote a matrix of the same dimensions as the calibration $q$, each entry of which is sampled from a random normal distribution.
-Sample with replacement to produce a surrogate dataset $N^{X,(1)},N^{Y,(1)}$, estimate $\hat p^{(1)},\hat q^{(1)}$ using the surrogate dataset and the method outlined above, and define $\hat \Theta^{(1)} \triangleq \Theta(\hat p^{(1)},\hat p^{(1)}\hat q^{(1)})$.
-Repeat this process to obtain $\hat \Theta^{(2)} \triangleq \Theta(\hat p^{(2)},\hat p^{(2)}\hat q^{(2)})$.
-Now, let us take $\tilde q^{(1)}$ to be the most extremal vertex of $\hat \Theta^{(1)}$ in the direction $d$ and $\tilde q^{(2)}$ to be the most extremal vertex of $\hat \Theta^{(2)}$ in the direction $-d$.
-The tuple $d,\tilde q^{(1)},\tilde q^{(2)}$ is said to be a \textbf{Bootstrapped Rotationally Uniform eXtremal Paired (BRUXP)} sample.
-If we adopt this same procedure without the resampling-with-replacement aspect, we will refer to it as \textbf{Rotationally Uniform eXtremal Paired (RUXP)}.
-\end{definition}
-
-For each BRUXP sample, the pair $\tilde q^{(1)},\tilde q^{(2)}$ gives for the range of what $q^*$ might be along the direction $d$.
-The variability amongst many such samples indicates our uncertainty about the true $\hat q$.
-
-We cannot make general claims about the bootstrap in this case, due to the usual difficulties with bootstrap and extremal statistics (cf.\ \cite{andrews2000inconsistency}).
-However, we do at least know that this procedure correctly handles the identifiability issue in the asymptotic limit.
-In particular, subject to mild reguarity conditions, we can show that $\inf_{q\in \Theta(\hat p, \hat p \hat q)} |q-q^*|\rightarrow 0$ as we gather sufficient data.
-That is, with sufficient data BRUXP samples are guaranteed to represent the uncertainty we face from identifiability concerns.
-An exact statement may be found in Appendix \ref{sec:proof}.
-
-The the variation amongst the BRUXP samples is necessarily due to a combination of effects from insufficient data and effects from the identifiability issues.
-To separate issues, we recommend that one also look at Rotationally Uniform eXtremal Paired (RUXP) samples (i.e.\ BRUXP samples taken without using the sampling-with-replacement procedure).
-The variability among these samples corresponds primarily to uncertainty due to identifiability concerns, and ignores variability due to insufficient data.
-
-Code for this procedure is available at https://github.com/jacksonloper/markov-link-method.
-All calculations in this paper can be replicated by going through a tutorial-style ipython notebook that can be found at the same address.
-
-\section{Empirical results}
+For our next simulation, we will dig deeper into the question of when $q$ can be perfectly estimated.
+To this end, we will consider cases in which the number of subpopulations is four, and there are six possible outcomes for both measurement tools.
+In the general case, this will make it \emph{impossible} to correctly determine $q$ without joint measurement.
+However, for certain values of $q$ the truth can be recovered.
+To see this, we will look at a variety of values of $q$.
+On the one hand, we will consider the case that $X,Y$ are independent, i.e.
+$q(y|x)=q(y|x')$ for every $x,x'$.
+On the other hand, we will consider the case that $X,Y$ are deterministically related, in particular $X=Y$.
+We will also consider every $q$ ``in-between'' these two extremes.
+These in-between $q$s are found by linear interpolation.
+For each such $q$ we will generate datasets of various sizes and try to estimate $q$ from the datasets using the MLM.
+The results are to be found in Figure \ref{fig:estimatingq}.
+These results show how in the two extreme cases it may be possible to recover $q$, but for in-between cases it is indeed not possible.
+This is true even in the limit of infinite data.
+We leave a complete mathematical understanding of this result to future work.
+For the purposes of this paper, we content ourselves with noting that if $X$ and $Y$ can take on one of $2^k$ values and are related to each other by an invertible deterministic function, then both the fact that the relationship is deterministic and the exact specification of the invertible function can be determined with only $k+1$ properly chosen subpopulations (at least in the asymptotic limit of infinite data).
+This result is proven in Appendix \ref{sec:miracle}.
 
 \figgy{
-\begin{minipage}[c]{0.15\textwidth}
-\figgygr{.15}{../images/Nlx}
-\end{minipage}\hfill
-\begin{minipage}[c]{0.15\textwidth}
-\figgygr{.15}{../images/Nly}
-\end{minipage}\hfill
-\begin{minipage}[c]{0.3\textwidth}
+\hfill{}\figgygr{.7}{../images/estimatingq}\hfill{}
+
+\caption{In our third simulation we fix a particular $p$ and consider a range of $q$s.
+For each possible $q$ we generate a certain number of samples per subpopulation and then try to use the resulting dataset to estimate $q$.
+On the one hand, we consider the case that $q$ causes $X,Y$ to be independent; in thie case we can get a good estimate of $q$.
+In some sense this is because $q(y|x)=q(y|x')$ for every $x,x'$, and so there are simply fewer values to estimate.
+On the other hand, we consider the case that $q$ causes $X,Y$ to be deterministically bound, i.e.
+$q(y|x)=1$ if $x=y$ and $q(y|x)=0$ otherwise.
+In this case it is also possible to estimate $q$ (given enough data).
+In-between, it is not possible to exactly determine $q$, even in the asymptotic limit of infinite data.\label{fig:estimatingq}}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Empirical results for cell-types}
+
+We now turn to an application of the MLM to real data.
+
+\figgy{
+\hfill{}\figgygr{1.0}{../images/allenNlx}\hfill{}
+
+\hfill{}\figgygr{.5}{../images/allenNly}\hfill{}
+
 \caption{Input: two tables.
@@
 Technique II gives a less complete analysis, but yields additional electrophysiological data that may be of interest.
-For each subpopulation and each technique, data was gathered to estimate the distribution of different cell-types.
+For each subpopulation and each technique, data was gathered to estimate the distribution of different results.
 These are shown above.
-Each row corresponds to a different cell-type.
-Each column corresponds to a different subpopulation.
-Each entry gives the number of a particular cell-type found within a particular population, according to a particular technique.
+Each row corresponds to a different subpopulation.
+Each column corresponds to a different measurement value.
+Each entry of the heatmaps indicate the number of specimens measured to have a particular measurement value within a particular subpopulation (as measured by the corresponding technique).
+Notice that tool I has many more different measurement results that it can return; this is why there are many more columns in the first heatmap.
 \label{fig:alleninput}}
-\end{minipage}\hfill
 }
@@
 Each modality would take a cell and determine what ``type'' of cell it was.
-However, as part of that process it would destroy the cell.
+However, each technique destroyed the cell in the process of measuring it.
 
 Best efforts were made to use biological intuition to calibrate the methods.
-For example, both methods have a notion of a ``Lamp5 Egln3\_1'' cell-type.
-If a cell was designated as ``Lamp5 Egln3\_1'' celltype using one method, the hope was that it would also be given the same designation if it was processed using the other method.
+For example, the $Y$ method has a notion of a ``Lamp5'' cell type.
+The $X$ method has refined this type into many sub-types, such as ``Lamp5 Pdlim5'' and ``Lamp5 Slc35d3.''  If a cell was designated as ``Lamp5 Pdlim5'' under the $X$ method, the hope was that it be given the ``Lamp5'' type under the $Y$ method.
 The two methods were designed to achieve this goal.
 However, each method has its own biases and errors, and it was not obvious whether this effort was successful.
-In particular, it seemed clear that in some cases cells labelled one way with one method would get labelled another way with another method, but it was not clear how often this occurred.
+In particular, it seemed clear that in some cases a cell labelled one way with one method would get labelled quite differently with another method, but it was not clear how often this occurred.
 
@@
 For each subpopulation and each method, many specimens were sampled and their cell-types determined.
-If the methods were perfectly calibrated, we would expect that both methods would yield the same distribution of cell-types in each subpopulation.
+The result of this process was two tables, shown in Figure \ref{fig:alleninput}.
+While it seemed clear that these tables should say something about the calibration, it was not obvious how to best use this information.
+It was for this purpose that the MLM was developed.
 
-Towards this effort, data was collected for each subpopulation and each method.
-The result of this process was two tables, shown in Figure \ref{fig:alleninput}.
-Perhaps not surprisingly, it was found that the distribution of cell-types appeared different under the two modalities.
-Indeed, it was found necessary in method II to designate some cells with certain ``nxx'' types that indicated different kinds of uncertainty about the cell-type.
-Thus even the set of cell-types was not the same between the two groups.
-Clearly the methods were not perfectly calibrated -- but how big of a problem was it?
-It is not obvious to know just by staring at Figure \ref{fig:alleninput}.
-A more quantitative method was needed.
-
-\subsection{Estimation}
+\subsection{Results}
 
 \figgy{
-\figgygr{0.8}{../images/qhat}
-\caption{The $\hat q$ estimated by the MLM method with $\kappa=.01$.
-Notice the mass along a somewhat diagonally-oriented slice of the calibration.
-This corresponds to cell-types which appear to be designated with the same cell-type by both methods.
-The large region on the right corresponds to the ``nxx'' types in Tool II; these indicate different forms of uncertainty about the type of the cell.\label{fig:allenout}}
+\hfill{}\figgygr{0.3}{../images/allenqhat}\hfill{}
+\caption{The $\hat q$ estimated by the MLM.\label{fig:allenout}}
 }
 
-The first step of the Markov Link Method is to estimate the calibration with the estimator $\hat q$ according to equations (\ref{eq:pseudocountp}) and (\ref{eq:mlmq}).
-Taking $\kappa=.01$, we obtained a calibration found in Figure \ref{fig:allenout}.
-From this, it appears that there are many cells which would be identified with the same cell type in either method.
-For example, according to $\hat q$, if a cell is classified as type ``VISp L5 Endou'' by tool I, it appears there is a 91\% chance it will be classified the same way by tool II.
+Using the Markov Link Method, we obtained an estimate $\hat q$ for the true calibration.
+This calibration is shown in Figure \ref{fig:allenout}.
+From this figure, it appears that the two methods may be fairly well-calibrated.
+For example, according to $\hat q$, if a cell is classified as type ``Lamp5 Pdlim5'' by tool I, it appears there is a 99.9\% chance it will be classified with the ``Lamp5'' type under tool II.
 However, there are other cell-types which seem to have more ambiguity.
-Lest we make too many conclusions without considering the uncertainty in our estimates of $q^*$, let us proceed to the criticism phase of the MLM.
+Yet others seem to perhaps be classified incorrectly \emph{most} of the time; we see that ``Vip Igfbp6'' type is most often classified as a ``Pvalb'' type instead of a ``Vip'' type.
 
-\subsection{Model criticism}
+Everything in the preceeding paragraph should be taken with a significant grain of salt; it is quite treacherous to directly interpret the estimator $\hat q$, because we know that this estimator for the calibration may suffer from errors.
+This is both due to insufficient data and to identifiability concerns (these are discussed in detail in Section \ref{sec:mlm}).
+For a qualitative assessment, we instead advocate looking at the variability visualizations which the MLM produces.
 
-The Markov Link Method assumption suggests a natural estimator for the distribution of tool II results ($Y$) under each subpopulation ($\ell$).
-We have denoted this distribution $h^*$, and the MLM estimates this distribution with $h^*(y|\ell)\approx \sum_x \hat p(x|\ell) \hat q(y|\ell)$.
-A more traditional method would be to estimate $h^*$ using a pseudocount method (cf.\ \cite{flach2012machine}).
-We can compare these methods using held-out log likelihood.
-That is, we train both methods using a subset of our data.
-Then we compute the negative log likelihood of the held out data under both models.
-This indicates how concisely the model can represent the held out data.
-If the MLM method is much worse, it may be a sign that the MLM assumption is poor.
-Smaller values are better:
+For example, in the right side of Figure \ref{fig:allenboot} we show the variability visualization for specimens which are measured as being ``Vip Igfbp6'' by tool I.
+Whereas in the original point estimate $\hat q$ we found that ``Vip Igfbp6'' seemed to be strongly associated with the ``Pvalb'' type, this figure shows that in reality we are not quite sure what types these specimens may be associated with.
+Each row corresponds to a different plausible distribution on $y$.
+As we can see, they truly run the gamut.
+In the left side of the same figure we also show the same visualization applied to the row of $\hat q$ corresponding to specimens measured as ``Lamp5 Pdlim5'' by tool I.
+We see very little variability in this visualization, indicating significant confidence that the ``Lamp5 Pdlim5'' type of tool I is truly well-calibrated to match the ``Lamp5'' type of tool II.
 
-\begin{itemize}
-    \item Traditional estimator: 3.96 nats per cell
-    \item MLM estimator: 3.70 nats per cell
-\end{itemize}
-
-This is encouraging.
-Our model is as good at predicting held-out data as a more traditional pseudocount estimator (which makes no assumptions on the conditional independence structure of the data).
-In fact, the MLM estimate is able to outperform the traditional estimate, presumably because it can leverage the data in $N^X$ to help us learn about the distributions on $Y$.
-At the very least, it is not a clear sign that the MLM assumption is inappropriate for this data.
-
-\subsection{Uncertainty characterization}
-
-We obtained BRUXP samples and RUXP samples for this dataset.  These can be used to understand our uncertainty about the calibration $q^*$ in an number of ways.
+In cases where the visualization suggests significant variability, the MLM then suggests what subpopulations we should look at in order to resolve the ambiguities.
+For example, to resolve the ambiguities about ``Vip Igfbp6'' specimens, it appears it would be most important to obtain samples of subpopulations of cells which have many cells measured as ``Pvalb'' by technique II and very few cells measured as ``Sncg.''
 
 \figgy{
-\figgygr{0.8}{../images/ruxquant}
-\caption{Here we plot a histogram of $|\tilde q^{(1)}_i- \tilde q^{(2)}_i|$ among fifty BRUXP samples $\{d_i,\tilde q^{(1)}_i,\tilde q^{(2)}_i\}_{i\in 1\cdots 50}$.
-Each distance estimates the extent of our uncertainty about the calibration $q$ along the direction $d$.
-We can compare this with the extent of the uncertainty we had before we applied our method, i.e.
-the diameter of the set of probability matrices with these dimensions.
-In this case the diameter of that set is exactly 14.
-In this sense, the MLM has enabled us to roughly half our uncertainty about the calibration.
-\label{fig:ruxquant}}
+\hfill{}\figgygr{0.4}{../images/q1rLamp5Pdlim5}
+\figgygr{0.4}{../images/q1rVipIgfbp6}\hfill{}
+\caption{The MLM provides \emph{variability visualizations} for each row of the calibration $q$.
+In this figure we show two such examples.
+To produce the visualization we must first select a row of $q$.
+This corresponds to particular measurement value that can be obtained from tool I.
+In this case we select ``Lamp5 Pdlim5'' to produce the left visualization and ``Vip Igfbp6'' to produce the right visualization.
+The bottom of each visualization shows the value of $\hat q(y|x)$ for the selected type $x$ (it shows this through a heatmap with a single row).
+Here $\hat q$ is the Markov Link Method estimator we have described before.
+Above this original $\hat q$ value, we show a heatmap with many other possible examples for what else this row might look like.
+The variability among these possibilities gives a qualitative measure of our uncertainty about that particular row of the calibration $q$.
+\label{fig:allenboot}}
 }
 
-The distances between the paired vector in a BRUXP sample give us one kind of insight into the extent of our uncertainty.
-Recall that for each direction $d$ the BRUXP sample comprises the extremal vertex in the direction $d$ as well as the extremal vertex in the direction $-d$.
-The distance between these two vertices gives a kind of measure for the extent of our uncertainty.
-We plot a histogram of these distances in Figure \ref{fig:ruxquant}.
-This figure suggest that this extent is around 7 (although we caution that general results on polytope diameter estimation are not encouraging, cf.
-\citep{brieden1998approximation}).
-By contrast, the diameter of the entire space of valid probability matrices is $14$.
-In this sense, the Markov Link Method enables us to half our uncertainty about the calibration.
+The MLM also yields estimates for $h^*$.
+As in our simulation section, these quantities could also be estimated using more traditional pseudocount methods.
+We can compare the MLM estimates with the traditional methods, using the method of log-likelihood on held-out data.
+We find that both methods achieve a rate of 1.3 nats per entry on held out data.
+This gives us confidence that the MLM assumption might be right; if the MLM assumption was wrong our estimation process could force $h$ to take on an incorrect value.
+This is no proof that the MLM assumption is correct for this data, but it is a good sanity-check.
 
-\figgy{
-\figgygr{0.6}{../images/VipRspo4}
-\caption{Each column represents a possible manner in which method I and method II are calibrated for cells which are designated as type `Vip Rspo4' by method I.
-Together, the columns form a notion of our uncertainty about this aspect of the calibration.
-This plot shows that cells designated as `Vip Rspo4' by method I may be designated by method II as taking on various `Vip'-related celltypes and or some of the `nxx' celltypes.
-We cannot be sure which of these columns more faithfully represents the true calibration between the methods.
-\label{fig:ruxqual}}
-}
-
-Qualitatively, we can inspect the BRUXP samples by focusing on a single value of $X$ at a time.
-For example, in Figure \ref{fig:ruxqual} we focus on the distribution of $Y$ given that method I measures a specimen as having the type `Vip Rspo4.'   In this figure we look at various BRUXP samples, lining up all the pairs into a single heatmap.
-We see that cells designated as type `Vip Rspo4' by method I may be associated with one of the `Vip'-related celltypes in method II or an `n53' or `n57' type (recall that these types indicate a group of possible cell types, of which `Vip Rspo4' is one).
-For example, it may be that method II would identify such cells as `Vip Htr1f' cells.
-This is potentially concerning.
-On the other hand, the plot above suggests that these cells are \emph{not} ever identified as `Vip Lmol\_1' cells.
-That is, method II is able to accurately discern between cells designated as `Vip Rspo4' by method I and cells designated as `Vip Lmol\_1' by method I.
-This is encouraging.
-To further resolve the various remaining ambiguities in the calibration, we would need to perform additional experiments involving a subpopulation that might be likely to include `Vip Rspo4' cells but no `Vip Htr1f' cells.
-
-Let us next look at the BRUXP samples for cells designated by tool I as having the `VISp L5 Endou' type; these samples are fairly unequivocal.
-Figure \ref{fig:ruxqual2} gives strong support for the notion that the methods are very well-calibrated with respect to this cell-type.
-In all of the BRUXP samples we see that a cell designated as `VISp L5 Endou' by method I would be likely to be designated with the same type with method II.
-
-\figgy{
-\figgygr{0.6}{../images/Endou}
-\caption{Each column represents a possible manner in which the methods are calibrated for cells designated by method I as being type 'VISp L5 Endou.'  Notice that in every column there is strong signal at the corresponding type in method II.
-This suggests a good calibration.
-\label{fig:ruxqual2}}
-}
-
-In the interests of trying to disentangle our uncertainty due to insufficient data from our uncertainty due to identifiability, we repeat the histogram-of-distances analysis on RUXP samples.
-Recall that RUXP samples are just like BRUXP samples, only we do not perform the resampling with replacement.
-This RUXP histogram gives us a lower bound on the diameter of the set $\Theta(\hat p,\hat q)$.
-We expect this diameter to be smaller than the extent we measured in the first analysis: the RUXP analysis only captures uncertainty due to identifiability whereas the BRUXP analysis also captures our uncertainty due to insufficient data.
-This is indeed what we find.
-Figure \ref{fig:ruxquant2} shows the result of the histogram-of-distances analysis on the RUXP samples.
-We see distances are generally closer to 6 and never as high as 7.
-Recall that 7 was the typical distance found in the original BRUXP analysis.
-
-\figgy{
-\figgygr{0.6}{../images/ruxquant2}
-\caption{Here we plot a histogram of $|\tilde q^{(1)}_i- \tilde q^{(2)}_i|$ among fifty RUXP samples $\{d_i,\tilde q^{(1)}_i,\tilde q^{(2)}_i\}_{i\in 1\cdots 50}$.
-Each distance estimates the extent of our uncertainty about the calibration $q$ along the direction $d$ that is due to identifiability issues alone.
-These distances do not measure the extent of our uncertainty due to insufficient data.
-We can compare this with the kinds of distances we saw in Figure \label{fig:ruxquant2}, which were generally larger because they measured uncertainty due to both insufficient data and identifiability issues.
-\label{fig:ruxquant2}}
-}
-
-With more samples, we expect that we could get our original BRUXP analysis in Figure \ref{fig:ruxquant} to look more like the RUXP analysis in Figure \ref{fig:ruxquant2}.
-However, it appears that more samples might not be sufficient to bring the diameter of our uncertainty below 6.
-For that, we would need different \emph{kinds} of experiments, using different subpopulations.
-For example, as we saw above, we could reduce some of the identifiability concerns with experiments using a new subpopulation that included `Vip Rspo4' cells but not `Vip Htr1f.'
-
-Finally, in this special case there is an additional analysis that is appropriate.
-The team producing method I and method II designed these methods to have corresponding types.
-The hope was that a cell designated as `Vip Rspo4' in method I would also be designated as `Vip Rspo4' in method II.
-As we have seen above, we cannot be exactly sure what the calibration is.
-Is it perhaps possible that the calibration is nearly perfect?
-We can place a kind of upper bound on the precision of the calibration by looking for the \emph{best possible} calibration that is consistent with the data we have observed.
-In particular, let $d$ denote the direction of ``perfect calibration'' in which method I and method II agree perfectly.
-We can find the value of $q \in \Theta(\hat p,\hat p \hat q)$ which is most extreme in this direction.
-
-\figgy{
-\figgygr{.8}{../images/bestq}
-\caption{The value of $q$ that yields the most precise correspondence between the two cell types while remaining consistent with the observed data.
-This gives a kind of upper bound on how well the methods may be calibrated.
-As we have seen throughout, there is an extent to which we simply cannot be sure what the calibration is.
-Within all this uncertainty, this figure shows something like the rosiest possible picture of how well the methods might be calibrated.
-The strong diagonal components in this calibration show a good correspondence, i.e.
-cells designated as a particular type under method I are likely to be associated with the same type in method II.
-We caution that there is no reason to suppose that this \emph{is} the true calibration, but this figure shows that the data does not rule out that possibility.\label{fig:bestq}}
-}
-
-We show this `best possible $q$' in Figure \ref{fig:bestq}.
-The result is very encouraging: this best $q$ suggests a strong calibration between the two methods.
-We emphasize that our method does not prove this calibration by any means; as we have seen, the data only allow us to know that the calibration up to a considerable degree of uncertainty.
-However, we can conclude that the observed data does not \emph{rule out} the possibility that the methods are actually quite well calibrated on the whole.
-Of course, even for this best $q$ we do see certain miscalibrations that might be worth further investigation.
-For example, it appears that cells designated as `Vip Serpinf1' by method I might be designated as type `Lamp5 Slc35d3' by method II.
+Finally, the MLM provides quantitative values to give a sense for the the error in our estimators, namely $\Delta_p,\Delta_q,\Delta_h,\bar d$.
+These can be difficult to interpret because they are single numerical quantities that summarize error in the whole of $p,q,h$.
+As we have seen from our visualizations, certain aspects of $q$ may be fairly tightly identified with very little error whereas others may not be clear at all.
+Numerical figures must somehow ``average'' over all these complications.
+For this reason, the qualitative visualizations may be more helpful.
+However, if a quantitative measurement is needed, they are readily available: we find that $\Delta_p=6.7,\Delta_q=50.3,\Delta_h=3.7,\bar d=18.0$.
 
@@
 However, assuming this death rate can be roughly measured, it can be adjusted for, yielding a different but equally meaningful assumption about the data.
-Indeed, if the MLM method gives insensible results, this could actually serve as a useful clue that this disproportionate cell death is happening.
-
-Another potential direction for future work is suggested by the `nxx' types found in method II of our empirical investigation.
-These types are meant to indicate some uncertainty about the exact cell type.
-Each one is associated with a family of cell types.
-Our analysis did not need to rely on this information to proceed.
-However, making better use of it could considerably improve our results.
-Developing a rigorous framework to incorporate these kinds of structured data would be a worthy endeavor.
+Indeed, if the MLM gives insensible results, this could actually serve as a useful clue that this disproportionate cell death is happening.
 
@@
 By probing this set carefully, we can learn what the data actually has to say and what experiments we need to do to learn more.
+
+\section{The Markov Link Method}
+
+\label{sec:mlm}
+
+Here we describe in detail the Markov Link Method, the identifiability problems which it must overcome, and some theoretical guarantees for the method.
+For the benefit of the reader, we repeat the basic inputs to the method:
+
+\begin{itemize}
+    \item Samples $(\ell_1,X_1)\cdots (\ell_n,X_n)$ sampled such that $\mathbb{P}(X_i=x|\ell_i=\ell)= p^*(x|\ell)$
+    \item Samples $(\ell_{n+1},Y_{n+1})\cdots (\ell_{n+m},Y_{n+m})$ sampled such that $\mathbb{P}(Y_i=x|\ell_i=\ell)= h^*(x|\ell)$
+\end{itemize}
+
+We assume that $\ell,X,Y$ are discrete random variables with finite support. We also assume that the MLM assumption holds, i.e.\ $h^*(x|\ell)=\sum_y p^*(x|\ell)q^*(x|y)$ for some unknown $q^*$.
+
+The method proceeds as follows:
+
+\begin{enumerate}
+\item Summarize the input data, with two matrices:
+%
+\begin{align*}
+N^X_{\ell x} &= |\{i\leq n:\ \ell_i=\ell,X_i=x\}|\\
+N^Y_{\ell y} &= |\{i>n:\ \ell_i=\ell,Y_i=x\}|
+\end{align*}
+%
+That is, $N^X_{\ell x}$ indicates the number of specimens from subpopulation $\ell$ which yielded measurement $x$ under tool I.
+Likewise $N^Y_{\ell y}$ indicates the number of specimens from subpopulation $\ell$ which yielded measurement $y$ under tool II.
+
+\item Produce point estimates. 
+\begin{itemize}
+\item For $p$, we use a simple pseudocount estimator, namely
+\[
+\hat p = \argmax_p \sum_{\ell,x} (N^X_{\ell x}+1) \log p(x|\ell)
+\] 
+\item For $h$, we take a similar idea.
+However, in this case, we make use of the Markov Link Assumption to increase the effficiency of our estimator.
+Let $S$ denote the set of values of $h$ which are consistent with $p$ and the Markov Link Assumption, i.e.
+\[
+S=\left\{h:\ \exists q:\ \sum_x \hat p(x|\ell)q(y|x)=h(y|\ell) \quad \forall \ell,y\right\}
+\]
+And then we define
+\[
+\hat h = \argmax_{h\in S} \sum_{\ell,y} (N^Y_{\ell y}+1) \log h(x|\ell)
+\] 
+\item Finally, we estimate $q$.  Towards this end, we define $\Theta$ as the set of values of $q$ which are consistent with $p,h$ and the Markov Link Assumption.  That is, let 
+\begin{align}\label{eq:thetadef}
+\Theta(p,h) \triangleq \left\{q:\ \sum_x p(x|\ell)q(y|x)= h(y|\ell) \quad \forall \ell,y \right\}
+\end{align}
+Any $q$ inside $\Theta$ should be considered equally plausible, since it yields the same distribution on the observable data.
+However, for the purposes of determining variability due to lack of data, we take a very particular estimator of $q$, namely the so-called ``analytic center'' of $\Theta$.
+\[
+\hat q = \argmax_{q\in\Theta(\hat p,\hat h)} \sum_{x,y} \log q(y|x)
+\]
+This gives us a single estimator for $q$, which we will use below.  
+\end{itemize}
+Several remarks are in order about these point estimates.  
+\begin{itemize}
+\item The pseudocount.
+An obvious alternative to the pseudocount estimator would be to simply use the maximum likelihood estimator, i.e.
+$\argmax_p \sum_{\ell,x} N^X_{\ell x} \log p(x|\ell)$.
+However, in the datasets we examined, there was simply not enough data to support this kind of estimator.
+Some of the subpopulations had very few samples, requiring some degree of smoothing.
+Pseudocounts are just one among a host of ways this smoothing could have been achieved.
+However, we chose pseudocounts because they are easy to interpret both mathematically and by scientists.
+We simply find the maximum likelihood estimator \emph{as if} the data had been that given by the pseudocount-augmented values.
+We here chose a pseudocount value of 1, but in practice an expert can meaningfully evaluate what kinds of pseudocounts each sample category should take.
+%
+\item We do use the Markov Link Assumption to estimate $h$.
+We note that an alternative way to estimate $h$ would be to simply use a pseudocount estimator just as we did to estimate $p$.
+However, if the Markov Link Assumption holds, we should be able to use this fact to get a superior estimators by enforcing this constraint.
+We saw in simulations that this does indeed occur at least in some cases.
+%
+\item We don't use the Markov Link Assumption to estimate $p$.
+The process outlined above follows three distinct stages: estimate $p$, estimate $h$, then estimate $q$.
+This process has the advantage that each individual problem is convex, with a unique solution which can be found with classic methods from convex optimization (see Appendix \ref{sec:convexopt} for details).
+However, there is a fundamentally different way we could seek our point estimates.
+We could try to \emph{simultaneously} optimize all of these objects to maximize the likelihood of the observables (subject to the Markov Link Assumption).
+Unfortunately, we were unable to get rigorous guarantees for the asymptotic properties of such a method.
+It is for this reason that we took this multi-stage approach.
+\end{itemize}
+%
+\item Produce quantitative estimates for the variance of our estimators, using the bootstrap.
+We a series of $K$ surrogate datasets $(N^{X,(1)},N^{Y,(1)})\cdots (N^{X,(K)},N^{Y,(K)})$ by sampling with replacement within each subpopulation from the original data.
+For each surrogate $(N^{X,(i)},N^{Y,(i)})$ we use a procedure similar to the one described above to produce point estimates $\hat p^{(i)},\hat q^{(i)},\hat h^{(i)}$.
+The only difference between the bootstrap point estimators and the original point estimators is that we use no pseudocount for the bootstrap estimators.
+This is because the pseudocount can in some cases make the estimator variance look smaller than it is.
+Finally, for each variable, $p,q,h$, we compute the average total variation distance between our original point estimates and these bootstrap estimates.
+Here by total variation distance we mean for example that $\TV{\hat q- \hat q^{(i)}} \triangleq \frac{1}{2}\sum_{x,y}|\hat q(y|x)-\hat q^{(i)}(y|x)|$.
+We denote these variation distance averages by $\Delta_p,\Delta_q,\Delta_h$.
+%
+\item Produce quantitative estimates for our uncertainty about $q^*$ due to identifiability issues.
+In some cases it may be that $q^*$ is not identifiable from the data.
+That is, even in the limit of an infinite number of specimens, it may be impossible to determine the exact value of $q^*$.
+This is because we can only observe $q^*$ through the distribution on $Y|\ell$ (i.e.\ $h^*$).
+Thus, if there are two values $q_1,q_2$ such that $h^*(y|\ell)=\sum_x p^*(x|\ell)q_1(y|x)=\sum_x p^*(x|\ell)q_2(y|x)$, then without joint measurement on $(X,Y)$ we would never be able to tell if $q^*=q_1$ or $q^*=q_2$.
+Nonetheless, even if $q^*$ cannot be determined exactly, it is possible to put bounds on what values of $q^*$ are consistent with the data: the Markov Link Method assumption does say that $q^* \in \Theta(p^*,h^*)$, where $\Theta$ is given by Equation (\ref{eq:thetadef}).
+This can tell us something about what the calibration $q^*$ looks like.
+What is more, subject to mild conditions, we can show that in the limit of infinite data, $q^*$ is very close to some point in $\Theta(\hat p,\hat q)$.
+This is the subject of Theorem \ref{thm:mainthm} in Appendix \ref{sec:mainthm}.
+To understand our uncertainty about $q^*$ due to identifiability, we must therefore understand the size of this set.
+To do so, we use the simple total variation diameter approximation procedure (see Appendix \ref{sec:diameterestimation} for details) to estimate the diameter for the set associated with each of our bootstrap samples, $d^{(i)}\approx \mathrm{diam}(\Theta(\hat p^{(i)},\hat h^{(i)}))$.
+To estimate the error due to identifiability in our particular case, we use the median of these diameter estimates, which we denote by $\bar d$.
+%
+\item Produce qualitative variability visualizations.
+Fix any value of $x$.
+Now consider any bootstrap estimate $\hat p^{(i)},\hat q^{(i)},\hat h^{(i)}$.
+For each $y$ we can compute the value of $q\in \Theta(\hat p^{(i)},\hat q^{(i)})$ with the maximal value of $q(y|x)$.
+We call this an ``extremal bootstrap estimate,'' \
+\[
+q^{(i,x,y)} = \sup_{q\in\Theta(\hat p^{(i)},\hat q^{(i)})} q(y|x)
+\]
+We can look at all of these estimates simulatenously by forming a square matrix $M_{y_1,y_2} = q^{(i,x,y_1)}(y_2|x)$ and plotting this matrix as a heatmap.
+If there are no identifiability issues concerning the estimation of $q^*$, every row of the matrix will be the same.
+If there are significant identifiability issues, different rows may be quite different.
+We can repeat this process for each bootstrap estimate; this yields a very tall and thin matrix formed by stacking all the matrices together.
+If there is sufficient data, we expect each block of this matrix to be approximately the same (since the bootstrap resampling will have little effect).
+Thus, if there is sufficient data and there are no identifiability issues every row of the entire tall matrix will be similar.
+For visualization, we can squish the matrix so that each row is much thinner than each column.
+The result is a single image that captures the variability for one particular row of $q$.
+
+\end{enumerate}
+
+\section{Relation to prior work}
+
+Our main goal is to achieve calibration without joint measurement.
+The main technical obstacle is what is known in the statistics world as an `identifiability problem.'  Due to this problem, we often simply \emph{cannot} directly estimate the calibration that we are interested in.
+The calibration is simply not ``identifiable.''  Nonetheless, all is not lost.
+Our main contribution is to show that one can bound the extent of this identifiability problem.
+In short, the fundamental problem cannot generally be vanquished, but we can put it in its place.
+
+Our identifiability analysis stands on the shoulders of a long history of turning probabilistic assumptions into bounds on unidentifiable parameters.
+The core idea of the MLM is to take an assumption (the `Markov Link assumption,' which we will define shortly) and use it to place bounds on an unidentifiable quantity (namely the calibration $q^*$).
+When these bounds are fairly tight on all sides, we see that much can be learned even though what we want isn't identifiable.
+Much of the prior literature in this kind of direction comes from research into causality.
+For example, in \cite{bonet2001instrumentality} Bonet uses polytopes not unlike the ones seen here to explore whether a variable can be used as an instrument.
+The famous Clauser-Horne-Shimony-Holt inequality was designed to help answer causality questions in quantum physics, but it also sheds light on what distributions are consistent with certain assumptions \cite{clauser1969proposed}.
+Indeed the physics literature has contributed many key inequalities (cf.
+\cite{chaves2014inferring}, \cite{kela2017semidefinite}, and the references therein).
+Perhaps the closest work to this one would be \cite{makarov1982estimates}, which used two marginal distributions to get bounds on a property of the joint distribution (namely the distribution of the sum).
+We advance this approach to a more general-purpose technique, both by using many subpopulations to refine our estimates and by considering the entire space of possible joint distributions instead of simply a particular aspect of the joint.
+
+It is perhaps worth mentioning that the actual theorem presented here has probably been derived before (although we did not find it in our review of the literature).
+Other results in the same flavor as the one presented here might also find a practical use for modern problems.
+There is a treasure-trove of ideas in the causality literature; this wealth has not yet been brought fully to bear on the challenging and important problems of calibration for modern experimental modalities.
+Our primary intention with this article is to bring attention to the practical utility of these kinds of result.
 
@@
 
-\section{Proof of the theorem}
+\section{Mathematical results}
 
-\label{sec:proof}
-
-For the benefit of the reader, we here repeat the statement of our theorem in more explicit terms.
+\label{sec:mainthm}
 
@@
 Recall that in this norm matrices satisfy a Cauchy-Schwarz like equality, $\EN{pq}\leq\EN{p}\EN{q}$.
-Also recall that $\UN{a} \leq EN{a} \leq \sqrt{n}\UN{a}$ where $n$ is the number of entries in $a$.
+Also recall that $\UN{a} \leq \EN{a} \leq \sqrt{n}\UN{a}$ where $n$ is the number of entries in $a$.
 \item Let $T_{a,b}$ denote the transition matrix polytope, i.e. the set of $a\times b$ matrices whose rows sum to 1 and whose entries are all positive.
@@
 \item Let $q^*\in T_{|\Omega_X|,|\Omega_Y|}$.
+\item Let $h^*=p^*q^*$.
+Here by $p^* q^*$ we intend the matrix multiplication of $p^*$ and $q^*$.
+We will go back and forth between thinking of $p^*,q^*,h^*$ as matrices and thinking of them as distributions.
 \item We require the matrix $q^*$ has strictly positive entries, $q^*_{xy}\geq c>0$.
-\item We require that the rows of $p^*$ are linearly independent.  
-\item Let $\hat p$ denote an empirical transition matrix drawn by obtaining $N_{X,\ell}$ samples for each row of $p^*$, i.e.
-we have samples $(\ell_1,X_1) \cdots (\ell_1,X_n)$ such that $\mathbb{P}(X_i=x) = p^*_{\ell_i,x}$, $N_{X,\ell}=\sum_{i=1}^n \mathbb{I}_{\ell_i=\ell}$, and $\hat p_{\ell x} = \sum_{i=1}^n \mathbb{I}_{X_i=x,\ell_i=\ell} / N_{X,\ell}$.
-\item Let $\hat h$ denote an empirical transition matrix drawn by obtainin $N_{Y,\ell}$ samples for each row of $h^*=p^*q^*$.  
+\item For each $\ell$ draw $n_\ell$ samples from $p^*_\ell$ and $m_\ell$ samples from $h^*_\ell$.  Let $N^X_{\ell x},N^Y_{\ell y}$ summarize the resulting number of each kind of sample we obtained.
+\item We require that there are fewer rows of $p^*$ than columns, and each row is linearly independent.
+If there are more rows than columns, then there is generally no identifiability issue so the theorem below is not relevant.
+If there are fewer rows than columns but the rows are not independent, then the result applies after enough rows are combined.
+\item Let $\hat p_{\ell x} = (N^{X}_{\ell,x}+1) / (\sum_{x'}N^{X}_{\ell,x'}+1)$
+\item Let
+\[
+L(q,\kappa) = \sum_{y,\ell} (N^Y_{\ell y}+1) \log \left(\sum_x \hat p(x|\ell)q(y|x)\right)
+    +\kappa\sum_{x,y} \log q(y|x)
+\]
+and take $\hat q = \lim_{\kappa\rightarrow 0} \argmax_q L(q,\kappa)$.
+\item Let $\hat \Theta=\{q:\ \hat p\hat q=\hat pq\} \cap T_{|\Omega_X|,|\Omega_Y|}$.
 \end{itemize}
+\vspace{.1in}
 
-Now fix any $\kappa>0$.  Let
-\begin{align}
-\hat q = \argmax_q \left(\sum_\ell N_{Y,\ell}\sum_{y}\hat h(y|\ell)\log\left(\sum_{x}\hat p(x|\ell)q(y|x)\right) + \kappa \sum_{xy} \log q(y|x)\right)\label{eq:mindef}
-\end{align}
-and $\hat \Theta=\{q:\ \hat p\hat q=\hat pq\} \cap T_{|\Omega_X|,|\Omega_Y|}$.  
-\vspace{.1in}
-\begin{thm}
-If $N_{X,\ell},N_{Y,\ell}\rightarrow \infty$ in such a way that $N_{Y,\ell'}/\sum_{\ell}N_{Y,\ell} \geq \rho>0$ for each $\ell'$, then $\inf_{q\in \hat\Theta} \UN{q^*-q}\rightarrow 0$ in probability.
+\begin{thm}\label{thm:mainthm}
+If $n_\ell, m_\ell\rightarrow \infty$ in such a way that $m_\ell/m_{\ell'} \geq \rho>0$ for each $\ell,\ell'$, then $\inf_{q\in \hat\Theta} \UN{q^*-q}\rightarrow 0$ in probability.
 \end{thm}
@@
 It is well-known that $\hat p\rightarrow p^*$ in probability (in both the uniform or the Euclidean norm, which are of course equivalent in this case).
-It is easy (albeit technical) to see that the same goes for $\hat p\hat q\rightarrow h^*$ (see Lemma \ref{lem:easyconsistency}).
-Thus, intuitively, the difficulty is this: by ensuring $\UN{\hat p - p^*},\EN{\hat p - p^*},\UN{\hat p\hat q - h^*},\EN{\hat p\hat q - h^*}$ sufficiently small, can we find some $\tilde q\in \hat \Theta$ so that $\UN{\tilde q - q^*}$ is arbitrarily small?
+It is easy to see that the condition on $m_\ell$ together with the strict positivity of $q^*$ guarantee that the same goes for $\hat p\hat q\rightarrow h^*$.
+Thus, intuitively, the difficulty is this: by ensuring $\EN{\hat p - p^*},\EN{\hat p\hat q - h^*}$ sufficiently small, can we find some $\tilde q\in \hat \Theta$ so that $\EN{\tilde q - q^*}$ is arbitrarily small?
 It turns out we can.
 
-Recall that $c>0$ is the smallest value of $q^*_{xy}$.
+There is a simple method to find such a $\tilde q$: we take $\tilde q$ to be the Euclidean projection of $q^*$ to the plane defined by $\{q:\ \hat p\hat q=\hat pq\}$.
+If $p^* \approx \hat p$ and $\hat p \hat q \approx p^* q^*$, it is easy to see that we can can ensure this projection will not carry us a very large distance, and so $\tilde q$ and $q^*$ will be close, as desired.
+However, it might carry us to a point outside of the transition matrix polytope; in particular, after projection we might see some negative values.
+Fortunately, because we have insisted that the truth is strictly positive, we are guaranteed that this is not a problem in the asymptotic limit.
+
+Let us make this argument formal.
+We haved defined $c>0$ as the smallest value of $q^*_{xy}$.
 Fix any $\epsilon<c, p^*,q^*$.
@@
 
-\begin{lem}
-\label{lem:easyconsistency} If $N_{X,\ell},N_{Y,\ell}\rightarrow \infty$ in such a way that $N_{Y,\ell'}/\sum_{\ell}N_{Y,\ell} \geq \rho>0$ for each $\ell'$, then $\UN{p^*q^*-\hat p\hat q},\EN{p^*q^*-\hat p\hat q}\rightarrow 0$ in probability.
-\end{lem}
-\begin{proof}
-Our first task is to make a short study of the continuity of KL divergences on categorical distributions when the probabilities are bounded away from zero.
-Recall that we have insisted $q^*_{xy}\geq c>0$ for every $x,y$ -- and this also means $(pq^*)_{\ell y}\geq c$ for every $\ell,y$, since each row of $p$ is itself a probability distribution.
-Moreover, observe that the KL-divergence on $|\Omega_Y|$-dimensional distributions, $\kldiv{\hat r}{\tilde r} \triangleq \sum_{y} \hat r_y \log \hat r_y / \tilde r_y$, is \emph{uniformly} continuous on the space of such distributions whose minimum probability is greater than any fixed positive constant.
-It follows that the map $h,p,q \mapsto D(h_\ell||(pq)_\ell)$ is also uniformly continuous on a space where $h$ and $q$ are strictly greater than some fixed positive constant.
+\section{Total Variation Diameter Estimation}
 
-With this in hand, the remainder of the proof follows naturally, using the well-known results that empirical distributions are consistent, i.e.\ $\hat p \rightarrow p^*$ and $\hat h \rightarrow p^*q^*$ in probability.
+\label{sec:diameterestimation}
 
-Fix any $\epsilon,\pi$.
-Let $\delta$ the modulus of continuity in the norm $\UN{\cdot}$ at level $\epsilon\rho$ for the map $h,p,q \mapsto \kldiv{h_\ell}{(pq)_\ell}$ restricted to the domain where $h,q>c/2$.
-Select $N$ large enough so that $\frac{1}{N_{Y,\ell}} \kappa |\Omega_X||\Omega_Y| \log \frac{1}{c} < \epsilon$ for each $\ell$ and so that with probability at least $\pi$ we have that $\hat h,\hat p$ so that $\UN{\hat h-p^*q^*},\UN{\hat p-p^*}\leq \delta,\UN{\hat h - p^*q^*}<c/2$.
-Then, with probability $\pi$, we must have
-%
+We follow the algorithm of \cite{brieden2001deterministic} in estimating of the diameter of a convex polytope $C$.
+We note that diameter estimation is a difficult problem, and if the polytope is particularly pathological and high-dimensional it may be quite difficult to estimate the diameter accurately.
+Nonetheless, it is a practical solution to the difficulty at hand.
+
+Let $C \in \mathbb{R}^{n}$ denote a convex polytope (note that the $n$ here does not correspond to the number of samples $n$ in the main body of this paper).
+We will be interested in a certain kind of diameter of $C$, namely
 \begin{align*}
-|\kldiv{\hat h_\ell}{(\hat p \hat q)_\ell} - \kldiv{h^*_\ell}{(\hat p \hat q)_\ell}| \leq \rho \epsilon \\
-\kldiv{\hat h_\ell}{(\hat p q^*)_\ell} = |\kldiv{\hat h_\ell}{(\hat p q^*)_\ell} - \kldiv{(p^*q^*)_\ell}{(p^* q^*)_\ell}| \leq \rho\epsilon 
+\diam(C) =& \frac{1}{2}\sup_{x,y\in\Theta} \sum_{i=1}^n |x_i-y_i| \\
+              =& \frac{1}{2}\max_{z\in\{-1,1\}^n} \left(\sup_{x,y\in\Theta} \sum_{i=1}^n z_i(x_i-y_i)\right)\\
+              =& \frac{1}{2}\max_{z\in\{-1,1\}^n} \left((\sup_{x\in\Theta} z \cdot x) - (\inf_{x\in\Theta} z \cdot x) \right)
 \end{align*}
-%
-Now, since $\hat q$ is defined as the maximizer of a certain quantity (Equation \ref{eq:mindef}), we may be sure that it is greater than the same quantity evaluated at $q=q^*$.  That is,
-%
-\begin{align*}
-0 \leq& \sum_\ell N_{Y,\ell}\sum_{y}\hat h(y|\ell)\log\frac{\sum_{x}\hat p(x|\ell)\hat q(y|x)}{\sum_{x}\hat p(x|\ell)q^*(y|x)} + \kappa \sum_{xy} \log \frac{\hat q(y|x)}{q^*(y|x)}\\
- =& \sum_\ell N_{Y,\ell}\left(\kldiv{\hat h_\ell}{(\hat pq^*)_\ell} - \kldiv{\hat h_\ell}{(\hat p\hat q)_\ell}\right) + \kappa \sum_{xy} \log \frac{\hat q(y|x)}{q^*(y|x))}
-\end{align*}
-%
-Applying our continuity results, it follows that
-%
-\begin{gather*}
- \sum_\ell N_{Y,\ell}\left(\kldiv{\hat h^*_\ell}{(\hat p\hat q)_\ell}\right)  \leq 2\left(\sum_\ell N_{Y,\ell}\right)\epsilon  + \kappa |\Omega_X||\Omega_Y| \log \frac{1}{c} 
-\end{gather*}
-%
-Note that the left-hand summands are all positive.  So, in particular, it follows that for each $\ell'$, applying the uniformity condition $\rho$, we have
-%
-\begin{gather*}
-\kldiv{\hat h^*_{\ell'}}{(\hat p\hat q)_{\ell'}}  \leq 2\epsilon  + \frac{1}{N_{Y,\ell'}} \kappa |\Omega_X||\Omega_Y| \log \frac{1}{c} \leq 3\epsilon
-\end{gather*}
 
-That is, we have shown convergence of probability in the KL sense: for any $\epsilon,\pi$ we can find $N$ high enough so that $\kldiv{\hat h^*_{\ell'}}{(\hat p\hat q)_{\ell'}}<\epsilon$ with at least probability $\pi$.
-This, in turn yields convergence in probability in the Euclidean or uniform metrics by Pinsker's inequality.
-\end{proof}
+For any fixed $z$, the inner maximization and minimization problems can be solved using linear programming.
+We can therefore estimate the diameter by taking every possible $z$ and solving this problem.
+In practice, this may be computationally prohibitive.
+If so, we can randomly subsample values of $z$ to obtain a lower bound on the diameter.
+
+We can apply this to estimate the diameter of our set of interest, namely
+\[
+\Theta(p,h) = \left\{q:\ \sum_x p(x|\ell)q(y|x)=h(y|\ell)\right\}
+\]
+In this case $\Theta$ is a space of conditional probability distributions.  However, the constraints that define $\Theta$ are all linear, and so the same method can be applied.
+
+\section{Convex optimization problems}
+
+\label{sec:convexopt}
+
+To be determined
+
+\section{Permutation matrices are perfectly identified}
+
+\label{sec:miracle}
+
+To be written
 

### 2018-07-02 markovlink/__init__.py

@@ -1,7 +1,13 @@
 from . import estimation
 from . import exploration
+from . import mlm
+from . import misc
+from . import methodI
+from . import methodII
+from . import methodIII
+from . import majorization
 
-train_q = estimation.train_q
-sample_bootstrap_rux = exploration.sample_bootstrap_rux
-train_test_split = exploration.train_test_split
-get_rux_tildetheta=exploration.get_rux_tildetheta
+point_estimates = mlm.point_estimates
+bootstrap = mlm.bootstrap
+diameter_estimates = mlm.diameter_estimates
+qualitative_one_row = mlm.qualitative_one_row

### 2018-07-02 markovlink/estimation.py

@@ -1,60 +1,10 @@
 import numpy as np
-
 import sys
+import scipy as sp
+import scipy.special
+import itertools
+from . import misc
+import time
 
-def emupdate(p,q,kappa,Nly):
-    d1,d2=p.shape
-    d2,d3=q.shape
-    pi = p.reshape((d1,d2,1)) * q.reshape((1,d2,d3))
-    pi=pi/np.sum(pi,axis=1,keepdims=True)
-    
-    L=pi*Nly.reshape((d1,1,d3))
-    L=np.sum(L,axis=0) + kappa
-    
-    L = L / np.sum(L,axis=1,keepdims=True)
-    
-    return L
-
-
-def check_transition_matrix(a):
-    return (np.abs(np.sum(a,axis=1) - 1).max()<1e-9) and (np.min(a)>-1e-9)
-
-def train_q(p,Nly,iterations=5000,kappa=.0001,verbose=False):
-    '''
-    Uses EM algorithm to find the qhat which optimizes of the markov link method objective.
-
-    - p -- estimate empirical distribution of X|l (each row sums to 1.0)
-    - Nly -- Nly[l,y] is the number of times samples from subpopulation l yielded Y=y 
-        (each row sums to the number of samples for subpopulation l under technique II)
-    - kappa -- regularization 
-    - verbose -- if true, prints out updates as it trains to indicate progress
-    '''
-
-    Nly = np.require(Nly,dtype=np.float)
-    p=np.require(p,dtype=np.float)
-
-    assert len(Nly.shape)==2,"Nly should be a matrix"
-    assert len(p.shape)==2,"p should be a matrix"
-    d1,d2=Nly.shape
-    d1p,d3=Nly.shape
-    assert d1==d1p,'The first dimensions of Nly and p should be the same'
-    assert np.abs(np.sum(p,axis=1)-1).max()<1e-9,"The rows of p should sum to 1 and all entries should be positive"
-
-    # initialize qxy
-    qxy = np.ones((p.shape[1],Nly.shape[1])) / Nly.shape[1]
-
-    # run emupdate repeatedly
-    if verbose:
-        # ersatz tqdm replacement (didn't want to include tqdm as a requirement)
-
-        update_interval=1+iterations//20
-        for i in range(iterations): 
-            if i%update_interval==0:
-                sys.stdout.write('(%d/%d)'%(i,iterations))
-                sys.stdout.flush()
[…]

### 2018-07-02 markovlink/exploration.py

@@ -4,48 +4,63 @@ import numpy as np
 import numpy.random as npr
 import sys
 
-from . import estimation
+from . import misc
 
-def get_rux_tildetheta(d):
-    q1=np.zeros(d.shape)
-    q1[np.r_[0:d.shape[0]],np.argmax(d,axis=1)]=1
-    
-    q2=np.zeros(d.shape)
-    q2[np.r_[0:d.shape[0]],np.argmax(-d,axis=1)]=1
+###################################
+########## QUALITATIVE-1-ROW ######
+###################################
 
-    return d,q1,q2
+def qualitative_one_row(phat,qhat,x):
+    szL,szX=phat.shape
+    szX,szY=qhat.shape
 
-def sample_bootstrap_rux(Nlx,Nly,verbose=True,kappa=.01,use_cvxopt=False,direc=None,bootstrap=True):
+    block=np.zeros((szY,szY))
 
-    if direc is None:
-        direc=npr.randn(Nlx.shape[1],Nly.shape[1])
+    for y in range(szY):
+        direc=np.zeros((szX,szY))
+        direc[x,y]=1
+        block[y] = find_extremal(phat,qhat,-direc)[x]
 
-    # do a resample, get extremal
-    if bootstrap:
-        Nlx2=resample(Nlx)
-        Nly2=resample(Nly)
-    else:
-        Nlx2=Nlx
-        Nly2=Nly
-    plx = (Nlx2+1)/np.sum(Nlx2+1,axis=1,keepdims=True)
-    qxy=estimation.train_q(plx,Nly2,kappa=kappa)
-    OPT1=find_extremal(plx,qxy,direc,use_cvxopt=use_cvxopt)
+    return block
 
-    # do another resample, get another extremal
-    if bootstrap:
-        Nlx2=resample(Nlx)
-        Nly2=resample(Nly)
-    else:
-        Nlx2=Nlx
-        Nly2=Nly
-    plx = (Nlx2+1)/np.sum(Nlx2+1,axis=1,keepdims=True)
-    qxy=estimation.train_q(plx,Nly2,kappa=kappa)
-    OPT2=find_extremal(plx,qxy,-direc,use_cvxopt=use_cvxopt)
+##################################
+########## DIAMETER ESTIMAT ######
+##################################
 
-    return direc,OPT1,OPT2
+def diameter_estimation(phat,qhat,n_samps=50):
+    '''
[…]

### 2018-07-02 markovlink/majorization.py

@@ -1,0 +1,57 @@
+import numpy as np
+import sys
+import scipy as sp
+import scipy.special
+import itertools
+from . import misc
+import time
+
+def train_q_by_majorization(p,Nly,kappa,q=None,n_train=10):
+    '''
+    Maximizes
+
+    L(q) = sum_{l,y} Nly log(sum_x plx qxy) + sum_{x,y} kappa log(qxy)
+
+    by a majorization algorithm
+
+    '''
+
+    # initialize
+    Nly=np.require(Nly)
+    l,x =p.shape
+    l,y = Nly.shape
+
+    assert misc.check_transition_matrix(p),"The rows of p should sum to 1 and all entries should be positive"
+    if q is None:
+        q=np.ones((x,y),dtype=np.float64)/y  
+    else:
+        assert misc.check_transition_matrix(q),"The rows of q should sum to 1 and all entries should be positive"
+
+    # run iterations
+    for i in range(n_train):
+        # update
+        q = train_mlm_em_update(Nly,p,q,kappa)
+       
+    return q
+
+def train_mlm_em_update(Nly,p,q,kappa):
+    '''
+    Uses a majorization technique to find a new value of q with a higher value of 
+    
+        L(q) = np.sum(Nlx * np.log(p @ q)) 
+    
+    starting from q as an initial guess.
+
+    '''
+    l,x =p.shape
+    x,y = q.shape
+    
+    # compute the majorization
+    pi = p.reshape((l,x,1)) * q.reshape((1,x,y))
+    pi = pi /np.sum(pi,axis=1,keepdims=True)
+    
+    # make updates
+    newq = np.sum(Nly.reshape((l,1,y)) * pi,axis=0).reshape(x,y)+kappa
+    newq = newq/np.sum(newq,axis=1,keepdims=True)
+    
+    return newq.reshape((x,y))

### 2018-07-02 markovlink/methodI.py

@@ -1,0 +1,90 @@
+import numpy as np
+import sys
+import scipy as sp
+import scipy.special
+import itertools
+from . import misc
+import time
+from . import estimation
+
+def train_q_primaldual_initialize(p,Nly,kappa,q):
+    szL,szY=Nly.shape
+
+    lam = Nly / (p@q)
+    gamma = -np.sum(p.T*(q@lam.T),axis=1) - szY*kappa
+
+    return gamma,lam
+
+def train_q_primaldual_residual(p,Nly,kappa,q,gamma,lam):
+    szL,szY=Nly.shape
+    szL,szX=p.shape
+
+    return np.r_[(q*(p.T@lam+gamma.reshape((-1,1)))).ravel()+kappa,np.sum(q,axis=1)-1,(lam*(p@q)-Nly).ravel()]
+
+def train_q_primaldual_update(q,gamma,lam,direc,s):
+    szL,szY=lam.shape
+    szX,szY=q.shape
+
+    q=q+s*direc[:szX*szY].reshape(q.shape)
+    gamma=gamma+s*direc[szX*szY:szX*szY+szX]
+    lam=lam+s*direc[szX*szY+szX:].reshape(lam.shape)
+
+    return q,gamma,lam
+
+def train_q_primaldual_jacobian(p,Nly,kappa,q,gamma,lam):
+    szL,szY=Nly.shape
+    szL,szX=p.shape
+
+    def iQ(x,y):
+        return np.ravel_multi_index((x,y),q.shape)
+    def iG(x):
+        return szX*szY+x
+    def iL(l,y):
+        return szX*szY + szX + np.ravel_multi_index((l,y),Nly.shape)
+
+    jac = np.zeros((szX*szY + szX + szL*szY,szX*szY + szX + szL*szY))
+
+    # rQ
+    for x in range(szX):
+        for y in range(szY):
+            jac[iQ(x,y),iQ(x,y)] = np.sum(lam[:,y]*p[:,x])+gamma[x]
+            jac[iQ(x,y),iG(x)] = q[x,y]
+            for l in range(szL):
+                jac[iQ(x,y),iL(l,y)] = q[x,y]*p[l,y]
+
+    # rG
+    for x in range(szX):
+        for y in range(szY):
+            jac[iG(x),iQ(x,y)]=1
+
+    # rL
+    for l in range(szL):
+        for y in range(szY):
+            jac[iL(l,y),iL(l,y)]=np.sum(p[l]*q[:,y])
+            for x in range(szX):
[…]

### 2018-07-02 markovlink/methodII.py

@@ -1,0 +1,68 @@
+import numpy as np
+import sys
+import scipy as sp
+import scipy.special
+import itertools
+from . import misc
+import time
+from . import estimation
+
+def train_q_primaldual_initialize(p,Nly,kappa,q):
+    szL,szY=Nly.shape
+
+    lam = Nly / (p@q)
+    gamma = -np.sum(p.T*(q@lam.T),axis=1) - szY*kappa
+
+    return gamma,lam
+
+def train_q_primaldual_residual(p,Nly,kappa,q,gamma,lam):
+    szL,szY=Nly.shape
+    szL,szX=p.shape
+
+    return np.r_[(p.T@lam+gamma.reshape((-1,1))).ravel()+kappa/q.ravel(),np.sum(q,axis=1)-1,(lam*(p@q)-Nly).ravel()]
+
+def train_q_primaldual_update(q,gamma,lam,direc,s):
+    szL,szY=lam.shape
+    szX,szY=q.shape
+
+    q=q+s*direc[:szX*szY].reshape(q.shape)
+    gamma=gamma+s*direc[szX*szY:szX*szY+szX]
+    lam=lam+s*direc[szX*szY+szX:].reshape(lam.shape)
+
+    return q,gamma,lam
+
+def train_q_primaldual_jacobian(p,Nly,kappa,q,gamma,lam):
+    szL,szY=Nly.shape
+    szL,szX=p.shape
+
+    def iQ(x,y):
+        return np.ravel_multi_index((x,y),q.shape)
+    def iG(x):
+        return szX*szY+x
+    def iL(l,y):
+        return szX*szY + szX + np.ravel_multi_index((l,y),Nly.shape)
+
+    jac = np.zeros((szX*szY + szX + szL*szY,szX*szY + szX + szL*szY))
+
+    # rQ
+    for x in range(szX):
+        for y in range(szY):
+            jac[iQ(x,y),iQ(x,y)] = -kappa/q[x,y]**2
+            jac[iQ(x,y),iG(x)] = 1
+            for l in range(szL):
+                jac[iQ(x,y),iL(l,y)] = p[l,y]
+
+    # rG
+    for x in range(szX):
+        for y in range(szY):
+            jac[iG(x),iQ(x,y)]=1
+
+    # rL
+    for l in range(szL):
+        for y in range(szY):
+            jac[iL(l,y),iL(l,y)]=np.sum(p[l]*q[:,y])
[…]

### 2018-07-02 markovlink/methodIII.py

@@ -1,0 +1,157 @@
+import numpy as np
+import sys
+import scipy as sp
+import scipy.special
+import itertools
+from . import misc
+import time
+from . import estimation
+
+def guess_gamma(Nly,p,q,kappa):
+    h=p@q
+
+    szX,szY=q.shape
+    szL,szX=p.shape
+    p=p.reshape((szL,szX,1))
+    q=q.reshape((1,szX,szY))
+    Nly=Nly.reshape((szL,1,szY))
+    pi = Nly*p*q
+    pi = pi/np.sum(pi,axis=1,keepdims=True)
+    newq=np.sum(pi*Nly,axis=0) + kappa
+    return np.sum(newq,axis=1)
+
+def gam_objective(Nly,p,q,kappa,gamma):
+    return np.sum(Nly*np.log(p@q)) + kappa*np.sum(np.log(q)) - np.sum(q.T@gamma)
+
+def calculate_surrogate_grad_and_hessian(Nly,p,q,kappa,gamma):
+    szX,szY=q.shape
+    szL,szX=p.shape
+
+    h=p@q
+
+    hess=np.zeros((szY,szX,szX))
+    for y in range(szY):
+        A = (np.sqrt(Nly)/h)[:,y][:,None]*p # ls x xs
+        hess[y] = - kappa*np.diag(q[:,y]**(-2)) - A.T@A
+
+    grad = p.T@(Nly/h) + (kappa/q) - gamma[:,None]
+
+    # perversely, the gradient means we should take a gradient step in the direction of grad.T
+    # because lam is (L x Y) whereas grad is (Y x L)
+
+    return grad,hess
+
+def calculate_surrogate_search_direction(Nly,p,q,kappa,gamma):
+    grad,hess=calculate_surrogate_grad_and_hessian(Nly,p,q,kappa,gamma,)
+    return -np.array([np.linalg.solve(h,g) for (h,g) in zip(hess,grad.T)]).T
+    # return -np.array([sp.sparse.linalg.lsqr(h,g,damp=0.4,iter_lim=2000)[0] for (h,g) in zip(hess,grad.T)]).T
+
+def update_q_by_surrogate(Nly,p,q,kappa,gamma,searchdirection,beta=.8):
+    szX,szY=q.shape
+    szL,szX=p.shape
+
+    curobjective = gam_objective(Nly,p,q,kappa,gamma) # want to make this big
+
+    # go forward, but not too much
+    s=1.0
+    working=False
+    while not working:
+        newq = q + searchdirection*s
[…]

### 2018-07-02 markovlink/misc.py

@@ -1,0 +1,14 @@
+
+import numpy as np
+import sys
+import scipy as sp
+
+def pnn(x): 
+    sys.stdout.write(str(x) + " ")
+    sys.stdout.flush()
+
+def check_transition_matrix(a):
+    return (np.abs(np.sum(a,axis=1) - 1).max()<1e-9) and (np.min(a)>-1e-9)
+
+def totalvardist(a,b):
+    return .5*np.sum(np.abs(a-b))

### 2018-07-02 markovlink/mlm.py

@@ -1,0 +1,75 @@
+from . import estimation
+from . import exploration
+from . import misc
+import numpy as np
+from . import majorization
+
+def point_estimates(Nlx,Nly,pseudocount=1.0):
+    '''
+    Get point estimates by the Markov Link Method
+    '''
+
+    p = (Nlx + pseudocount) / np.sum(Nlx+pseudocount,axis=1,keepdims=True)
+
+    kappa=.001
+    q=majorization.train_q_by_majorization(p,Nly+pseudocount,kappa,n_train=500)
+
+    h=p@q
+
+    return p,q,p@q
+
+def bootstrap(Nlx,Nly,phat,qhat,n_straps=50,verbose=False):
+    '''
+    Draws surrogate datasets for Nlx and Nly, with replacement, and then gets point estimates
+    for the surrogate datasets using the Markov Link Method.
+
+    Also estimates the Deltap,Deltaq,Deltah
+    '''
+
+    hhat=phat @ qhat
+
+    straps=[]
+
+    dsts=np.zeros((n_straps,3))
+
+    if verbose:
+        misc.pnn("%d bootstraps:"%n_straps)
+    for i in range(n_straps):
+        if verbose:
+            misc.pnn(i)
+        Nlx2=exploration.resample(Nlx)
+        Nly2=exploration.resample(Nly)
+
+        p,q,h =point_estimates(Nlx2,Nly2,pseudocount=0.0)
+
+        dsts[i,0] = .5*np.sum(np.abs(p-phat))
+        dsts[i,1] = .5*np.sum(np.abs(q-qhat))
+        dsts[i,2] = .5*np.sum(np.abs(h-hhat))
+
+        straps.append(dict(Nlx=Nlx2,Nly=Nly2,estimates=(p,q,h)))
+
+    return straps,np.mean(dsts,axis=0)
+
+def diameter_estimates(straps,verbose=False):
+    diams=[]
+
+    if verbose:
+        misc.pnn("%d bootstraps to consider:"%len(straps))
+    for i,strap in enumerate(straps):
+        if verbose:
+            misc.pnn(i)
+        p,q,h = strap['estimates']
+        diams.append(exploration.diameter_estimation(p,q)[0])
+
+    return diams
+
+def qualitative_one_row(straps,x,verbose=False):
+    Q1Rs=[]
+    if verbose:
+        misc.pnn("%d bootstraps to consider:"%len(straps))
+    for i,strap in enumerate(straps):
[…]

### 2018-07-19 paper/paper.tex
Changed
Latest and greatest

@@
 \newcommand{\figgygr}[2]{\includegraphics[width=#1\textwidth]{#2}}
-% \newcommand{\figgygr}[2]{picture here}
+\newcommand{\figgygrs}[3]{\begin{tabular}{c}\includegraphics[width=#1\textwidth]{#2}\\#3\end{tabular}}
+\newcommand{\figgygrsTable}[3]{\begin{tabular}{c}{#2}\\#3\end{tabular}}
+\newcommand{\figgygrsH}[3]{\begin{tabular}{c}\includegraphics[width=#1\textwidth,angle=270]{#2}\\#3\end{tabular}}
+
+% \newcommand{\figgygr}[2]{pic}
+% \newcommand{\figgygrs}[3]{\begin{tabular}{c}pic\\#3\end{tabular}}
+% \newcommand{\figgygrsTable}[3]{\begin{tabular}{c}{#2}\\#3\end{tabular}}
+% \newcommand{\figgygrsH}[3]{\begin{tabular}{c}pic\\#3\end{tabular}}
+
 \newcommand{\figgy}[1]{\begin{figure}\fbox{\begin{minipage}{\textwidth}#1\end{minipage}}\end{figure}}
@@
 Let us say we obtain measurement result $x$ from one tool on a particular specimen.
-The number $q(y|x)$ indicates the probability of obtaining result $y$ from a second tool applied to measure the same specimen.
+The number $q^*(y|x)$ indicates the probability of obtaining result $y$ from a second tool applied to measure the same specimen.
 One way to learn the calibration is to measure the same specimens with both tools.
@@
 If each subpopulation captures a different slice of the overall population, we can obtain tight bounds on the true calibration.
-This is true even if each subpopulation is highly heterogeneous.
+This is true even if the subpopulations are highly heterogeneous and overlapping.
 By integrating information from all the subpopulations we can make rigorous deductions about what the calibration might be.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The Markov Link Method assumption}
+\figgy{
+\hfill{}\figgygrs{.5}{../images/bigpicture}{(a) Setup: each specimen can be \\ measured with either tool I or tool II, \\ but can't be measured with both.\\ How do we calibrate the tools?}
+\hfill{}\figgygrs{.4}{../images/plate}{(b) Key modeling assumption: \\ the calibration $q$ is the same \\ for every subpopulation. }\hfill{}
 
-The MLM is only applicable if a certain assumption is met.
-In this section we will articulate this assumption as clearly as possible.
-Let us say we are considering the specimens of a large population.
-For example, each specimen in the large population might be a human cell.
-Each specimen could also be be a piece of metal which needs to be tested.
-We will assume there are three basic properties of interest for each specimen $i$:
-
-\begin{enumerate}
-\item $\ell_i$, the subpopulation.
-We assume that we can identify a variety of these subpopulations.
-They are not required to be disjoint subpopulations; it is only necessary that each subpopulation captures a different slice of the overall population.
-For example, we could define subpopulations by looking at cells in different parts of the body or cells with different sizes.
-\item $X_i$, the result of measuring a specimen with tool I.  For example, perhaps tool I takes a picture of the cell with a destructive electron microscopy method.
-\item $Y_i$, the result of measuring a specimen with tool II.  For example, perhaps tool II measures the RNA expression of the cell with a destructive sequencing method.
-\end{enumerate}
-
-We here consider the case that ``joint measurement'' is impossible or impractical.
-In terms of the notation above, that means that for any given specimen we can either observe $\ell_i, X_i$ or $\ell_i,Y_i$.
-We can never observe $\ell_i,X_i,Y_i$ for any specimen $i$.
-We would like to estimate as much as possible about the distribution of $X,Y|\ell$.
-In this paper, we will consider the case that this estimation is facilitated by an additional assumption:
-
-\figgy{
-\hfill{}\figgygr{.7}{../images/bigpicture}\hfill{}
-\caption{We consider the case that for each specimen there are two ways we might measure it.
-We divide the specimens into two groups.
-The specimens on the left are measured with technique I and the specimens on the right are measured with technique II.
-In the end, we are left wondering: what would have happened if we had measured the specimens on the left with technique II?
-Or what if we had measured the specimens on the right with technique I?
-The Markov Link Method gives a way to answer this problem using a subpopulation analysis.
+\caption{{\bf The Markov Link Method assumption makes calibration possible, even without joint measurement}.
+In subfigure (a) we measure some neurons with tool I and others with tool II.
+How can we calibrate the tools?
+That is, how can we know what we might have seen if we had measured the first group with tool II or the second group with tool I?
+The Markov Link Method (MLM) gives a way to get a partial answer, using a subpopulation analysis.
+In subfigure (b) we present a plate diagram describing our beliefs about the conditional independencies in this situation; cf.\ \cite{koller2009probabilistic} for a thorough explanation of this type of diagram.
+For each specimen we are interested in $\ell$ (subpopulation), $X$ (tool-I measurement), and $Y$ (tool-II measurement).
+For $n$ samples we can see $(\ell,X)$ and for $m$ samples we can see $(\ell,Y)$.
+However, for the first group of $n$ samples we cannot observe $Y$, and for the second group of $m$ samples we canot observe $X$.
+This is indicated in the diagram by the grayed-out circles.
+This lack of joint measurement makes it hard to learn the variable distributions.
+The MLM has a key assumption that makes it possible: we assume $X$ is sufficiently detailed to \emph{statistically isolate} $Y$ from $\ell$.
+This assumption is reflected in the arrows of the diagram; all paths from $\ell$ variables to $Y$ variables pass through $X$ variables.
+The assumption allows us to learn much about the important probability distributions, as parametrized by the tool-I measurement distribution $p(x|\ell)$ and the calibration $q(y|x)$.
 \label{fig:bigpicture}}
 }
+
+This paper will proceed in the following sections:
+\begin{itemize}
+    \item Description of how to apply the Markov Link Method, using an example to explain the process
+    \item A variety of examples where the method may apply
+    \item The method's performance in simulation
+    \item The method's results on real-world data
+    \item Relation to prior work
+    \item Discussion
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%            _           
+%  _ __ ___ | |_ __ ___  
+% | '_ ` _ \| | '_ ` _ \ 
+% | | | | | | | | | | | |
+% |_| |_| |_|_|_| |_| |_|
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{The Markov Link Method}
+
+\label{sec:mlm}
+
+We begin with an example.
+
+Imagine that we have a large collection of human cells.  We will call each cell a `specimen.'  For each specimen we have:
+%
+\begin{enumerate}
+    \item $\ell$, the sampling strategy used to obtain the specimen.
+We will also call $\ell$ the specimen ``subpopulation,'' for each strategy should capture a different subpopulation of the overall set of specimens.
+For example, we could have a few strategies based on selecting cells of particular sizes, a few strategies based on selecting cells from particular places in the body, and a few strategies based on based on selecting cells with particular proteomic markers.
+Each additional sampling strategy shows us a different subpopulation of the cells and strengthens what kinds of inferences are possible.
+    \item $X$, the cell's `transcriptomic type,' as measured by looking at the transcriptomic activity in the cell.  
+    \item $Y$, the cell's `morphological type,' as measured by looking at an image of the cell.
+\end{enumerate}
+%
+Our task is to reconcile the two different notions of `cell-type' indicated by $X$ and $Y$, by constructing what we call a `calibration.'  This calibration should tell us what transcriptomic types are associated with what morphological types.
+For example, we might like to be able to say ``This cell has transcriptomic type 3, therefore it probably has morphological type F.''  We can make math of this idea by defining a calibration as a conditional probability $q^*(y|x,\ell)$.
+This object indicate the likelihood that the morphological type is $y$ given that the transcriptomic type is $x$ and the specimen was sampled with strategy $\ell$.
+In these terms, our task is to learn the calibration $q^*$.
+
+Learning $q^*$ is hard when ``joint measurement'' is impossible or impractical.
+By this we mean that for any given specimen we can either observe $\ell, X$ or $\ell,Y$.
+We can never observe $\ell,X,Y$ for any specimen.
+This makes it much more difficult to estimate the calibration; the fundamental difficulty is sketched in Figure \ref{fig:bigpicture}.
+To make calibration possible without joint measurement, we need an assumption additional assumption.
+The Markov Link Method supplies this assumption in the form of a conditional independence assumption.
+This makes it possible to learn about the calibration.
+
+To make it possible to learn $q^*$ even without joint measurement, the MLM makes the assumption that the calibration is the same for every sampling strategy:
 
@@
 \begin{center}
-\textbf{The Markov Link Method Assumption}\\$\mathbb{P}(Y=y|X=x,\ell)=q^*(y|x)$ for every value of $\ell$.
+\textbf{The Markov Link Method Assumption}\\$q^*(y|x,\ell)=q^*(y|x)$ for each $\ell$
 \end{center}
@@
 
-Intuitively, this signifies that the manner in which $X$ predicts $Y$ would be the same for each subpopulation.
-In the language of statistics, the assumption is that the variable $X$ is `sufficient' for $\ell$.
-If this assumption is not met, then the method presented here is not applicable.
-If it \emph{is} met, it opens the door for new methods to estimate three important objects:
+This assumption can be understood through the following thought experiment.
+Let us say we are told that a cell has transcriptomic type $x$.
+We could then make predictions about the morphological type $y$.
+Now we are told new information: the cell was obtained via sampling strategy $\ell$.
+Does this change our predictions about the morphological type?
+If so, the MLM assumption is violated.
+However, if $X$ is sufficiently informative then learning the sampling strategy will not change our predictions about the morphological type.
+In this case, the MLM assumption holds.
 
-\begin{itemize} 
-    \item The calibration $q^*(y|x)$ 
-    \item The distribution of tool-I measurement values found in each subpopulation, $p^*(x|\ell)$ 
-    \item The distribution of tool-II measurement values found in each subpopulation,  $h^*(y|\ell)$.  
-\end{itemize}
-
-Under the MLM assumption, it is straightforward to see that these three objects are all bound together by the formula $h^*(y|\ell) = \sum_x p^*(x|\ell)q^*(y|x)$ for each $\ell,y$.
-Using this equation, we will see that it is possible to estimate properties of the calibration $q^*$ even without access to joint measurement.
-It also makes it possible to perform more efficient estimation of $h^*$.
-Theoretically we could also get more efficient estimation of $p^*$, but we did not find plausible cases where this actually occured.
+If the MLM assumption holds, a frequentist point of view suggests the following procedure for estimating the calibration $q^*$ from data:
 
 \figgy{
-\hfill{}\figgygr{.7}{../images/plate}\hfill{}
-\caption{Another way to articulate the Markov Link Method is to look at the graphical model associated with the MLM assumption.
-This reveals that we have two kinds of data.
-In one case (above), we can observe $\ell,X$, but $Y$ is hidden.
-For other kinds of data (below) we can observe $\ell,Y$, but $X$ is hidden.
-All of this data is joined by the unknown parameters $p,q$.
-Here $p$ indicates the distribution of tool-I measurement values from each subpopulation, and $q$ indicates the calibration.
-Our task is to use these different kinds of data to determine the parameters $p,q$ as best as we can.
-Once these parameters are known, notice that distribution of tool-II measurement values for each subpopulation, which we denote $h$, can be determined immediately using the Markov Link Assumption: $h^*(y|\ell) = \sum_x p^*(x|\ell)q^*(y|x)$.
-\label{fig:platediag}}
+\hfill{}\figgygr{.9}{../images/qtildesamps}\hfill{}
+\caption{{\bf Estimating the calibration and measuring our uncertainty}.
+\label{fig:qtildesamps} On the left we visualize our best guess for true the calibration between two different tools for measuring cell-type.
+Tool I uses one approach to obtain a fine-grained classification of cells, whereas tool II uses a different approach and only gives coarse type indication.
+We can visualize our estimated calibration with a heatmap, formed by drawing a rectangle for each $x,y$ with a color indicating the magnitude of $\tilde q(y|x)$.
+For example, examine the bright yellow square on the `Pvalb Gpr149' row and the `unknown' column.
+This suggests that if a cell has the `Pvalb Gpr149' type according to tool I, it is extremely likely to be assigned the `unknown' type by tool II.
+However, our best guess might not be correct.
+On the right we show many plausible alternatives for the `Pvalb Gpr149' row of the calibration.
+Each possibility is drawn from a 95\% confidence region for calibration.
+Some of the alternatives suggest that some `Pvalb Gpr149' cells might not be classified as `unknown,' but might instead be correctly identified as being of the `Pvalb' type.
+The data we have cannot tell us which of these alternatives is closest to the truth.
+More data and more subpopulations would be necessary to reduce the variability within this confidence region.
+However, we can have some confidence that `Pvalb Gpr149' cells are not being classified as `Meis2' or `Vip', insofar as none of the plausible calibrations give high probability to these tool II types.}
 }
 
+\begin{enumerate}
+    \item Summarize our observed data in two matrices:
+    \begin{align*}
+    N^X_{\ell x} &= |\{i\leq n:\ \ell_i=\ell,X_i=x\}|\\
+    N^Y_{\ell y} &= |\{j\leq m:\ \ell_j=\ell,Y_j=x\}|
+    \end{align*}
+    That is, $N^X_{\ell x}$ indicates the number of specimens from subpopulation $\ell$ with transcriptomic type $x$.
+Likewise $N^Y_{\ell y}$ indicates the number of specimens from subpopulation $\ell$ which with morphological type $y$.
+
+    \item Define a confidence region for calibration $q^*$.
+To do this we incorporate a nuisance parameter: let $p^*(x|\ell)$ denote the probability that a cell sampled with strategy $\ell$ will have transcriptomic type $x$.
+We form a joint simultaneous confidence region for all of $p^*,q^*$ using the following likelihood ratio statistic:
+    \[
+    L(p,q)=\frac{1}{\Omega_X}\sum_{\ell,x} N^X_{\ell x} \log \frac{\hat p(x|\ell)}{p(x|\ell)} + \frac{1}{\Omega_Y}\sum_{\ell,y} N^Y_{\ell y} \log \sum_x \frac{\hat h(y|\ell)}{p(x|\ell) q(y|x)}
+    \]
+    where $\hat p(y|x)=N^X_{\ell x} / \sum_{x'}N^X_{\ell x'}$ and $\hat h(y|x)=N^Y_{\ell y} / \sum_{y'}N^Y_{\ell y'}$ denote empirical conditional distributions.
+We then define our confidence region as
+    \[
+    R = \{(p,q):\ L(p,q) < k\}
+    \]
+    and choose $k$ to ensure any desired coverage.
+For example, to produce a 95\% confidence region, we would like to choose $k$ so that the probability that $p^*,q^* \in R$ is at least 95\%.
+We refer the reader to Appendix \ref{sec:confinterval} for details on choosing $k$.
+
+    \item Determine whether the confidence region is empty.
+If the region is empty then we can reject the MLM assumption hypothesis with a significance level of one minus the region coverage.
+Assuming we fail to reject, we can continue:
+
+    \item Sketch out the the confidence region, using examples.
+In most nontrivial cases the region $R$ is high-dimensional and difficult to inspect directly.
+We therefore produce examples from this region to try to help the user understand what values of $q$ are plausible.
+In particular, we sample $Z_{x,y} \sim \mathrm{Exponential}(1)$ for each $x,y$ and solve the problem
+    \[
+    \max_{p,q \in R} \quad \sum_{x,y} Z_{x,y} \log q(y|x) \\
+    \]
+    This problem does not have a closed-form solution, but can generally be solved using an iterative procedure; we describe our method in detail in Appendix \ref{sec:optproblems}.
+The result is a particular value of $q$ which lies within $R$.
+In practice, we find these examples are almost always on the boundary of $R$.
+We can then repeat this process by resampling $Z$.
+By repeating this process many times, we can get a large collection of plausible values of $q$.
+For any fixed value of $x$ we can visualize these samples by using a heatmap to plot $q(\cdot |x)$ for all the samples.
+See Figure \ref{fig:qtildesamps} for an example.
+
+    \item Produce a point estimate.
+It is not generally possible to produce a consistent point estimate.
+Indeed, there are many situations in which the confidence region $R$ does not concentrate even in the limit of infinite data.
+Nonetheless, if a point estimate is desired, one may pick point estimates $\tilde p,\tilde q$ by solving the minimization problem $\min L(p,q)$.
+We caution that this problem does not always have a unique solution, so the result may depend upon the minimization strategy used.
+Precise details for our algorithm may be found in Appendix \ref{sec:optproblems}.
+\end{enumerate}
+
+A Bayesian point of view would suggest a different procedure; this point of view may provide a useful complementary perspective and is worthy of future research.
+However, we caution that it is nontrivial to choose reasonable prior distributions for $p^*,q^*$.
+Priors which may seem uninformative (such as the Dirichlet distribution) can actually carry significant and unanticipated weight, due to the high-dimensional nature of the problem.
+We give an example in \ref{sec:bayesian}.
+
+The frequentist confidence region allows us to avoid making any modeling assumptions.
+However, we are then faced with the task of understanding a high-dimensional confidence region.
+As described above, our solution is to look at examples from inside the region.
+We emphasize that these examples are not drawn uniformly within the confidence region; instead the examples illustrate something more like the boundary of the region, sketching out what might be possible.
+We concede that this method is somewhat ad-hoc and hope that future research may discover a superior method to visualize this high-dimensional confidence region.
+For now, it appears to be a practical solution.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                 _           
+%   _____  ____ _ _ __ ___  _ __ | | ___  ___ 
+%  / _ \ \/ / _` | '_ ` _ \| '_ \| |/ _ \/ __|
+% |  __/>  < (_| | | | | | | |_) | |  __/\__ \
+%  \___/_/\_\__,_|_| |_| |_| .__/|_|\___||___/
+%                          |_|               
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Examples where the Markov Link Method may apply}
+
+\label{sec:examples}
 
@@
 Even if the error is correlated, the MLM assumption may still hold.
-For example, imagine that the $Y$ error is correlated with the absolute reliability of the part, this may pose no problem if that reliability is adequately measured by $X$.
+For example, imagine that the $Y$ error is correlated with the absolute reliability of the part; this may pose no problem if that reliability is adequately measured by $X$.
 
@@
 The result is two completely independent classifications of cells, one for each way of looking at the cell.
-MLM allows us to estimate the relationship between those two classification systems, yielding a wholistic understanding of the different types of cells.
+MLM allows us to estimate the relationship between those two classification systems, yielding a holistic understanding of the different types of cells.
 In this case, $\ell$ might indicate some side information such as where in the body the cell was found, $X$ would indicate a detailed classification of the cell according to its transcriptomics, and $Y$ would indicate a coarser classification of a cell according to its morphology.
@@
 $X$ would indicate the image and $Y$ would indicate a caption.
-If caption is largely determined by the picture $X$, the MLM assumption may hold.
+If the caption is largely determined by the picture $X$, the MLM assumption may hold.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%      _                 _       _   _                 
+%  ___(_)_ __ ___  _   _| | __ _| |_(_) ___  _ __  ___ 
+% / __| | '_ ` _ \| | | | |/ _` | __| |/ _ \| '_ \/ __|
+% \__ \ | | | | | | |_| | | (_| | |_| | (_) | | | \__ \
+% |___/_|_| |_| |_|\__,_|_|\__,_|\__|_|\___/|_| |_|___/
+%                                                    
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The Markov Link Method: input and output}
 
-The Markov Link Method is a procedure from estimating the distribution $X,Y|\ell$ from data, under the Markov Link Method assumption.
-Details on the exact algorithm and convergence results are given in Section \ref{sec:mlm}.
-For now, let us content ourselves with looking at the input and the output of the MLM:
+\section{Simulation results}
 
-\begin{description}
-    \item[Input] We assume we are given $n$ samples using tool I and $m$ samples using tool II.
+\label{sec:simulations}
+
+We use simulations to test the capabilities of the Markov Link Method to estimate a calibration.  We will:
+
 \begin{itemize}
-        \item Samples $(\ell_1,X_1)\cdots (\ell_n,X_n)$ sampled such that $\mathbb{P}(X_i=x|\ell_i=\ell)= p^*(x|\ell)$
-        \item Samples $(\ell_{n+1},Y_{n+1})\cdots (\ell_{n+m},Y_{n+m})$ sampled such that $\mathbb{P}(Y_i=x|\ell_i=\ell)= h^*(x|\ell)$
+    \item Pick parameters to be `ground truth': 
+    \begin{itemize}
+        \item A calibration, $q^*(y|x)$.  
+        \item A tool-I measurement distribution, $p^*(x|\ell)$.  Recall that this indicates the probability that tool I measures $x$ for a specimen gathered with sampling strategy $\ell$. 
+        \item A tool-II measurement distribution, $h^*(y|\ell)$.
+This indicates the probability that tool II measures $y$ for a specimen gathered with sampling strategy $\ell$.
+We assume the the MLM assumption holds, and so this can be calculated from $q^*,p^*$ by the formula $h^*(y|\ell)=\sum_x p^*(x|\ell)q(y|x)$.
     \end{itemize}
-    We assume that $\ell,X,Y$ are discrete random variables with finite support.
-We also assume that the MLM assumption holds, i.e.\ $h^*(x|\ell)=\sum_y p^*(x|\ell)q^*(x|y)$ for some unknown $q^*$.
-It would be interesting to generalize the method to non-discrete random variables, but we leave this for future work.
-Note that if the random variables are low-dimensional and continuous, the MLM can still be applied by discretizing the continuous space into a finite number of bins.
-    \item[Output] Our goal is to estimate $p^*,q^*,h^*$ from the samples.  In this effort, we return:
-    \begin{itemize}
-        \item Point estimates, $\hat p,\hat q,\hat h$.  
-        \item Quantitative estimates of the variability of these point estimates due to insufficient data, $\Delta_p,\Delta_q,\Delta_h$.  
-        \item An quantitative estimate of the variability of $\hat q$ due to so-called `identifiability' issues, $\bar d$.  
-        \item Variability visualizations indicating a set of ``plausible'' values for each row of $q$; this gives a qualitative counterpart to the quantitative measures of variability.
+    \item Select a number of samples to take per subpopulation, $s$, and produce a simulated dataset for each subpopulation $\ell$:
+    \begin{align*}
+    (N^X_{\ell 1},N^X_{\ell 2}\cdots N^X_{\ell \Omega_X}) \sim \mathrm{Multinomial}(s,p^*(\cdot | \ell))\\
+    (N^Y_{\ell 1},N^Y_{\ell 2}\cdots N^Y_{\ell \Omega_Y}) \sim \mathrm{Multinomial}(s,h^*(\cdot | \ell))
+    \end{align*}
+    \item Apply the Markov Link Method to the generated samples to obtain a calibration estimate $\tilde q$.
+    \item Compute the error of the estimate $\tilde q$, as measured by a kind of averaged total variation distance, namely
+    \[
+    \frac{1}{2\Omega_{\ell}} \sum_{\ell x} |\tilde q(y|x) - q^*(y|x)|
+    \]
+    This distance ranges between zero and one.
+A distance of zero indicates that $\tilde q=q^*$, and a distance of one indicates that the two calibrations place mass on completely disjoint sets, i.e.
+$q(y|x)=0$ whenever $\tilde q(y|x)>0$ and vice-versa.
 \end{itemize}
-    Precise definitions and mathematical results are given in section \ref{sec:mlm}.
-\end{description}
+
+\figgy{
+\hfill{}\figgygrs{.4}{../images/simulationIandIIpq.png}{(a) `Ground truth' for $p^*,q^*$.}\hfill{}
+\hfill{}\figgygrs{.5}{../images/simulationsIandII.png}{(b) MLM performance \\ in estimating $q^*$.}\hfill{}
+
+\caption{{\bf The Markov Link Method accurately estimates the calibration, as long as there are enough subpopulations}.
+\label{fig:simsIandII}  We pick parameters as `ground truth' for $p^*,q^*$; these are as shown in subfigure (a) as heatmaps.
+We then create simulated datasets using these parameter values, use the Markov Link Method to produce its estimator, and see and measure the total variation error of the estimator.
+This error ranges from zero to one, where zero indicates no error and one indicates complete disagreement.
+In subfigure (b) we see that the MLM is able to estimate the ground truth with negligible error, as long as it has enough samples.
+However, we also consider another simulation: we merge some of the subpopulations in the ground truth parameters so that they become indistingiushable.
+In this degraded setup, the MLM is unable to learn the true calibration no matter how many samples we have.
+}
+}
+
+For the first simulation, we will look at a case in which the MLM performs well.
+We will assume there are twenty different samping strategies, and tools I and II both measure specimens as being in one of six categories.
+Our particular choices of ground truth can be found in Figure \ref{fig:simsIandII}.
+This figure also shows that the MLM is able to accurately estimate the ground truth calibration $q^*$, as long as there are enough samples.
+
+We also used this simulation to test the coverage of the confidence region.
+Recall that the `actual coverage' of a confidence interval indicates the probability that a confidence region contains the ground truth parameters.
+This coverage depends upon the particular choice of ground truth.
+The `nominal' coverage indicates the intended coverage of the interval, and may be more or less than the actual coverage.
+We estimated the actual coverage for the ground truth we have chosen.
+Taking 200 sample datasets with $s=80$ samples per subpopulation, we found that an MLM confidence region with 95\% nominal coverage has an actual coverage of about $99\%$.
+In this sense our region is `conservative,' which was somewhat by design.
+We refer the reader to Appendix \ref{sec:confinterval} for details.
+We also investigated other values of $s$, finding that the model became less conservative with more samples.
+For example, for datasets $s=800$ with the actual coverage appears to be approximately $97.5\%$.
+
+In our second simulation, we consider a case where the MLM does not succeed in accurately estimating the calibration.
+We use the same ground truth parameters as in first simulation, but we will merge groups of sampling strategies.
+In particular, we will divide the sampling strategies into four groups, each with five members.
+We will pretend that we cannot distinguish subpopulations within any given group.
+In this case, Figure \ref{fig:simsIandII} shows that the Markov Link Method cannot accurately estimate the calibration, no matter how many samples we have.
+This is due to a fundamental identifiability issue which is discussed in detail in Appendix \ref{sec:casestudies}.
+
+Despite the fact that we could not get an accurate point estimate for the calibration in the second simulation, the confidence regions appear to remain on solid ground.
+We took 200 simulated datasets with $s=400$ samples per subpopulation (1600 samples total) and computed MLM confidence regions with $95\%$ nominal coverage.
+We found an actual coverage of about $96.5\%$.
+
+\figgy{
+\hfill{}\figgygr{.7}{../images/simulationIII}\hfill{}
+
+\caption{{\bf The error of the MLM estimator depends on the nature of $q^*$.}  In our third simulation we fix a particular $p^*$ and consider a range of $q^*$s.
+For each possible $q^*$ we generate a certain number of samples per subpopulation and then try to use the resulting dataset to estimate $q^*$.
+On the one hand, we consider the case that $q^*$ causes $X,Y$ to be independent; in this case we can get a good estimates of $q$ with enough data.
+On the other hand, we consider the case that $q$ causes $X,Y$ to be deterministically bound, i.e.
+$q^*(y|x)=1$ if $x=y$ and $q(y|x)=0$ otherwise.
+In this case it is also possible to estimate $q$ with enough data.
+In-between, it is not possible to exactly determine $q^*$, even in the asymptotic limit of infinite data.\label{fig:estimatingq}}
+}
+
+For our final simulation, we will dig deeper into the question of when the calibration can be perfectly estimated with enough data.
+To this end, we will consider cases in which the number of subpopulations is four, and there are six possible outcomes for both measurement tools.
+For many choices of $p^*,q^*$, this will make it \emph{impossible} to correctly determine the calibration $q^*$ without joint measurement.
+However, for certain values of $q^*$ the truth can be recovered.
+To see this, we will look at a variety of calibrations.
+On the one hand, we will consider the case that $X,Y$ are independent, i.e.
+$q^*(y|x)=q^*(y|x')$ for every $x,x'$.
+On the other hand, we will consider the case that $X,Y$ are deterministically related, in particular $X=Y$.
+We will also consider every calibration ``in-between'' these two extremes.
+These in-between calibrations are found by linear interpolation.
+For each calibration we will generate datasets of various sizes and try to estimate the calibration from the datasets using the MLM.
+In Figure \ref{fig:estimatingq}, we see that in the two extreme cases it may be possible to recover the calibration.
+However, for in-between cases it is indeed not possible.
+This is true even in the limit of infinite data.
+We leave a complete mathematical understanding of this result to future work.
+For the purposes of this paper, we content ourselves by noting a special case of particular interest: that $X$ and $Y$ can take on one of $2^k$ values and are related to each other by an invertible deterministic function.
+In this case, with enough enough samples of only $k+1$ different sampling strategies, we can determine both that the relationship is deterministic and the exact specification of the invertible function.
+This result is proven in Appendix \ref{sec:casestudies}, Theorem \ref{thm:miracle}.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                 _     _       _        
+%  _ __ ___  __ _| | __| | __ _| |_ __ _ 
+% | '__/ _ \/ _` | |/ _` |/ _` | __/ _` |
+% | | |  __/ (_| | | (_| | (_| | || (_| |
+% |_|  \___|\__,_|_|\__,_|\__,_|\__\__,_|
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Simulation results}
 
-Before we show the results of the MLM on real data, let us look at some simulation results to see what the MLM can and cannot achieve.  In particular, we will:
+\section{Empirical results for cell-types}
 
-\begin{itemize}
-    \item Pick a set of possible parameter values for $p^*,q^*,h^*$.
-    \item Generate data according to these distributions, namely $n$ samples according to $p^*$ and $m$ samples according to $h^*$.  
-    \item Pretend to forget the true values of $p^*,q^*,h^*$.
-    \item Apply the Markov Link Method to the generated samples.
-    \item Compare the results of the MLM with the original true values.
-\end{itemize}
-
-Our purpose is to see how well the method performs if all of its assumptions are met. We will measure performance two ways:
-
-\begin{enumerate}
-    \item First, we will measure the error of the estimators $\hat p,\hat q,\hat h$, i.e.\ how much they differ from the true generating distributions $p^*,q^*,h^*$.
-For $h$ we will compare this measurement error against a standard pseudocount estimator (we will not do the same for $p$, because the MLM estimator for $p$ \emph{is} a pseudocount estimator).
-We refer the reader to Section \ref{sec:mlm} for details on this estimator.
-    \item Second, we will look at how well the MLM has estimated its uncertainty about the estimates.
-We know that there will always be some amount of error.
-For this reason, the MLM offers two tools to estimate the amount of error that may be present.
-In this section we will particularly focus on the quantities $\Delta_p,\Delta_q,\Delta_h$ (which approximately indicate variability due to insufficient data) and $\bar d$ (which approximately indicates variability due to so-called identifiability concerns).
-For a precise explanation of these two sources of variability, we refer the reader to Section \ref{sec:mlm}.
-\end{enumerate}
-
-For our first simulation, we will look at a case in which the MLM performs well.
-We consider twenty different subpopulations (which we will label $\{1,2,3,\cdots 20\}$).
-We will assume that tool I always returns an integer measurement in $\{1,2,3,4,5,6\}$ and tool II returns a measurement in $\{1,2,3,4,5,6\}$.
-We will assume we have one hundred samples for each subpopulation for each method.
-Some results of this simulation can be seen in Figure \ref{fig:simulation1gt}.
-This figure shows that the MLM gives us an edge in estimating $h$, performing better than a traditional pseudocount estimator.
-It also allows us to estimate $q$ with decent accuracy (which is only possible because of the Markov Link Assumption).
-The figure also shows that the $\Delta$ variability measures seem to accurately predict the estimation error.
-The $\bar d$ for this simulation was identically zero, because there were no identifiability issues in this case; this is related to the fact that there were more subpopulations than there were possible values that tool I could return.
+We now apply the MLM to real data.
 
 \figgy{
-\hfill{}\figgygr{.7}{../images/simulation1Groundtruth}\hfill{}
-
-\vspace{.1in}
-
-\hfill{}\begin{tabular}{r|p{.3in}p{.7in}p{1.0in}}
-Variable & MLM error & Pseudocount error & $\Delta$ \\
-\hline
-$p$ & 1.5 & N/A & 1.6 \\
-$q$ & 0.6 & N/A & 0.6 \\
-$h$ & 0.9 & 1.4 & 1.0 \\ 
-\end{tabular}\hfill{}
-
-\caption{For our first simulation, we imagine that the true values of $p^*,q^*,h^*$ are indicated by the first row of heatmaps above.
-The color of each entry indicates a probability between 0 and 1, with 0 being represented by black and 1 being represented by white.
-For example, the entry at the third row and the fifth column of the $q$ plot indicates $\mathbb{P}(Y=5|X=3)=q^*(5|3)$.
-The second row of heatmaps indicates the estimates for these quantities returned by the MLM.
-We can also look at this more quantitatively, as shown in the table.
-For each variable, $p,q,h$, we can compute the total variation distance between our estimates and the truth.
-For $h$ we can also compare the error we get when using a more standard pseudocount estimator.
-Finally, in the right-most column of the table, we show the $\Delta$ variability measures for each variable.
-These are intended to roughly indicate the level of error we would expect due to insufficient data.
-They are caclulated using a bootstrap method.
-We can see that the MLM variability measured do seem to give a good indication of the actual error level.
-\label{fig:simulation1gt}}
-}
-
-For our second simulation, we will look at a case where the MLM does a good job of estimating $p,h$, but does not succeed in accurately estimating $q$.
-We use the same data as in the previous simulation, but we will merge groups of subpopulations into four larger subpopulations.
-This will make it easier to estimate $h(y|\ell),p(x|\ell)$, since we have fewer subpopulations and for each one we will have five times as much data.
-The results are shown in Figure \ref{fig:simulation2gt}.
-As we would expect, estimation error of $p,h$ improves.
-However, estimation of $q$ does not improve at all.
-In fact, it is considerably worse.
-To understand this, we run a further simulation with the same true distributions but a virtually infinite number of samples.
-Even in this case, we see that accurate estimation of $q$ is still not achieved; we end up with a final error of $0.79$.
-This significant error is due to a fundamental identifiability issue which is discussed in detail in Section \ref{sec:mlm}.
-Accurate estimation of $q$ is simply impossible in this case, regardless of the number of samples.
-This is related to the fact that the number of subpopulations is smaller than the number of different values that tool I can report.
-What is worse, the $\Delta_q$ reports only $.01$ for the predicted error of $\hat q$.
-It is for this reason we must also look at $\bar d$, which approximates the error due to identifiability.
-For this simulation we calculate that $\bar d=3.5$, which captures the fact that our estimator $\hat q$ may suffer from nontrivial identifiability errors.
-Indeed, it appears to substantially overestimate the total error.
-This problem is somewhat unavoidable without additional assumptions.
-A precise statement of the theoretical guarantees regarding the ability of the MLM to estimate its own error due to identifiability issues can be found in Section \ref{sec:mlm}.
-I
-
-\figgy{
-$500$ samples per population:
-\hfill{}\begin{tabular}{r|p{.7in}p{1.0in}}
-Variable & MLM error & $\Delta$ \\
-\hline
-$p$ & 0.12 & 0.16 \\
-$q$ & 1.33 & 1.19 \\
-$h$ & 0.17 & 0.15 \\ 
-\end{tabular}\hfill{}
-
-\vspace{.2in}
-
-$10^{10}$ samples per population:
-\hfill{}\begin{tabular}{r|p{.7in}p{1.0in}}
-Variable & MLM error & $\Delta$ \\
-\hline
-$p$ & $10^{-7}$ & $10^{-3}$ \\
-$q$ & $0.79$ & $0.01$ \\
-$h$ & $10^{-3}$ & $10^{-3}$ \\
-\end{tabular}\hfill{}
-
-\caption{Our second simulation is similar to our first.
-However, instead of twenty subpopulations, we will assume we only have five subpopulations with four times the amount of data for each subpopulation.
-This makes estimating $p,h$ easier, but it doesn't make estimating $q$ any easier.
-\label{fig:simulation2gt}  Even when we give ourselves a virtually unlimited number of samples, perfect estimation of $q$ is just not possible.
-The $\Delta$ measures of the MLM do \emph{not} accurately reflect this fact.
-It is for this reason that the MLM additionally provide the quantity $\bar d$.
-For this simulation $\bar d=3.5$, which allows us to know that there are identifiability concerns, and also to bound the magnitude of those concerns.}
-}
-
-For our next simulation, we will dig deeper into the question of when $q$ can be perfectly estimated.
-To this end, we will consider cases in which the number of subpopulations is four, and there are six possible outcomes for both measurement tools.
-In the general case, this will make it \emph{impossible} to correctly determine $q$ without joint measurement.
-However, for certain values of $q$ the truth can be recovered.
-To see this, we will look at a variety of values of $q$.
-On the one hand, we will consider the case that $X,Y$ are independent, i.e.
-$q(y|x)=q(y|x')$ for every $x,x'$.
-On the other hand, we will consider the case that $X,Y$ are deterministically related, in particular $X=Y$.
-We will also consider every $q$ ``in-between'' these two extremes.
-These in-between $q$s are found by linear interpolation.
-For each such $q$ we will generate datasets of various sizes and try to estimate $q$ from the datasets using the MLM.
-The results are to be found in Figure \ref{fig:estimatingq}.
-These results show how in the two extreme cases it may be possible to recover $q$, but for in-between cases it is indeed not possible.
-This is true even in the limit of infinite data.
-We leave a complete mathematical understanding of this result to future work.
-For the purposes of this paper, we content ourselves with noting that if $X$ and $Y$ can take on one of $2^k$ values and are related to each other by an invertible deterministic function, then both the fact that the relationship is deterministic and the exact specification of the invertible function can be determined with only $k+1$ properly chosen subpopulations (at least in the asymptotic limit of infinite data).
-This result is proven in Appendix \ref{sec:miracle}.
-
-\figgy{
-\hfill{}\figgygr{.7}{../images/estimatingq}\hfill{}
-
-\caption{In our third simulation we fix a particular $p$ and consider a range of $q$s.
-For each possible $q$ we generate a certain number of samples per subpopulation and then try to use the resulting dataset to estimate $q$.
-On the one hand, we consider the case that $q$ causes $X,Y$ to be independent; in thie case we can get a good estimate of $q$.
-In some sense this is because $q(y|x)=q(y|x')$ for every $x,x'$, and so there are simply fewer values to estimate.
-On the other hand, we consider the case that $q$ causes $X,Y$ to be deterministically bound, i.e.
-$q(y|x)=1$ if $x=y$ and $q(y|x)=0$ otherwise.
-In this case it is also possible to estimate $q$ (given enough data).
-In-between, it is not possible to exactly determine $q$, even in the asymptotic limit of infinite data.\label{fig:estimatingq}}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Empirical results for cell-types}
-
-We now turn to an application of the MLM to real data.
-
-\figgy{
-\hfill{}\figgygr{1.0}{../images/allenNlx}\hfill{}
-
-\hfill{}\figgygr{.5}{../images/allenNly}\hfill{}
-
-\caption{Input: two tables.
-The Allen Institute had access to various cre/lox cell selection techniques.
-Each technique pulls out a different group of cells.
-Once the cells were selected, they were then either subjected to technique I or technique II.
-Technique I gives a very complete analysis of the gene expression of the cell.
-Technique II gives a less complete analysis, but yields additional electrophysiological data that may be of interest.
-For each subpopulation and each technique, data was gathered to estimate the distribution of different results.
-These are shown above.
-Each row corresponds to a different subpopulation.
-Each column corresponds to a different measurement value.
-Each entry of the heatmaps indicate the number of specimens measured to have a particular measurement value within a particular subpopulation (as measured by the corresponding technique).
-Notice that tool I has many more different measurement results that it can return; this is why there are many more columns in the first heatmap.
-\label{fig:alleninput}}
+\hfill{}\figgygr{.95}{../images/allenNlxy}\hfill{}
+\caption{{\bf The input to the Markov Link Method: data about each tool with many sampling strategies but without any joint measurement}.
+Here we show a portion of the real-world data to which we applied the MLM.
+The Allen Institute gathered neurons from the visual cortex of mice, using a variety of cre/lox-based sampling strategies (cf.\ \cite{tasic2017shared}).
+Each strategy was designed to bring out different subpopulations of cells.
+For each sampling strategy, one classification protocol (tool I) was used to estimate the some of the neurons' cell-types, and another protocol (tool II) was used to estimate the other neurons' cell-types.
+Tool I recognizes 104 different types of cells.
+Tool II has a coarser notion of cell-type, distinguishing only 10 types.
+For each sampling strategy and for each tool, we tabulated the number of cells assigned to each type.
+Above we show a subset of these results through heatmaps; the color of each square indicates the number of specimens sampled with a particular technique that were found to have a particular type according to a particular tool.
+Using this kind of data, our task is to calibrate the two classification protocols.
+That is, we want to be able to ask question of the following form: ``if a neuron is classified as being of type `Peri Kcnj8' by tool I, how might it have been classified by tool II?'' \label{fig:alleninput}}
 }
@@
 
-Our motivation for this problem arose from looking at Allen Institute cell-type assignment of cells, performed using two different experimental techniques (also called experimental ``modalities'').
-Each modality would take a cell and determine what ``type'' of cell it was.
-However, each technique destroyed the cell in the process of measuring it.
+Our motivation for this problem arose from looking at Allen Institute cell-type assignment of cells, performed using two different experimental tools.
+Each tool takes a neuron and determines what type of neuron it is.
+However, both tools destroy the cell in the process of measuring it.
+The first tool uses a standard single-cell RNA sequencing pipeline to determine transcriptomic expression in the cell, and determine the cell-type based on these results.
+The second tool also uses transcriptomic information, but additionally obtains electrophysiological and morphological properties of the neuron.
+This comes at the cost of a degraded transcriptomic signal, requiring new methods to estimate the cell-type.
+We refer the reader to \citep{tasic2017shared} for details on the two methods.
 
 Best efforts were made to use biological intuition to calibrate the methods.
-For example, the $Y$ method has a notion of a ``Lamp5'' cell type.
-The $X$ method has refined this type into many sub-types, such as ``Lamp5 Pdlim5'' and ``Lamp5 Slc35d3.''  If a cell was designated as ``Lamp5 Pdlim5'' under the $X$ method, the hope was that it be given the ``Lamp5'' type under the $Y$ method.
+For example, the tool II protocol has a notion of a ``Lamp5'' cell type.
+The tool I method refines this idea into many sub-types, such as ``Lamp5 Pdlim5'' and ``Lamp5 Slc35d3.''  If a cell was designated as ``Lamp5 Pdlim5'' under the tool I, the hope was that it be given the ``Lamp5'' type by tool II.
 The two methods were designed to achieve this goal.
 However, each method has its own biases and errors, and it was not obvious whether this effort was successful.
-In particular, it seemed clear that in some cases a cell labelled one way with one method would get labelled quite differently with another method, but it was not clear how often this occurred.
+In particular, it seemed clear that sometimes a cell labelled one way with one method would get labelled quite differently with another method, but it was not clear how often this occurred.
 
-Fortunately, there was a kind of information that seemed like it might help determine whether the two methods were properly calibrated: sub-populations.
+Fortunately, there was a kind of information that seemed like it might help determine whether the two methods were properly calibrated: a variety of sampling strategies.
 Using a cre/lox system (cf.\ \citep{tasic2017shared}) they were able to pick out specific, overlapping subpopulations of neurons.
@@
 For each subpopulation and each method, many specimens were sampled and their cell-types determined.
-The result of this process was two tables, shown in Figure \ref{fig:alleninput}.
+The result of this process was two tables, parts of which are shown in Figure \ref{fig:alleninput}.
 While it seemed clear that these tables should say something about the calibration, it was not obvious how to best use this information.
@@
 
-\figgy{
-\hfill{}\figgygr{0.3}{../images/allenqhat}\hfill{}
-\caption{The $\hat q$ estimated by the MLM.\label{fig:allenout}}
-}
-
-Using the Markov Link Method, we obtained an estimate $\hat q$ for the true calibration.
-This calibration is shown in Figure \ref{fig:allenout}.
-From this figure, it appears that the two methods may be fairly well-calibrated.
-For example, according to $\hat q$, if a cell is classified as type ``Lamp5 Pdlim5'' by tool I, it appears there is a 99.9\% chance it will be classified with the ``Lamp5'' type under tool II.
-However, there are other cell-types which seem to have more ambiguity.
-Yet others seem to perhaps be classified incorrectly \emph{most} of the time; we see that ``Vip Igfbp6'' type is most often classified as a ``Pvalb'' type instead of a ``Vip'' type.
-
-Everything in the preceeding paragraph should be taken with a significant grain of salt; it is quite treacherous to directly interpret the estimator $\hat q$, because we know that this estimator for the calibration may suffer from errors.
-This is both due to insufficient data and to identifiability concerns (these are discussed in detail in Section \ref{sec:mlm}).
-For a qualitative assessment, we instead advocate looking at the variability visualizations which the MLM produces.
-
-For example, in the right side of Figure \ref{fig:allenboot} we show the variability visualization for specimens which are measured as being ``Vip Igfbp6'' by tool I.
-Whereas in the original point estimate $\hat q$ we found that ``Vip Igfbp6'' seemed to be strongly associated with the ``Pvalb'' type, this figure shows that in reality we are not quite sure what types these specimens may be associated with.
-Each row corresponds to a different plausible distribution on $y$.
-As we can see, they truly run the gamut.
-In the left side of the same figure we also show the same visualization applied to the row of $\hat q$ corresponding to specimens measured as ``Lamp5 Pdlim5'' by tool I.
-We see very little variability in this visualization, indicating significant confidence that the ``Lamp5 Pdlim5'' type of tool I is truly well-calibrated to match the ``Lamp5'' type of tool II.
-
-In cases where the visualization suggests significant variability, the MLM then suggests what subpopulations we should look at in order to resolve the ambiguities.
-For example, to resolve the ambiguities about ``Vip Igfbp6'' specimens, it appears it would be most important to obtain samples of subpopulations of cells which have many cells measured as ``Pvalb'' by technique II and very few cells measured as ``Sncg.''
+As described in Section \ref{sec:mlm}, to apply the MLM we first produced an approximate 95\% confidence region for $p^*,q^*$.
+We found that this confidence region was non-empty, signifying that we could not reject the hypothesis that the MLM assumption holds.
+We then used the Markov Link Method of producing a variety of example calibrations inside the confidence region.
+We then visualized these examples one cell-type at a time.
+That is, we first select a tool-I cell-type $x$ (such as $x=$`Lamp5 Egln3 1').
+Then for any calibration $q$, we can consider the vector $q_x$ formed by looking at $q(y|x)$ for every value of $y$.
+We can then visualize many calibrations by plotting many of stacking these vectors together into a heatmap.
+The result for various choices of $X$ may be found in Figure \ref{fig:allenboot}.
 
 \figgy{
-\hfill{}\figgygr{0.4}{../images/q1rLamp5Pdlim5}
-\figgygr{0.4}{../images/q1rVipIgfbp6}\hfill{}
-\caption{The MLM provides \emph{variability visualizations} for each row of the calibration $q$.
-In this figure we show two such examples.
-To produce the visualization we must first select a row of $q$.
-This corresponds to particular measurement value that can be obtained from tool I.
-In this case we select ``Lamp5 Pdlim5'' to produce the left visualization and ``Vip Igfbp6'' to produce the right visualization.
-The bottom of each visualization shows the value of $\hat q(y|x)$ for the selected type $x$ (it shows this through a heatmap with a single row).
-Here $\hat q$ is the Markov Link Method estimator we have described before.
-Above this original $\hat q$ value, we show a heatmap with many other possible examples for what else this row might look like.
-The variability among these possibilities gives a qualitative measure of our uncertainty about that particular row of the calibration $q$.
+\hfill{}\figgygr{0.9}{../images/vviz}\hfill{}
+\caption{{\bf Estimating the calibration and measuring our uncertainty on real data.}  Here we present visualizations of the confidence region found by the Markov Link Method applied to Allen Institute data.
+The colors indicates the value of $q(y|x)$ for a particular cell-types $x,y$ and a particular calibration $q$.
+All of the calibrations are examples from a 95\% confidence interval.
+For each individual plot, $x$ is fixed, each column indicates a particular value of $y$, and each row indicates a different plausible $q$.
+Each plot helps us answer the question: how might specimens labelled with type $x$ by tool I be labelled by tool II?
+For example, consider the `Lamp5 Pdlim5' plot.
+For every calibration (i.e.\ every row of the plot), we see a bright yellow in the `Lamp5' column.
+This gives some confidence that specimens assigned type `Lamp5 Pdlim5' by tool I will be assigned type `Lamp5' by tool II.
+By contrast, consider the `Lamp5 Egln3 1' plot.
+This plot shows we have very little knowledge about the calibration for cells identified as this type by tool I.
+Each calibration in the confidence region gives a completely different picture.
+The data we have simply isn't sufficient to help us know this aspect of the calibration.
 \label{fig:allenboot}}
@@
 
-The MLM also yields estimates for $h^*$.
-As in our simulation section, these quantities could also be estimated using more traditional pseudocount methods.
-We can compare the MLM estimates with the traditional methods, using the method of log-likelihood on held-out data.
-We find that both methods achieve a rate of 1.3 nats per entry on held out data.
-This gives us confidence that the MLM assumption might be right; if the MLM assumption was wrong our estimation process could force $h$ to take on an incorrect value.
-This is no proof that the MLM assumption is correct for this data, but it is a good sanity-check.
+Inspecting these visualizations we can see different aspects of our knowledge about how well-calibrated the two tools are:
+\begin{itemize}
+    \item For some tool-I types, such as `Lamp5 Pdlim5,' the two methods appear very well-calibrated.
+All the examples from the confidence region suggest that cells measured with type `Lamp5 Pdlim5' are likely to be assigned type `Lamp5' by tool II.
+    \item For some types, such as `Lamp5 Egln 1,' we simply have no idea.  There is tremendous variability within the confidence region; each example looks different.  
+    \item For some types, such as `Vip Crispld2 2,' it appears that cells of this type are difficult for tool II to understand.
+In particular, for most example calibration they have a reasonably high probability of being assigned the  `unknown' type.
+However, we cannot be sure of the the exact depth of this issue, because in some plausible example calibrations it appears that these cells are in fact correctly labelled as being of the `Vip' type fairly often.
+More data is necessary.
+\end{itemize}
+The variability in the confidence regions suggests what we need to do in order to more closely determine the value of the calibration.
+For example, if we could develop more unique sampling strategies which will include `Lamp5 Egln 1' cells, this would help us resolve our ambiguity about this aspect of the calibration.
+Indeed, going back to the original data, it is easy to see why this ambiguity appeared in the first place.
+Cells of the `Lamp5 Egln 1' type only appear in any number when using the sampling strategies `Gad2-IRES-Cre' and 'Slc32a1-IRES-Cre.'  Both of those sampling strategies yield a fairly similar mix of types when measured with tool II.
+To get a better resolution of the calibration, we would need a sampling strategy that included `Lamp5 Egln 1' cells but represents a significantly different slice of the overal population.
+For a particular proposed experiment, simulations such as those found in Section \ref{sec:simulations} can be used to determine how many samples might be required to get an accurate estimate of the calibration.
 
-Finally, the MLM provides quantitative values to give a sense for the the error in our estimators, namely $\Delta_p,\Delta_q,\Delta_h,\bar d$.
-These can be difficult to interpret because they are single numerical quantities that summarize error in the whole of $p,q,h$.
-As we have seen from our visualizations, certain aspects of $q$ may be fairly tightly identified with very little error whereas others may not be clear at all.
-Numerical figures must somehow ``average'' over all these complications.
-For this reason, the qualitative visualizations may be more helpful.
-However, if a quantitative measurement is needed, they are readily available: we find that $\Delta_p=6.7,\Delta_q=50.3,\Delta_h=3.7,\bar d=18.0$.
-
-\section{Conclusions}
-
-When joint measurement is impossible, it can be difficult to calibrate two methods against each other or understand how they may be related.
-Here we show that a simple Markov assumption can make it possible to actually learn quite a lot.
-Although the exact relationship may not be identifiable, we can rigorously bound our uncertainty.
-We have proposed the Markov Link Method as procedure to estimate the calibration and understand our uncertainty regarding that estimate.
-We investigated a real-world calibration problem; the MLM gave bounds on the accuracy of the calibration and also the suggested directions for future experiments to further refine our uncertainty about this calibration.
-Code is published at https://github.com/jacksonloper/markov-link-method, including a tutorial-style ipython notebook detailing every calculation used in this paper.
-
-The Markov assumption is of course not the only one that we could have used, and may not be valid in every case; future work may be to investigate others.
-For example, it has been speculated that some cell types tend to die more often in one experimental modality than another, and these cells are simply excised from the data without comment.
-This would violate our assumptions.
-However, assuming this death rate can be roughly measured, it can be adjusted for, yielding a different but equally meaningful assumption about the data.
-Indeed, if the MLM gives insensible results, this could actually serve as a useful clue that this disproportionate cell death is happening.
-
-Once we accept that what we're interested in may not be fully identifiable, any of a wide variety of assumptions can help us obtain practical bounds.
-Although we may not be able to learn exactly what we want, we can learn a set of possibilities.
-By probing this set carefully, we can learn what the data actually has to say and what experiments we need to do to learn more.
-
-\section{The Markov Link Method}
-
-\label{sec:mlm}
-
-Here we describe in detail the Markov Link Method, the identifiability problems which it must overcome, and some theoretical guarantees for the method.
-For the benefit of the reader, we repeat the basic inputs to the method:
-
-\begin{itemize}
-    \item Samples $(\ell_1,X_1)\cdots (\ell_n,X_n)$ sampled such that $\mathbb{P}(X_i=x|\ell_i=\ell)= p^*(x|\ell)$
-    \item Samples $(\ell_{n+1},Y_{n+1})\cdots (\ell_{n+m},Y_{n+m})$ sampled such that $\mathbb{P}(Y_i=x|\ell_i=\ell)= h^*(x|\ell)$
-\end{itemize}
-
-We assume that $\ell,X,Y$ are discrete random variables with finite support. We also assume that the MLM assumption holds, i.e.\ $h^*(x|\ell)=\sum_y p^*(x|\ell)q^*(x|y)$ for some unknown $q^*$.
-
-The method proceeds as follows:
-
-\begin{enumerate}
-\item Summarize the input data, with two matrices:
-%
-\begin{align*}
-N^X_{\ell x} &= |\{i\leq n:\ \ell_i=\ell,X_i=x\}|\\
-N^Y_{\ell y} &= |\{i>n:\ \ell_i=\ell,Y_i=x\}|
-\end{align*}
-%
-That is, $N^X_{\ell x}$ indicates the number of specimens from subpopulation $\ell$ which yielded measurement $x$ under tool I.
-Likewise $N^Y_{\ell y}$ indicates the number of specimens from subpopulation $\ell$ which yielded measurement $y$ under tool II.
-
-\item Produce point estimates. 
-\begin{itemize}
-\item For $p$, we use a simple pseudocount estimator, namely
-\[
-\hat p = \argmax_p \sum_{\ell,x} (N^X_{\ell x}+1) \log p(x|\ell)
-\] 
-\item For $h$, we take a similar idea.
-However, in this case, we make use of the Markov Link Assumption to increase the effficiency of our estimator.
-Let $S$ denote the set of values of $h$ which are consistent with $p$ and the Markov Link Assumption, i.e.
-\[
-S=\left\{h:\ \exists q:\ \sum_x \hat p(x|\ell)q(y|x)=h(y|\ell) \quad \forall \ell,y\right\}
-\]
-And then we define
-\[
-\hat h = \argmax_{h\in S} \sum_{\ell,y} (N^Y_{\ell y}+1) \log h(x|\ell)
-\] 
-\item Finally, we estimate $q$.  Towards this end, we define $\Theta$ as the set of values of $q$ which are consistent with $p,h$ and the Markov Link Assumption.  That is, let 
-\begin{align}\label{eq:thetadef}
-\Theta(p,h) \triangleq \left\{q:\ \sum_x p(x|\ell)q(y|x)= h(y|\ell) \quad \forall \ell,y \right\}
-\end{align}
-Any $q$ inside $\Theta$ should be considered equally plausible, since it yields the same distribution on the observable data.
-However, for the purposes of determining variability due to lack of data, we take a very particular estimator of $q$, namely the so-called ``analytic center'' of $\Theta$.
-\[
-\hat q = \argmax_{q\in\Theta(\hat p,\hat h)} \sum_{x,y} \log q(y|x)
-\]
-This gives us a single estimator for $q$, which we will use below.  
-\end{itemize}
-Several remarks are in order about these point estimates.  
-\begin{itemize}
-\item The pseudocount.
-An obvious alternative to the pseudocount estimator would be to simply use the maximum likelihood estimator, i.e.
-$\argmax_p \sum_{\ell,x} N^X_{\ell x} \log p(x|\ell)$.
-However, in the datasets we examined, there was simply not enough data to support this kind of estimator.
-Some of the subpopulations had very few samples, requiring some degree of smoothing.
-Pseudocounts are just one among a host of ways this smoothing could have been achieved.
-However, we chose pseudocounts because they are easy to interpret both mathematically and by scientists.
-We simply find the maximum likelihood estimator \emph{as if} the data had been that given by the pseudocount-augmented values.
-We here chose a pseudocount value of 1, but in practice an expert can meaningfully evaluate what kinds of pseudocounts each sample category should take.
-%
-\item We do use the Markov Link Assumption to estimate $h$.
-We note that an alternative way to estimate $h$ would be to simply use a pseudocount estimator just as we did to estimate $p$.
-However, if the Markov Link Assumption holds, we should be able to use this fact to get a superior estimators by enforcing this constraint.
-We saw in simulations that this does indeed occur at least in some cases.
-%
-\item We don't use the Markov Link Assumption to estimate $p$.
-The process outlined above follows three distinct stages: estimate $p$, estimate $h$, then estimate $q$.
-This process has the advantage that each individual problem is convex, with a unique solution which can be found with classic methods from convex optimization (see Appendix \ref{sec:convexopt} for details).
-However, there is a fundamentally different way we could seek our point estimates.
-We could try to \emph{simultaneously} optimize all of these objects to maximize the likelihood of the observables (subject to the Markov Link Assumption).
-Unfortunately, we were unable to get rigorous guarantees for the asymptotic properties of such a method.
-It is for this reason that we took this multi-stage approach.
-\end{itemize}
-%
-\item Produce quantitative estimates for the variance of our estimators, using the bootstrap.
-We a series of $K$ surrogate datasets $(N^{X,(1)},N^{Y,(1)})\cdots (N^{X,(K)},N^{Y,(K)})$ by sampling with replacement within each subpopulation from the original data.
-For each surrogate $(N^{X,(i)},N^{Y,(i)})$ we use a procedure similar to the one described above to produce point estimates $\hat p^{(i)},\hat q^{(i)},\hat h^{(i)}$.
-The only difference between the bootstrap point estimators and the original point estimators is that we use no pseudocount for the bootstrap estimators.
-This is because the pseudocount can in some cases make the estimator variance look smaller than it is.
-Finally, for each variable, $p,q,h$, we compute the average total variation distance between our original point estimates and these bootstrap estimates.
-Here by total variation distance we mean for example that $\TV{\hat q- \hat q^{(i)}} \triangleq \frac{1}{2}\sum_{x,y}|\hat q(y|x)-\hat q^{(i)}(y|x)|$.
-We denote these variation distance averages by $\Delta_p,\Delta_q,\Delta_h$.
-%
-\item Produce quantitative estimates for our uncertainty about $q^*$ due to identifiability issues.
-In some cases it may be that $q^*$ is not identifiable from the data.
-That is, even in the limit of an infinite number of specimens, it may be impossible to determine the exact value of $q^*$.
-This is because we can only observe $q^*$ through the distribution on $Y|\ell$ (i.e.\ $h^*$).
-Thus, if there are two values $q_1,q_2$ such that $h^*(y|\ell)=\sum_x p^*(x|\ell)q_1(y|x)=\sum_x p^*(x|\ell)q_2(y|x)$, then without joint measurement on $(X,Y)$ we would never be able to tell if $q^*=q_1$ or $q^*=q_2$.
-Nonetheless, even if $q^*$ cannot be determined exactly, it is possible to put bounds on what values of $q^*$ are consistent with the data: the Markov Link Method assumption does say that $q^* \in \Theta(p^*,h^*)$, where $\Theta$ is given by Equation (\ref{eq:thetadef}).
-This can tell us something about what the calibration $q^*$ looks like.
-What is more, subject to mild conditions, we can show that in the limit of infinite data, $q^*$ is very close to some point in $\Theta(\hat p,\hat q)$.
-This is the subject of Theorem \ref{thm:mainthm} in Appendix \ref{sec:mainthm}.
-To understand our uncertainty about $q^*$ due to identifiability, we must therefore understand the size of this set.
-To do so, we use the simple total variation diameter approximation procedure (see Appendix \ref{sec:diameterestimation} for details) to estimate the diameter for the set associated with each of our bootstrap samples, $d^{(i)}\approx \mathrm{diam}(\Theta(\hat p^{(i)},\hat h^{(i)}))$.
-To estimate the error due to identifiability in our particular case, we use the median of these diameter estimates, which we denote by $\bar d$.
-%
-\item Produce qualitative variability visualizations.
-Fix any value of $x$.
-Now consider any bootstrap estimate $\hat p^{(i)},\hat q^{(i)},\hat h^{(i)}$.
-For each $y$ we can compute the value of $q\in \Theta(\hat p^{(i)},\hat q^{(i)})$ with the maximal value of $q(y|x)$.
-We call this an ``extremal bootstrap estimate,'' \
-\[
-q^{(i,x,y)} = \sup_{q\in\Theta(\hat p^{(i)},\hat q^{(i)})} q(y|x)
-\]
-We can look at all of these estimates simulatenously by forming a square matrix $M_{y_1,y_2} = q^{(i,x,y_1)}(y_2|x)$ and plotting this matrix as a heatmap.
-If there are no identifiability issues concerning the estimation of $q^*$, every row of the matrix will be the same.
-If there are significant identifiability issues, different rows may be quite different.
-We can repeat this process for each bootstrap estimate; this yields a very tall and thin matrix formed by stacking all the matrices together.
-If there is sufficient data, we expect each block of this matrix to be approximately the same (since the bootstrap resampling will have little effect).
-Thus, if there is sufficient data and there are no identifiability issues every row of the entire tall matrix will be similar.
-For visualization, we can squish the matrix so that each row is much thinner than each column.
-The result is a single image that captures the variability for one particular row of $q$.
-
-\end{enumerate}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%            _                               _    
+%  _ __  _ __(_) ___  _ ____      _____  _ __| | __
+% | '_ \| '__| |/ _ \| '__\ \ /\ / / _ \| '__| |/ /
+% | |_) | |  | | (_) | |   \ V  V / (_) | |  |   < 
+% | .__/|_|  |_|\___/|_|    \_/\_/ \___/|_|  |_|\_\
+% |_|                                              
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-Our main goal is to achieve calibration without joint measurement.
+In this paper we sought to achieve calibration without joint measurement.
 The main technical obstacle is what is known in the statistics world as an `identifiability problem.'  Due to this problem, we often simply \emph{cannot} directly estimate the calibration that we are interested in.
-The calibration is simply not ``identifiable.''  Nonetheless, all is not lost.
-Our main contribution is to show that one can bound the extent of this identifiability problem.
-In short, the fundamental problem cannot generally be vanquished, but we can put it in its place.
+The calibration is simply not ``identifiable.''  Our main contribution is to produce a an approximate confidence region for the parameter of interest.
 
-Our identifiability analysis stands on the shoulders of a long history of turning probabilistic assumptions into bounds on unidentifiable parameters.
-The core idea of the MLM is to take an assumption (the `Markov Link assumption,' which we will define shortly) and use it to place bounds on an unidentifiable quantity (namely the calibration $q^*$).
-When these bounds are fairly tight on all sides, we see that much can be learned even though what we want isn't identifiable.
-Much of the prior literature in this kind of direction comes from research into causality.
-For example, in \cite{bonet2001instrumentality} Bonet uses polytopes not unlike the ones seen here to explore whether a variable can be used as an instrument.
+The necessity to deal with such identifiability issues has become of increasing concern as nature of model parameters and data has grown more intricate.
+There are many examples in the literature.
+How much can we hope to learn about the parameters of complex biological reaction networks, if we can only observe part of the process (cf.\ \cite{raue2009structural})?
+Given what we can observe in practice, how much can we ever hope to learn about how cancers divide and spread through the body (cf.
+\cite{brouwer2017parameter,luebeck2002multistage})?
+Do the tests we have enable us to actually learn the parameters of our models for how materials bend and break (cf.
+\cite{hartmann2018identifiability})?
+At their heart, these are all questions of estimation the face of potential unidentifiability.
+
+This work is inspired by a large literature of examples where assumptions are used to bound potentially unidentifiable parameters.
+Some of this literature comes from the field of causality.
+For example, in \cite{bonet2001instrumentality} Bonet produces regions not unlike the ones seen here to explore whether a variable can be used as an instrument.
 The famous Clauser-Horne-Shimony-Holt inequality was designed to help answer causality questions in quantum physics, but it also sheds light on what distributions are consistent with certain assumptions \cite{clauser1969proposed}.
-Indeed the physics literature has contributed many key inequalities (cf.
+More generally, the physics literature has contributed many key assumptions that bound unidentifiable parameters (cf.
 \cite{chaves2014inferring}, \cite{kela2017semidefinite}, and the references therein).
 Perhaps the closest work to this one would be \cite{makarov1982estimates}, which used two marginal distributions to get bounds on a property of the joint distribution (namely the distribution of the sum).
-We advance this approach to a more general-purpose technique, both by using many subpopulations to refine our estimates and by considering the entire space of possible joint distributions instead of simply a particular aspect of the joint.
+We advance this approach to a more general-purpose technique, both by using many subpopulations to closely refine our estimates and by considering the entire space of possible joint distributions instead of simply a particular property of the joint.
 
-It is perhaps worth mentioning that the actual theorem presented here has probably been derived before (although we did not find it in our review of the literature).
-Other results in the same flavor as the one presented here might also find a practical use for modern problems.
-There is a treasure-trove of ideas in the causality literature; this wealth has not yet been brought fully to bear on the challenging and important problems of calibration for modern experimental modalities.
-Our primary intention with this article is to bring attention to the practical utility of these kinds of result.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                       _           _                 
+%   ___ ___  _ __   ___| |_   _ ___(_) ___  _ __  ___ 
+%  / __/ _ \| '_ \ / __| | | | / __| |/ _ \| '_ \/ __|
+% | (_| (_) | | | | (__| | |_| \__ \ | (_) | | | \__ \
+%  \___\___/|_| |_|\___|_|\__,_|___/_|\___/|_| |_|___/
+%                                                    
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Discussion}
+
+When joint measurement is unavailable, it can be difficult to calibrate two measurement tools.
+
+In this paper we see that a conditional independence assumption can help overcome this difficulty.
+We proposed a procedure, the Markov Link Method (MLM), to use this assumption to estimate the calibration and measure uncertainty.
+Code is published at https://github.com/jacksonloper/markov-link-method, including a tutorial-style ipython notebook detailing every computation made in this paper.
+
+To understand our uncertainty, we use to a confidence region; this comes with all the usual attendent caveats.
+For example, we here take the view that calibrations inside the confidence region are somehow `plausible' or `likely.'  As has been pointed out numerous time, such a statement cannot be philosophically justified without some at least a hint of Bayesian reasoning (cf.\ \cite{morey2016fallacy}).
+We leave a more rigorous Bayesian treatment of this problem for future research.
+The key problem will be to find a suitable prior; this search comes with certain dangers which we are not sure how to circumvent.
+We give an example of these dangers in Appendix \ref{sec:bayesian}.
+
+Theory aside, the MLM appears to provide reasonable estimates in simulation and give useful insight to real data.
+Applied to neuronal cell-type data, the MLM clarified how two different cell-type classification systems are related.
+We saw that some aspects of the two systems seem well-calibrated, but others we are less sure about.
+The nature of the variability in the confidence region suggested directions for experiments which would further refine our understanding.
+
+In future work, it may be interesting to consider additional assumptions that might help us estimate calibrations.
+For example, as it stands the Markov Link Method only works if each tool can only return one of a relatively small number of measurement values.
+In some cases, we may believe that similar measurement values should have similar probabilities.
+Such smoothness assumptions would make it possible to apply the MLM to measurement tools which can return many different values.
+
+More generally, it is clear that good assumptions can help us get real insight for tough problems.
+Even if these assumptions are not sufficient to allow us to perfectly identify our object of interest, confidence regions can often be constructed.
+By probing these regions carefully, we can learn what the data actually has to say and what experiments will help us learn more.
 
@@
 
-\section{Mathematical results}
+\section{Confidence interval details}
 
-\label{sec:mainthm}
+Recall that we have an $\Omega_\ell\times\Omega_X$ matrix $N^X$ and an $\Omega_\ell \times \Omega_Y$ matrix $N^Y$, sampled according to
+\begin{align*}
+(N^X_{\ell 1},N^X_{\ell 2}\cdots N^X_{\ell \Omega_X}) \sim \mathrm{Multinomial}(n_\ell,p^*(\cdot | \ell))\\
+(N^Y_{\ell 1},N^Y_{\ell 2}\cdots N^Y_{\ell \Omega_Y}) \sim \mathrm{Multinomial}(m_\ell,h^*(\cdot | \ell))
+\end{align*}
+where $p^*(x|\ell)$ is some conditional distribution and there is some $q^*(y|x)$ such that $h^*(y|\ell)=\sum_x p^*(x|\ell)q^*(y|x)$.  We have then defined
+\[
+L(p,q)=\frac{1}{\Omega_X}\sum_{\ell,x} N^X_{\ell x} \log \frac{N^X_{\ell x}}{n_\ell p(x|\ell)} + \frac{1}{\Omega_Y}\sum_{\ell,y} N^Y_{\ell y} \log  \frac{N^Y_{\ell y}}{m_\ell \sum_x p(x|\ell) q(y|x)}
+\]
+We then constructed the confidence region
+\[
+R = \{p,q:\ L(p,q) \leq k\}
+\]
+This is sometimes called so-called likelihood ratio confidence interval, and its properties are fairly well-understood.
+We refer the reader to \cite{brazzale2007applied} for a useful exposition.
+The essential point is that in the asymptotic limit the distribution of $L(p^*,q^*)$ depends only upon the number of zero probability values in the parameters $p^*,h^*$.
+For any number of zero probability values, the asymptotic distributions are known exactly; they are scaled sums of $\chi^2$ distributions, one for each subpopulation and each tool.
+The difficulty is of course that we do not know how many probability value are zero.
+This can can be circumvented if one is only interested in a conservative confidence region, i.e.\ one whose actual coverage is at least as great as its nominal coverage.
+In particular, if we assume that \emph{no} values are zero we can a confidence interval which is guaranteed to be conservative; this is an immediate result of the fact that $\chi^2$ distributions with more degrees of freedom stochastically dominate those with fewer.
+Unfortunately, using these $\chi^2$ distributions can be quite inadvisble in low-sample regimes.
+For instance, one may readily see that if some values of $p^*,h^*$ are small (but nonzero), then the distribution of $L(p^*,q^*)$ will be much more similar to sums of $\chi^2$ distributions with fewer degrees of freedom -- at least until the number of samples grows very very large.
 
-\begin{itemize}
-\item Let $\UN{\cdot}$ denote the uniform norm (i.e.\ the maximum absolute value) and $\EN{\cdot}$ denote the Euclidean norm (i.e.\ the square root of the sum of the squares).
-In the case of matrices, this Euclidean norm goes by the name of the Frobenius norm.
-Recall that in this norm matrices satisfy a Cauchy-Schwarz like equality, $\EN{pq}\leq\EN{p}\EN{q}$.
-Also recall that $\UN{a} \leq \EN{a} \leq \sqrt{n}\UN{a}$ where $n$ is the number of entries in $a$.
-\item Let $T_{a,b}$ denote the transition matrix polytope, i.e. the set of $a\times b$ matrices whose rows sum to 1 and whose entries are all positive.
-\item Let $|\Omega_\ell|,|\Omega_X|,|\Omega_Y| \in \mathbb{N}$.
-\item Let $p^*\in T_{|\Omega_\ell|,|\Omega_X|}$.
-\item Let $q^*\in T_{|\Omega_X|,|\Omega_Y|}$.
-\item Let $h^*=p^*q^*$.
-Here by $p^* q^*$ we intend the matrix multiplication of $p^*$ and $q^*$.
-We will go back and forth between thinking of $p^*,q^*,h^*$ as matrices and thinking of them as distributions.
-\item We require the matrix $q^*$ has strictly positive entries, $q^*_{xy}\geq c>0$.
-\item For each $\ell$ draw $n_\ell$ samples from $p^*_\ell$ and $m_\ell$ samples from $h^*_\ell$.  Let $N^X_{\ell x},N^Y_{\ell y}$ summarize the resulting number of each kind of sample we obtained.
-\item We require that there are fewer rows of $p^*$ than columns, and each row is linearly independent.
-If there are more rows than columns, then there is generally no identifiability issue so the theorem below is not relevant.
-If there are fewer rows than columns but the rows are not independent, then the result applies after enough rows are combined.
-\item Let $\hat p_{\ell x} = (N^{X}_{\ell,x}+1) / (\sum_{x'}N^{X}_{\ell,x'}+1)$
-\item Let
+We find it practical and accurate to choose $k$ by using the data itself.
+In particular, we choose a pseudocount $c$ and take $\bar p(y|x)=N^X_{\ell x} +c /n_\ell +c$ and $\bar h(y|\ell)=N^Y_{\ell y} +c / m_\ell +c$ to give estimates of $p^*,h^*$.
+For a nominal coverage of $1-\alpha$, we then choose $k$ so that
 \[
-L(q,\kappa) = \sum_{y,\ell} (N^Y_{\ell y}+1) \log \left(\sum_x \hat p(x|\ell)q(y|x)\right)
-    +\kappa\sum_{x,y} \log q(y|x)
+\mathbb{P}\left(\frac{1}{\Omega_X}\sum_{\ell,x} N^X_{\ell x} \log \frac{N^X_{\ell x}}{n_\ell \bar p(x|\ell)} + \frac{1}{\Omega_Y}\sum_{\ell,y} N^Y_{\ell y} \log  \frac{N^Y_{\ell y}}{m_\ell \bar h(y|\ell)}<k\right)=\alpha
 \]
-and take $\hat q = \lim_{\kappa\rightarrow 0} \argmax_q L(q,\kappa)$.
-\item Let $\hat \Theta=\{q:\ \hat p\hat q=\hat pq\} \cap T_{|\Omega_X|,|\Omega_Y|}$.
-\end{itemize}
-\vspace{.1in}
+when $N^X,N^Y$ are drawn from $\bar p,\bar h$ respectively.
+Asymptotically $\bar p\rightarrow p^*$ and $\bar h\rightarrow h^*$, so it seems reasonable to hope that this will give the correct coverage.
+Simulations appear to bear this out, though a further mathematical inquiry is certainly warranted.
 
-\begin{thm}\label{thm:mainthm}
-If $n_\ell, m_\ell\rightarrow \infty$ in such a way that $m_\ell/m_{\ell'} \geq \rho>0$ for each $\ell,\ell'$, then $\inf_{q\in \hat\Theta} \UN{q^*-q}\rightarrow 0$ in probability.
-\end{thm}
+The pseudocount bears some discussion.
+We introduce this pseudocount in order to err on the side of a slightly conservative confidence region.
+As we said before, when the parameters $p^*,h^*$ have many small values it is generally possible to produce a smaller confidence region for any given level of coverage.
+We therefore apply a pseudocount to give a slight boost to events that appear to have small probability.
+In practice we find that this encourages the confidence region to err on the side of being conservative, giving us greater confidence that the true parameter lies within the confience region.
 
-\begin{proof}
-It is well-known that $\hat p\rightarrow p^*$ in probability (in both the uniform or the Euclidean norm, which are of course equivalent in this case).
-It is easy to see that the condition on $m_\ell$ together with the strict positivity of $q^*$ guarantee that the same goes for $\hat p\hat q\rightarrow h^*$.
-Thus, intuitively, the difficulty is this: by ensuring $\EN{\hat p - p^*},\EN{\hat p\hat q - h^*}$ sufficiently small, can we find some $\tilde q\in \hat \Theta$ so that $\EN{\tilde q - q^*}$ is arbitrarily small?
-It turns out we can.
+\label{sec:confinterval}
 
-There is a simple method to find such a $\tilde q$: we take $\tilde q$ to be the Euclidean projection of $q^*$ to the plane defined by $\{q:\ \hat p\hat q=\hat pq\}$.
-If $p^* \approx \hat p$ and $\hat p \hat q \approx p^* q^*$, it is easy to see that we can can ensure this projection will not carry us a very large distance, and so $\tilde q$ and $q^*$ will be close, as desired.
-However, it might carry us to a point outside of the transition matrix polytope; in particular, after projection we might see some negative values.
-Fortunately, because we have insisted that the truth is strictly positive, we are guaranteed that this is not a problem in the asymptotic limit.
+\section{The Bayesian Alternative}
 
-Let us make this argument formal.
-We haved defined $c>0$ as the smallest value of $q^*_{xy}$.
-Fix any $\epsilon<c, p^*,q^*$.
-Let the right inverse of a matrix be defined by $a^\dagger \triangleq a^T (aa^T)^{-1}$.
-Note that since $p^*$ has linearly independent rows, this is well-defined and continuous in a small neighborhood around $p^*$.
-Let $M=\EN{(p^*)^\dagger}$.
-Find $\delta$ small enough so that if $\UN{p-p^*}<\delta$ then $\EN{p^\dagger}<2M$.
-Taking a further smaller $\delta$ if necessary, ensure that if $\UN{p^*-p}<\delta$ then $\EN{p^*-p}$ is less than $\epsilon / 4M\sqrt{|\Omega_X||\Omega_Y|}$.
-Now fix any $\hat p,\hat q$ with $\UN{\hat p-p^*}<\delta$ and $\EN{\hat p \hat q - p^*q^*}<\epsilon/4M$.  Take
-\[
-\tilde q = q^* + \hat p^\dagger \hat p (\hat q - q^*)
-\] 
-Then we make the following observations:
-\begin{itemize}
-  \item Let us compute $\EN{\tilde q -q^*}$. We have
-   \begin{align*}
-   \EN{\tilde q -q^*} &= \EN{\hat p^\dagger \hat p (\hat q - q^*)} \leq 2M \EN{\hat p \hat q - \hat p q^*} \\
-        &\leq 2M \EN{\hat p \hat q - p^* q^*} + 2M\EN{(p^* - \hat p)q^*}\\
-        &\leq 2M \frac{\epsilon}{4M} + \frac{2M\epsilon}{4M\sqrt{|\Omega_X||\Omega_Y|}}\sqrt{|\Omega_X||\Omega_Y|}\UN{q^*} \leq \epsilon 
-   \end{align*}
-  \item $\hat p \tilde q = \hat p q^* + \hat p \hat q - \hat p q^* = \hat p \hat q$
-  \item The rows of $\tilde q$ sum to 1.
-This is easy to see, because the rows of $q^*$ sum to 1 and the rows of $\hat q$ sum to 1, and so $\tilde q \mathbf{1} = q^* \mathbf{1} + \hat p^\dagger \hat p (\hat q - q^*) \mathbf{1} = \mathbf{1} + 0$ as desired.
-  \item The entries of $\tilde q$ are positive.
-Indeed, the the smallest value of $q^*$ is $c$, and we have already argued that $\UN{\tilde q - q^*}\leq \epsilon$.
-Thus the smallest value of $\tilde q$ is at least $c-\epsilon$, and we have required $\epsilon<c$.
-\end{itemize}
-Thus $\UN{\tilde q - q^*}<\epsilon$ and $\tilde q \in \hat \Theta$.
+Todo
 
-In conclusion, we see that by taking $\hat p$ sufficiently close to $p^*$ and $\hat p \hat q$ sufficiently close to $p^*q^*$, we can ensure that the set $\hat \Theta$ contains a close which is arbitrarily close to the true $q^*$.
-Since $\hat p$ and $\hat p \hat q$ are themselves consistent estimators, this completes the proof.
-\end{proof}
-
-\section{Total Variation Diameter Estimation}
-
-\label{sec:diameterestimation}
-
-We follow the algorithm of \cite{brieden2001deterministic} in estimating of the diameter of a convex polytope $C$.
-We note that diameter estimation is a difficult problem, and if the polytope is particularly pathological and high-dimensional it may be quite difficult to estimate the diameter accurately.
-Nonetheless, it is a practical solution to the difficulty at hand.
-
-Let $C \in \mathbb{R}^{n}$ denote a convex polytope (note that the $n$ here does not correspond to the number of samples $n$ in the main body of this paper).
-We will be interested in a certain kind of diameter of $C$, namely
-\begin{align*}
-\diam(C) =& \frac{1}{2}\sup_{x,y\in\Theta} \sum_{i=1}^n |x_i-y_i| \\
-              =& \frac{1}{2}\max_{z\in\{-1,1\}^n} \left(\sup_{x,y\in\Theta} \sum_{i=1}^n z_i(x_i-y_i)\right)\\
-              =& \frac{1}{2}\max_{z\in\{-1,1\}^n} \left((\sup_{x\in\Theta} z \cdot x) - (\inf_{x\in\Theta} z \cdot x) \right)
-\end{align*}
-
-For any fixed $z$, the inner maximization and minimization problems can be solved using linear programming.
-We can therefore estimate the diameter by taking every possible $z$ and solving this problem.
-In practice, this may be computationally prohibitive.
-If so, we can randomly subsample values of $z$ to obtain a lower bound on the diameter.
-
-We can apply this to estimate the diameter of our set of interest, namely
-\[
-\Theta(p,h) = \left\{q:\ \sum_x p(x|\ell)q(y|x)=h(y|\ell)\right\}
-\]
-In this case $\Theta$ is a space of conditional probability distributions.  However, the constraints that define $\Theta$ are all linear, and so the same method can be applied.
+\label{sec:bayesian}
 
@@
 
-\label{sec:convexopt}
+\label{sec:optproblems}
 
-To be determined
+Todo
 
-\section{Permutation matrices are perfectly identified}
+\section{Identifiability case studies}
 
-\label{sec:miracle}
+\label{sec:casestudies}
 
-To be written
+Our analysis of the identifiability issues associated with the Markov Link Method is not completely satisfactory.
+We are able to make rigorous guarantees about consistency for any given true value of $p^*,q^*$, but we are not content with our understanding of the situations in which identifiability is and is not a serious problem.
 
+Here we present two suggestive case studies which we hope may inspire future research.  In both cases we are able to prove something of interest -- but not quite as much as we might hope.
+
+\subsection{Permutation matrices}
+
+Let us consider the case that $\Omega_\ell = \{1,2,\cdots k\}$ and $\Omega_X=\Omega_Y=\{1,2\cdots,2^{k-1}\}$.
+That is, there are $k$ separate subpopulations, and both tool I and tool II can return one of $2^{k-1}$ possible values.
+Let us furthermore assume that
+\[
+q^*(y|x)=\begin{cases}1 \quad \mathrm{if}\ x=y\\0\quad \mathrm{else}\end{cases}
+\]
+and  $p^*(x|\ell)=A_{\ell,x}$, where this matrix is given by
+\[
+A=2^{2-k}\left(\begin{array}{ccccccccccc}
+0 & 1 & 0 & 1 & 0 & 1 & \cdots & 0 & 1 & 0 & 1\\
+0 & 0 & 1 & 1 & 0 & 0 & \cdots & 0 & 0 & 1 & 1\\
+0 & 0 & 0 & 0 & 1 & 1 & \cdots & 1 & 1 & 1 & 1\\
+0 & 0 & 0 & 0 & 0 & 0 & \cdots & 1 & 1 & 1 & 1\\
+ &  &  &  &  &  & \vdots\\
+0 & 0 & 0 & 0 & 0 & 0 & \cdots & 1 & 1 & 1 & 1\\
+1 & 0 & 1 & 0 & 1 & 0 & \cdots & 1 & 0 & 1 & 0
+\end{array}\right)
+\]
+That is, the $x$th column of the first $k-1$ rows is simply the binary expansion of the number $x-1$, and the last row alternates $1$s and $0$s.
+Now let us say we have perfect knowledge of $p^*(x|\ell)$ and $h^*(y|\ell)=\sum_x p^*(y|\ell)q^*(y|x)$.
+Notice that due to the simple structure of $q^*$ we obtain $h^*(y|\ell)=A^{\ell,y}$.
+However, let us imagine we know nothing about the true value of $q^*$.
+
+How much can we say about $q^*$, if we only had knowledge of $p^*$ and $h^*$?
+On the one hand, we observe that in the absence of any other constraints, the object $q^*$ has $2^{2k-3}$ degrees of freedom.
+This is because there are $2^{k-1}$ values of $\ell$ and for each subpopulation $q^*(\cdot|\ell)$ must lie in the $2^{k-2}$-dimensional simplex on $2^{k-1}$ atoms.
+On the other hand, we see that the Markov Link Assumption gives us $k\times(2^{k-1}-1)$ linear constraints on the value of $q$.
+Indeed, for each subpopulation in $1\cdots k$ and each value of $y \in 1 \cdots 2^{k-1}$ we have an equation of the form
+\[
+\sum_x p(y|\ell)q(y|x)=h(y|\ell)
+\]
+Of these $k\times2^{k-1}$ constraints, $k$ of them are redundant with the fact that $\sum_y q(y|x)=1$.
+Thus, altogether, the Markov Link Assumption together with approximate knowledge of $p,h$ gives us $k\times(2^{k-1}-1)$ linear constraints.
+It would follow that $q$ would have $2^{2k-3} - k\times(2^{k-1}-1)$ degrees of freedom yet remaining.
+
+In conclusion, a simple degrees-of-freedom counting argument would suggest that there will be substantial ambiguity about what value $q$ might take on, if our only knowledge about $q$ is that it must satisfy $\sum_x p^*(y|\ell)q(y|x)=h^*(y|\ell)$.
+Indeed, we have \emph{exponentially many} more degrees of freedom than we have constraints.
+
+However, the reality is that $q$ is exactly determined by $p^*,h^*$.
+This is possible because there are inequality constraints which also govern $q$, namely $q(y|x)\geq 0$.
+Thus, while a simple degrees-of-freedom counting argument might suggest that we would have substantial identifiability issues in this problem, the reality is quite the opposite.
+This idea is made rigorous in the following theorem.
+
+\begin{thm}\label{thm:miracle}
+Let $p^*,h^*$ be as they are defined above.  Then there is exactly one $q$ that is consistent with $p^*,h^*$ and the Markov Link assumption.  That is, $q^*$ is the only possible $q$ satisfying
+\begin{gather*}
+\sum_y q(y|x)=1 \\
+\sum_x A_{\ell,x}q(y|x)=A_{\ell,y}\\
+q(y|x)\geq 0
+\end{gather*}
+\end{thm} 
+\begin{proof}
+We prove by recursion.  First take the case $k=2$.  In this case the result holds trivially, since $X,Y\in\{1\}$.
+
+Now consider a general case $k>2$.  Let us focus on the constraints implied by the second-to-last row population.  It is straightforward to see that these constraints imply
+\begin{align*}
+0=&q(y|x) \qquad \forall y\leq 2^{k-2},x>2^{k-2}
+\end{align*}
+Indeed, for each $y\leq 2^{k-2}$ we obtain a constraint showing that $\sum_{x>2^{k-2}}q(y|x)=0$, which yields that in fact $q(y|x)=0$ for every $x>2^{k-2}$ and every $y\leq 2^{k-2}$.
+
+It follows that for $y\leq 2^{k-2}$ our original constraints may be rewritten as
+\[
+\sum_{x\leq2^{k-2}} A_{\ell x} q(x|y) = A_{\ell y} \qquad \forall y\leq 2^{k-2}
+\]
+This is an example of our problem with $k$ one smaller.
+Applying our inductive hypothesis, we may thus obtain that $q(y|x)=q^*(y|x)$ for the first $2^{k-2}$ values of $x,y$.
+Moreover, since $\sum_{y\leq 2^{k-2}} q^*(y|x)=1$, we see that $q$ must also satisfy $q(y|x)=0$ for $y>2^{k-2}$ and $x\leq 2^{k-2}$.
+Thus we have seen that $q=q^*$ for all entries except those in which $x,y\geq 2^{k-2}$.
+
+For $x,y \geq2^{k-2}$ we linearly combine equations concerning the first, last, and second to last rows of $A$ with factors of $1,1,-1$ respectively.
+We obtain constraints showing that $\sum_{x\leq2^{k-2}}q(y|x)=0$ for each $y>2^{k-2}$.
+We can then use the same reasoning to obtain that $q=q^*$ for the remaining values of $x,y$.
+\end{proof}
+
+This result is somewhat robust to slight perturbations in $p^*,h^*$.
+In particular, if we have some $\hat p\approx p^*$ and $\hat h\approx h^*$ then at each stage of the argument we can replace statements of the form $q(y|x)=0$ with statements of the form $q(y|x)\leq \epsilon$.
+Applying this with the kinds of arguments above will show that we can be sure that $\hat q$ is arbitrarily close to $q^*$ if we know that $\hat p,\hat h$ are sufficiently close to $p^*,h^*$.
+
+%%%%%%%% TODO!
+% {\huge still need to prove this}
+
+However, it turns out that the relationship between $q$ and $p^*,h^*$ is not robust in every situation.  In the next section we will see that it can in fact be quite discontinuous:
+
+\subsection{A discontinuity}
+
+Let us consider the case that $\Omega_\ell = \{1\}$ and $\Omega_X=\Omega_Y=\{1,2\}$.
+That is, there is only one population (no subpopulations) and both tool I and tool II can return one of $2$ possible values.
+We will now consider two possiblities:
+\begin{enumerate}
+\item First let us take the case
+    \begin{itemize}
+    \item $\mathbb{P}(X=1)=p^*(1)=0$
+    \item $\mathbb{P}(X=2)=p^*(2)=1$
+    \item $\mathbb{P}(Y=1)=h^*(1)=0$
+    \item $\mathbb{P}(Y=2)=h^*(2)=1$
+    \end{itemize}
+    In this case the MLM assumption $\sum_x p^*(x)q(y|x)=h^*(y)$ can be used to prove that $q(1|2)=0,q(2|2)=1$, but we now have \emph{absolutely no} knowledge of $q(1|1),q(2|1)$.
+This is because we simply never observed the case $X=1$ (it occurs with probability zero), and so we cannot possibly have any knowledge about $q(y|x)$ for $x=1$.
+\item Now let us take a slight variation:
+    \begin{itemize}
+    \item $\mathbb{P}(X=1)=p^*(1)=0.01$
+    \item $\mathbb{P}(X=2)=p^*(2)=0.99$
+    \item $\mathbb{P}(Y=1)=h^*(1)=0$
+    \item $\mathbb{P}(Y=2)=h^*(2)=1$
+    \end{itemize}
+    In this case we can again prove that $q(1|2)=0,q(2|2)=1$, but we can also prove that $q(1|1)=0,q(2|1)=1$.  
+\item Now we take yet another slight variation:
+    \begin{itemize}
+    \item $\mathbb{P}(X=1)=p^*(1)=0.01$
+    \item $\mathbb{P}(X=2)=p^*(2)=0.99$
+    \item $\mathbb{P}(Y=1)=h^*(1)=0.01$
+    \item $\mathbb{P}(Y=2)=h^*(2)=0.99$
+    \end{itemize}
+    In this case we can prove that $q(1|2)\leq 1/99$ and $q(2|1)\geq 1-1/99$, but we again cannot prove almost anything about $q(2|1)$.
+In particular, it is easy to produce cases in which $q(2|1)=0$ and other cases in which $q(2|1)=1$.
+
+\end{enumerate}
+
+The disturbing thing about this example is that by making infinitesimal perturbations to $p^*$ we can pass from uncertainty to complete certainty back to uncertainty.
+It is for this reason that in this paper we refuse to ever treat $\hat p,\hat h$ as fixed and given, always considering the space of perturbations around any such values.
+
+It is worth noting that these kinds of problems essentially vanish if the true $q^*$ is bounded away from zero i.e.\ $q^*(y|x)>c$ for every $x,y$ for some $c>0$.
+If this holds, together with a certain linear independence assumption, we can guarantee that the true $q^*$ is close to the set $\Theta(\hat p,\hat h)$ when $\hat p,\hat h$ are good approximations to $p^*,h^*$ (recall that we defined $\Theta$ in Equation (\ref{eq:thetadef})).
+In particular:\vspace{.1in}
+
+\begin{thm}\label{thm:mainthm}
+Fix any $q^*$ satisfying $q^*(y|x)>c$ for some $c>0$.
+Let us further assume that the matrix $B^*_{\ell,x}=p^*(x|\ell)$ has linearly independent rows.
+Fix any $p^*$ and define $h^*(y|\ell)=\sum_x p^*(x|\ell)q^*(y|x)$.
+Then by taking any $\hat p,\hat h$ sufficiently close to $p^*,h^*$ we can ensure that $q^*$ is arbitrarily close to some point in the set
+\begin{align}
+\hat \Theta \triangleq \left\{q\in T:\ \sum_x \hat p(x|\ell)q(y|x)= \hat  h(y|\ell) \quad \forall \ell,y \right\}
+\end{align}
+where by $T$ we mean the transition matrix polytope, $T=\{q:\ q(y|x)\geq 0, \sum_y q(y|x)=1\}$.
+\end{thm}
+\begin{proof}
+Let $\hat A$ denote the affine plane $\hat A=\{q:\ \sum_x \hat p(x|\ell)q(y|x)= \hat  h(y|\ell)\}$.  Thus $\hat \Theta = T \cap \hat A$.
+
+Now fix any $\hat p,\hat h$.
+Now let $\tilde q$ be the orthogonal Euclidean projection of $q^*$ to the affine space $A$.
+That is, $\tilde q$ is minimizes a sum-of-squares difference to $q^*$ among all the points in $A$.
+The linearly independent rows of $B^*$ allow us to bound the spectral norm of the right-pseudoinverse of $\hat B_{\ell,x}=\hat p(x|\ell)$, by taking $\hat p$ sufficiently close to $p^*$.
+If we furthermore require $\hat h$ sufficiently close to $h^*$, we can use this to ensure the projection distance is small.
+That is, we can force $\tilde q$ to be arbitrarily close to $q^*$.
+Using the fact that $q^*(y|x)>c$ we can thereby furthermore insure that $\tilde q(y|x)\geq0$.
+Finally, it is straightforward to see that the projection leaves the constraint $\sum_y q(y|x)=1$ unchanged.
+Thus $\tilde q\in \hat \Theta$ and $\tilde q$ is arbitrarily close to $q^*$.
+\end{proof}
+
+This theorem is certainly a step in the right direction, but we emphasize that it does require some assumptions.
+It is our opinion that the linear independence requirement is fairly mild (if it is not met then subpopulations can simply be merged together).
+However, the positivity requirement is quite troubling.
+In many cases of interest the true calibration $q^*$ has genuine zeros: pairs of measurements between the two tools which are fundamentally incompatible.
+In this case such a theorem cannot be applied.
+
+However, it may be that the above theorem's requirement ($q(x|y)>c>0$) is much stronger than is necessary.
+A precise understanding of the discontinuity example above has remained elusive.
+Better understanding could lead to a much more accurate and practical estimates.
+We leave it for future work.
 \end{document}

### 2018-07-19 markovlink/__init__.py

@@ -2,12 +2,11 @@ from . import estimation
 from . import exploration
 from . import mlm
 from . import misc
-from . import methodI
-from . import methodII
-from . import methodIII
-from . import majorization
+from . import minorization
+from . import logslines
+from . import polytopes
+from . import centering
 
 point_estimates = mlm.point_estimates
-bootstrap = mlm.bootstrap
-diameter_estimates = mlm.diameter_estimates
-qualitative_one_row = mlm.qualitative_one_row
+uncertainty_assessment = mlm.uncertainty_assessment
+hypothesis_test = mlm.hypothesis_test

### 2018-07-19 markovlink/centering.py

@@ -1,0 +1,131 @@
+'''
+Tools for finding the value of q which is closest to the analytic center of the polytope defined by
+
+          p@q = h
+'''
+
+import numpy as np
+import sys
+import scipy as sp
+import scipy.special
+import itertools
+from . import misc
+import time
+import scipy.sparse
+from . import logslines
+from . import polytopes
+
+def rundual(c,A,b,x,y,dualchunks,tol):
+    for i in range(dualchunks):
+        # get newton step
+        grad,hess=logslines.dual_grad_hess(c,A,b,y)
+        sd=-np.linalg.solve(hess,grad)
+
+        # in some cases newton direction may not be a search direction
+        # because of floating point error.  so we just give up.
+        # there's no way we're going to get better because floating point
+        # issues are going to slam us.
+        if np.sum(sd*grad)>=0:
+            return y
+        else:
+            stepsize=logslines.dual_simple_backtracking_for_stepsize(c,A,b,y,sd,initsize=1.0,grad=grad)
+            y=y+sd*stepsize
+
+            gap=logslines.dual_objective(c,A,b,y) - logslines.primal_objective(c,A,b,x)
+            if gap<tol:
+                return y
+
+    return y
+
+def runprimal(c,A,b,x,y,primalchunks,tol):
+    for i in range(primalchunks):
+        grad=c/x
+        sd=logslines.primal_newton_direction(c,A,b,x,method='eliminationII')
+
+
+        # in some cases newton direction may not be a search direction
+        # because of floating point error. 
+        # there's no way we're going to get better because floating point
+        # issues are going to slam us.
+        if np.sum(sd*grad)<=0:
+            return x
+        else:
+            stepsize=logslines.primal_simple_backtracking_for_stepsize(c,A,b,x,sd,feas_mult=.9)
+            x=x+sd*stepsize
+
+            gap=logslines.dual_objective(c,A,b,y) - logslines.primal_objective(c,A,b,x)
+            if gap<tol:
+                return x
+
+    return x
+
+def centerq(p,hstar,q=None,tol=.0001,dualchunks=50,primalchunks=10,precondition=True):
+    '''
+    Solves the problem
+
+    max sum(log(q))
+
[…]

### 2018-07-19 markovlink/exploration.py

@@ -1,3 +1,8 @@
+'''
+Some tools for bootstrap and exploring the polytopes of interest
+'''
+
+
 import scipy as sp
 import scipy.optimize
 import numpy as np
@@ -5,159 +10,37 @@ import numpy.random as npr
 import sys
 
 from . import misc
+from . import minorization
 
-###################################
-########## QUALITATIVE-1-ROW ######
-###################################
-
-def qualitative_one_row(phat,qhat,x):
-    szL,szX=phat.shape
-    szX,szY=qhat.shape
-
-    block=np.zeros((szY,szY))
-
-    for y in range(szY):
-        direc=np.zeros((szX,szY))
-        direc[x,y]=1
-        block[y] = find_extremal(phat,qhat,-direc)[x]
-
-    return block
-
-##################################
-########## DIAMETER ESTIMAT ######
-##################################
-
-def diameter_estimation(phat,qhat,n_samps=50):
+def construct_interval_samples(Nlx,Nly,n_extremals=50,alpha=.05,verbose=True):
     '''
-    Estimates the total variation diameter (that is, one half of the L1 diameter) of the polytope
-
-    Theta(phat,hhat) = {q: phat@q= hhat}
-
-    where hhat = phat @ qhat
-
-
+    Construct extremal samples to help us explore the confidence interval
     '''
+    szL,szX=Nlx.shape
+    szL,szY=Nly.shape
 
-    szL,szX=phat.shape
-    szX,szY=qhat.shape
+    Ms=np.zeros((n_extremals,szX,szY))
 
-    hhat= phat@qhat
+    if verbose:
+        misc.pnn("%d samples:"%n_extremals)
+    for i in range(n_extremals):
+        if verbose:
+            misc.pnn(i)
 
-    diam=0
-    samps=[]
+        # random initial conditions
+        p=np.array([npr.dirichlet(np.ones(szX)) for l in range(szL)])
+        q=np.array([npr.dirichlet(np.ones(szY)) for l in range(szX)])
 
-    for i in range(n_samps):
-        direc=(npr.randn(szX,szY)>0)*2-1 # L1 test directions
-        newq1 = find_extremal(phat,qhat,direc)
-        newq2 = find_extremal(phat,qhat,-direc)
-        samps.append([newq1,newq2])
-        dst=misc.totalvardist(newq1,newq2)
-        diam =np.max([diam,dst])
[…]

### 2018-07-19 markovlink/logslines.py

@@ -1,0 +1,317 @@
+'''
+Here we consider problems of the form
+
+   max_x     f(x)=sum(c*log(x))
+   subj      A@x=b
+
+where c>=0.  This problem also admits the dual, unconstrained problem 
+
+  min_lam   f*(lam) = sum(c*(log(c)-1)) - sum(c*log(A.T@lam)) + sum(lam*b)
+
+In the comments we will also refer to the lagrangian,
+
+    L(x,lam) = sum(c*log(x)) - sum(lam*(A@x-b))
+
+'''
+
+import numpy as np
+import sys
+import scipy as sp
+import scipy.special
+import itertools
+from . import misc
+import time
+import scipy.sparse
+
+class InfeasibleException(Exception):
+    pass
+
+def roundtrip_from_feasible_primal(c,A,b,x,tol=1e-10):
+    '''
+    Given a feasible x, we try to guess a value of lam, and then
+    use lam to try to get a guess of x.  If x is optimal we will
+    end up where we started.  
+    '''
+    if np.abs(A@x-b).max()>tol:
+        raise InfeasibleException("x not feasible")
+
+    return get_primal_from_dual(c,A,estimate_dual_from_primal(c,A,x))
+
+def estimate_gap_from_feasible_primal(c,A,b,x,tol=1e-10):
+    '''
+    Given a feasible x, we guess a value of lam and try
+    to use that to get a certificate of the duality gap
+    '''
+    lam = estimate_dual_from_primal(c,A,x)
+    return dual_objective(c,A,b,lam) - primal_objective(c,A,b,x,tol=tol)
+
+def primal_objective(c,A,b,x,tol=1e-10):
+    '''
+    Compute the primal objective, f
+    '''
+    if np.abs(A@x-b).max()>tol:
+        raise InfeasibleException("x not feasible")
+    if (x<=0).any():
+        raise InfeasibleException("x negative")
+
+    return np.sum(c*np.log(x))
+
+def lagrangian(c,A,b,x,lam):
+    return np.sum(c*np.log(x)) - np.sum(lam*(A@x-b))
+
+def dual_objective(c,A,b,lam):
+    '''
+    Compute the dual objective, f*
+    '''
+
+    xi = A.T@lam
+
+    if (xi>0).all():
+        return np.sum(c*(np.log(c)-1))-np.sum(c*np.log(A.T@lam)) +np.sum(lam*b)
+    else:
[…]

### 2018-07-19 markovlink/majorization.py

@@ -1,57 +1,0 @@
-import numpy as np
-import sys
-import scipy as sp
-import scipy.special
-import itertools
-from . import misc
-import time
-
-def train_q_by_majorization(p,Nly,kappa,q=None,n_train=10):
-    '''
-    Maximizes
-
-    L(q) = sum_{l,y} Nly log(sum_x plx qxy) + sum_{x,y} kappa log(qxy)
-
-    by a majorization algorithm
-
-    '''
-
-    # initialize
-    Nly=np.require(Nly)
-    l,x =p.shape
-    l,y = Nly.shape
-
-    assert misc.check_transition_matrix(p),"The rows of p should sum to 1 and all entries should be positive"
-    if q is None:
-        q=np.ones((x,y),dtype=np.float64)/y  
-    else:
-        assert misc.check_transition_matrix(q),"The rows of q should sum to 1 and all entries should be positive"
-
-    # run iterations
-    for i in range(n_train):
-        # update
-        q = train_mlm_em_update(Nly,p,q,kappa)
-       
-    return q
-
-def train_mlm_em_update(Nly,p,q,kappa):
-    '''
-    Uses a majorization technique to find a new value of q with a higher value of 
-    
-        L(q) = np.sum(Nlx * np.log(p @ q)) 
-    
-    starting from q as an initial guess.
-
-    '''
-    l,x =p.shape
-    x,y = q.shape
-    
-    # compute the majorization
-    pi = p.reshape((l,x,1)) * q.reshape((1,x,y))
-    pi = pi /np.sum(pi,axis=1,keepdims=True)
-    
-    # make updates
-    newq = np.sum(Nly.reshape((l,1,y)) * pi,axis=0).reshape(x,y)+kappa
-    newq = newq/np.sum(newq,axis=1,keepdims=True)
-    
-    return newq.reshape((x,y))

### 2018-07-19 markovlink/methodI.py

@@ -1,90 +1,0 @@
-import numpy as np
-import sys
-import scipy as sp
-import scipy.special
-import itertools
-from . import misc
-import time
-from . import estimation
-
-def train_q_primaldual_initialize(p,Nly,kappa,q):
-    szL,szY=Nly.shape
-
-    lam = Nly / (p@q)
-    gamma = -np.sum(p.T*(q@lam.T),axis=1) - szY*kappa
-
-    return gamma,lam
-
-def train_q_primaldual_residual(p,Nly,kappa,q,gamma,lam):
-    szL,szY=Nly.shape
-    szL,szX=p.shape
-
-    return np.r_[(q*(p.T@lam+gamma.reshape((-1,1)))).ravel()+kappa,np.sum(q,axis=1)-1,(lam*(p@q)-Nly).ravel()]
-
-def train_q_primaldual_update(q,gamma,lam,direc,s):
-    szL,szY=lam.shape
-    szX,szY=q.shape
-
-    q=q+s*direc[:szX*szY].reshape(q.shape)
-    gamma=gamma+s*direc[szX*szY:szX*szY+szX]
-    lam=lam+s*direc[szX*szY+szX:].reshape(lam.shape)
-
-    return q,gamma,lam
-
-def train_q_primaldual_jacobian(p,Nly,kappa,q,gamma,lam):
-    szL,szY=Nly.shape
-    szL,szX=p.shape
-
-    def iQ(x,y):
-        return np.ravel_multi_index((x,y),q.shape)
-    def iG(x):
-        return szX*szY+x
-    def iL(l,y):
-        return szX*szY + szX + np.ravel_multi_index((l,y),Nly.shape)
-
-    jac = np.zeros((szX*szY + szX + szL*szY,szX*szY + szX + szL*szY))
-
-    # rQ
-    for x in range(szX):
-        for y in range(szY):
-            jac[iQ(x,y),iQ(x,y)] = np.sum(lam[:,y]*p[:,x])+gamma[x]
-            jac[iQ(x,y),iG(x)] = q[x,y]
-            for l in range(szL):
-                jac[iQ(x,y),iL(l,y)] = q[x,y]*p[l,y]
-
-    # rG
-    for x in range(szX):
-        for y in range(szY):
-            jac[iG(x),iQ(x,y)]=1
-
-    # rL
-    for l in range(szL):
-        for y in range(szY):
-            jac[iL(l,y),iL(l,y)]=np.sum(p[l]*q[:,y])
-            for x in range(szX):
[…]

### 2018-07-19 markovlink/methodII.py

@@ -1,68 +1,0 @@
-import numpy as np
-import sys
-import scipy as sp
-import scipy.special
-import itertools
-from . import misc
-import time
-from . import estimation
-
-def train_q_primaldual_initialize(p,Nly,kappa,q):
-    szL,szY=Nly.shape
-
-    lam = Nly / (p@q)
-    gamma = -np.sum(p.T*(q@lam.T),axis=1) - szY*kappa
-
-    return gamma,lam
-
-def train_q_primaldual_residual(p,Nly,kappa,q,gamma,lam):
-    szL,szY=Nly.shape
-    szL,szX=p.shape
-
-    return np.r_[(p.T@lam+gamma.reshape((-1,1))).ravel()+kappa/q.ravel(),np.sum(q,axis=1)-1,(lam*(p@q)-Nly).ravel()]
-
-def train_q_primaldual_update(q,gamma,lam,direc,s):
-    szL,szY=lam.shape
-    szX,szY=q.shape
-
-    q=q+s*direc[:szX*szY].reshape(q.shape)
-    gamma=gamma+s*direc[szX*szY:szX*szY+szX]
-    lam=lam+s*direc[szX*szY+szX:].reshape(lam.shape)
-
-    return q,gamma,lam
-
-def train_q_primaldual_jacobian(p,Nly,kappa,q,gamma,lam):
-    szL,szY=Nly.shape
-    szL,szX=p.shape
-
-    def iQ(x,y):
-        return np.ravel_multi_index((x,y),q.shape)
-    def iG(x):
-        return szX*szY+x
-    def iL(l,y):
-        return szX*szY + szX + np.ravel_multi_index((l,y),Nly.shape)
-
-    jac = np.zeros((szX*szY + szX + szL*szY,szX*szY + szX + szL*szY))
-
-    # rQ
-    for x in range(szX):
-        for y in range(szY):
-            jac[iQ(x,y),iQ(x,y)] = -kappa/q[x,y]**2
-            jac[iQ(x,y),iG(x)] = 1
-            for l in range(szL):
-                jac[iQ(x,y),iL(l,y)] = p[l,y]
-
-    # rG
-    for x in range(szX):
-        for y in range(szY):
-            jac[iG(x),iQ(x,y)]=1
-
-    # rL
-    for l in range(szL):
-        for y in range(szY):
-            jac[iL(l,y),iL(l,y)]=np.sum(p[l]*q[:,y])
[…]

### 2018-07-19 markovlink/methodIII.py

@@ -1,157 +1,0 @@
-import numpy as np
-import sys
-import scipy as sp
-import scipy.special
-import itertools
-from . import misc
-import time
-from . import estimation
-
-def guess_gamma(Nly,p,q,kappa):
-    h=p@q
-
-    szX,szY=q.shape
-    szL,szX=p.shape
-    p=p.reshape((szL,szX,1))
-    q=q.reshape((1,szX,szY))
-    Nly=Nly.reshape((szL,1,szY))
-    pi = Nly*p*q
-    pi = pi/np.sum(pi,axis=1,keepdims=True)
-    newq=np.sum(pi*Nly,axis=0) + kappa
-    return np.sum(newq,axis=1)
-
-def gam_objective(Nly,p,q,kappa,gamma):
-    return np.sum(Nly*np.log(p@q)) + kappa*np.sum(np.log(q)) - np.sum(q.T@gamma)
-
-def calculate_surrogate_grad_and_hessian(Nly,p,q,kappa,gamma):
-    szX,szY=q.shape
-    szL,szX=p.shape
-
-    h=p@q
-
-    hess=np.zeros((szY,szX,szX))
-    for y in range(szY):
-        A = (np.sqrt(Nly)/h)[:,y][:,None]*p # ls x xs
-        hess[y] = - kappa*np.diag(q[:,y]**(-2)) - A.T@A
-
-    grad = p.T@(Nly/h) + (kappa/q) - gamma[:,None]
-
-    # perversely, the gradient means we should take a gradient step in the direction of grad.T
-    # because lam is (L x Y) whereas grad is (Y x L)
-
-    return grad,hess
-
-def calculate_surrogate_search_direction(Nly,p,q,kappa,gamma):
-    grad,hess=calculate_surrogate_grad_and_hessian(Nly,p,q,kappa,gamma,)
-    return -np.array([np.linalg.solve(h,g) for (h,g) in zip(hess,grad.T)]).T
-    # return -np.array([sp.sparse.linalg.lsqr(h,g,damp=0.4,iter_lim=2000)[0] for (h,g) in zip(hess,grad.T)]).T
-
-def update_q_by_surrogate(Nly,p,q,kappa,gamma,searchdirection,beta=.8):
-    szX,szY=q.shape
-    szL,szX=p.shape
-
-    curobjective = gam_objective(Nly,p,q,kappa,gamma) # want to make this big
-
-    # go forward, but not too much
-    s=1.0
-    working=False
-    while not working:
-        newq = q + searchdirection*s
[…]

### 2018-07-19 markovlink/minorization.py

@@ -1,0 +1,275 @@
+import numpy as np
+import sys
+import scipy as sp
+import scipy.stats
+import scipy.special
+import itertools
+from . import misc
+import time
+from . import logslines
+from . import polytopes
+import numpy.random as npr
+
+class OutOfTime(Exception):
+    pass
+class OutOfIterations(Exception):
+    pass
+
+def LL_ph(Nlx,Nly,p,h):
+    '''
+    We return log likelihood of Nlx,Nly under p,h
+
+        sum(Nlx*log(p)) + sum(Nly*log(h))
+
+    with the convention that 0*log(0/0) = 0
+    '''
+
+
+    # compute the log likelihood for Nlx
+    good=(Nlx>0)
+    Nlxratio = np.sum(Nlx[good]*np.log(p[good]))
+    
+    # compute the log likelihood for Nly    
+    good=(Nly>0)
+    Nlyratio = np.sum(Nly[good]*np.log(h[good]))
+
+    # done!
+    return Nlxratio+Nlyratio
+
+def LLR(Nlx,Nly,p,h):
+    '''
+    LLR =  mean(Nlx*log(phat/p)) + mean(Nly*log(hhat/h))
+
+    with the convention that 0*log(0/0) = 0
+    '''
+
+    # compute empirical averages
+    phat = Nlx/np.sum(Nlx,axis=1,keepdims=True)
+    hhat = Nly/np.sum(Nly,axis=1,keepdims=True)
+
+    # compute the log likelihood ratios for Nlx
+    good=(Nlx>0)
+    Nlxratio = np.mean(Nlx[good]*np.log(phat[good]/p[good]))
+    
+    # compute the log likelihood ratios for Nly    
+    good=(Nly>0)
+    Nlyratio = np.mean(Nly[good]*np.log(hhat[good]/h[good]))
+
+    # done!
+    return Nlxratio+Nlyratio
+
+def estimate_PPF(Nlx,Nly,alpha,n_samps=1000):
+    '''
+    Get an approximate estimate for the value of PPF so that 
+
+        P(LLR(Nlx,Nly,pstar,hstar) > PPF) = alpha
+
+    when Nly is drawn from pstar 
+         Nly is drawn from hstar
+
+    '''
+
+    # we assume the TRUTH is given by pseudocounted empirical distributions
+    # (the pseudocounts end up giving us a more conservative estimator)
+    # we then look at samples from the distribution of LLR
+    samps=PPF_samples(Nlx,Nly)
+
+    # we estimate our object of interest based on these samples
[…]

### 2018-07-19 markovlink/misc.py

@@ -2,6 +2,8 @@
 import numpy as np
 import sys
 import scipy as sp
+import numpy.random as npr
+
 
 def pnn(x): 
     sys.stdout.write(str(x) + " ")
@@ -11,4 +13,41 @@ def check_transition_matrix(a):
     return (np.abs(np.sum(a,axis=1) - 1).max()<1e-9) and (np.min(a)>-1e-9)
 
 def totalvardist(a,b):
-    return .5*np.sum(np.abs(a-b))
+    '''
+    Returns the average total variation distance between two discrete conditional
+    distributions, parameterized by a transition matrix.  
+    '''
+    return .5*np.mean(np.sum(np.abs(a-b),axis=1),axis=0)
+
+def log_likelihood(Nlx,p):
+    good=(Nlx>0)
+    return np.sum(Nlx[good]*np.log(p[good]))
+
+
+def resample(tmx):
+    '''
+    Resamples each row of tmx, with replacement
+    '''
+    tmx2=np.zeros(tmx.shape)
+    for z in range(len(tmx)):
+        p=tmx[z]
+        p=p/np.sum(p)
+        tmx2[z]=npr.multinomial(np.sum(tmx[z]),p)
+    return tmx2
+
+
+def sample(N,p):
+    '''
+    Returns a matrix formed by sampling
+
+        Multinomial(N[1],p[1])
+        Multinomial(N[2],p[2])
+                .
+                .
+                .
+    '''
+    tmx2=np.zeros(p.shape)
+    for z in range(len(N)):
+        tmx2[z]=npr.multinomial(np.sum(N[z]),p[z])
+    return tmx2
+

### 2018-07-19 markovlink/mlm.py

@@ -2,74 +2,36 @@ from . import estimation
 from . import exploration
 from . import misc
 import numpy as np
-from . import majorization
+from . import minorization
 
-def point_estimates(Nlx,Nly,pseudocount=1.0):
+def hypothesis_test(Nlx,Nly,alpha=.05,maxtime=120):
+    p,q=point_estimates(Nlx,Nly)
+    PPF=minorization.LLR(Nlx,Nly,p,p@q)
+    PPF = minorization.estimate_PPF(Nlx,Nly,alpha)
+
+    p,q,niter=minorization.find_feasible_point(Nlx,Nly,PPF,maxtime=maxtime)
+
+    return "fail to reject"
+
+def point_estimates(Nlx,Nly,maxtime=10):
     '''
-    Get point estimates by the Markov Link Method
-    '''
-
-    p = (Nlx + pseudocount) / np.sum(Nlx+pseudocount,axis=1,keepdims=True)
-
-    kappa=.001
-    q=majorization.train_q_by_majorization(p,Nly+pseudocount,kappa,n_train=500)
-
-    h=p@q
-
-    return p,q,p@q
-
-def bootstrap(Nlx,Nly,phat,qhat,n_straps=50,verbose=False):
-    '''
-    Draws surrogate datasets for Nlx and Nly, with replacement, and then gets point estimates
-    for the surrogate datasets using the Markov Link Method.
-
-    Also estimates the Deltap,Deltaq,Deltah
+    Get point estimates for the calibration
     '''
 
-    hhat=phat @ qhat
+    # find a q inside an approximate 95% confidence interval
+    PPF = minorization.estimate_PPF(Nlx,Nly,.05)
+    p,q,niter=minorization.find_feasible_point(Nlx,Nly,PPF,maxtime=maxtime)
 
-    straps=[]
+    # keep seeking a q that better optimizes the likelihood of the data,
+    # for twice the number of iterations it took to find the feasible point:
+    p,q = minorization.train_mlm_minorized(Nlx,Nly,p,q,0,1,niter*2)
 
-    dsts=np.zeros((n_straps,3))
+    return p,q
 
-    if verbose:
-        misc.pnn("%d bootstraps:"%n_straps)
-    for i in range(n_straps):
-        if verbose:
-            misc.pnn(i)
-        Nlx2=exploration.resample(Nlx)
-        Nly2=exploration.resample(Nly)
[…]

### 2018-07-19 markovlink/plotting.py

@@ -1,0 +1,20 @@
+def Ms_plot(Ms,x,additional_rows=None,additional_labels=None):
+    import matplotlib.pylab as plt
+
+    idx=5
+
+    plt.gcf().add_axes([0,.6,1,1])
+    plt.pcolor(Ms[:,x],vmin=0,vmax=1); plt.yticks([]); plt.xticks([])
+    plt.ylabel("$q$s from the confidence interval")
+    plt.title("Inspecting $q(\\cdot | %d)$"%idx)
+
+    plt.gcf().add_axes([0,.3,1.0,.2])
+    plt.pcolor(qtilde[idx,None],vmin=0,vmax=1); plt.yticks([]); plt.xticks([])
+    plt.ylabel("$\\tilde q$")
+
+    plt.gcf().add_axes([0,0,1.0,.2])
+    mappable=plt.pcolor(qxy_star[idx,None],vmin=0,vmax=1); plt.yticks([]); plt.xticks([])
+    plt.ylabel("$q^*$")
+
+    plt.gcf().add_axes([1.1,0,.1,1.6])
+    plt.colorbar(mappable,cax=plt.gca())

### 2018-07-19 markovlink/polytopes.py

@@ -1,0 +1,269 @@
+'''
+Defining and exploring polytopes involved in the MLM
+'''
+
+import scipy as sp
+import scipy.optimize
+import numpy as np
+import numpy.random as npr
+import sys
+from . import misc
+
+def precondition_constraint(A,b,tol=1e-10,return_preconditioner=False):
+    '''
+    Transforms the constraint 
+
+       Ax=b --> Vx=b2
+
+    where U is unitary.  We do this by writing 
+
+        Ax= UEVx = b
+
+    And replacing this with Vx = E^-1 U^T b.  That is, we premultiply both sides by
+
+        E^-1 U^T
+
+    We drop singular values less than tol
+    '''
+
+    U,e,V = np.linalg.svd(A)
+
+    e=e[e>tol]
+
+    V=V[:len(e)]
+
+    if return_preconditioner:
+        return V, (U.T@b)[:len(e)]/e,np.diag(1/e)@U.T[:len(e)]
+    else:
+        return V, (U.T@b)[:len(e)]/e
+
+
+def project_to_polytope(A,b,y,feasiblepoint=None):
+    '''
+    Project y to the polytope
+
+    Ax=b
+    x>=0
+    '''
+
+    '''
+    cvxopt wants this in the form
+
+    qp(P, q, G=None, h=None, A=None, b=None)
+
+    minimize    (1/2)*x'*P*x + q'*x 
+    subject to  G*x <= h      
+                A*x = b.
+
+    In our case we have 
+
+    - G=-I_n
+    - h=-0_n
+    - A =A
+    - b=b
+    - P=I_n
+    - q = -y
+
+    ''' 
+
+    import cvxopt
+
+    szM,szN=A.shape
+    com=lambda x: cvxopt.matrix(x)
+    
+    ident=np.eye(szN)
+    h=np.zeros((szN,1))
+    q=-y.reshape((-1,1))
+
+
+    if feasiblepoint is not None:
+        rez = cvxopt.solvers.qp(com(ident),com(q),com(-ident),com(h),com(A),com(b),
+            initvals=dict(x=com(feasiblepoint)))
+    else:
+        rez = cvxopt.solvers.qp(com(ident),com(q),com(-ident),com(h),com(A),com(b))
+
+    return rez
+
+
+def q_with_fixedph(plx,hly,precondition=True,preconditiontol=1e-10):
+    '''
+    make explicit the equality constraints 
+    that define {q: p @ q = p @ qstar, q@1=1}
+    '''
[…]

### 2018-08-06 paper/paper.tex
Confidence intervals, which we're now ignoring...

@@
 
-\newcommand{\figgygr}[2]{\includegraphics[width=#1\textwidth]{#2}}
-\newcommand{\figgygrs}[3]{\begin{tabular}{c}\includegraphics[width=#1\textwidth]{#2}\\#3\end{tabular}}
+% \newcommand{\figgygr}[2]{\includegraphics[width=#1\textwidth]{#2}}
+% \newcommand{\figgygrs}[3]{\begin{tabular}{c}\includegraphics[width=#1\textwidth]{#2}\\#3\end{tabular}}
+% \newcommand{\figgygrsTable}[3]{\begin{tabular}{c}{#2}\\#3\end{tabular}}
+% \newcommand{\figgygrsH}[3]{\begin{tabular}{c}\includegraphics[width=#1\textwidth,angle=270]{#2}\\#3\end{tabular}}
+
+\newcommand{\figgygr}[2]{pic}
+\newcommand{\figgygrs}[3]{\begin{tabular}{c}pic\\#3\end{tabular}}
 \newcommand{\figgygrsTable}[3]{\begin{tabular}{c}{#2}\\#3\end{tabular}}
-\newcommand{\figgygrsH}[3]{\begin{tabular}{c}\includegraphics[width=#1\textwidth,angle=270]{#2}\\#3\end{tabular}}
-
-% \newcommand{\figgygr}[2]{pic}
-% \newcommand{\figgygrs}[3]{\begin{tabular}{c}pic\\#3\end{tabular}}
-% \newcommand{\figgygrsTable}[3]{\begin{tabular}{c}{#2}\\#3\end{tabular}}
-% \newcommand{\figgygrsH}[3]{\begin{tabular}{c}pic\\#3\end{tabular}}
+\newcommand{\figgygrsH}[3]{\begin{tabular}{c}pic\\#3\end{tabular}}
 
@@
 
-\title{Markov Link Method for calibrating without joint measurement, including the case of destructive measurements}
+\title{The Markov Link Method: a nonparametric approach to combine observations from multiple experiments}
 \author{Jackson Loper, Osnat Penn, Trygve Bakken, David Blei, Liam Paninski}
@@
 \begin{abstract}
-A proliferation of new experimental tools has left a serious gap: calibration.
-Two thermometers can be calibrated against each other by simply measuring the same bodies of water with both thermometers, but the problem is much harder for many modern tools.
-One common problem is that we do not have measurements from the same ``body of water'' for both tools.
-We propose the Markov Link Method (MLM) as a way to overcome this difficulty.
-This method produces consistent estimators that tightly bound the calibration, i.e.
-the conditional distribution of one tool's measurement given another tool's measurement.
-It achieves this without any measurement data from both tools applied to the same ``bodies of water.'' Moreover, MLM can be applied even if we cannot make any assumptions about what calibrations we might expect to see; instead, it uses a subpopulation-based conditional independence assumption.
-We evaluate MLM on a pair of single-cell RNA techniques, obtaining a calibration between the tools.
+Using observations from multiple experiments is essential to building larger scientific theories.
+For example, we might analyze different aspects of human neurons, such as morphology, transcriptomics, or electrophysiology.
+A strong theory would account for each perspective as well as the relationship between the perspectives.
+Unfortunately, we rarely have a single experiment that simultaneously measures every viewpoint, uses a sufficient sample size, and gets sufficient measurement precision.
+In practice, we must therefore conduct different experiments and find a way to link the results.
+The Markov Link Method (MLM) offers an algorithm to reason about these `links.' The method is nonparametric: it makes no assumptions about what the link should look like.
+We evaluate the MLM on a pair of single-cell RNA techniques and gain new insight about the link between the two methods.
 \end{abstract}
 
-The modern setting is rife with experimental measurement tools, and it can be very frustrating to understand how the output of these tools relate to one another.
-This problem is known as ``calibration'' or ``zeroing.''  A calibration tells us what readings we should expect from one tool, given the reading we obtained from another tool.
-Calibration additionally must give uncertainty bounds for how much we can trust those expectations  \cite{bipm2008international}.
-Calibration between measurement tools allows us to combine experimental results from different labs and different methodologies into larger scientific theories.
+The last century has seen a proliferation of experimental techniques.
+Different perspectives arise from different techniques (and even from different labs performing the `same' protocol).
+Ideally, this yields theories that account for all perspectives.
+Instead, sometimes, this leaves us with many perspectives and no idea how they are connected.
 
-Formally, a calibration is simply a conditional distribution.
-We will denote it by $q^*(y|x)$.
-Let us say we obtain measurement result $x$ from one tool on a particular specimen.
-The number $q^*(y|x)$ indicates the probability of obtaining result $y$ from a second tool applied to measure the same specimen.
-One way to learn the calibration is to measure the same specimens with both tools.
-We call this ``joint measurement.''  Unfortunately, calibrations are often required even when joint measurement is unavailable.
-For example, if the measurement tool significantly alters the specimen being measured, joint measurement is simply impossible.
-In other cases, it may be expensive or impractical.
+Our goal is to connect different perspectives.  To make this idea rigorous, we assume each experiment has two aspects:
+%
+\begin{itemize}
+    \item a \emph{sampling procedure} determines what we study
+    \item a \emph{measurement procedure} defines the perspective we take in studying it.
+\end{itemize}
+%
+If two experiments differ in both aspects, it may be difficult to gain insight by combining their data.
+For example, imagine one experiment performs a transcriptomic measurement of liver cells and another applies a morphological measurement of heart cells.
+Without additional experiments or knowledge, looking at these two experiments together may not be very helpful.
+On the other hand, if two experiments have some aspect in common, there is something to be learned.
+In this sense, a collection of experiments can be understood as a network: a pair of experiments are connected if they have some aspect in common.
+In some cases, we can also connect two experiments if they differ in both aspects but we have prior knowledge (e.g.\ if the sampling procedures are different, but a reweighting can be used to correct for the difference).
+When many experiments can be joined into a common network, much can be learned.
 
-We here propose the Markov Link Method (MLM) to estimate calibrations between tools.
-The MLM can be applied without any joint measurement.
-The key idea is to use multiple subpopulations of specimens.
-If each subpopulation captures a different slice of the overall population, we can obtain tight bounds on the true calibration.
-This is true even if the subpopulations are highly heterogeneous and overlapping.
-By integrating information from all the subpopulations we can make rigorous deductions about what the calibration might be.
-MLM also gives suggestions about which further subpopulations might be helpful to study in order to further refine our knowledge of the true calibration.
+We here propose the The Markov Link Method (MLM), an algorithm that uses a network of experiments to learn the relationship between two different perspectives.
+We formalize this through what we call the `measurement link' parameters, $q^*(y|x)$.
+The meaning of these parameters is defined as follows.
+Let us say we obtain measurement result $x$ from one measurement procedure.
+The number $q^*(y|x)$ indicates the probability of obtaining result $y$ from the second measurement procedure applied to measure the same specimen.
+The link tells us how the properties measured by each procedure are associated.
+For example, the link lets us answer questions such as ``I have observed the transcriptome of this cell -- what morphologies might this cell have?''
+
+If we have a single large experiment that measures each specimen with both measurement procedures, we can estimate the link without the MLM.
+In this case it is straightforward to study the link between the two perspectives, because we can examine specimens through both perspectives simultaneously.
+However, in many cases we have some experiments in which specimens are measured with one procedure, other experiments using the other procedure, and few or no experiments where specimens are measured with both procedures.
+This is particularly true if the measurement techniques destroy or alter the specimen in the process of measuring it.
+
+In this paper we focus particularly on the case that we have a network of experiments based on several sampling procedures and exactly two different measurement procedures.
+We also suppose that the measurement procedures return categorical observations, i.e.\ a `measurement' assigns a specimen to one of a finite set of categories.
+The MLM could also generalize to numerical observations, but in that case it would be appropriate to incorporate some knowledge of smoothness (e.g.\ through a Gaussian assumption).
+Categorical data requires no additional modeling assumptions, so for simplicity we will here focus on the categorical case.
+This setup, sketched in Figure \ref{fig:bigpicture}, is perhaps the simplest example that highlights all the important assumptions and properties of the MLM.
 
 \figgy{
-\hfill{}\figgygrs{.5}{../images/bigpicture}{(a) Setup: each specimen can be \\ measured with either tool I or tool II, \\ but can't be measured with both.\\ How do we calibrate the tools?}
-\hfill{}\figgygrs{.4}{../images/plate}{(b) Key modeling assumption: \\ the calibration $q$ is the same \\ for every subpopulation. }\hfill{}
+\hfill{}\figgygrs{.5}{../images/bigpicture}{(a) Setup: each specimen can be \\ measured with one of two procedures, \\ but can't be measured with both.\\ How do we study the relationship \\ between the procedures?}
+\hfill{}\figgygrs{.4}{../images/plate}{(b) Key modeling assumption: \\ the link $q$ is the same for \\ every sampling procedure. }\hfill{}
 
-\caption{{\bf The Markov Link Method assumption makes calibration possible, even without joint measurement}.
-In subfigure (a) we measure some neurons with tool I and others with tool II.
-How can we calibrate the tools?
-That is, how can we know what we might have seen if we had measured the first group with tool II or the second group with tool I?
-The Markov Link Method (MLM) gives a way to get a partial answer, using a subpopulation analysis.
-In subfigure (b) we present a plate diagram describing our beliefs about the conditional independencies in this situation; cf.\ \cite{koller2009probabilistic} for a thorough explanation of this type of diagram.
-For each specimen we are interested in $\ell$ (subpopulation), $X$ (tool-I measurement), and $Y$ (tool-II measurement).
-For $n$ samples we can see $(\ell,X)$ and for $m$ samples we can see $(\ell,Y)$.
-However, for the first group of $n$ samples we cannot observe $Y$, and for the second group of $m$ samples we canot observe $X$.
-This is indicated in the diagram by the grayed-out circles.
-This lack of joint measurement makes it hard to learn the variable distributions.
-The MLM has a key assumption that makes it possible: we assume $X$ is sufficiently detailed to \emph{statistically isolate} $Y$ from $\ell$.
-This assumption is reflected in the arrows of the diagram; all paths from $\ell$ variables to $Y$ variables pass through $X$ variables.
-The assumption allows us to learn much about the important probability distributions, as parametrized by the tool-I measurement distribution $p(x|\ell)$ and the calibration $q(y|x)$.
+\caption{{\bf The Markov Link Method (MLM) links different measurement procedures by using a set of different sampling procedures.}  In subfigure (a) we have neurons gathered using many different sampling procedures.
+We select half of the neurons and determine what type of neuron they are, using transcriptomic measurements.
+We classify the second half of neurons based on a morphological properties.
+The MLM gives a way to link these two types of measurement and reconcile the two notions of cell-type.
+In subfigure (b) we present a plate diagram articulating the statistical assumption that makes it possible (cf.\ \cite{koller2009probabilistic} for a thorough explanation of this type of diagram).
+For each neuron in each group, this diagram considers the sampling procedure ($\ell$), the transcriptomic type ($X$), and the morphological type ($Y$).
+The assumption is that the transcriptomic type is sufficiently detailed to \emph{statistically isolate} the transcriptomic type from the sampling procedure.
+This is reflected in the arrows of the diagram; all paths from $\ell$ variables to $Y$ variables pass through $X$ variables.
+The assumption allows us to estimate the link, $q$.
 \label{fig:bigpicture}}
 }
-
-This paper will proceed in the following sections:
-\begin{itemize}
-    \item Description of how to apply the Markov Link Method, using an example to explain the process
-    \item A variety of examples where the method may apply
-    \item The method's performance in simulation
-    \item The method's results on real-world data
-    \item Relation to prior work
-    \item Discussion
-\end{itemize}
 
@@
 
-We begin with an example.
-
-Imagine that we have a large collection of human cells.  We will call each cell a `specimen.'  For each specimen we have:
+Imagine that we have a large collection of human cells.  For each cell we are interested in 
 %
 \begin{enumerate}
-    \item $\ell$, the sampling strategy used to obtain the specimen.
-We will also call $\ell$ the specimen ``subpopulation,'' for each strategy should capture a different subpopulation of the overall set of specimens.
-For example, we could have a few strategies based on selecting cells of particular sizes, a few strategies based on selecting cells from particular places in the body, and a few strategies based on based on selecting cells with particular proteomic markers.
-Each additional sampling strategy shows us a different subpopulation of the cells and strengthens what kinds of inferences are possible.
+    \item $\ell$, the sampling procedure used to obtain the specimen.
+For example, we can use select different subpopulations of cells based on proteomic markers.
+For each marker we might have a different sampling procedure designed to select cells with that marker.
     \item $X$, the cell's `transcriptomic type,' as measured by looking at the transcriptomic activity in the cell.  
@@
 %
-Our task is to reconcile the two different notions of `cell-type' indicated by $X$ and $Y$, by constructing what we call a `calibration.'  This calibration should tell us what transcriptomic types are associated with what morphological types.
-For example, we might like to be able to say ``This cell has transcriptomic type 3, therefore it probably has morphological type F.''  We can make math of this idea by defining a calibration as a conditional probability $q^*(y|x,\ell)$.
-This object indicate the likelihood that the morphological type is $y$ given that the transcriptomic type is $x$ and the specimen was sampled with strategy $\ell$.
-In these terms, our task is to learn the calibration $q^*$.
-
-Learning $q^*$ is hard when ``joint measurement'' is impossible or impractical.
-By this we mean that for any given specimen we can either observe $\ell, X$ or $\ell,Y$.
-We can never observe $\ell,X,Y$ for any specimen.
-This makes it much more difficult to estimate the calibration; the fundamental difficulty is sketched in Figure \ref{fig:bigpicture}.
-To make calibration possible without joint measurement, we need an assumption additional assumption.
-The Markov Link Method supplies this assumption in the form of a conditional independence assumption.
-This makes it possible to learn about the calibration.
-
-To make it possible to learn $q^*$ even without joint measurement, the MLM makes the assumption that the calibration is the same for every sampling strategy:
+Our task is to reconcile the two different notions of `cell-type' indicated by $X$ and $Y$.
+We would like to know what transcriptomic types are associated with what morphological types.
+For example, we might like to say ``This cell has transcriptomic type 3, therefore it probably has morphological type F.''  Our main assumption is that these kinds of associations are unaffected by the choice of sampling procedure.
+In particular:
 
@@
 \begin{center}
-\textbf{The Markov Link Method Assumption}\\$q^*(y|x,\ell)=q^*(y|x)$ for each $\ell$
+\textbf{The Markov Link Method Assumption}\\$\mathbb{P}(Y|X,\ell)=q^*(y|x)$ for each $\ell$
 \end{center}
@@
 
-This assumption can be understood through the following thought experiment.
-Let us say we are told that a cell has transcriptomic type $x$.
-We could then make predictions about the morphological type $y$.
-Now we are told new information: the cell was obtained via sampling strategy $\ell$.
-Does this change our predictions about the morphological type?
+Here $\mathbb{P}(Y|X,\ell)$ indicates the probability that the morphological type is $y$ given that the transcriptomic type is $x$ and the specimen was sampled with strategy $\ell$.
+The assumption is that this probability actually does not depend upon $\ell$.
+This can be understood intuitively through a thought experiment.
+Let us imagine we have attained perfect understanding of cellular biology.
+Someone selects a cell using one of a set of sampling procedures and measures the transcriptomic activity of the cell.
+We are then told the cell's transcriptomic activity (but we are not told the sampling procedure used to gather the specimen).
+Using our perfect knowledge of the physical systems, we could then make predictions about a measurement of the cell's morphology.
+Now we are told new information: we are told the sampling procedure used to gather the specimen.
+Would this substantially change our predictions?
 If so, the MLM assumption is violated.
-However, if $X$ is sufficiently informative then learning the sampling strategy will not change our predictions about the morphological type.
-In this case, the MLM assumption holds.
+However, if the transcriptomic measurement is sufficiently informative, learning the sampling strategy would not substantially change our predictions about the cell morphology.
+As long as we can find a measurement procedure that is sufficiently informative in this way, we can apply the MLM.
+A list of examples where the MLM might apply may be found in Section \ref{sec:examples}.
 
-If the MLM assumption holds, a frequentist point of view suggests the following procedure for estimating the calibration $q^*$ from data:
+Our task is to estimate the link, $q^*$, from a collection of experiments.
+We suppose that for each sampling procedure we have two experiments: one that measured transcriptomics and another that measured morphology.
+We have no experiments in which both types were measured for the same specimens.
+This setup is sketched in Figure \ref{fig:bigpicture}.
+Different setups could also be considered.
+For example, we might additionally have a small experiment in which both measurement procedures could be applied to the same specimen.
+We might have more than two measurement procedures.
+We might have measurement procedures that yield numerical observations instead of categorical ones.
+Generalizing the method to all these cases should be straightforward, but we leave it for future work.
 
-\figgy{
-\hfill{}\figgygr{.9}{../images/qtildesamps}\hfill{}
-\caption{{\bf Estimating the calibration and measuring our uncertainty}.
-\label{fig:qtildesamps} On the left we visualize our best guess for true the calibration between two different tools for measuring cell-type.
-Tool I uses one approach to obtain a fine-grained classification of cells, whereas tool II uses a different approach and only gives coarse type indication.
-We can visualize our estimated calibration with a heatmap, formed by drawing a rectangle for each $x,y$ with a color indicating the magnitude of $\tilde q(y|x)$.
-For example, examine the bright yellow square on the `Pvalb Gpr149' row and the `unknown' column.
-This suggests that if a cell has the `Pvalb Gpr149' type according to tool I, it is extremely likely to be assigned the `unknown' type by tool II.
-However, our best guess might not be correct.
-On the right we show many plausible alternatives for the `Pvalb Gpr149' row of the calibration.
-Each possibility is drawn from a 95\% confidence region for calibration.
-Some of the alternatives suggest that some `Pvalb Gpr149' cells might not be classified as `unknown,' but might instead be correctly identified as being of the `Pvalb' type.
-The data we have cannot tell us which of these alternatives is closest to the truth.
-More data and more subpopulations would be necessary to reduce the variability within this confidence region.
-However, we can have some confidence that `Pvalb Gpr149' cells are not being classified as `Meis2' or `Vip', insofar as none of the plausible calibrations give high probability to these tool II types.}
-}
+In the specific case we will focus on, all of the experimental data can be summarized with two matrices, $N^X$ and $N^Y$.
+Here $N^X_{\ell x}$ indicates the number of specimens gathered with procedure $\ell$ that have transcriptomic type $x$.
+Likewise $N^Y_{\ell y}$ indicates the number of specimens gathered with procedure $\ell$ that have morphological type $y$.
 
-\begin{enumerate}
-    \item Summarize our observed data in two matrices:
-    \begin{align*}
-    N^X_{\ell x} &= |\{i\leq n:\ \ell_i=\ell,X_i=x\}|\\
-    N^Y_{\ell y} &= |\{j\leq m:\ \ell_j=\ell,Y_j=x\}|
-    \end{align*}
-    That is, $N^X_{\ell x}$ indicates the number of specimens from subpopulation $\ell$ with transcriptomic type $x$.
-Likewise $N^Y_{\ell y}$ indicates the number of specimens from subpopulation $\ell$ which with morphological type $y$.
+Taking a frequentist point of view, the Markov Link Method uses this data to produce two kinds of objects:
+\begin{itemize}
+    \item $\tilde q(y|x)$, a point estimate for the link parameter $q^*(y|x)$.
+To obtain this we must incorporate nuisance parameters: let $p^*(x|\ell)$ denote the probability that a cell sampled with procedure $\ell$ will have transcriptomic type $x$.
+The log likelihood of the data is then given by
+    \[
+    L(p,q)=\left(\sum_{\ell,x} N^X_{\ell x} \log p(x|\ell) + \sum_{\ell,y} N^Y_{\ell y} \log \left(\sum_x p(x|\ell) q(y|x)\right)\right)
+    \]
+    We use a maximum likelihood approach to build our point estimators:
+    \[
+    \tilde p,\tilde q \in \argmax_{p,q} L(p,q)
+    \]
+    We caution that this optimization problem does not always have a unique maximizer, so the result may depend upon the maximization strategy used.
+Precise details for our optimization algorithms may be found in Appendix \ref{sec:optproblems}.
+This non-uniqueness is related certain so-called `identifiability' issues: in some cases it may be impossible to obtain a perfect point estimate even with an infinite number of samples.
+We discuss this issue at length in Appendix \ref{sec:casestudies}.
+In light of these concerns, it is particularly important to also consder our uncertainty:
+    %
+    \item $C_{x,y}$, a 95\% confidence interval for $q^*(y|x)$.  We obtain these with profile likelihood approach.  The profile likelihood is defined as
+    \[
+    L_{x y}(\xi)=\max_{\substack{p,q \\ q(y|x)=\xi}} L(p,q)
+    \]
+    This is the maximum possible likelihood of the data under the constraint that $q(y| x)=\xi$.  We then take $L^* = \sup_{p,q} L(p,q)$ and define a confidence region for $q^*( y| x)$ by
+    \[
+    R_{x y} = \left\{\xi:\ L^* - L(\xi) < k\right\}
+    \]
+    We choose $k$ to achieve the desired coverage.
+That is, we would like to choose $k$ so that $q^*(y|x) \in R_{x y}$ with probability at least 95\%.
+We refer the reader to Appendix \ref{sec:confinterval} for details on choosing $k$.
+Other coverages besides 95\% can also be produced by the same method.
+Finally, for simplicity of visualization we transform the set $R_{x y}$ (which could be composed of several disjoint intervals) into a simple interval by taking $C_{x,y} = [\min R_{x,y},\max R_{x,y}]$.
+Since this interval is a superset of the region, it will certainly contain $q^*(y|x)$ as long the region does.
+\end{itemize}
 
-    \item Define a confidence region for calibration $q^*$.
-To do this we incorporate a nuisance parameter: let $p^*(x|\ell)$ denote the probability that a cell sampled with strategy $\ell$ will have transcriptomic type $x$.
-We form a joint simultaneous confidence region for all of $p^*,q^*$ using the following likelihood ratio statistic:
-    \[
-    L(p,q)=\frac{1}{\Omega_X}\sum_{\ell,x} N^X_{\ell x} \log \frac{\hat p(x|\ell)}{p(x|\ell)} + \frac{1}{\Omega_Y}\sum_{\ell,y} N^Y_{\ell y} \log \sum_x \frac{\hat h(y|\ell)}{p(x|\ell) q(y|x)}
-    \]
-    where $\hat p(y|x)=N^X_{\ell x} / \sum_{x'}N^X_{\ell x'}$ and $\hat h(y|x)=N^Y_{\ell y} / \sum_{y'}N^Y_{\ell y'}$ denote empirical conditional distributions.
-We then define our confidence region as
-    \[
-    R = \{(p,q):\ L(p,q) < k\}
-    \]
-    and choose $k$ to ensure any desired coverage.
-For example, to produce a 95\% confidence region, we would like to choose $k$ so that the probability that $p^*,q^* \in R$ is at least 95\%.
-We refer the reader to Appendix \ref{sec:confinterval} for details on choosing $k$.
-
-    \item Determine whether the confidence region is empty.
-If the region is empty then we can reject the MLM assumption hypothesis with a significance level of one minus the region coverage.
-Assuming we fail to reject, we can continue:
-
-    \item Sketch out the the confidence region, using examples.
-In most nontrivial cases the region $R$ is high-dimensional and difficult to inspect directly.
-We therefore produce examples from this region to try to help the user understand what values of $q$ are plausible.
-In particular, we sample $Z_{x,y} \sim \mathrm{Exponential}(1)$ for each $x,y$ and solve the problem
-    \[
-    \max_{p,q \in R} \quad \sum_{x,y} Z_{x,y} \log q(y|x) \\
-    \]
-    This problem does not have a closed-form solution, but can generally be solved using an iterative procedure; we describe our method in detail in Appendix \ref{sec:optproblems}.
-The result is a particular value of $q$ which lies within $R$.
-In practice, we find these examples are almost always on the boundary of $R$.
-We can then repeat this process by resampling $Z$.
-By repeating this process many times, we can get a large collection of plausible values of $q$.
-For any fixed value of $x$ we can visualize these samples by using a heatmap to plot $q(\cdot |x)$ for all the samples.
-See Figure \ref{fig:qtildesamps} for an example.
-
-    \item Produce a point estimate.
-It is not generally possible to produce a consistent point estimate.
-Indeed, there are many situations in which the confidence region $R$ does not concentrate even in the limit of infinite data.
-Nonetheless, if a point estimate is desired, one may pick point estimates $\tilde p,\tilde q$ by solving the minimization problem $\min L(p,q)$.
-We caution that this problem does not always have a unique solution, so the result may depend upon the minimization strategy used.
-Precise details for our algorithm may be found in Appendix \ref{sec:optproblems}.
-\end{enumerate}
-
-A Bayesian point of view would suggest a different procedure; this point of view may provide a useful complementary perspective and is worthy of future research.
+A Bayesian point of view would suggest a different procedure, one which might provide a useful complementary perspective.
+This is worthy of future research.
 However, we caution that it is nontrivial to choose reasonable prior distributions for $p^*,q^*$.
-Priors which may seem uninformative (such as the Dirichlet distribution) can actually carry significant and unanticipated weight, due to the high-dimensional nature of the problem.
+Priors which may seem uninformative (such as the Dirichlet distribution) can have significant unanticipated consequences, due to the high-dimensional nature of the problem.
 We give an example in \ref{sec:bayesian}.
 
-The frequentist confidence region allows us to avoid making any modeling assumptions.
-However, we are then faced with the task of understanding a high-dimensional confidence region.
-As described above, our solution is to look at examples from inside the region.
-We emphasize that these examples are not drawn uniformly within the confidence region; instead the examples illustrate something more like the boundary of the region, sketching out what might be possible.
-We concede that this method is somewhat ad-hoc and hope that future research may discover a superior method to visualize this high-dimensional confidence region.
-For now, it appears to be a practical solution.
+The frequentist confidence interval allows us to avoid making any modeling assumptions (beyond the basic Markov Link Method assumption).
+However, it does come with some drawbacks.
+For example, we here take the view that values inside the confidence interval are somehow `plausible' or `likely.'  Such a statement cannot be philosophically justified without some at least a hint of Bayesian reasoning (cf.\ \cite{morey2016fallacy}).
+It is also important to note that, as is typical for confidence intervals, we make no corrections for multiple testing.
+We therefore expect that a small proportion of our confidence intervals will not contain the truth.
+It is therefore important to consider all of the intervals produced and be wary of `cherry picking' only those intervals that support a particular conclusion.
+However, if cherry picking is desired, a false discovery rate correction may be applied (cf.\ \cite{benjamini2005false}).
+A related drawback is that the confidence intervals are not always logically consistent with one another.
+For example, it can happen that that $C_{x,y_1}=[0,1]$ and $C_{x,y_2}=[.2,1.0]$.
+These intervals don't really make sense together, in that if $q(y_2|x)\geq .2$ then certainly $q(y_2|x)\leq .8$.
+A possible remedy would be to construct a simultaneous confidence region for all the parameters concerning a fixed value of $x$.
+Analyzing the statistical properties of these high-dimensional confidence regions would be nontrivial; we leave it for future research.
 
@@
 For example, imagine that the $Y$ error is correlated with the absolute reliability of the part; this may pose no problem if that reliability is adequately measured by $X$.
-
-    \item Combining knowledge across experimental modalities: morphology and transcriptomics.
-There are different ways to think about the different types of cells in an organism.
-A traditional approach is to classify cells based on what they look like (cf.\ \cite{steinman1973identification,bloomfield1982physiological}).
-A more modern approach is to assay the cell's transcriptome (cf.\ \citep{tasic2017shared}).
-Unfortunately, modern high-resolution cell photography and single-cell sequencing technologies are both destructive.
-As a result, we can't always get both kinds of data for the same specimens.
-For cells native to regions full of diverse cell-types, it is thus quite hard to grasp the correspondence between these different kinds of classification systems.
-The result is two completely independent classifications of cells, one for each way of looking at the cell.
-MLM allows us to estimate the relationship between those two classification systems, yielding a holistic understanding of the different types of cells.
-In this case, $\ell$ might indicate some side information such as where in the body the cell was found, $X$ would indicate a detailed classification of the cell according to its transcriptomics, and $Y$ would indicate a coarser classification of a cell according to its morphology.
-We expect that cell morphology is largely a function of cell transcriptomics.
-Thus, as long as the $X$ measurement is sufficiently detailed, we expect that any correlations between $Y$ and $\ell$ would be explained by $X$.
-That is, the MLM assumption holds.
 
@@
 
-\section{Simulation results}
+\section{Simulations}
 
@@
 
-We use simulations to test the capabilities of the Markov Link Method to estimate a calibration.  We will:
+\figgy{
+\hfill{}\figgygrs{.4}{../images/simulationI}{(a) Estimation error, \\ varying number of samples \\ and number of sampling procedures.}\hfill{}
+\hfill{}\figgygrs{.5}{../images/simulationIconf}{(b) Confidence intervals, \\ varying number of samples \\ and number of sampling procedures.}\hfill{}
+
+\hfill{}\figgygrs{.7}{../images/simulationII}{(c) Estimation error, varying number \\ of samples and strength of link dependency.}\hfill{}
+
+\caption{{\bf When can the MLM accurately estimate the link?} \label{fig:simI}  We use simulations to test the MLM.
+In subfigure (a) we look at the error of the MLM link estimator in many different situations, varying the numbers of samples, sampling procedures, and the link parameters.
+In each trial our `ground truth' link parameters picked uniformly at random from the parameter space.
+The MLM estimator accurately recovers the ground truth from simulated data, as long as we have enough samples \emph{and} enough distinct sampling procedures.
+In subfigure (b) we consider the MLM's 95\% confidence intervals.
+For a selection of trials we choose a random value of of $x,y$ and then compare the true link parameter $q^*(y|x)$ with the MLM confidence interval $C_{x,y}$.
+We show the interval (in black) centered around the ground truth (in red).
+Since the link parameters lie in $[0,1]$, the interval is always contained within $\pm1$ of the ground truth (this larger region is shown in gray).
+The interval mostly contain the truth, and get narrower with more samples as long as there are enough distinct sampling procedures.
+In subfigure (c) we focus on the case of exactly $4$ sampling procedures.
+In this case it is only possible to get very accurate estimates if the link takes on a particular form.
+If the link is indepenent (i.e.\ $X,Y$ are independent) or if the link is invertible and deterministic (e.g.\ $X=Y$) then we can determine what the link is.
+However, if the truth is more complex, small numbers of sampling procedures may make it impossible to determine the true link; this issue is discussed in detail in Section \ref{sec:casestudies}.}
+}
+
+We used simulations to test the two parts of the MLM: the point estimators $\tilde q(y|x)$ and confidence intervals $C_{x,y}$.
+For any fixed value of the ground truth parameters, $(p^*,q^*)$, we would ordinarily hope for two things:
+\begin{description}
+    \item[Estimator convergence] As we obtain more samples, $\tilde q(y|x)$ should converge to $q^*(y|x)$ for each $x,y$.
+    \item[Conservative coverage] $\mathbb{P}(q^*(y|x) \in C_{x,y})\geq95\%$ for each $x,y$.
+That is, for each choice of $x,y$ the confidence interval $C_{x,y}$ should contain $q^*(y|x)$ with probability at least 95\%.\footnote{Note that we are not producing a `simultaneous' interval.
+In a simultaneous interval, there would be a 95\% chance that $q^*(y|x)\in C_{x,y}$ for \emph{every} $x,y$.
+The key difference is the change from `each' (in our intervals) to `every' (in simultaneous intervals).
+Thus, it may be that $q^*(y|x)\notin C_{x,y}$ for a few values of $x,y$.
+For \emph{most} values of $x,y$ the true parameters should be contained inside of their corresponding confidence intervals.}
+\end{description}
+Our simulations yielded good news and bad news.
+The good news: the MLM generally achieves conservative coverage.
+In many cases, the MLM also achieves estimator convergence.
+The bad news: in many cases the MLM fails to achieve estimator convergence.
+Indeed, we can show that an `identifiability issue' blocks any possibility of success in some cases; we detail this problem in Appendix \ref{sec:casestudies}.
+The following two simulations give a good overview of the relevant issues:
 
 \begin{itemize}
-    \item Pick parameters to be `ground truth': 
-    \begin{itemize}
-        \item A calibration, $q^*(y|x)$.  
-        \item A tool-I measurement distribution, $p^*(x|\ell)$.  Recall that this indicates the probability that tool I measures $x$ for a specimen gathered with sampling strategy $\ell$. 
-        \item A tool-II measurement distribution, $h^*(y|\ell)$.
-This indicates the probability that tool II measures $y$ for a specimen gathered with sampling strategy $\ell$.
-We assume the the MLM assumption holds, and so this can be calculated from $q^*,p^*$ by the formula $h^*(y|\ell)=\sum_x p^*(x|\ell)q(y|x)$.
-    \end{itemize}
-    \item Select a number of samples to take per subpopulation, $s$, and produce a simulated dataset for each subpopulation $\ell$:
-    \begin{align*}
-    (N^X_{\ell 1},N^X_{\ell 2}\cdots N^X_{\ell \Omega_X}) \sim \mathrm{Multinomial}(s,p^*(\cdot | \ell))\\
-    (N^Y_{\ell 1},N^Y_{\ell 2}\cdots N^Y_{\ell \Omega_Y}) \sim \mathrm{Multinomial}(s,h^*(\cdot | \ell))
-    \end{align*}
-    \item Apply the Markov Link Method to the generated samples to obtain a calibration estimate $\tilde q$.
-    \item Compute the error of the estimate $\tilde q$, as measured by a kind of averaged total variation distance, namely
+    \item How does MLM performance depend on the number of samples and the number of sampling procedures?
+We suppose we have two measurement procedures which return categorical measurements among six categories.
+Thus the first procedure yields $X\in \{1,2,3,4,5,6\}$ and the second yields $Y\in \{1,2,3,4,5,6\}$.
+To see how the method performs in different circumstances, we will run many trials.
+In each trial we fix the number of sampling procedures and pick `ground truth' parameters uniformly at random from the parameter space.
+We then simulate a dataset from these ground truth parameters, fixing the total number of samples (spreading these samples equally among all combinations of sampling procedure and measurement procedure).
+Finally, we apply the MLM to the simulated data to get the point estimates $\tilde q(y|x)$ and the confidence intervals $C_{x y}$.
+We measure the overall estimator convergence using a kind of total variation distance:
     \[
-    \frac{1}{2\Omega_{\ell}} \sum_{\ell x} |\tilde q(y|x) - q^*(y|x)|
+    \mathrm{Error}(\tilde q,q^*)=\frac{1}{12}\sum_{x=1}^{6}\sum_{y=1}^{6} |\tilde q(y|x) - q^*(y|x)|
     \]
-    This distance ranges between zero and one.
-A distance of zero indicates that $\tilde q=q^*$, and a distance of one indicates that the two calibrations place mass on completely disjoint sets, i.e.
-$q(y|x)=0$ whenever $\tilde q(y|x)>0$ and vice-versa.
-\end{itemize}
+    This error ranges between zero and one.
+Zero error indicates that $\tilde q=q^*$.
+An error of one indicates that the estimate has completely incorrect beliefs about the probability mass, i.e.
+$q^*(y|x)=0$ whenever $\tilde q(y|x)>0$ and vice-versa.
+We also look at the MLM confidence intervals and see how often they cover the true parameters.
 
-\figgy{
-\hfill{}\figgygrs{.4}{../images/simulationIandIIpq.png}{(a) `Ground truth' for $p^*,q^*$.}\hfill{}
-\hfill{}\figgygrs{.5}{../images/simulationsIandII.png}{(b) MLM performance \\ in estimating $q^*$.}\hfill{}
+    The results are summarized in Figure \ref{fig:simI}.
+In trials with more samples, the estimator generally has lower error.
+However, to make the error actually converge to zero we need at least six distinct sampling procedures.
+This figure also shows that the confidence intervals work correctly regardless of of the number of samples or sampling procedures; they include the ground truth with high probability.
+A more precise analysis of these intervals suggests that they achieve conservative coverage (see Appendix \ref{sec:confinterval} for details).
+With many samples and sampling procedures, these intervals are small and concentrated around the truth.
+With fewer samples or sampling procedures, the intervals are sometimes forced to be larger.
+However, our uncertainty is not the same for every parameter.
+In some cases we are able to obtain a very tight confidence interval for one particular parameter even though the overall estimator error is high.
 
-\caption{{\bf The Markov Link Method accurately estimates the calibration, as long as there are enough subpopulations}.
-\label{fig:simsIandII}  We pick parameters as `ground truth' for $p^*,q^*$; these are as shown in subfigure (a) as heatmaps.
-We then create simulated datasets using these parameter values, use the Markov Link Method to produce its estimator, and see and measure the total variation error of the estimator.
-This error ranges from zero to one, where zero indicates no error and one indicates complete disagreement.
-In subfigure (b) we see that the MLM is able to estimate the ground truth with negligible error, as long as it has enough samples.
-However, we also consider another simulation: we merge some of the subpopulations in the ground truth parameters so that they become indistingiushable.
-In this degraded setup, the MLM is unable to learn the true calibration no matter how many samples we have.
-}
-}
-
-For the first simulation, we will look at a case in which the MLM performs well.
-We will assume there are twenty different samping strategies, and tools I and II both measure specimens as being in one of six categories.
-Our particular choices of ground truth can be found in Figure \ref{fig:simsIandII}.
-This figure also shows that the MLM is able to accurately estimate the ground truth calibration $q^*$, as long as there are enough samples.
-
-We also used this simulation to test the coverage of the confidence region.
-Recall that the `actual coverage' of a confidence interval indicates the probability that a confidence region contains the ground truth parameters.
-This coverage depends upon the particular choice of ground truth.
-The `nominal' coverage indicates the intended coverage of the interval, and may be more or less than the actual coverage.
-We estimated the actual coverage for the ground truth we have chosen.
-Taking 200 sample datasets with $s=80$ samples per subpopulation, we found that an MLM confidence region with 95\% nominal coverage has an actual coverage of about $99\%$.
-In this sense our region is `conservative,' which was somewhat by design.
-We refer the reader to Appendix \ref{sec:confinterval} for details.
-We also investigated other values of $s$, finding that the model became less conservative with more samples.
-For example, for datasets $s=800$ with the actual coverage appears to be approximately $97.5\%$.
-
-In our second simulation, we consider a case where the MLM does not succeed in accurately estimating the calibration.
-We use the same ground truth parameters as in first simulation, but we will merge groups of sampling strategies.
-In particular, we will divide the sampling strategies into four groups, each with five members.
-We will pretend that we cannot distinguish subpopulations within any given group.
-In this case, Figure \ref{fig:simsIandII} shows that the Markov Link Method cannot accurately estimate the calibration, no matter how many samples we have.
-This is due to a fundamental identifiability issue which is discussed in detail in Appendix \ref{sec:casestudies}.
-
-Despite the fact that we could not get an accurate point estimate for the calibration in the second simulation, the confidence regions appear to remain on solid ground.
-We took 200 simulated datasets with $s=400$ samples per subpopulation (1600 samples total) and computed MLM confidence regions with $95\%$ nominal coverage.
-We found an actual coverage of about $96.5\%$.
-
-\figgy{
-\hfill{}\figgygr{.7}{../images/simulationIII}\hfill{}
-
-\caption{{\bf The error of the MLM estimator depends on the nature of $q^*$.}  In our third simulation we fix a particular $p^*$ and consider a range of $q^*$s.
-For each possible $q^*$ we generate a certain number of samples per subpopulation and then try to use the resulting dataset to estimate $q^*$.
-On the one hand, we consider the case that $q^*$ causes $X,Y$ to be independent; in this case we can get a good estimates of $q$ with enough data.
-On the other hand, we consider the case that $q$ causes $X,Y$ to be deterministically bound, i.e.
-$q^*(y|x)=1$ if $x=y$ and $q(y|x)=0$ otherwise.
-In this case it is also possible to estimate $q$ with enough data.
-In-between, it is not possible to exactly determine $q^*$, even in the asymptotic limit of infinite data.\label{fig:estimatingq}}
-}
-
-For our final simulation, we will dig deeper into the question of when the calibration can be perfectly estimated with enough data.
-To this end, we will consider cases in which the number of subpopulations is four, and there are six possible outcomes for both measurement tools.
-For many choices of $p^*,q^*$, this will make it \emph{impossible} to correctly determine the calibration $q^*$ without joint measurement.
-However, for certain values of $q^*$ the truth can be recovered.
-To see this, we will look at a variety of calibrations.
-On the one hand, we will consider the case that $X,Y$ are independent, i.e.
+    \item How does estimator convergence depends upon the link parameters themselves?
+In each trial of this simulation we use four sampling procedures and our measurement procedures yield one of six categories.
+In the previous simulations we saw that estimator convergence was generally impossible in this situation, due to the small number of sampling procedures.
+However, in our previous simulations we picked the link parameters uniformly from parameter space.
+Now we will be more choosey.
+On one extreme, we will produce trials where the link makes $X,Y$ independent, i.e.
 $q^*(y|x)=q^*(y|x')$ for every $x,x'$.
-On the other hand, we will consider the case that $X,Y$ are deterministically related, in particular $X=Y$.
-We will also consider every calibration ``in-between'' these two extremes.
-These in-between calibrations are found by linear interpolation.
-For each calibration we will generate datasets of various sizes and try to estimate the calibration from the datasets using the MLM.
-In Figure \ref{fig:estimatingq}, we see that in the two extreme cases it may be possible to recover the calibration.
-However, for in-between cases it is indeed not possible.
-This is true even in the limit of infinite data.
+On the other extreme, we will have trials where $X,Y$ are deterministically related by the equation $X=Y$.
+We will also consider every link ``in-between'' these two extremes (found by convex combinations).
+Figure \ref{fig:simI} shows that estimator convergence is actually possible in the two extreme cases.
+However, estimator convergence fails for the in-between cases.
 We leave a complete mathematical understanding of this result to future work.
-For the purposes of this paper, we content ourselves by noting a special case of particular interest: that $X$ and $Y$ can take on one of $2^k$ values and are related to each other by an invertible deterministic function.
-In this case, with enough enough samples of only $k+1$ different sampling strategies, we can determine both that the relationship is deterministic and the exact specification of the invertible function.
+For the purposes of this paper, we content ourselves by noting one interesting case: let the measurements return one of $2^k$ categories and let us use only $k+1$ carefully chosen sampling procedures.
+Now suppose the link defines any invertible deterministic function between $X$ and $Y$.
+In this case, with enough enough samples, we can determine both that the relationship is deterministic and the exact specification of the invertible function.
 This result is proven in Appendix \ref{sec:casestudies}, Theorem \ref{thm:miracle}.
@@
 
-We now apply the MLM to real data.
-
 \figgy{
 \hfill{}\figgygr{.95}{../images/allenNlxy}\hfill{}
-\caption{{\bf The input to the Markov Link Method: data about each tool with many sampling strategies but without any joint measurement}.
-Here we show a portion of the real-world data to which we applied the MLM.
-The Allen Institute gathered neurons from the visual cortex of mice, using a variety of cre/lox-based sampling strategies (cf.\ \cite{tasic2017shared}).
+\caption{{\bf The input to the Markov Link Method: one experiment for each measurement procedure and each sampling procedure.}  Here we show a portion of the real-world data to which we applied the MLM.
+The Allen Institute gathered neurons from the visual cortex of mice, using a variety of cre/lox-based sampling procedure (cf.\ \cite{tasic2017shared}).
 Each strategy was designed to bring out different subpopulations of cells.
-For each sampling strategy, one classification protocol (tool I) was used to estimate the some of the neurons' cell-types, and another protocol (tool II) was used to estimate the other neurons' cell-types.
-Tool I recognizes 104 different types of cells.
-Tool II has a coarser notion of cell-type, distinguishing only 10 types.
-For each sampling strategy and for each tool, we tabulated the number of cells assigned to each type.
-Above we show a subset of these results through heatmaps; the color of each square indicates the number of specimens sampled with a particular technique that were found to have a particular type according to a particular tool.
+Neurons were measured to determine their `type,' using one of two procedures: Standard or Patch.
+Standard recognizes 104 different types of neurons.
+Patch has a coarser notion of cell-type, distinguishing only 10 types.
+For each experiment, we tabulated the number of cells assigned to each type.
+Above we show a subset of these results through heatmaps; the color of each square indicates the number of specimens found to have a particular type.
 Using this kind of data, our task is to calibrate the two classification protocols.
@@
 
-Our motivation for this problem arose from looking at Allen Institute cell-type assignment of cells, performed using two different experimental tools.
-Each tool takes a neuron and determines what type of neuron it is.
-However, both tools destroy the cell in the process of measuring it.
-The first tool uses a standard single-cell RNA sequencing pipeline to determine transcriptomic expression in the cell, and determine the cell-type based on these results.
-The second tool also uses transcriptomic information, but additionally obtains electrophysiological and morphological properties of the neuron.
+Our motivation for this problem arose from the Allen Institute's investigation of neuronal cell-types.
+The institute has a variety of methods for examining a neuron and determining what kind of neuron it is.
+However, many of these procedures destroy the neurons in the process of measuring them.
+We'll look at two procedures in particular, which we'll call `Standard' and 'Patch.'  Standard uses a single-cell RNA sequencing pipeline to determine transcriptomic expression in the cell, and determine the cell-type based on these results.
+Patch also uses transcriptomic information from RNA sequencing, but Patch additionally obtains electrophysiological and morphological properties of the neuron.
 This comes at the cost of a degraded transcriptomic signal, requiring new methods to estimate the cell-type.
@@
 Best efforts were made to use biological intuition to calibrate the methods.
-For example, the tool II protocol has a notion of a ``Lamp5'' cell type.
-The tool I method refines this idea into many sub-types, such as ``Lamp5 Pdlim5'' and ``Lamp5 Slc35d3.''  If a cell was designated as ``Lamp5 Pdlim5'' under the tool I, the hope was that it be given the ``Lamp5'' type by tool II.
+For example, Patch has a notion of a ``Lamp5'' cell type.
+Standard gives a more granular analysis, dividing this type into many sub-types, such as ``Lamp5 Pdlim5'' and ``Lamp5 Slc35d3.''  If a cell was designated as ``Lamp5 Pdlim5'' using Standard, the hope was that it be given the ``Lamp5'' type by Patch.
 The two methods were designed to achieve this goal.
@@
 
-Fortunately, there was a kind of information that seemed like it might help determine whether the two methods were properly calibrated: a variety of sampling strategies.
-Using a cre/lox system (cf.\ \citep{tasic2017shared}) they were able to pick out specific, overlapping subpopulations of neurons.
-Each subpopulation was expected to contain different proportions of the different cell-types.
-For each subpopulation and each method, many specimens were sampled and their cell-types determined.
+Fortunately, there was a kind of information that seemed like it might help determine whether the two methods were properly calibrated: a variety of sampling procedures.
+Using a cre/lox system (cf.\ \citep{tasic2017shared}) they were able to pick out various overlapping subpopulations of neurons using different procedures.
+Each procedure was expected to yield different proportions of the different cell-types.
+For each sampling procedure and eaach measurement procedure, many specimens were sampled and their cell-types determined.
 The result of this process was two tables, parts of which are shown in Figure \ref{fig:alleninput}.
@@
 
-As described in Section \ref{sec:mlm}, to apply the MLM we first produced an approximate 95\% confidence region for $p^*,q^*$.
-We found that this confidence region was non-empty, signifying that we could not reject the hypothesis that the MLM assumption holds.
-We then used the Markov Link Method of producing a variety of example calibrations inside the confidence region.
-We then visualized these examples one cell-type at a time.
-That is, we first select a tool-I cell-type $x$ (such as $x=$`Lamp5 Egln3 1').
-Then for any calibration $q$, we can consider the vector $q_x$ formed by looking at $q(y|x)$ for every value of $y$.
-We can then visualize many calibrations by plotting many of stacking these vectors together into a heatmap.
-The result for various choices of $X$ may be found in Figure \ref{fig:allenboot}.
-
 \figgy{
-\hfill{}\figgygr{0.9}{../images/vviz}\hfill{}
-\caption{{\bf Estimating the calibration and measuring our uncertainty on real data.}  Here we present visualizations of the confidence region found by the Markov Link Method applied to Allen Institute data.
-The colors indicates the value of $q(y|x)$ for a particular cell-types $x,y$ and a particular calibration $q$.
-All of the calibrations are examples from a 95\% confidence interval.
-For each individual plot, $x$ is fixed, each column indicates a particular value of $y$, and each row indicates a different plausible $q$.
-Each plot helps us answer the question: how might specimens labelled with type $x$ by tool I be labelled by tool II?
-For example, consider the `Lamp5 Pdlim5' plot.
-For every calibration (i.e.\ every row of the plot), we see a bright yellow in the `Lamp5' column.
-This gives some confidence that specimens assigned type `Lamp5 Pdlim5' by tool I will be assigned type `Lamp5' by tool II.
-By contrast, consider the `Lamp5 Egln3 1' plot.
-This plot shows we have very little knowledge about the calibration for cells identified as this type by tool I.
-Each calibration in the confidence region gives a completely different picture.
-The data we have simply isn't sufficient to help us know this aspect of the calibration.
+\hfill{}\figgygr{0.9}{../images/allenout}\hfill{}
+\caption{{\bf MLM estimation of the link, with confidence interval.}  The central plot shows a portion of the MLM estimator applied to the Allen Institute data.
+We look at a subselection of Standard types and all of the Patch types.
+For each combination $x,y$ we draw a rectangle whose color indicates the value of the estimator $\tilde q(y|x)$.
+We also show our confidence about these estimators.
+On the left we indicate the lower bounds indicated by the confidence interval, on the right we indicate upper bounds.
+For some parameters our intervals are much tighter than others.
+For example, it appears we have high confidence the Standard type `Vip Rspo4' is highly associated with the Patch type `Vip.'  By constrast, we have almost no idea what is associated with the Standard type `Meis2.'  The upper confidence interval bounds suggest that $q^*(y|\mathrm{Meis2})$ could be nearly 1 for $y$ equal to `Lamp5,' 'Pvalb,' 'Sncg,' 'Sst,' or `Visp' Patch types.
+Obviously it cannot be 1 for all of those types simultaneously, since $\sum_y q^*(y|\mathrm{Meis2})=1$, but the data simply doesn't tell us which $y$ carries the mass.
 \label{fig:allenboot}}
@@
 
-Inspecting these visualizations we can see different aspects of our knowledge about how well-calibrated the two tools are:
-\begin{itemize}
-    \item For some tool-I types, such as `Lamp5 Pdlim5,' the two methods appear very well-calibrated.
-All the examples from the confidence region suggest that cells measured with type `Lamp5 Pdlim5' are likely to be assigned type `Lamp5' by tool II.
-    \item For some types, such as `Lamp5 Egln 1,' we simply have no idea.  There is tremendous variability within the confidence region; each example looks different.  
-    \item For some types, such as `Vip Crispld2 2,' it appears that cells of this type are difficult for tool II to understand.
-In particular, for most example calibration they have a reasonably high probability of being assigned the  `unknown' type.
-However, we cannot be sure of the the exact depth of this issue, because in some plausible example calibrations it appears that these cells are in fact correctly labelled as being of the `Vip' type fairly often.
-More data is necessary.
-\end{itemize}
+We applied the MLM to this data to obtain point estimates $\tilde q(y|x)$ and confidence intervals $C_{x y}$.
+In Figure \ref{fig:allenboot} we visualize these objects for selected values of $x,y$.
+Each parameter has a slightly different story.
+For example, for the Standard type $x=$`Vip Rspo4' and the Patch type $y=$`Vip,' we have that $C_{x,y}=[.92,1.0]$.
+This supports the idea that the true link satisfies $q^*(y|x)\geq .92$: at least 92\% of the cells classified as type `Vip Rspo4' by the Standard method will be classified as `Vip' by the Patch method.
+However, for other types there is more ambiguity.
+For the Standard type $x=$`Lamp5 Egln 1' and the Patch types $y=$`Lamp' we have $C_{x y}=[.42,1.0]$.
+This is a much wider confidence interval.
+It does say that at least 42\% of the cells with Standard type `Lamp5 Egln1 1' are given the `Lamp5' Patch type.
+But what about the other 58\%?
+It might be that they're also assigned the `Lamp5' Patch type, but the data doesn't tell us for sure.
+Looking at the confidence intervals for other Patch types, we conclude that `Lamp5 Egln 1' types are not ususally classified with Patch types `CR,' 'L6b,' `Sst,' `VISp,' `Vip,' or `Unknown': the upper limits on $q^*(y'|x)$ are small for each of these Patch types.
+The intervals show that many of these `Lamp5 Egln 1' cells \emph{might} be assigned the Patch types 'Pvalb' or 'Sncg.'  The data we have does not give us a definitive answer.
+
 The variability in the confidence regions suggests what we need to do in order to more closely determine the value of the calibration.
-For example, if we could develop more unique sampling strategies which will include `Lamp5 Egln 1' cells, this would help us resolve our ambiguity about this aspect of the calibration.
+For example, if we could develop more unique sampling strategies which will include `Lamp5 Egln 1' cells, this would help us resolve our ambiguity about this aspect of the link.
 Indeed, going back to the original data, it is easy to see why this ambiguity appeared in the first place.
-Cells of the `Lamp5 Egln 1' type only appear in any number when using the sampling strategies `Gad2-IRES-Cre' and 'Slc32a1-IRES-Cre.'  Both of those sampling strategies yield a fairly similar mix of types when measured with tool II.
-To get a better resolution of the calibration, we would need a sampling strategy that included `Lamp5 Egln 1' cells but represents a significantly different slice of the overal population.
+Cells of the `Lamp5 Egln 1' type only appear in any number when using the sampling strategies `Gad2-IRES-Cre' and 'Slc32a1-IRES-Cre.'  Both of those sampling strategies yield a fairly similar mix of types when measured with Patch.
+To get a better resolution of the calibration, we would need a sampling strategy that included `Lamp5 Egln 1' cells but represents a significantly different slice of the overall population.
 For a particular proposed experiment, simulations such as those found in Section \ref{sec:simulations} can be used to determine how many samples might be required to get an accurate estimate of the calibration.
@@
 
-In this paper we sought to achieve calibration without joint measurement.
-The main technical obstacle is what is known in the statistics world as an `identifiability problem.'  Due to this problem, we often simply \emph{cannot} directly estimate the calibration that we are interested in.
-The calibration is simply not ``identifiable.''  Our main contribution is to produce a an approximate confidence region for the parameter of interest.
+The task of this paper was infer the link between different measurement procedures, so that experimental data can be meaningfully combined.
+There is an enormous literature on this subject.
+For example, when experiments are performed in batches, the exact measurement procedures can vary slightly between batches.
+The entire field of `batch effects' is devoted to handling these problems.
+The general approach is to use some knowledge of the procedures to make modeling assumptions about the links.
+These assumptions give us a way to estimate the link (cf.\ \cite{johnson2007adjusting}).
+If different measurement procedures yield results in the same space, we can also implicitly articulate these kinds of assumptions by placing a metric on the space.
+We suppose that different measurement procedures should yield results that are `close' in this metric.
+We can then use optimal transport techniques to produce a link based on these assumptions (cf.\ \cite{tabak2018explanation}).
+From the most general point of view, we are engaged in meta-analysis; we refer the reader to \cite{borenstein2011introduction} for a general introduction to the field.
+In this broader context, the main distinguishing characteristics of this paper are two-fold: we place no assumptions on the nature of the link and we take the identifiability issues seriously.\footnote{We refer the reader to \cite{walter2014identifiability} for an introduction to what we mean by identifiability.
+We discuss the identifiability issues for our problem in particular in Appendix \ref{sec:casetstudies}.}  Our only assumption is that the sampling procedures can be statistically isolated from the link.
+This assumption says nothing about what the link itself is.
+We then use profile likelihood confidence intervals to deal with any identifiability issues.
+These intervals account for uncertainty due both to small sample sizes and to identifiability issues.
 
-The necessity to deal with such identifiability issues has become of increasing concern as nature of model parameters and data has grown more intricate.
-There are many examples in the literature.
-How much can we hope to learn about the parameters of complex biological reaction networks, if we can only observe part of the process (cf.\ \cite{raue2009structural})?
-Given what we can observe in practice, how much can we ever hope to learn about how cancers divide and spread through the body (cf.
-\cite{brouwer2017parameter,luebeck2002multistage})?
-Do the tests we have enable us to actually learn the parameters of our models for how materials bend and break (cf.
-\cite{hartmann2018identifiability})?
-At their heart, these are all questions of estimation the face of potential unidentifiability.
-
-This work is inspired by a large literature of examples where assumptions are used to bound potentially unidentifiable parameters.
+The main technical contribution of this paper was figuring out how to use the MLM assumption to get confidence intervals that worked in practice.
+In this we were inspired by a large literature of examples where assumptions are used to bound potentially unidentifiable parameters.
 Some of this literature comes from the field of causality.
@@
 
-When joint measurement is unavailable, it can be difficult to calibrate two measurement tools.
+It can be difficult to understand how two different measurement procedures are related to each other.
+This makes it difficult to understand how to combine knowledge from experiments which use different procedures.
+The problem is particularly tricky if we cannot look at the same specimen through the perpsective of both procedures simultaneously.
 
-In this paper we see that a conditional independence assumption can help overcome this difficulty.
-We proposed a procedure, the Markov Link Method (MLM), to use this assumption to estimate the calibration and measure uncertainty.
+In this paper we suggest one way to overcome these difficulties.
+We formalize the `relationship' between two measurement procedures through the notion of a `measurement link.'  Given that we have measured a specimen with one procedure and obtained the outcome $x$, the link parameter $q(y|x)$ tells us the probability that we would obtain outcome $y$ if we measured the same specimen with the second procedure.
+We present the Markov Link Method assumption, which roughly states that the link can be statistically isolated from the sampling procedures.
+We provide the Markov Link Method (MLM) algorithm that uses this assumption to estimate the link and measure our uncertainty.
 Code is published at https://github.com/jacksonloper/markov-link-method, including a tutorial-style ipython notebook detailing every computation made in this paper.
 
-To understand our uncertainty, we use to a confidence region; this comes with all the usual attendent caveats.
-For example, we here take the view that calibrations inside the confidence region are somehow `plausible' or `likely.'  As has been pointed out numerous time, such a statement cannot be philosophically justified without some at least a hint of Bayesian reasoning (cf.\ \cite{morey2016fallacy}).
-We leave a more rigorous Bayesian treatment of this problem for future research.
-The key problem will be to find a suitable prior; this search comes with certain dangers which we are not sure how to circumvent.
-We give an example of these dangers in Appendix \ref{sec:bayesian}.
-
-Theory aside, the MLM appears to provide reasonable estimates in simulation and give useful insight to real data.
+The MLM appears to provide reasonable values in simulation and gives useful insight to real data.
 Applied to neuronal cell-type data, the MLM clarified how two different cell-type classification systems are related.
 We saw that some aspects of the two systems seem well-calibrated, but others we are less sure about.
-The nature of the variability in the confidence region suggested directions for experiments which would further refine our understanding.
+The nature of the variability in the confidence intervals suggested directions for experiments which would further refine our understanding.
 
-In future work, it may be interesting to consider additional assumptions that might help us estimate calibrations.
-For example, as it stands the Markov Link Method only works if each tool can only return one of a relatively small number of measurement values.
+In future work, it would be interesting to use additional assumptions to help us estimate calibrations.
+For example, as it stands the Markov Link Method will only yield narrow confidence intervals if each measurement tool returns one of a finite number of measurement outcomes.
 In some cases, we may believe that similar measurement values should have similar probabilities.
 Such smoothness assumptions would make it possible to apply the MLM to measurement tools which can return many different values.
+The frequentist point of view allows us to phrase these kinds of assumptions as hard constraints that we believe in, instead of guesses about what might be more or less likely.
+In some cases these assumptions are easier to think about than prior distributions, as we see in Appendix \ref{sec:bayesian}.
 
 More generally, it is clear that good assumptions can help us get real insight for tough problems.
-Even if these assumptions are not sufficient to allow us to perfectly identify our object of interest, confidence regions can often be constructed.
+Even if these assumptions are not sufficient to allow us to perfectly identify our object of interest, confidence regions can be constructed.
 By probing these regions carefully, we can learn what the data actually has to say and what experiments will help us learn more.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                   __ _     _                     
+%   ___ ___  _ __  / _(_) __| | ___ _ __   ___ ___ 
+%  / __/ _ \| '_ \| |_| |/ _` |/ _ \ '_ \ / __/ _ \
+% | (_| (_) | | | |  _| | (_| |  __/ | | | (_|  __/
+%  \___\___/|_| |_|_| |_|\__,_|\___|_| |_|\___\___|
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Confidence interval details}
 
-Recall that we have an $\Omega_\ell\times\Omega_X$ matrix $N^X$ and an $\Omega_\ell \times \Omega_Y$ matrix $N^Y$, sampled according to
+\label{sec:confinterval}
+
+In the main text of this paper, Section \ref{sec:mlm}, we outline a method for producing confidence intervals for the link parameters $q^*(\tilde y|\tilde x)$ for a specific choice of $\tilde x,\tilde y$.
+Here we give full details on this method and discuss its properties in depth.
+
+Consider experiments yielding a $\Omega_\ell\times\Omega_X$ matrix $N^X$ and an $\Omega_\ell \times \Omega_Y$ matrix $N^Y$, sampled according to
 \begin{align*}
@@
 \end{align*}
-where $p^*(x|\ell)$ is some conditional distribution and there is some $q^*(y|x)$ such that $h^*(y|\ell)=\sum_x p^*(x|\ell)q^*(y|x)$.  We have then defined
+where $p^*(x|\ell)$ is some conditional distribution and there is some $q^*(y|x)$ such that $h^*(y|\ell)=\sum_x p^*(x|\ell)q^*(y|x)$.  We have then defined the log likelihood as 
 \[
-L(p,q)=\frac{1}{\Omega_X}\sum_{\ell,x} N^X_{\ell x} \log \frac{N^X_{\ell x}}{n_\ell p(x|\ell)} + \frac{1}{\Omega_Y}\sum_{\ell,y} N^Y_{\ell y} \log  \frac{N^Y_{\ell y}}{m_\ell \sum_x p(x|\ell) q(y|x)}
+L(N;p,q)=\left(\sum_{\ell,x} N^X_{\ell x} \log p(x|\ell) + \sum_{\ell,y} N^Y_{\ell y} \log \left(\sum_x p(x|\ell) q(y|x)\right)\right)
 \]
-We then constructed the confidence region
+In the main text we supressed the dependence of $L$ the data $N=(N^X,N^Y)$, but it will be important here so we will emphasize it.  We define the profile likelihood as
 \[
-R = \{p,q:\ L(p,q) \leq k\}
+L(N;\xi)=\max_{\substack{p,q \\ q(\tilde y|\tilde x)=\xi}} L(p,q)
 \]
-This is sometimes called so-called likelihood ratio confidence interval, and its properties are fairly well-understood.
-We refer the reader to \cite{brazzale2007applied} for a useful exposition.
-The essential point is that in the asymptotic limit the distribution of $L(p^*,q^*)$ depends only upon the number of zero probability values in the parameters $p^*,h^*$.
-For any number of zero probability values, the asymptotic distributions are known exactly; they are scaled sums of $\chi^2$ distributions, one for each subpopulation and each tool.
-The difficulty is of course that we do not know how many probability value are zero.
-This can can be circumvented if one is only interested in a conservative confidence region, i.e.\ one whose actual coverage is at least as great as its nominal coverage.
-In particular, if we assume that \emph{no} values are zero we can a confidence interval which is guaranteed to be conservative; this is an immediate result of the fact that $\chi^2$ distributions with more degrees of freedom stochastically dominate those with fewer.
-Unfortunately, using these $\chi^2$ distributions can be quite inadvisble in low-sample regimes.
-For instance, one may readily see that if some values of $p^*,h^*$ are small (but nonzero), then the distribution of $L(p^*,q^*)$ will be much more similar to sums of $\chi^2$ distributions with fewer degrees of freedom -- at least until the number of samples grows very very large.
+We take $L^*=\max_{p,q} L(N;p,q)$ and define the sets
+\[
+R(N,k) = \left\{\xi:\ L^* - L(\xi) < k\right\}
+\]
+Our task is to find $k$ such that $q^*(\tilde y|\tilde x)$ is contained in $R(N,k)$ with probability at least 95\% if $N$ is sampled according to $p^*,h^*$.
+If we can do this, then $R(N,k)$ would constitute what is called a (conservative) 95\% confidence region for the parameter $q^*(\tilde y|\tilde x)$.
+We refer the reader to \cite{brazzale2007applied} for a helpful introduction to these kinds of confidence regions and some of the difficulties that can arise in using them.
 
-We find it practical and accurate to choose $k$ by using the data itself.
-In particular, we choose a pseudocount $c$ and take $\bar p(y|x)=N^X_{\ell x} +c /n_\ell +c$ and $\bar h(y|\ell)=N^Y_{\ell y} +c / m_\ell +c$ to give estimates of $p^*,h^*$.
-For a nominal coverage of $1-\alpha$, we then choose $k$ so that
+The difficulty is that $p^*,q^*$ is unknown.
+So for any particular choice of $k$, we cannot determine the probability that $q^*(\tilde y|\tilde x)$ is contained in $R(N,k)$.
+In other situations this difficulty can sometimes be overcome by applying asymptotic normality results to argue that the the relevant quantities are actually invariant to the ground truth parameters.
+For example, in many situations the distribution of the relevant quantities can be asymptotically approximated using properties of $\chi^2$ distributions.
+In our case these normality results are unlikely to be useful.
+There are two difficulties.
+First, the ground truth parameters may lie against the edges of the viable parameter space (e.g.
+if $q^*(\tilde y|\tilde x)=0$); this invalidates many of the standard arguments for asymptotic normality.
+Second, the parameters $q^*(\tilde y|\tilde x)$ are not always identifiable; this can also create some strange artifacts.
+
+Our approach is to use the data, $N$, to guide our choice of $k$.
+We define $k=k(N)$ in the following way.
+Our first step is to find a collection of plausible possibilities for the values of $p^*,h^*$.
+To do so, we define the following distribution for $P,H$.
+%
+\begin{itemize}
+    \item Let $P,\hat H$ denote a sample from the posterior of $p^*,h^*$ under a uniform prior.
+That is, consider the distribution of $p^*,h^*,N$ defined by sampling $p^*,h^*$ uniformly from parameter space and then sampling $N$ from $p^*,h^*$.
+The objects $P,\hat H$ are samples from this distribution when conditioned on the observed values of $N^X,N^Y$.
+    \item Let $H$ denote a projection of $\hat H$ to the set of parameters which are consistent with the Markov Link Method assumption.  In particular, 
     \[
-\mathbb{P}\left(\frac{1}{\Omega_X}\sum_{\ell,x} N^X_{\ell x} \log \frac{N^X_{\ell x}}{n_\ell \bar p(x|\ell)} + \frac{1}{\Omega_Y}\sum_{\ell,y} N^Y_{\ell y} \log  \frac{N^Y_{\ell y}}{m_\ell \bar h(y|\ell)}<k\right)=\alpha
+    Q \in \argmax_{p,q} \sum_{\ell y} \hat H(x_\ell) \log \sum_x p(x|\ell) q(y|x)
     \] 
-when $N^X,N^Y$ are drawn from $\bar p,\bar h$ respectively.
-Asymptotically $\bar p\rightarrow p^*$ and $\bar h\rightarrow h^*$, so it seems reasonable to hope that this will give the correct coverage.
-Simulations appear to bear this out, though a further mathematical inquiry is certainly warranted.
+    and $H(y|\ell)=\sum_x P(x|\ell) Q(y|x)$.  Notice that maximization problem for $Q$ may have no unique solution, but the problem is convex and the value of $H$ is uniquely defined.
+\end{itemize}
+%
+We now define our collection of plausible possibilities as those which are `credible' under this distribution.
+In particular, let $l$ denote the fifth percentile of the variable $L(N;P,H)$, holding $N$ fixed at its observed value.
+That is $\mathbb{P}(L(N;P,H)<l|N)<5\%$.
+We take all $P,H$ The set of $P,H$
+Finally, we define $k(N)$ as the smallest value of $k$ such that that $V(k)\geq 95\%$ for every sample of $P,H$ such that $L(N;P,H)$ is above the fifth percentile  credible interval.
+Note that the computation of $k(N)$ involves optimization problems and difficult sums.
+Many of these problems do not have closed form solutions.
+We discuss the exact details of our numerical strategies in Appendix \ref{sec:optproblems}.
 
-The pseudocount bears some discussion.
-We introduce this pseudocount in order to err on the side of a slightly conservative confidence region.
-As we said before, when the parameters $p^*,h^*$ have many small values it is generally possible to produce a smaller confidence region for any given level of coverage.
-We therefore apply a pseudocount to give a slight boost to events that appear to have small probability.
-In practice we find that this encourages the confidence region to err on the side of being conservative, giving us greater confidence that the true parameter lies within the confience region.
+    \item Define $\Theta(p,h) \triangleq \{q:\ h(y|\ell)=\sum_x p(x|\ell)q(y|x) \}$.  
+    \item Let $Q_\mathrm{low} = \argmin_{q\in\Theta(P,H)} q(\tilde y|\tilde x)$ and $Q_\mathrm{high} = \argmax_{q\in\Theta(P,H)} q(\tilde y|\tilde x)$.
+    \item Let $V(k)$ denote the probability that the interval $[Q_\mathrm{low},Q_\mathrm{high}]$ is contained inside $R(N,k)$ if $N$ is sampled according to $P,H$.
 
-\label{sec:confinterval}
+The 95\% MLM confidence interval for $q^*(y|x)$ is defined as $R(N,k(N))$.
+Let us consider why this approach might work.
+Observe first that $q^* \in \Theta(p^*,h^*)$.
+Thus, letting $q^*_\mathrm{low} = \argmin_{q\in\Theta(p^*,h^*)} q^*(\tilde y|\tilde x)$ and $q^*_\mathrm{high} = \argmax_{q\in\Theta(p^*,h^*)} q(\tilde y|\tilde x)$ we know that $q^*(\tilde y|\tilde x)$ lies in the interval $[q^*_\mathrm{low},q^*_\mathrm{high}]$.
+We could therefore obtain a conservative confidence interval if we could pick $k$ so that $[q^*_\mathrm{low}(y|x),q^*_\mathrm{high}(y|x)]$ is contained within $R_{x,y}(N,k)$ with probability at least 95\% when $N$ is sampled according to $p^*,h^*$.
+Notice that $q^*_\mathrm{low},q^*_\mathrm{high}$ are deterministic functions of $p^*,h^*$, so we would not need to know $q^*$ in order to determine this probability.
+This is a good thing because the parameter $q^*$ is not always identifiable even with infinite data.
+However, the parameters $p^*,h^*$ can be estimated, and our uncertainty about them reasonably well described using a credible interval.
+We therefore pick the smallest $k$ so that $[Q_\mathrm{low},Q_\mathrm{high}] \subset R(N,k)$ with probability at least 95\% when $N$ is sampled from any of a collection of estimates for $p^*,h^*$.
+We hope that this $k$ also ensures that $[q^*_\mathrm{low}(y|x),q^*_\mathrm{high}(y|x)] \subset R_{x,y}(N,k)$ with probability at least 95\% when $N$ is sampled from $p^*,h^*$.
+
+It is far from obvious when this approach yields a conservative 95\% confidence region.
+That is, we do not know that $q^*(\tilde y|\tilde x)$ is inside $R_{x y}(N,k(N))$ with probability at least 95\% for experiments with any possible choice of $p^*,q^*$ and any possible number of samples.
+The geometric and statistical relationships among all the relevant objects is not easily understood.
+There could be some pathological edge cases where the method fails dramatically.
+In the future it would be interesting to discover troublesome cases.
+
+For the time being, we content ourselves with the fact that the method performs adequately in simulations.
 
@@
 
-Todo
-
 \label{sec:bayesian}
 
-\section{Convex optimization problems}
+The main task of this paper is to understand how samples of $N=(N^X,N^Y)$ should inform our knowledge about the link parameters $q^*(y|x)$, when $N$ is sampled according to $p^*,h^*$ and $h^*(y|\ell)=\sum_x p^*(x|\ell)q^*(y|x)$ (see Appendix \ref{sec:confinterval} for a precise description of the probability model).
+
+We took a frequentist approach, but there are certain reasons a bayesian approach might be more appealing.
+The complications of non-simulatenous confidence intervals would vanish.
+The complicated procedure for estimating the crucial quantity $k$ (see Appendix \ref{sec:confinterval}) would no longer be necessary.
+We could replace confidence intervals with bayesian credible intervals, which have a philosophically appealing interpretation.
+
+The difficulty is that we would not like to place any modeling assumptions on $q^*$.
+Indeed, we know that $q^*$ is not always an identifiable parameter, and so there may be \emph{no opportunity} for the data to correct any erroneous assumptions we make in specifying our prior belief.
+Even seemingly `noninformative' priors can actually have very strong implications for the credible intervals.
+
+Here we present an example where a noninformative prior results in profoundly erroneous conclusions.
+
+There is one way we could apply Bayesian reasoning without placing any assumptions on $q^*$.
+However, we were unable to figure out how to overcome the computational issues involved.
+The idea is that it is straightforward to articulate sensible prior beliefs about $p^*,h^*$.
+These are simple categorical parameters which we know are identifiable from data; Dirichlet priors are frequently employed with great success.
+We can apply this prior with a Bayesian perspective to obtain posterior samples $P,H$ by conditioning our distribution for $p^*,h^*$ upon the additional knowledge which we have:
+%
+\begin{itemize}
+    \item We have observed the matrices $N^X,N^Y$, which were sampled from $p^*,h^*$.
+    \item There exists some $q^*$ such that $h^*(y|\ell)=\sum_x p^*(x|\ell)q^*(y|x)$.
+\end{itemize}
+%
+Notice that in doing so we make no assumptions about what $q^*$ might be like -- only that it exists.
+A grasp of this conditional distribution would enable us to get credible intervals for $q^*$ by solving simple linear programming problems.
+Indeed, let $\Theta(p,h) = \{q:\ h(y|\ell)=\sum_x p(x|\ell)q(y|x) \}$.
+For each conditional sample $P,H$ and any choice of $x,y$ we could calculate
+%
+\begin{align*}
+Q_\mathrm{low}(P,H) &= \argmin_{q\in\Theta(P,H)} q(y|x)\\
+Q_\mathrm{high}(P,H) &= \argmax_{q\in\Theta(P,H)} q(y|x)
+\end{align*}
+%
+Let $\hat q_\mathrm{low}$ denote the 5th percentile of $Q_\mathrm{low}(P,H)$ and $\hat q_\mathrm{high}$ denote the 95th percentile of $Q_\mathrm{high}(P,H)$.
+Then $[\hat q_\mathrm{low},\hat q_\mathrm{high}]$ is a credible interval for $q^*(y|x)$.
+In particular, consider any prior distribution on $p^*,q^*,h^*$ that is consistent with our prior beliefs on $p^*,h^*$.
+There may be an infinite family of such distributions, each member reflecting a different prior belief we might have about $q^*$.
+For \emph{any} choice inside this family, we would be guaranteed a high posterior probability for the event $q^*(y|x) \in [\hat q_\mathrm{low},\hat q_\mathrm{high}]$.
+If the relevant computational barriers could be overcome, this would be a good direction for future research.
+
+\section{Numerical issues}
 

### 2018-08-07 paper/paper.tex
Back to bayes

@@
 
-% \newcommand{\figgygr}[2]{\includegraphics[width=#1\textwidth]{#2}}
-% \newcommand{\figgygrs}[3]{\begin{tabular}{c}\includegraphics[width=#1\textwidth]{#2}\\#3\end{tabular}}
+\newcommand{\figgygr}[2]{\includegraphics[width=#1\textwidth]{#2}}
+\newcommand{\figgygrs}[3]{\begin{tabular}{c}\includegraphics[width=#1\textwidth]{#2}\\#3\end{tabular}}
+\newcommand{\figgygrsTable}[3]{\begin{tabular}{c}{#2}\\#3\end{tabular}}
+\newcommand{\figgygrsH}[3]{\begin{tabular}{c}\includegraphics[width=#1\textwidth,angle=270]{#2}\\#3\end{tabular}}
+
+% \newcommand{\figgygr}[2]{pic}
+% \newcommand{\figgygrs}[3]{\begin{tabular}{c}pic\\#3\end{tabular}}
 % \newcommand{\figgygrsTable}[3]{\begin{tabular}{c}{#2}\\#3\end{tabular}}
-% \newcommand{\figgygrsH}[3]{\begin{tabular}{c}\includegraphics[width=#1\textwidth,angle=270]{#2}\\#3\end{tabular}}
-
-\newcommand{\figgygr}[2]{pic}
-\newcommand{\figgygrs}[3]{\begin{tabular}{c}pic\\#3\end{tabular}}
-\newcommand{\figgygrsTable}[3]{\begin{tabular}{c}{#2}\\#3\end{tabular}}
-\newcommand{\figgygrsH}[3]{\begin{tabular}{c}pic\\#3\end{tabular}}
+% \newcommand{\figgygrsH}[3]{\begin{tabular}{c}pic\\#3\end{tabular}}
 
@@
 
-If we have a single large experiment that measures each specimen with both measurement procedures, we can estimate the link without the MLM.
-In this case it is straightforward to study the link between the two perspectives, because we can examine specimens through both perspectives simultaneously.
+If we have a single large experiment that measures each specimen with both measurement procedures, it is straightforward to estimate the link.
+We can examine specimens through both perspectives simultaneously.
 However, in many cases we have some experiments in which specimens are measured with one procedure, other experiments using the other procedure, and few or no experiments where specimens are measured with both procedures.
@@
     \item $\ell$, the sampling procedure used to obtain the specimen.
-For example, we can use select different subpopulations of cells based on proteomic markers.
+For example, we can select different subpopulations of cells based on proteomic markers.
 For each marker we might have a different sampling procedure designed to select cells with that marker.
@@
 Someone selects a cell using one of a set of sampling procedures and measures the transcriptomic activity of the cell.
-We are then told the cell's transcriptomic activity (but we are not told the sampling procedure used to gather the specimen).
+We are then told the cell's transcriptomic activity (but we are not told which of the sampling procedured was used to gather the specimen).
 Using our perfect knowledge of the physical systems, we could then make predictions about a measurement of the cell's morphology.
-Now we are told new information: we are told the sampling procedure used to gather the specimen.
+Now we are told new information: we are told which sampling procedure was used to gather the specimen.
 Would this substantially change our predictions?
@@
 Likewise $N^Y_{\ell y}$ indicates the number of specimens gathered with procedure $\ell$ that have morphological type $y$.
+For each transcriptomic type $x$ and morphological type $y$, the Markov Link Method uses the data to produce two objects:
+%
+\begin{enumerate}
+    \item $\hat q(y|x)$ -- a point estimate for the link parameter $q^*(y|x)$
+    \item $C_{x,y}$ -- a credible interval which contains the true link parameter $q^*(y|x)$ with high probability
+\end{enumerate}
+%
+To produce these objects we proceeds as follows.
+Let $\tilde P(\cdot |\ell) \sim \mathrm{Dirichlet}(N^X_\ell+1)$ and $\tilde H(\cdot |\ell) \sim \mathrm{Dirichlet}(N^Y_\ell+1)$ for each $\ell$.
+For any pair $h,h'$ let
+\[
+D(h|h') = \sum_{\ell,y} h(y|\ell) \log \frac{h(y|\ell)}{h'(y|\ell)}
+\]
+denote the Kullback-Leibler divergence from $h$ to $h'$.
+For each $\tilde P,\tilde H$, let $E(\tilde P,\tilde H)$ denote the minimum value of $D(\tilde H|h')$ among all $h'$ such that there exists $q$ satisfying $h'(y|\ell)=\sum_x \tilde P(y|\ell)q(y|x)$.
+Let $P,H$ denote variables carrying the distribution of $\tilde P,\tilde H$ after conditioned on $E(\tilde P,\tilde H)=0$.
+Let $\Theta(p,h)=\{q:\ h(y|\ell) = \sum_x p(x|\ell)q(y|x)\}$.
+We then define the MLM objects as follows:
+\begin{enumerate}
+    \item $\hat q\triangleq\mathbb{E}[\argmin_{q\in \Theta(P,H)} \sum_{x,y} q(y|x)^2]$.
+    \item The lower bound of $C_{x,y}$ is defined as the largest number $c$ such that $\min_{q\in \Theta(P,H)} q(y|x) \geq c$ with probability 97.5\%.
+Likewise the upper bound is the smallest number $c$ such that $\max_{q\in \Theta(P,H)} q(y|x) \leq c$ with probability 97.5\%.
+\end{enumerate} 
+%
+Code to compute these objects is published at https://github.com/jacksonloper/markov-link-method, including a tutorial-style ipython notebook detailing every computation made in this paper.
 
-Taking a frequentist point of view, the Markov Link Method uses this data to produce two kinds of objects:
-\begin{itemize}
-    \item $\tilde q(y|x)$, a point estimate for the link parameter $q^*(y|x)$.
-To obtain this we must incorporate nuisance parameters: let $p^*(x|\ell)$ denote the probability that a cell sampled with procedure $\ell$ will have transcriptomic type $x$.
-The log likelihood of the data is then given by
-    \[
-    L(p,q)=\left(\sum_{\ell,x} N^X_{\ell x} \log p(x|\ell) + \sum_{\ell,y} N^Y_{\ell y} \log \left(\sum_x p(x|\ell) q(y|x)\right)\right)
-    \]
-    We use a maximum likelihood approach to build our point estimators:
-    \[
-    \tilde p,\tilde q \in \argmax_{p,q} L(p,q)
-    \]
-    We caution that this optimization problem does not always have a unique maximizer, so the result may depend upon the maximization strategy used.
-Precise details for our optimization algorithms may be found in Appendix \ref{sec:optproblems}.
-This non-uniqueness is related certain so-called `identifiability' issues: in some cases it may be impossible to obtain a perfect point estimate even with an infinite number of samples.
-We discuss this issue at length in Appendix \ref{sec:casestudies}.
-In light of these concerns, it is particularly important to also consder our uncertainty:
+Our approach arises naturally from a Bayesian point of view.
+To understand them, first notice that there are several `nuisance parameters' we must consider in order to estimate the link $q^*$.
+Let $p^*(x|\ell)$ denote the probability that a cell sampled with procedure $\ell$ will have transcriptomic type $x$ and $h^*(y|\ell) \triangleq \sum_x p^*(x|\ell)q(y|x)$ denote the probability such a cell will have morphological type $y$.
+Note that the distribution of the data $N^X,N^Y$ is entirely determined by $p^*,h^*$.
+In particular, if someone changed $q^*$ without changing $p^*,h^*$ there would be no way for us to know.
+If this can occur, $q^*$ is said to be `nonidentifiable.'   Therefore, instead of trying to directly estimate $q^*$ (which may be impossible to accurately estimate) we focus on three quantities which can be estimated from data:
 %
-    \item $C_{x,y}$, a 95\% confidence interval for $q^*(y|x)$.  We obtain these with profile likelihood approach.  The profile likelihood is defined as
-    \[
-    L_{x y}(\xi)=\max_{\substack{p,q \\ q(y|x)=\xi}} L(p,q)
-    \]
-    This is the maximum possible likelihood of the data under the constraint that $q(y| x)=\xi$.  We then take $L^* = \sup_{p,q} L(p,q)$ and define a confidence region for $q^*( y| x)$ by
-    \[
-    R_{x y} = \left\{\xi:\ L^* - L(\xi) < k\right\}
-    \]
-    We choose $k$ to achieve the desired coverage.
-That is, we would like to choose $k$ so that $q^*(y|x) \in R_{x y}$ with probability at least 95\%.
-We refer the reader to Appendix \ref{sec:confinterval} for details on choosing $k$.
-Other coverages besides 95\% can also be produced by the same method.
-Finally, for simplicity of visualization we transform the set $R_{x y}$ (which could be composed of several disjoint intervals) into a simple interval by taking $C_{x,y} = [\min R_{x,y},\max R_{x,y}]$.
-Since this interval is a superset of the region, it will certainly contain $q^*(y|x)$ as long the region does.
-\end{itemize}
+\begin{align*}
+q^*_{\mathrm{mid}} &\triangleq \argmin_{q\in \Theta(p^*,h^*)} \sum_{x,y} q(y|x)^2\\
+q^*_{\mathrm{lo},x,y} &\triangleq \min_{q\in \Theta(p^*,h^*)} q(y|x) \\
+q^*_{\mathrm{hi},x,y} &\triangleq \max_{q\in \Theta(p^*,h^*)} q(y|x) 
+\end{align*}
+%
+Note that all three of these optimization problems have unique solutions and would be straightforward to solve if $p^*,h^*$ were known; optimization details can be found in Appendix \ref{sec:optproblems}.
+Moreover, regardless of any identifiability issues for $q^*$ itself, the objects $q^*_{\mathrm{lo},x,y}$ and $q^*_{\mathrm{hi},x,y}$ are always identifiable for every $(x,y)$ and define bounds on the true value of $q^*(y|x)$.
 
-A Bayesian point of view would suggest a different procedure, one which might provide a useful complementary perspective.
-This is worthy of future research.
-However, we caution that it is nontrivial to choose reasonable prior distributions for $p^*,q^*$.
-Priors which may seem uninformative (such as the Dirichlet distribution) can have significant unanticipated consequences, due to the high-dimensional nature of the problem.
-We give an example in \ref{sec:bayesian}.
+The MLM algorithm therefore adopts a Bayesian point of view to estimate $q^*_{\mathrm{mid}}$ and create a credible interval for $q^*_{\mathrm{lo},x,y},q^*_{\mathrm{hi},x,y}$.
+We begin with uniform prior beliefs for the parameters of $p^*,h^*$.
+That is, our prior is that $p^*(\cdot|\ell),h^*(\cdot|\ell)$ carry $\mathrm{Dirichlet}(1)$ distributions for each $\ell$.
+Following the Bayesian philosophy, we would like to incorporate any additional knowledge through conditioning.
+We have two important pieces of knowledge we would like to incorporate: that the MLM assumption holds (i.e.\ there is some $q^*$ so that $\sum_x p^*(x|\ell)q^*(y|x)=h^*(y|\ell)$) and we have observed the data $N^X,N^Y$.
+By conditioning on this knowledge, we obtain the posterior distribution carried by the variables $P,H$ described above.
+In terms of these variables, the MLM estimator $\hat q$ is simply the posterior mean for $q^*_{\mathrm{mid}}$ and the MLM credible interval $C_{x,y}$ is an interval which contains $q^*_{\mathrm{lo},x,y},q^*_{\mathrm{hi},x,y}$ with high probability.
 
-The frequentist confidence interval allows us to avoid making any modeling assumptions (beyond the basic Markov Link Method assumption).
-However, it does come with some drawbacks.
-For example, we here take the view that values inside the confidence interval are somehow `plausible' or `likely.'  Such a statement cannot be philosophically justified without some at least a hint of Bayesian reasoning (cf.\ \cite{morey2016fallacy}).
-It is also important to note that, as is typical for confidence intervals, we make no corrections for multiple testing.
-We therefore expect that a small proportion of our confidence intervals will not contain the truth.
-It is therefore important to consider all of the intervals produced and be wary of `cherry picking' only those intervals that support a particular conclusion.
-However, if cherry picking is desired, a false discovery rate correction may be applied (cf.\ \cite{benjamini2005false}).
-A related drawback is that the confidence intervals are not always logically consistent with one another.
-For example, it can happen that that $C_{x,y_1}=[0,1]$ and $C_{x,y_2}=[.2,1.0]$.
-These intervals don't really make sense together, in that if $q(y_2|x)\geq .2$ then certainly $q(y_2|x)\leq .8$.
-A possible remedy would be to construct a simultaneous confidence region for all the parameters concerning a fixed value of $x$.
-Analyzing the statistical properties of these high-dimensional confidence regions would be nontrivial; we leave it for future research.
+In practice, we were not able to find a way to compute the credible interval and posterior means exactly.
+It seems difficult to even sample the conditional random variables $P,H$.
+Common approaches to this type of problem involve Markov-Chain Monte Carlo and Variational Inference, but we were unable to make these approaches work in practice.
+It seems nontrivial to work with the condition $E(p,h)=0$ that formalizes the MLM assumption.
+We instead take a somewhat na\"ive approach.
+We start with $\tilde P,\tilde H$, which carry the distribution of $p^*,h^*$ conditioned on the observed data (but not conditioned on the fact that the MLM assumption holds).
+We then simply project $\tilde H$ to the space of parameters which are consistent with the MLM assumption.
+That is, we take $\hat H$ to be the solution of the minimization problem $\min_{h'}D(\tilde H|h')$, subject to the constraint that $E(\tilde P,h')=0$.
+Optimization details can be found in Appendix \ref{sec:optproblems}.
+We hope that the distribution of $\tilde P,\hat H$ will approximate the desired posterior $P,H$.
+We then use samples from this distribution to compute what we need.
+The implications of this approximation are difficult to reason about; we hope further study will clarify when this is a reasonable procedure.
+In the mean time, we content ourselves that it is a practical solution to a difficult problem.
+
+A frequentist point of view might provide a valuable counterpoint to our Bayesian analysis.
+However, we were unable to produce effective confidence intervals for the link parameters.
+There are at least two difficulties: the link is unidentifiable and the true parameter values may lie on the edge of the parameter space.
+These challenges make traditional appproaches based on asymptotic normality difficult to apply.
+Perhaps future work could uncover a way to overcome these challenges.
 
@@
 The MLM estimator accurately recovers the ground truth from simulated data, as long as we have enough samples \emph{and} enough distinct sampling procedures.
-In subfigure (b) we consider the MLM's 95\% confidence intervals.
+In subfigure (b) we consider the MLM's credible intervals.
 For a selection of trials we choose a random value of of $x,y$ and then compare the true link parameter $q^*(y|x)$ with the MLM confidence interval $C_{x,y}$.
@@
 Since the link parameters lie in $[0,1]$, the interval is always contained within $\pm1$ of the ground truth (this larger region is shown in gray).
-The interval mostly contain the truth, and get narrower with more samples as long as there are enough distinct sampling procedures.
+The intervals mostly contain the truth, and get narrower with more samples as long as there are enough distinct sampling procedures.
 In subfigure (c) we focus on the case of exactly $4$ sampling procedures.
 In this case it is only possible to get very accurate estimates if the link takes on a particular form.
-If the link is indepenent (i.e.\ $X,Y$ are independent) or if the link is invertible and deterministic (e.g.\ $X=Y$) then we can determine what the link is.
+If the link is independent (i.e.\ $X,Y$ are independent) or if the link is invertible and deterministic (e.g.\ $X=Y$) then we can determine what the link is.
 However, if the truth is more complex, small numbers of sampling procedures may make it impossible to determine the true link; this issue is discussed in detail in Section \ref{sec:casestudies}.}
@@
 
-We used simulations to test the two parts of the MLM: the point estimators $\tilde q(y|x)$ and confidence intervals $C_{x,y}$.
-For any fixed value of the ground truth parameters, $(p^*,q^*)$, we would ordinarily hope for two things:
+We used simulations to test the two parts of the MLM: the point estimators $\hat q(y|x)$ and credible intervals $C_{x,y}$.  We would ordinarily hope for two things:
 \begin{description}
-    \item[Estimator convergence] As we obtain more samples, $\tilde q(y|x)$ should converge to $q^*(y|x)$ for each $x,y$.
-    \item[Conservative coverage] $\mathbb{P}(q^*(y|x) \in C_{x,y})\geq95\%$ for each $x,y$.
-That is, for each choice of $x,y$ the confidence interval $C_{x,y}$ should contain $q^*(y|x)$ with probability at least 95\%.\footnote{Note that we are not producing a `simultaneous' interval.
-In a simultaneous interval, there would be a 95\% chance that $q^*(y|x)\in C_{x,y}$ for \emph{every} $x,y$.
-The key difference is the change from `each' (in our intervals) to `every' (in simultaneous intervals).
-Thus, it may be that $q^*(y|x)\notin C_{x,y}$ for a few values of $x,y$.
-For \emph{most} values of $x,y$ the true parameters should be contained inside of their corresponding confidence intervals.}
+    \item[Estimator convergence] As we obtain more samples, $\hat q(y|x)$ should converge to the true link $q^*(y|x)$ for each $x,y$.
+    \item[Conservative coverage] $q^*(y|x) \in C_{x,y}$ with high probability.
 \end{description}
@@
 The bad news: in many cases the MLM fails to achieve estimator convergence.
-Indeed, we can show that an `identifiability issue' blocks any possibility of success in some cases; we detail this problem in Appendix \ref{sec:casestudies}.
+Indeed, we can show that the identifiability issues block any possibility of success in some cases; we detail this problem in Appendix \ref{sec:casestudies}.
 The following two simulations give a good overview of the relevant issues:
@@
 We then simulate a dataset from these ground truth parameters, fixing the total number of samples (spreading these samples equally among all combinations of sampling procedure and measurement procedure).
-Finally, we apply the MLM to the simulated data to get the point estimates $\tilde q(y|x)$ and the confidence intervals $C_{x y}$.
+Finally, we apply the MLM to the simulated data to get the point estimates $\tilde q(y|x)$ and the credible intervals $C_{x y}$.
 We measure the overall estimator convergence using a kind of total variation distance:
@@
 $q^*(y|x)=0$ whenever $\tilde q(y|x)>0$ and vice-versa.
-We also look at the MLM confidence intervals and see how often they cover the true parameters.
+We also look at the MLM credible intervals and see how often they cover the true parameters.
 
@@
 This figure also shows that the confidence intervals work correctly regardless of of the number of samples or sampling procedures; they include the ground truth with high probability.
-A more precise analysis of these intervals suggests that they achieve conservative coverage (see Appendix \ref{sec:confinterval} for details).
-With many samples and sampling procedures, these intervals are small and concentrated around the truth.
+With many samples and sampling procedures, the intervals are small and concentrated around the truth.
 With fewer samples or sampling procedures, the intervals are sometimes forced to be larger.
 However, our uncertainty is not the same for every parameter.
-In some cases we are able to obtain a very tight confidence interval for one particular parameter even though the overall estimator error is high.
+In some cases we are able to obtain a very tight credible interval for one particular parameter even though the overall estimator error is high.
 
-    \item How does estimator convergence depends upon the link parameters themselves?
+    \item How does estimator convergence depend upon the link parameters themselves?
 In each trial of this simulation we use four sampling procedures and our measurement procedures yield one of six categories.
@@
 This result is proven in Appendix \ref{sec:casestudies}, Theorem \ref{thm:miracle}.
+\end{itemize}
 
@@
 Using this kind of data, our task is to calibrate the two classification protocols.
-That is, we want to be able to ask question of the following form: ``if a neuron is classified as being of type `Peri Kcnj8' by tool I, how might it have been classified by tool II?'' \label{fig:alleninput}}
+That is, we want to be able to ask question of the following form: ``if a neuron is classified as being of type `Peri Kcnj8' by Standard, how might it have been classified by Patch?'' \label{fig:alleninput}}
 }
@@
 However, many of these procedures destroy the neurons in the process of measuring them.
-We'll look at two procedures in particular, which we'll call `Standard' and 'Patch.'  Standard uses a single-cell RNA sequencing pipeline to determine transcriptomic expression in the cell, and determine the cell-type based on these results.
-Patch also uses transcriptomic information from RNA sequencing, but Patch additionally obtains electrophysiological and morphological properties of the neuron.
+We'll look at two procedures in particular, which we'll call `Standard' and 'Patch.'  Standard uses a single-cell RNA sequencing pipeline to determine transcriptomic expression in the cell and determines the cell-type based on these results.
+Patch also uses transcriptomic information from RNA sequencing.
+Patch additionally obtains electrophysiological and morphological properties of the neuron.
 This comes at the cost of a degraded transcriptomic signal, requiring new methods to estimate the cell-type.
@@
 Each procedure was expected to yield different proportions of the different cell-types.
-For each sampling procedure and eaach measurement procedure, many specimens were sampled and their cell-types determined.
+For each sampling procedure and each measurement procedure, many specimens were sampled and their cell-types determined.
 The result of this process was two tables, parts of which are shown in Figure \ref{fig:alleninput}.
@@
 \caption{{\bf MLM estimation of the link, with confidence interval.}  The central plot shows a portion of the MLM estimator applied to the Allen Institute data.
-We look at a subselection of Standard types and all of the Patch types.
-For each combination $x,y$ we draw a rectangle whose color indicates the value of the estimator $\tilde q(y|x)$.
+We look at a subselection of Standard types ($x$) and all of the Patch types ($y$).
+For each combination $x,y$ we draw a rectangle whose color indicates the value of the estimator $\hat q(y|x)$.
 We also show our confidence about these estimators.
-On the left we indicate the lower bounds indicated by the confidence interval, on the right we indicate upper bounds.
+On the left we indicate the lower bounds indicated by the credible interval, on the right we indicate upper bounds.
 For some parameters our intervals are much tighter than others.
-For example, it appears we have high confidence the Standard type `Vip Rspo4' is highly associated with the Patch type `Vip.'  By constrast, we have almost no idea what is associated with the Standard type `Meis2.'  The upper confidence interval bounds suggest that $q^*(y|\mathrm{Meis2})$ could be nearly 1 for $y$ equal to `Lamp5,' 'Pvalb,' 'Sncg,' 'Sst,' or `Visp' Patch types.
+For example, it appears we have high confidence the Standard type `Vip Rspo4' is highly associated with the Patch type `Vip.'  By constrast, we have almost no idea what is associated with the Standard type `Meis2.'  The upper confidence interval bounds suggest that $q^*(y|\mathrm{Meis2})$ could be nearly 1 for many different Patch types.
 Obviously it cannot be 1 for all of those types simultaneously, since $\sum_y q^*(y|\mathrm{Meis2})=1$, but the data simply doesn't tell us which $y$ carries the mass.
@@
 However, for other types there is more ambiguity.
-For the Standard type $x=$`Lamp5 Egln 1' and the Patch types $y=$`Lamp' we have $C_{x y}=[.42,1.0]$.
-This is a much wider confidence interval.
-It does say that at least 42\% of the cells with Standard type `Lamp5 Egln1 1' are given the `Lamp5' Patch type.
-But what about the other 58\%?
-It might be that they're also assigned the `Lamp5' Patch type, but the data doesn't tell us for sure.
-Looking at the confidence intervals for other Patch types, we conclude that `Lamp5 Egln 1' types are not ususally classified with Patch types `CR,' 'L6b,' `Sst,' `VISp,' `Vip,' or `Unknown': the upper limits on $q^*(y'|x)$ are small for each of these Patch types.
-The intervals show that many of these `Lamp5 Egln 1' cells \emph{might} be assigned the Patch types 'Pvalb' or 'Sncg.'  The data we have does not give us a definitive answer.
+For the Standard type $x=$`Lamp5 Egln 1' and the Patch types $y=$`Lamp' we have $C_{x y}=[0,1.0]$.
+The data we have does not give us a definitive answer as to whether cells with Standard type `Lamp5 Egln 1' are being classified with Patch type `Lamp.'
 
@@
 To get a better resolution of the calibration, we would need a sampling strategy that included `Lamp5 Egln 1' cells but represents a significantly different slice of the overall population.
-For a particular proposed experiment, simulations such as those found in Section \ref{sec:simulations} can be used to determine how many samples might be required to get an accurate estimate of the calibration.
+For a particular proposed experiment, simulations such as those found in Section \ref{sec:simulations} can be used to determine how many samples might be required to get an accurate estimate of the link, $q^*$.
 
@@
 From the most general point of view, we are engaged in meta-analysis; we refer the reader to \cite{borenstein2011introduction} for a general introduction to the field.
-In this broader context, the main distinguishing characteristics of this paper are two-fold: we place no assumptions on the nature of the link and we take the identifiability issues seriously.\footnote{We refer the reader to \cite{walter2014identifiability} for an introduction to what we mean by identifiability.
-We discuss the identifiability issues for our problem in particular in Appendix \ref{sec:casetstudies}.}  Our only assumption is that the sampling procedures can be statistically isolated from the link.
+
+The main distinguishing characteristics of this paper are two-fold: we place no assumptions on the nature of the link and we take the identifiability issues seriously.
+We refer the reader to \cite{walter2014identifiability} for an introduction to what we mean by identifiability.
+We discuss the identifiability issues for our problem in particular in Appendix \ref{sec:casestudies}.
+Our only assumption is that the sampling procedures can be statistically isolated from the link.
 This assumption says nothing about what the link itself is.
-We then use profile likelihood confidence intervals to deal with any identifiability issues.
-These intervals account for uncertainty due both to small sample sizes and to identifiability issues.
+We take a Bayesian approach, but we only use our prior beliefs to estimate quantities which actually can be estimated, carefully avoiding the identifiability issues.
+We use these quantities to produce intervals which account for uncertainty due both to small sample sizes and to identifiability issues.
 
-The main technical contribution of this paper was figuring out how to use the MLM assumption to get confidence intervals that worked in practice.
+The main technical contribution of this paper was figuring out how to use the MLM assumption to get credible intervals that worked in practice.
 In this we were inspired by a large literature of examples where assumptions are used to bound potentially unidentifiable parameters.
@@
 In some cases, we may believe that similar measurement values should have similar probabilities.
-Such smoothness assumptions would make it possible to apply the MLM to measurement tools which can return many different values.
-The frequentist point of view allows us to phrase these kinds of assumptions as hard constraints that we believe in, instead of guesses about what might be more or less likely.
-In some cases these assumptions are easier to think about than prior distributions, as we see in Appendix \ref{sec:bayesian}.
+Such smoothness assumptions would make it possible to apply the MLM to measurement tools which can return a continuum of values.
 
 More generally, it is clear that good assumptions can help us get real insight for tough problems.
-Even if these assumptions are not sufficient to allow us to perfectly identify our object of interest, confidence regions can be constructed.
+Even if these assumptions are not sufficient to allow us to perfectly identify our object of interest, credible regions can be constructed.
 By probing these regions carefully, we can learn what the data actually has to say and what experiments will help us learn more.
@@
 
-\section{Confidence interval details}
+\section{Numerical issues}
 
-\label{sec:confinterval}
+\label{sec:optproblems}
 
-In the main text of this paper, Section \ref{sec:mlm}, we outline a method for producing confidence intervals for the link parameters $q^*(\tilde y|\tilde x)$ for a specific choice of $\tilde x,\tilde y$.
-Here we give full details on this method and discuss its properties in depth.
+There are three numerical problems which the MLM must solve.  Here we detail our method for solving each of them.
 
-Consider experiments yielding a $\Omega_\ell\times\Omega_X$ matrix $N^X$ and an $\Omega_\ell \times \Omega_Y$ matrix $N^Y$, sampled according to
+\begin{enumerate}
+
+    \item Projecting to the MLM assumption.
+Fix any parameters $P,\tilde H$.
+One step in the MLM involves projecting $\tilde H$ to the set of parameters which are consistent with $P$ and the MLM assumption.
+In particular, we defined
+    \[
+    D(h|h') = \sum_{\ell,y} h(y|\ell) \log \frac{h(y|\ell)}{h'(y|\ell)}
+    \]
+    and we needed to solve the problem
+    \[
+    \min_{h} D(\tilde H|h)
+    \]
+    subject to the constraint that there exists some $q$ such that $h(y|\ell)=\sum_x p(x|\ell)q(y|x)$.  Parametrizing valid $h$ through $q$, we obtain the problem
+    \[
+    \max_{q} \sum_{\ell,y} \tilde H(y|\ell) \log \left(\sum_x P(x|\ell) q(y|x)\right)
+    \]
+    Taking derivatives one can readily show that this problem is convex.  We solve it using an iterative method.  We initially guess that $q$ is uniform.  We then repeatedly make the updates
+    \begin{align*}
+    q(y|x)  &\propto q(y|x)\sum_{\ell}p(x|\ell)\frac{\tilde H(y|\ell)}{\sum_x p(x|\ell)q(y|x)}\\
+    \end{align*}
+    until convergence.  Our convergence criteria is that all parameters change less than $10^{-5}$ in a single iteration.
+
+    \item Linear programming.
+Fix any $P,H$.
+To deal with the identifiability issues, we defined $\Theta(p,h)=\{q:\ h(y|\ell) \triangleq \sum_x p(y|\ell)q(y|x)\}$.
+The MLM requires us to solve linear optimization problems within $\Theta$, such as
+    \[
+    \min_{q\in \Theta(P,H)} q(y|x) \\
+    \]
+    We solve these problems using the {\tt cvxopt} python package.
+
+    \item Quadratic programming.  The MLM also requires us to solve quadratic optimization problems within $\Theta$:
+    \[
+    \min_{q\in \Theta(P,H)} \sum q(y|x)^2 \\
+    \]
+    We solve these problems using the {\tt cvxopt} python package.
+
+\end{enumerate}
+
+\section{Identifiability}
+
+\label{sec:casestudies}
+
+The issue of identifiability comes up repeatedly throughout this paper.
+Here we give a brief overview of the fundamentals of this issue and how it effects us.
+We also present two suggestive case studies which we hope may inspire future research.
+In both cases we are able to prove something of interest -- but not quite as much as we might hope.
+
+Consider experiments yielding an $\Omega_\ell\times\Omega_X$ matrix $N^X$ and an $\Omega_\ell \times \Omega_Y$ matrix $N^Y$, sampled according to
 \begin{align*}
@@
 \end{align*}
-where $p^*(x|\ell)$ is some conditional distribution and there is some $q^*(y|x)$ such that $h^*(y|\ell)=\sum_x p^*(x|\ell)q^*(y|x)$.  We have then defined the log likelihood as 
-\[
-L(N;p,q)=\left(\sum_{\ell,x} N^X_{\ell x} \log p(x|\ell) + \sum_{\ell,y} N^Y_{\ell y} \log \left(\sum_x p(x|\ell) q(y|x)\right)\right)
-\]
-In the main text we supressed the dependence of $L$ the data $N=(N^X,N^Y)$, but it will be important here so we will emphasize it.  We define the profile likelihood as
-\[
-L(N;\xi)=\max_{\substack{p,q \\ q(\tilde y|\tilde x)=\xi}} L(p,q)
-\]
-We take $L^*=\max_{p,q} L(N;p,q)$ and define the sets
-\[
-R(N,k) = \left\{\xi:\ L^* - L(\xi) < k\right\}
-\]
-Our task is to find $k$ such that $q^*(\tilde y|\tilde x)$ is contained in $R(N,k)$ with probability at least 95\% if $N$ is sampled according to $p^*,h^*$.
-If we can do this, then $R(N,k)$ would constitute what is called a (conservative) 95\% confidence region for the parameter $q^*(\tilde y|\tilde x)$.
-We refer the reader to \cite{brazzale2007applied} for a helpful introduction to these kinds of confidence regions and some of the difficulties that can arise in using them.
+where $p^*(x|\ell)$ is some conditional distribution and there is some $q^*(y|x)$ such that $h^*(y|\ell)=\sum_x p^*(x|\ell)q^*(y|x)$.
+This is simply a restatement of the assumptions we have made throughout the data about how our parameters of interest $(p^*,q^*,h^*)$ are related to the data we can observe $(N^X,N^Y)$.
 
-The difficulty is that $p^*,q^*$ is unknown.
-So for any particular choice of $k$, we cannot determine the probability that $q^*(\tilde y|\tilde x)$ is contained in $R(N,k)$.
-In other situations this difficulty can sometimes be overcome by applying asymptotic normality results to argue that the the relevant quantities are actually invariant to the ground truth parameters.
-For example, in many situations the distribution of the relevant quantities can be asymptotically approximated using properties of $\chi^2$ distributions.
-In our case these normality results are unlikely to be useful.
-There are two difficulties.
-First, the ground truth parameters may lie against the edges of the viable parameter space (e.g.
-if $q^*(\tilde y|\tilde x)=0$); this invalidates many of the standard arguments for asymptotic normality.
-Second, the parameters $q^*(\tilde y|\tilde x)$ are not always identifiable; this can also create some strange artifacts.
+It is well-known that we can obtain arbitrarily good estimates of $p^*,h^*$ by taking enough samples (i.e.\ taking $n_\ell,m_\ell$ sufficiently high).
+Let us therefore imagine for a moment that we in fact have \emph{perfect knowledge} of $p^*,h^*$.
+Even so, the data does not necessarily tell us the value of the link parameters $q^*$.
+There may be many possible links, $q$, which are all equally consistent with $p^*,h^*$.
+Each such link yields the exact same distribution on the data we can observe, so there can be no way to use data to distinguish among them.
+This is known as a `nonidentifiability problem.'  Even with infinite data, we simply cannot identify exactly what the value of $q^*$ might be.
 
-Our approach is to use the data, $N$, to guide our choice of $k$.
-We define $k=k(N)$ in the following way.
-Our first step is to find a collection of plausible possibilities for the values of $p^*,h^*$.
-To do so, we define the following distribution for $P,H$.
-%
-\begin{itemize}
-    \item Let $P,\hat H$ denote a sample from the posterior of $p^*,h^*$ under a uniform prior.
-That is, consider the distribution of $p^*,h^*,N$ defined by sampling $p^*,h^*$ uniformly from parameter space and then sampling $N$ from $p^*,h^*$.
-The objects $P,\hat H$ are samples from this distribution when conditioned on the observed values of $N^X,N^Y$.
-    \item Let $H$ denote a projection of $\hat H$ to the set of parameters which are consistent with the Markov Link Method assumption.  In particular, 
-    \[
-    Q \in \argmax_{p,q} \sum_{\ell y} \hat H(x_\ell) \log \sum_x p(x|\ell) q(y|x)
-    \] 
-    and $H(y|\ell)=\sum_x P(x|\ell) Q(y|x)$.  Notice that maximization problem for $Q$ may have no unique solution, but the problem is convex and the value of $H$ is uniquely defined.
-\end{itemize}
-%
-We now define our collection of plausible possibilities as those which are `credible' under this distribution.
-In particular, let $l$ denote the fifth percentile of the variable $L(N;P,H)$, holding $N$ fixed at its observed value.
-That is $\mathbb{P}(L(N;P,H)<l|N)<5\%$.
-We take all $P,H$ The set of $P,H$
-Finally, we define $k(N)$ as the smallest value of $k$ such that that $V(k)\geq 95\%$ for every sample of $P,H$ such that $L(N;P,H)$ is above the fifth percentile  credible interval.
-Note that the computation of $k(N)$ involves optimization problems and difficult sums.
-Many of these problems do not have closed form solutions.
-We discuss the exact details of our numerical strategies in Appendix \ref{sec:optproblems}.
+We will now look at some examples:
 
-    \item Define $\Theta(p,h) \triangleq \{q:\ h(y|\ell)=\sum_x p(x|\ell)q(y|x) \}$.  
-    \item Let $Q_\mathrm{low} = \argmin_{q\in\Theta(P,H)} q(\tilde y|\tilde x)$ and $Q_\mathrm{high} = \argmax_{q\in\Theta(P,H)} q(\tilde y|\tilde x)$.
-    \item Let $V(k)$ denote the probability that the interval $[Q_\mathrm{low},Q_\mathrm{high}]$ is contained inside $R(N,k)$ if $N$ is sampled according to $P,H$.
+\subsection{A simple failure case}
 
-The 95\% MLM confidence interval for $q^*(y|x)$ is defined as $R(N,k(N))$.
-Let us consider why this approach might work.
-Observe first that $q^* \in \Theta(p^*,h^*)$.
-Thus, letting $q^*_\mathrm{low} = \argmin_{q\in\Theta(p^*,h^*)} q^*(\tilde y|\tilde x)$ and $q^*_\mathrm{high} = \argmax_{q\in\Theta(p^*,h^*)} q(\tilde y|\tilde x)$ we know that $q^*(\tilde y|\tilde x)$ lies in the interval $[q^*_\mathrm{low},q^*_\mathrm{high}]$.
-We could therefore obtain a conservative confidence interval if we could pick $k$ so that $[q^*_\mathrm{low}(y|x),q^*_\mathrm{high}(y|x)]$ is contained within $R_{x,y}(N,k)$ with probability at least 95\% when $N$ is sampled according to $p^*,h^*$.
-Notice that $q^*_\mathrm{low},q^*_\mathrm{high}$ are deterministic functions of $p^*,h^*$, so we would not need to know $q^*$ in order to determine this probability.
-This is a good thing because the parameter $q^*$ is not always identifiable even with infinite data.
-However, the parameters $p^*,h^*$ can be estimated, and our uncertainty about them reasonably well described using a credible interval.
-We therefore pick the smallest $k$ so that $[Q_\mathrm{low},Q_\mathrm{high}] \subset R(N,k)$ with probability at least 95\% when $N$ is sampled from any of a collection of estimates for $p^*,h^*$.
-We hope that this $k$ also ensures that $[q^*_\mathrm{low}(y|x),q^*_\mathrm{high}(y|x)] \subset R_{x,y}(N,k)$ with probability at least 95\% when $N$ is sampled from $p^*,h^*$.
-
-It is far from obvious when this approach yields a conservative 95\% confidence region.
-That is, we do not know that $q^*(\tilde y|\tilde x)$ is inside $R_{x y}(N,k(N))$ with probability at least 95\% for experiments with any possible choice of $p^*,q^*$ and any possible number of samples.
-The geometric and statistical relationships among all the relevant objects is not easily understood.
-There could be some pathological edge cases where the method fails dramatically.
-In the future it would be interesting to discover troublesome cases.
-
-For the time being, we content ourselves with the fact that the method performs adequately in simulations.
-
-\section{The Bayesian Alternative}
-
-\label{sec:bayesian}
-
-The main task of this paper is to understand how samples of $N=(N^X,N^Y)$ should inform our knowledge about the link parameters $q^*(y|x)$, when $N$ is sampled according to $p^*,h^*$ and $h^*(y|\ell)=\sum_x p^*(x|\ell)q^*(y|x)$ (see Appendix \ref{sec:confinterval} for a precise description of the probability model).
-
-We took a frequentist approach, but there are certain reasons a bayesian approach might be more appealing.
-The complications of non-simulatenous confidence intervals would vanish.
-The complicated procedure for estimating the crucial quantity $k$ (see Appendix \ref{sec:confinterval}) would no longer be necessary.
-We could replace confidence intervals with bayesian credible intervals, which have a philosophically appealing interpretation.
-
-The difficulty is that we would not like to place any modeling assumptions on $q^*$.
-Indeed, we know that $q^*$ is not always an identifiable parameter, and so there may be \emph{no opportunity} for the data to correct any erroneous assumptions we make in specifying our prior belief.
-Even seemingly `noninformative' priors can actually have very strong implications for the credible intervals.
-
-Here we present an example where a noninformative prior results in profoundly erroneous conclusions.
-
-There is one way we could apply Bayesian reasoning without placing any assumptions on $q^*$.
-However, we were unable to figure out how to overcome the computational issues involved.
-The idea is that it is straightforward to articulate sensible prior beliefs about $p^*,h^*$.
-These are simple categorical parameters which we know are identifiable from data; Dirichlet priors are frequently employed with great success.
-We can apply this prior with a Bayesian perspective to obtain posterior samples $P,H$ by conditioning our distribution for $p^*,h^*$ upon the additional knowledge which we have:
-%
-\begin{itemize}
-    \item We have observed the matrices $N^X,N^Y$, which were sampled from $p^*,h^*$.
-    \item There exists some $q^*$ such that $h^*(y|\ell)=\sum_x p^*(x|\ell)q^*(y|x)$.
-\end{itemize}
-%
-Notice that in doing so we make no assumptions about what $q^*$ might be like -- only that it exists.
-A grasp of this conditional distribution would enable us to get credible intervals for $q^*$ by solving simple linear programming problems.
-Indeed, let $\Theta(p,h) = \{q:\ h(y|\ell)=\sum_x p(x|\ell)q(y|x) \}$.
-For each conditional sample $P,H$ and any choice of $x,y$ we could calculate
+Consider the case that $\Omega_\ell = \Omega_Y = 2$ and $\Omega_X=3$.
+That is, there are $2$ separate sampling procedures, tool I recognizes 3 categories and tool II recognizes 2 categories.
+In particular, let us imagine that $p^*(x|\ell)=A_{\ell x}$ and $h^*(y|\ell)=B_{\ell y}$ where $A,B$ are matrices given by
 %
 \begin{align*}
-Q_\mathrm{low}(P,H) &= \argmin_{q\in\Theta(P,H)} q(y|x)\\
-Q_\mathrm{high}(P,H) &= \argmax_{q\in\Theta(P,H)} q(y|x)
+A&=\left(\begin{array}{ccc}
+\frac{1}{2} & \frac{1}{2} & 0\\
+\frac{2}{6} & \frac{1}{6} & \frac{3}{2}
+\end{array}\right) \\
+B&=\left(\begin{array}{cc}
+\frac{1}{2} & \frac{1}{2}\\
+\frac{2}{3} & \frac{1}{3}
+\end{array}\right)
 \end{align*}
-%
-Let $\hat q_\mathrm{low}$ denote the 5th percentile of $Q_\mathrm{low}(P,H)$ and $\hat q_\mathrm{high}$ denote the 95th percentile of $Q_\mathrm{high}(P,H)$.
-Then $[\hat q_\mathrm{low},\hat q_\mathrm{high}]$ is a credible interval for $q^*(y|x)$.
-In particular, consider any prior distribution on $p^*,q^*,h^*$ that is consistent with our prior beliefs on $p^*,h^*$.
-There may be an infinite family of such distributions, each member reflecting a different prior belief we might have about $q^*$.
-For \emph{any} choice inside this family, we would be guaranteed a high posterior probability for the event $q^*(y|x) \in [\hat q_\mathrm{low},\hat q_\mathrm{high}]$.
-If the relevant computational barriers could be overcome, this would be a good direction for future research.
 
-\section{Numerical issues}
-
-\label{sec:optproblems}
-
-Todo
-
-\section{Identifiability case studies}
-
-\label{sec:casestudies}
-
-Our analysis of the identifiability issues associated with the Markov Link Method is not completely satisfactory.
-We are able to make rigorous guarantees about consistency for any given true value of $p^*,q^*$, but we are not content with our understanding of the situations in which identifiability is and is not a serious problem.
-
-Here we present two suggestive case studies which we hope may inspire future research.  In both cases we are able to prove something of interest -- but not quite as much as we might hope.
+Rows correspond to different sampling procedures and columns correspond to a different measurement outcome.
+Now let $q^*(y|x)=C_{x y}$, another matrix.
+The Markov Link Method assumption then tells us that $A\times C=B$, where $\times$ indicates matrix multiplication.
+This corresponds to $\Omega_\ell \times \Omega_Y = 4$ equations.
+We also have a normalizing constraint that $\sum y q^*(y|x)=1$, which creates $\Omega_X=3$ additional equations.
+However, these normalizing constraints actually make two of the MLM assumption constraints redundant.
+In the end, we have $5$ constraining equations on the matrix $C$.
+However, the matrix $C$ contains six numbers.
+The result is a degree of freedom in $C$, corresponding to an aspect of $q^*$ that we simply cannot resolve.
+For example, here are two choices of $C$ which are both consistent with the equation $A\times C=B$:
+\[
+C=\left(\begin{array}{cc}
+0 & 1\\
+1 & 0\\
+1 & 0
+\end{array}\right)\qquad C=\left(\begin{array}{cc}
+1 & 0\\
+0 & 1\\
+\frac{2}{3} & \frac{1}{3}
+\end{array}\right)
+\]
 
@@
 
-Let us consider the case that $\Omega_\ell = \{1,2,\cdots k\}$ and $\Omega_X=\Omega_Y=\{1,2\cdots,2^{k-1}\}$.
+Consider the case that $\Omega_\ell = k$ and $\Omega_X=\Omega_Y=2^{k-1}$.
 That is, there are $k$ separate subpopulations, and both tool I and tool II can return one of $2^{k-1}$ possible values.
@@
 \]
-That is, the $x$th column of the first $k-1$ rows is simply the binary expansion of the number $x-1$, and the last row alternates $1$s and $0$s.
+That is, the $x$th column of the first $k-1$ rows is the binary expansion of the number $x-1$, and the last row alternates $1$s and $0$s.
 Now let us say we have perfect knowledge of $p^*(x|\ell)$ and $h^*(y|\ell)=\sum_x p^*(y|\ell)q^*(y|x)$.
@@
 \]
-This is an example of our problem with $k$ one smaller.
+This is an example of the same problem we started with -- except with $k$ one smaller.
 Applying our inductive hypothesis, we may thus obtain that $q(y|x)=q^*(y|x)$ for the first $2^{k-2}$ values of $x,y$.
@@
 
-%%%%%%%% TODO!
-% {\huge still need to prove this}
-
 However, it turns out that the relationship between $q$ and $p^*,h^*$ is not robust in every situation.  In the next section we will see that it can in fact be quite discontinuous:
 
-\subsection{A discontinuity}
+\subsection{Discontinuity}
 
-Let us consider the case that $\Omega_\ell = \{1\}$ and $\Omega_X=\Omega_Y=\{1,2\}$.
-That is, there is only one population (no subpopulations) and both tool I and tool II can return one of $2$ possible values.
+Consider the case that $\Omega_\ell = 1$ and $\Omega_X=\Omega_Y=2$.
+That is, there is only one sampling procedure (no subpopulations) and both tool I and tool II can return one of $2$ possible values.
 We will now consider two possiblities:
@@
 It is worth noting that these kinds of problems essentially vanish if the true $q^*$ is bounded away from zero i.e.\ $q^*(y|x)>c$ for every $x,y$ for some $c>0$.
-If this holds, together with a certain linear independence assumption, we can guarantee that the true $q^*$ is close to the set $\Theta(\hat p,\hat h)$ when $\hat p,\hat h$ are good approximations to $p^*,h^*$ (recall that we defined $\Theta$ in Equation (\ref{eq:thetadef})).
-In particular:\vspace{.1in}
+If this holds, together with a certain linear independence assumption, we can guarantee that the true $q^*$ is close to the set $\Theta(\hat p,\hat h)$ when $\hat p,\hat h$ are good approximations to $p^*,h^*$, where
+\[
+\Theta(p,h) \triangleq \left\{q\in T:\ \sum_x p(x|\ell)q(y|x)= h(y|\ell) \quad \forall \ell,y \right\}
+\]
+and by $T$ we mean the transition matrix polytope, $T=\{q:\ q(y|x)\geq 0, \sum_y q(y|x)=1\}$.  In particular:\vspace{.1in}
 
@@
 Fix any $p^*$ and define $h^*(y|\ell)=\sum_x p^*(x|\ell)q^*(y|x)$.
-Then by taking any $\hat p,\hat h$ sufficiently close to $p^*,h^*$ we can ensure that $q^*$ is arbitrarily close to some point in the set
-\begin{align}
-\hat \Theta \triangleq \left\{q\in T:\ \sum_x \hat p(x|\ell)q(y|x)= \hat  h(y|\ell) \quad \forall \ell,y \right\}
-\end{align}
-where by $T$ we mean the transition matrix polytope, $T=\{q:\ q(y|x)\geq 0, \sum_y q(y|x)=1\}$.
+Then by taking any $\hat p,\hat h$ sufficiently close to $p^*,h^*$ we can ensure that $q^*$ is arbitrarily close to some point in the set $\hat \Theta = \Theta(\hat p,\hat h)$.
 \end{thm}
@@
 However, the positivity requirement is quite troubling.
-In many cases of interest the true calibration $q^*$ has genuine zeros: pairs of measurements between the two tools which are fundamentally incompatible.
+In many cases of interest the true link $q^*$ has genuine zeros: pairs of measurements between the two tools which are fundamentally incompatible.
 In this case such a theorem cannot be applied.

### 2018-08-07 markovlink/__init__.py

@@ -1,12 +1,11 @@
 from . import estimation
-from . import exploration
 from . import mlm
 from . import misc
-from . import minorization
-from . import logslines
+from . import globalconf
+from . import localconf
 from . import polytopes
-from . import centering
+from . import approxbayes
 
-point_estimates = mlm.point_estimates
-uncertainty_assessment = mlm.uncertainty_assessment
-hypothesis_test = mlm.hypothesis_test
+estimates=approxbayes.estimates
+totalvardist=misc.totalvardist
+pnn=misc.pnn 

### 2018-08-07 markovlink/approxbayes.py

@@ -1,0 +1,94 @@
+from . import estimation
+from . import misc
+import numpy as np
+import numpy.random as npr
+from . import polytopes
+
+def posterior_samples(Nlx,Nly,nsamps=100,verbose=False,clipthresh=1e-6):
+    estimates=[]
+    if verbose:
+        misc.pnn("\n %d samples to produce:"%nsamps)
+    for i in range(nsamps):
+        if verbose:
+            misc.pnn(i)
+        p=np.array([npr.dirichlet(x+1) for x in Nlx])
+        htilde=np.array([npr.dirichlet(x+1) for x in Nly])
+
+        qtilde=estimation.train_mlm_fixedp(htilde,p)
+
+        # get the central q, for uniqueness
+        q = polytopes.find_central(p,qtilde,clipthresh=1e-6)
+        
+        estimates.append((p,q))
+
+    return estimates
+
+
+def credible_interval(estimates,xtilde,ytilde,alpha=.05,verbose=False):
+    lower,upper=get_extremes(estimates,xtilde,ytilde,verbose=verbose)
+
+    lower=np.percentile(lower[:,xtilde,ytilde],100*(alpha*.5))
+    upper=np.percentile(upper[:,xtilde,ytilde],100*(1-alpha*.5))
+
+    return np.clip(lower,0,1),np.clip(upper,0,1)
+
+def get_extremes(estimates,xtilde,ytilde,verbose=False):
+    szL,szX=estimates[0][0].shape
+    szX,szY=estimates[0][1].shape
+
+    lower=np.ones((len(estimates),szX,szY))
+    upper=np.ones((len(estimates),szX,szY))
+
+    good=np.ones(len(estimates),dtype=np.bool)
+
+    for i,(ptilde,qtilde) in enumerate(estimates):
+        if verbose:
+            misc.pnn(i)
+
+        direc=np.zeros((szX,szY))
+        direc[xtilde,ytilde]=1
+
+        lower[i]=polytopes.find_extremal(ptilde,qtilde,direc)
+        upper[i]=polytopes.find_extremal(ptilde,qtilde,-direc)
+
+    return lower,upper
+
+def credible_intervals(estimates,xys,alpha=.05,verbose=False,onfail='throw'):
+    CIs=np.ones((len(xys),2))*np.nan
+
+    if verbose:
+        misc.pnn("\n %d parameters look at:"%len(xys))
+
+    for i,(x,y) in enumerate(xys):
[…]

### 2018-08-07 markovlink/estimation.py

@@ -5,6 +5,211 @@ import scipy.special
 import itertools
 from . import misc
 import time
+import numpy.random as npr
 
-def pseudocount_point_estimate(Nlx,pseudocount=1.0):
-    return (Nlx+pseudocount)/np.sum(Nlx+pseudocount,axis=1,keepdims=True)
+def initialconds(Nlx,Nly,xytilde=None,pseudocount=1.0):
+    szL,szX=Nlx.shape
+    szL,szY=Nly.shape
+
+    phat=(Nlx+pseudocount)/np.sum(Nlx+pseudocount,axis=1,keepdims=True)
+
+    qhat=np.ones((szX,szY))/szY
+
+    if xytilde is not None:
+        xtilde,ytilde,val = xytilde
+        qhat[xtilde]=(1-val)/(szY-1)
+        qhat[xtilde,ytilde]=val
+
+    return phat,qhat
+
+def LL(Nlx,Nly,p,h):
+    '''
+    We return log likelihood of Nlx,Nly under p,h
+
+        sum(Nlx*log(p)) + sum(Nly*log(h))
+
+    with the convention that 0*log(0/0) = 0
+    '''
+
+    # compute the log likelihood for Nlx
+    good=(Nlx>0)
+    Nlxll = np.sum(Nlx[good]*np.log(p[good]))
+    
+    # compute the log likelihood for Nly    
+    good=(Nly>0)
+    Nlyll = np.sum(Nly[good]*np.log(h[good]))
+
+    # done!
+    return Nlxll+Nlyll
+
+def random_estimate(Nlx,Nly,niter=100):
+    '''
+    - Grabs a sample from the (Dirichlet) posterior on p,h
+    - Projects to MLM
+    '''
+
+    p=np.array([npr.dirichlet(x+1) for x in Nlx])
+    hhat=np.array([npr.dirichlet(x+1) for x in Nly])
+
+    return p,train_mlm_fixedp(hhat,niter,p)
+
+def random_estimates(Nlx,Nly,niter=100,nsamps=100,alpha=.05):
+    parms=[random_estimate(Nlx,Nly,niter=niter) for i in range(nsamps)]
+    lls = np.array([LL(Nlx,Nly,p,p@q) for (p,q) in parms])
+
+    good=np.where(lls>=np.percentile(lls,alpha*100))[0]
+
+    return [parms[x] for x in good]
+
[…]

### 2018-08-07 markovlink/exploration.py

@@ -1,83 +1,1 @@
-'''
-Some tools for bootstrap and exploring the polytopes of interest
-'''
-
-
-import scipy as sp
-import scipy.optimize
-import numpy as np
-import numpy.random as npr
-import sys
-
-from . import misc
-from . import minorization
-
-def construct_interval_samples(Nlx,Nly,n_extremals=50,alpha=.05,verbose=True):
-    '''
-    Construct extremal samples to help us explore the confidence interval
-    '''
-    szL,szX=Nlx.shape
-    szL,szY=Nly.shape
-
-    Ms=np.zeros((n_extremals,szX,szY))
-
-    if verbose:
-        misc.pnn("%d samples:"%n_extremals)
-    for i in range(n_extremals):
-        if verbose:
-            misc.pnn(i)
-
-        # random initial conditions
-        p=np.array([npr.dirichlet(np.ones(szX)) for l in range(szL)])
-        q=np.array([npr.dirichlet(np.ones(szY)) for l in range(szX)])
-
-        # random direction of search
-        c=npr.exponential(size=(szX,szY))
-
-        # find the most extreme q in the confidence interval in that direction
-        p,q=minorization.find_extremal_in_confidence_interval(Nlx,Nly,p,q,c,alpha)
-        Ms[i]=q
-
-    return Ms
-
-
-
-##################################
-########## TEST TRAIN SPLIT ######
-##################################
-
-
-def train_test_split(Nly,m):
-    '''
-    Input:
-    - an array of positive integers, Nly
-
-    Output:
-    - an array train,test
-
-    For each row, we hold out m of the counts from Nly
-    '''
-
-    assert (np.sum(Nly,axis=1)>=m).all(),"Not all rows of Nly have %d total count; can't hold out that much"%m
-
-    Nly=np.require(Nly)
-    Nshp=Nly.shape
-
-    # sample (l,Y) events, uniformly without replacement
-    # from the (l,Y) events indicated by Nly
-
-    # we iteratively select a random point from Nly
-    # and move it into our new datset
-    Nly=Nly.copy().ravel() # this will hold train
-    Nnew = np.zeros(len(Nly)) # this will hold test
-    for l in range(Nshp[0]):
-        for i in range(m):
-            # choose an event in the lth row of Nly to sample
[…]

### 2018-08-07 markovlink/globalconf.py

@@ -1,0 +1,253 @@
+import numpy as np
+import sys
+import scipy as sp
+import scipy.stats
+import scipy.special
+import itertools
+from . import misc
+import time
+from . import logslines
+from . import polytopes
+import numpy.random as npr
+from . import estimation
+
+
+
+
+#  _ _                        _                __ 
+# | | |_ __    __ _ _ __   __| |  _ __  _ __  / _|
+# | | | '__|  / _` | '_ \ / _` | | '_ \| '_ \| |_ 
+# | | | |    | (_| | | | | (_| | | |_) | |_) |  _|
+# |_|_|_|     \__,_|_| |_|\__,_| | .__/| .__/|_|  
+#                                |_|   |_|        
+
+def LLR(Nlx,Nly,p,h):
+    '''
+    log likelihood ratio:
+
+        LLR =  sum(Nlx*log(phat/p)) + sum(Nly*log(hhat/h))
+
+    with the convention that 0*log(0/0) = 0
+    '''
+
+    # compute empirical averages
+    phat = Nlx/np.sum(Nlx,axis=1,keepdims=True)
+    hhat = Nly/np.sum(Nly,axis=1,keepdims=True)
+
+    # compute the log likelihood ratios for Nlx
+    good=(Nlx>0)
+    Nlxratio = np.sum(Nlx[good]*np.log(phat[good]/p[good]))
+    
+    # compute the log likelihood ratios for Nly    
+    good=(Nly>0)
+    Nlyratio = np.sum(Nly[good]*np.log(hhat[good]/h[good]))
+
+    # done!
+    return Nlxratio+Nlyratio
+
+def estimate_PPF_chisq(alpha,phat,hhat,n_samps=1000):
+    '''
+    Use asymptotic normality to get an approximate estimate for the value of PPF so that 
+
+        P(LLR(Nlx,Nly,phat,hhat) > PPF) = alpha
+
+    In a way that doesn't depend upon phat,hhat
+
+    '''
+
+    szL,szX=phat.shape
+    szL,szY=hhat.shape
+
+    dof=szL*(szX+szY-2)
+
+    return sp.stats.chi2.ppf(alpha,dof)
+
+
+def estimate_PPF(alpha,splX,splY,phat,hhat,n_samps=1000):
+    '''
+    Use data to estimate for the value of PPF so that 
+
+        P(LLR(Nlx,Nly,phat,hhat) > PPF) = alpha
+
+    when Nlx is drawn from phat (with splX[ell] samples per subpop ell)
[…]

### 2018-08-07 markovlink/localconf.py

@@ -1,0 +1,178 @@
+import numpy as np
+import sys
+import scipy as sp
+import scipy.stats
+import scipy.special
+import itertools
+from . import misc
+import time
+from . import logslines
+from . import polytopes
+import numpy.random as npr
+from . import estimation
+from . import globalconf
+import functools
+import bisect
+
+
+
+
+###############################################
+#  _ _                        _                __ 
+# | | |_ __    __ _ _ __   __| |  _ __  _ __  / _|
+# | | | '__|  / _` | '_ \ / _` | | '_ \| '_ \| |_ 
+# | | | |    | (_| | | | | (_| | | |_) | |_) |  _|
+# |_|_|_|     \__,_|_| |_|\__,_| | .__/| .__/|_|  
+#                                |_|   |_|        
+###############################################
+
+def LL_prof(Nlx,Nly,niter,xytilde,pseudocount=1.0):
+    '''
+    We return (approximate) max log likelihood of Nlx,Nly of
+
+        sum(Nlx*log(p)) + sum(Nly*log(h))
+
+    with the convention that 0*log(0/0) = 0, subject
+    to the constraint that q[xtilde,ytilde]=val
+    '''
+
+    szL,szX=Nlx.shape
+    szL,szY=Nly.shape
+    xtilde,ytilde,val=xytilde
+
+
+    # get initial guess
+    p,q = estimation.initialconds(Nlx,Nly,xytilde,pseudocount=pseudocount)
+
+    # train
+    secondbestp,secondbestq=estimation.train_mlm_fixedpqxy(Nly,niter,xytilde,p,q=q)
+
+    # get ll
+    ll=estimation.LL(Nlx,Nly,secondbestp,secondbestp@secondbestq)
+
+    # done!
+    return secondbestp,secondbestq,ll
+
+def estimate_PPF(alpha,xtilde,ytilde,splX,splY,pstar,qstar,niter,n_samps=100):
+    '''
+    Say pstar,qstar is true.
+
+    Say you are going to produce a confidence interval defined by
+
+                C = {val: LL_prof(bestval)-LL_prof(val) < PPF}
+
+    What value of kk should you pick so that with probability (1-alpha) the
+    confidence interval contains the low and hi possibilities defined by pstar,qstar?
+
[…]

### 2018-08-07 markovlink/minorization.py

@@ -1,275 +1,0 @@
-import numpy as np
-import sys
-import scipy as sp
-import scipy.stats
-import scipy.special
-import itertools
-from . import misc
-import time
-from . import logslines
-from . import polytopes
-import numpy.random as npr
-
-class OutOfTime(Exception):
-    pass
-class OutOfIterations(Exception):
-    pass
-
-def LL_ph(Nlx,Nly,p,h):
-    '''
-    We return log likelihood of Nlx,Nly under p,h
-
-        sum(Nlx*log(p)) + sum(Nly*log(h))
-
-    with the convention that 0*log(0/0) = 0
-    '''
-
-
-    # compute the log likelihood for Nlx
-    good=(Nlx>0)
-    Nlxratio = np.sum(Nlx[good]*np.log(p[good]))
-    
-    # compute the log likelihood for Nly    
-    good=(Nly>0)
-    Nlyratio = np.sum(Nly[good]*np.log(h[good]))
-
-    # done!
-    return Nlxratio+Nlyratio
-
-def LLR(Nlx,Nly,p,h):
-    '''
-    LLR =  mean(Nlx*log(phat/p)) + mean(Nly*log(hhat/h))
-
-    with the convention that 0*log(0/0) = 0
-    '''
-
-    # compute empirical averages
-    phat = Nlx/np.sum(Nlx,axis=1,keepdims=True)
-    hhat = Nly/np.sum(Nly,axis=1,keepdims=True)
-
-    # compute the log likelihood ratios for Nlx
-    good=(Nlx>0)
-    Nlxratio = np.mean(Nlx[good]*np.log(phat[good]/p[good]))
-    
-    # compute the log likelihood ratios for Nly    
-    good=(Nly>0)
-    Nlyratio = np.mean(Nly[good]*np.log(hhat[good]/h[good]))
-
-    # done!
-    return Nlxratio+Nlyratio
-
-def estimate_PPF(Nlx,Nly,alpha,n_samps=1000):
-    '''
-    Get an approximate estimate for the value of PPF so that 
-
-        P(LLR(Nlx,Nly,pstar,hstar) > PPF) = alpha
-
-    when Nly is drawn from pstar 
-         Nly is drawn from hstar
-
-    '''
-
-    # we assume the TRUTH is given by pseudocounted empirical distributions
-    # (the pseudocounts end up giving us a more conservative estimator)
-    # we then look at samples from the distribution of LLR
-    samps=PPF_samples(Nlx,Nly)
-
-    # we estimate our object of interest based on these samples
[…]

### 2018-08-07 markovlink/misc.py

@@ -4,6 +4,22 @@ import sys
 import scipy as sp
 import numpy.random as npr
 
+def newdataset(Nlx,Nly,p,h):
+    '''
+    Constructs a new dataset based on the number of samples in Nlx,Nly
+    and the distributions in p,h
+    '''
+
+    szL,szX=Nlx.shape
+    szL,szY=Nly.shape
+
+    Nlx2=np.array([npr.multinomial(np.sum(Nlx[ell]),p[ell]) for ell in range(szL)])
+    Nly2=np.array([npr.multinomial(np.sum(Nly[ell]),h[ell]) for ell in range(szL)])
+
+    return Nlx2,Nly2
+
+def empiricals(Nlx,Nly):
+    return Nlx/np.sum(Nlx,axis=1,keepdims=True),Nly/np.sum(Nly,axis=1,keepdims=True)
 
 def pnn(x): 
     sys.stdout.write(str(x) + " ")
@@ -50,4 +66,40 @@ def sample(N,p):
     for z in range(len(N)):
         tmx2[z]=npr.multinomial(np.sum(N[z]),p[z])
     return tmx2
+
+
+def train_test_split(Nly,m):
+    '''
+    Input:
+    - an array of positive integers, Nly
+
+    Output:
+    - an array train,test
+
+    For each row, we hold out m of the counts from Nly
+    '''
+
+    assert (np.sum(Nly,axis=1)>=m).all(),"Not all rows of Nly have %d total count; can't hold out that much"%m
+
+    Nly=np.require(Nly)
+    Nshp=Nly.shape
+
+    # sample (l,Y) events, uniformly without replacement
+    # from the (l,Y) events indicated by Nly
+
+    # we iteratively select a random point from Nly
+    # and move it into our new datset
+    Nly=Nly.copy().ravel() # this will hold train
+    Nnew = np.zeros(len(Nly)) # this will hold test
+    for l in range(Nshp[0]):
+        for i in range(m):
+            # choose an event in the lth row of Nly to sample
+            idxs=np.r_[0:Nshp[1]]+Nshp[1]*l
+            idx=npr.choice(idxs,size=1,p=Nly[idxs]/np.sum(Nly[idxs]))[0]
+            Nly[idx]=Nly[idx]-1
+            Nnew[idx]=Nnew[idx]+1
+    Nnew=Nnew.reshape(Nshp)
+
+    return Nly.reshape(Nshp),Nnew.reshape(Nshp)
+
 

### 2018-08-07 markovlink/mlm.py

@@ -2,36 +2,10 @@ from . import estimation
 from . import exploration
 from . import misc
 import numpy as np
-from . import minorization
+import numpy.random as npr
+from . import globalconf
+from . import localconf
+from . import polytopes
+from . import approxbayes
 
-def hypothesis_test(Nlx,Nly,alpha=.05,maxtime=120):
-    p,q=point_estimates(Nlx,Nly)
-    PPF=minorization.LLR(Nlx,Nly,p,p@q)
-    PPF = minorization.estimate_PPF(Nlx,Nly,alpha)
 
-    p,q,niter=minorization.find_feasible_point(Nlx,Nly,PPF,maxtime=maxtime)
-
-    return "fail to reject"
-
-def point_estimates(Nlx,Nly,maxtime=10):
-    '''
-    Get point estimates for the calibration
-    '''
-
-    # find a q inside an approximate 95% confidence interval
-    PPF = minorization.estimate_PPF(Nlx,Nly,.05)
-    p,q,niter=minorization.find_feasible_point(Nlx,Nly,PPF,maxtime=maxtime)
-
-    # keep seeking a q that better optimizes the likelihood of the data,
-    # for twice the number of iterations it took to find the feasible point:
-    p,q = minorization.train_mlm_minorized(Nlx,Nly,p,q,0,1,niter*2)
-
-    return p,q
-
-def uncertainty_assessment(Nlx,Nly,alpha=.05,verbose=True,n_extremals=50):
-    '''
-    Get a measure of our uncertainty about the calibration, in the form of a collection of 
-    calibrations lying within an approximate (1-alpha) confidence interval
-    '''
-
-    return exploration.construct_interval_samples(Nlx,Nly,n_extremals=n_extremals,alpha=alpha,verbose=verbose)

### 2018-08-07 markovlink/polytopes.py

@@ -9,6 +9,9 @@ import numpy.random as npr
 import sys
 from . import misc
 
+class LinearProgrammingFailure(Exception):
+    pass
+
 def precondition_constraint(A,b,tol=1e-10,return_preconditioner=False):
     '''
     Transforms the constraint 
@@ -38,51 +41,6 @@ def precondition_constraint(A,b,tol=1e-10,return_preconditioner=False):
         return V, (U.T@b)[:len(e)]/e
 
 
-def project_to_polytope(A,b,y,feasiblepoint=None):
-    '''
-    Project y to the polytope
-
-    Ax=b
-    x>=0
-    '''
-
-    '''
-    cvxopt wants this in the form
-
-    qp(P, q, G=None, h=None, A=None, b=None)
-
-    minimize    (1/2)*x'*P*x + q'*x 
-    subject to  G*x <= h      
-                A*x = b.
-
-    In our case we have 
-
-    - G=-I_n
-    - h=-0_n
-    - A =A
-    - b=b
-    - P=I_n
-    - q = -y
-
-    ''' 
-
-    import cvxopt
-
-    szM,szN=A.shape
-    com=lambda x: cvxopt.matrix(x)
-    
-    ident=np.eye(szN)
-    h=np.zeros((szN,1))
-    q=-y.reshape((-1,1))
-
-
-    if feasiblepoint is not None:
-        rez = cvxopt.solvers.qp(com(ident),com(q),com(-ident),com(h),com(A),com(b),
-            initvals=dict(x=com(feasiblepoint)))
-    else:
-        rez = cvxopt.solvers.qp(com(ident),com(q),com(-ident),com(h),com(A),com(b))
-
-    return rez
 
 
 def q_with_fixedph(plx,hly,precondition=True,preconditiontol=1e-10):
@@ -211,7 +169,7 @@ def qh_with_fixed_p(p,szY,precondition=True,preconditiontol=1e-10):
 ########## EXTREMAL FINDING ######
 ##################################
 
-def find_extremal(plx,qxy,c,use_cvxopt=True):
+def find_extremal(plx,qxy,c,use_cvxopt=True,precondition=False):
     '''
     Consider the polytope of transition matrices q such that plx @ q = plx @ qxy.
     Find the most extremal vertex of this polytope in the direction of c.
@@ -224,7 +182,7 @@ def find_extremal(plx,qxy,c,use_cvxopt=True):
     assert c.shape[0]==np.prod(qxy.shape),"c should be the same shape as qxy"
     
[…]

### 2018-08-07 paper/paper.tex
credible/confidence typo

@@
 \hfill{}\figgygrs{.4}{../images/simulationI}{(a) Estimation error, \\ varying number of samples \\ and number of sampling procedures.}\hfill{}
-\hfill{}\figgygrs{.5}{../images/simulationIconf}{(b) Confidence intervals, \\ varying number of samples \\ and number of sampling procedures.}\hfill{}
+\hfill{}\figgygrs{.5}{../images/simulationIconf}{(b) Credible intervals, \\ varying number of samples \\ and number of sampling procedures.}\hfill{}
 
@@
 In subfigure (b) we consider the MLM's credible intervals.
-For a selection of trials we choose a random value of of $x,y$ and then compare the true link parameter $q^*(y|x)$ with the MLM confidence interval $C_{x,y}$.
+For a selection of trials we choose a random value of of $x,y$ and then compare the true link parameter $q^*(y|x)$ with the MLM interval $C_{x,y}$.
 We show the interval (in black) centered around the ground truth (in red).
@@
 However, to make the error actually converge to zero we need at least six distinct sampling procedures.
-This figure also shows that the confidence intervals work correctly regardless of of the number of samples or sampling procedures; they include the ground truth with high probability.
+This figure also shows that the intervals work correctly regardless of of the number of samples or sampling procedures; they include the ground truth with high probability.
 With many samples and sampling procedures, the intervals are small and concentrated around the truth.
@@
 \hfill{}\figgygr{0.9}{../images/allenout}\hfill{}
-\caption{{\bf MLM estimation of the link, with confidence interval.}  The central plot shows a portion of the MLM estimator applied to the Allen Institute data.
+\caption{{\bf MLM estimation of the link, with credible interval.}  The central plot shows a portion of the MLM estimator applied to the Allen Institute data.
 We look at a subselection of Standard types ($x$) and all of the Patch types ($y$).
@@
 For some parameters our intervals are much tighter than others.
-For example, it appears we have high confidence the Standard type `Vip Rspo4' is highly associated with the Patch type `Vip.'  By constrast, we have almost no idea what is associated with the Standard type `Meis2.'  The upper confidence interval bounds suggest that $q^*(y|\mathrm{Meis2})$ could be nearly 1 for many different Patch types.
+For example, it appears we have high confidence the Standard type `Vip Rspo4' is highly associated with the Patch type `Vip.'  By constrast, we have almost no idea what is associated with the Standard type `Meis2.'  The upper credible interval bounds suggest that $q^*(y|\mathrm{Meis2})$ could be nearly 1 for many different Patch types.
 Obviously it cannot be 1 for all of those types simultaneously, since $\sum_y q^*(y|\mathrm{Meis2})=1$, but the data simply doesn't tell us which $y$ carries the mass.
@@
 
-We applied the MLM to this data to obtain point estimates $\tilde q(y|x)$ and confidence intervals $C_{x y}$.
+We applied the MLM to this data to obtain point estimates $\tilde q(y|x)$ and credible intervals $C_{x y}$.
 In Figure \ref{fig:allenboot} we visualize these objects for selected values of $x,y$.
@@
 
-The variability in the confidence regions suggests what we need to do in order to more closely determine the value of the calibration.
+The variability in the credible regions suggests what we need to do in order to more closely determine the value of the calibration.
 For example, if we could develop more unique sampling strategies which will include `Lamp5 Egln 1' cells, this would help us resolve our ambiguity about this aspect of the link.
@@
 We saw that some aspects of the two systems seem well-calibrated, but others we are less sure about.
-The nature of the variability in the confidence intervals suggested directions for experiments which would further refine our understanding.
+The nature of the variability in the credible intervals suggested directions for experiments which would further refine our understanding.
 
 In future work, it would be interesting to use additional assumptions to help us estimate calibrations.
-For example, as it stands the Markov Link Method will only yield narrow confidence intervals if each measurement tool returns one of a finite number of measurement outcomes.
+For example, as it stands the Markov Link Method will only yield narrow credible intervals if each measurement tool returns one of a finite number of measurement outcomes.
 In some cases, we may believe that similar measurement values should have similar probabilities.

### 2018-08-07 paper/paper.tex
types typos typos

@@
 For each $\tilde P,\tilde H$, let $E(\tilde P,\tilde H)$ denote the minimum value of $D(\tilde H|h')$ among all $h'$ such that there exists $q$ satisfying $h'(y|\ell)=\sum_x \tilde P(y|\ell)q(y|x)$.
-Let $P,H$ denote variables carrying the distribution of $\tilde P,\tilde H$ after conditioned on $E(\tilde P,\tilde H)=0$.
+Let $P,H$ denote variables carrying the distribution of $\tilde P,\tilde H$ after conditioning on $E(\tilde P,\tilde H)=0$.
 Let $\Theta(p,h)=\{q:\ h(y|\ell) = \sum_x p(x|\ell)q(y|x)\}$.
@@
 Our approach arises naturally from a Bayesian point of view.
-To understand them, first notice that there are several `nuisance parameters' we must consider in order to estimate the link $q^*$.
+To understand how, first notice that there are several `nuisance parameters' we must consider in order to estimate the link $q^*$.
 Let $p^*(x|\ell)$ denote the probability that a cell sampled with procedure $\ell$ will have transcriptomic type $x$ and $h^*(y|\ell) \triangleq \sum_x p^*(x|\ell)q(y|x)$ denote the probability such a cell will have morphological type $y$.

### 2018-08-07 markovlink/polytopes.py

@@ -265,14 +265,14 @@ def cvxopt_linprog(c,A_eq,b_eq,initialcondition,gaptol=1e-5,verbose=True):
 
     p("cond=%f"%np.linalg.svd(A_eq)[1].min())
 
-    # first try glpk
-    p("gpkl")
-    rez = cvxopt.solvers.lp(com(c),com(G),com(g_eq),A=com(A_eq),b=com(b_eq),
-                            primalstart=dict(x=com(initialcondition),s=com(initialcondition)),
-                           solver='glpk')
-    if rez['status']=='optimal':
-        p("suc")
-        return np.array(rez['x'])
+    # # first try glpk
+    # p("gpkl")
+    # rez = cvxopt.solvers.lp(com(c),com(G),com(g_eq),A=com(A_eq),b=com(b_eq),
+    #                         primalstart=dict(x=com(initialcondition),s=com(initialcondition)),
+    #                        solver='glpk')
+    # if rez['status']=='optimal':
+    #     p("suc")
+    #     return np.array(rez['x'])
 
     # okay glpk didn't work.  try cvxopt's coneopt
     p("coneopt")

### 2018-08-10 paper/paper.tex
More elaborate explanation of method

@@
 %
-To produce these objects we proceeds as follows.
-Let $\tilde P(\cdot |\ell) \sim \mathrm{Dirichlet}(N^X_\ell+1)$ and $\tilde H(\cdot |\ell) \sim \mathrm{Dirichlet}(N^Y_\ell+1)$ for each $\ell$.
-For any pair $h,h'$ let
+To produce these estimates for the link, we take a bayesian point of view.
+It might seem sensible to start by specifying some prior beliefs about the link $q^*$.
+However, the link is in many cases `nonidentifiable';  we describe this problem in detail in Appendix \ref{sec:casestudies}.
+One consequence of nonidentifiability is that standard bayesian inferences would have no robustness to our precise choice of prior beliefs.
+To avoid these problems, we focus on quantities which can be identified from the data.
+In particular,
+\begin{itemize}
+    \item Let $p^*(x|\ell)$ denote the probability that a cell sampled with procedure $\ell$ will have transcriptomic type $x$.  These probabilities can be estimated from the data, $N^X$.
+    \item Let $h^*(y|\ell) \triangleq \sum_x p^*(x|\ell)q^*(y|x)$ denote the probability that a cell sampled with procedure $\ell$ will have morphological type $y$.
+These probabilities can be estimated from the data, $N^Y$.
+    \item Let $\Theta(p,h)=\{q:\ h(y|\ell) = \sum_x p(x|\ell)q(y|x)\}$ denote the set of links which are consistent with $p,h$.
+Notice that $q^*\in\Theta(p^*,h^*)$, but there may be other links $q\in\Theta(p^*,h^*)$.
+Any such link yields the same distribution on what we can observe, $N^X,N^Y$.
+Therefore there is simply no way to use data to distinguish among them.
+This is the essence of the nonidentifiability problem mentioned above.
+So, instead of trying to estimate this inestimable quantity, we will focus on three other quantities which can be identified from data.
+Let
+    \begin{itemize}
+        \item $q_{\mathrm{lo},x,y} \triangleq \min_{q\in \Theta(p^*,h^*)} q(y|x)$
+        \item $q_{\mathrm{hi},x,y} \triangleq \max_{q\in \Theta(p^*,h^*)} q(y|x)$ 
+        \item $q_{\mathrm{mid}} \triangleq \argmin_{q\in \Theta(p^*,h^*)} \sum_{x,y} q(y|x)^2$
+    \end{itemize}
+    Note that all three of these optimization problems have unique solutions and would be straightforward to solve if $p^*,h^*$ were known; optimization details can be found in Appendix \ref{sec:optproblems}.
+Morever, even when $q^*$ is nonidentifiable, we can still be assured that $q_{\mathrm{lo}}, q_{\mathrm{hi}}$ are identifiable and $q_{\mathrm{lo},x,y} \leq q^*(y|x) \leq q_{\mathrm{hi},x,y}$.
+We can also hope that $q_{\mathrm{mid}}$ will provide a single estimate which strikes a middle ground; in particular, $q_{\mathrm{mid}}$ is the link value which is closest to uniform among all links which are consistent with $p^*,h^*$.
+\end{itemize}  
+To summarize: it may not be possible to estimate $q^*$ from data, but we can always estimate $p^*,h^*$.
+We can also compute deterministic functions of $p^*,h^*$ which bound $q^*$, namely $q_{\mathrm{lo}},q_{\mathrm{hi}}$.
+We take a Bayesian point of view in estimating $p^*,h^*$.
+For prior beliefs, we assume a noninformative uniform prior $\mathbb{P}$, namely
 \[
-D(h|h') = \sum_{\ell,y} h(y|\ell) \log \frac{h(y|\ell)}{h'(y|\ell)}
+\mathbb{P}(p^*,h^*) \propto 1
 \]
-denote the Kullback-Leibler divergence from $h$ to $h'$.
-For each $\tilde P,\tilde H$, let $E(\tilde P,\tilde H)$ denote the minimum value of $D(\tilde H|h')$ among all $h'$ such that there exists $q$ satisfying $h'(y|\ell)=\sum_x \tilde P(y|\ell)q(y|x)$.
-Let $P,H$ denote variables carrying the distribution of $\tilde P,\tilde H$ after conditioning on $E(\tilde P,\tilde H)=0$.
-Let $\Theta(p,h)=\{q:\ h(y|\ell) = \sum_x p(x|\ell)q(y|x)\}$.
-We then define the MLM objects as follows:
+Following the bayesian philosophy, we then incorporate new knowledge by conditioning.
+We have two important pieces of knowledge we would like to incorporate.
+First, we have observed the data, $N^X,N^Y$.
+Second, we know from the MLM assumption that there exists \emph{some} value $q^*$ which is consistent with $p^*,h^*$.
+To make the second idea formal, let $D(h|h')$ denote the Kullback-Leibler divergence and $G(p,h) = \min_{h':\ \Theta(p,h')\neq \emptyset} D(h|h')$.
+$G(p,h)=0$ is a numerical way to indicate that the MLM assumption can hold.
+After incorporating this knowledge, we find ourselevs interested in the conditional distribution
+\[
+\mathbb{P}\left(p^*,h^* | N^X,N^Y,G(p^*,h^*)=0\right)
+\]
+We use this posterior to produce estimates for our quantities of interest:
 \begin{enumerate}
-    \item $\hat q\triangleq\mathbb{E}[\argmin_{q\in \Theta(P,H)} \sum_{x,y} q(y|x)^2]$.
-    \item The lower bound of $C_{x,y}$ is defined as the largest number $c$ such that $\min_{q\in \Theta(P,H)} q(y|x) \geq c$ with probability 97.5\%.
-Likewise the upper bound is the smallest number $c$ such that $\max_{q\in \Theta(P,H)} q(y|x) \leq c$ with probability 97.5\%.
+    \item $\hat q$ is calculated using posterior expectation:
+    \[
+    \hat q \triangleq \mathbb{E}\left[q_{\mathrm{mid}} | N^X,N^Y,G(p^*,h^*)=0\right]
+    \]
+    %
+    \item $C_{x,y}$ is calculated using credible intervals.  For each $x,y$, we define $C_{x,y}$ to be the largest interval such that 
+    \begin{align*}
+    \mathbb{P}\left(q_{\mathrm{lo},x,y} \in C_{x,y} | N^X,N^Y,G(p^*,h^*)=0\right)\geq 97.5\% \\
+    \mathbb{P}\left(q_{\mathrm{hi},x,y} \in C_{x,y} | N^X,N^Y,G(p^*,h^*)=0\right)\geq 97.5\% 
+    \end{align*}
 \end{enumerate}
-%
 Code to compute these objects is published at https://github.com/jacksonloper/markov-link-method, including a tutorial-style ipython notebook detailing every computation made in this paper.
 
-Our approach arises naturally from a Bayesian point of view.
-To understand how, first notice that there are several `nuisance parameters' we must consider in order to estimate the link $q^*$.
-Let $p^*(x|\ell)$ denote the probability that a cell sampled with procedure $\ell$ will have transcriptomic type $x$ and $h^*(y|\ell) \triangleq \sum_x p^*(x|\ell)q(y|x)$ denote the probability such a cell will have morphological type $y$.
-Note that the distribution of the data $N^X,N^Y$ is entirely determined by $p^*,h^*$.
-In particular, if someone changed $q^*$ without changing $p^*,h^*$ there would be no way for us to know.
-If this can occur, $q^*$ is said to be `nonidentifiable.'   Therefore, instead of trying to directly estimate $q^*$ (which may be impossible to accurately estimate) we focus on three quantities which can be estimated from data:
-%
-\begin{align*}
-q^*_{\mathrm{mid}} &\triangleq \argmin_{q\in \Theta(p^*,h^*)} \sum_{x,y} q(y|x)^2\\
-q^*_{\mathrm{lo},x,y} &\triangleq \min_{q\in \Theta(p^*,h^*)} q(y|x) \\
-q^*_{\mathrm{hi},x,y} &\triangleq \max_{q\in \Theta(p^*,h^*)} q(y|x) 
-\end{align*}
-%
-Note that all three of these optimization problems have unique solutions and would be straightforward to solve if $p^*,h^*$ were known; optimization details can be found in Appendix \ref{sec:optproblems}.
-Moreover, regardless of any identifiability issues for $q^*$ itself, the objects $q^*_{\mathrm{lo},x,y}$ and $q^*_{\mathrm{hi},x,y}$ are always identifiable for every $(x,y)$ and define bounds on the true value of $q^*(y|x)$.
-
-The MLM algorithm therefore adopts a Bayesian point of view to estimate $q^*_{\mathrm{mid}}$ and create a credible interval for $q^*_{\mathrm{lo},x,y},q^*_{\mathrm{hi},x,y}$.
-We begin with uniform prior beliefs for the parameters of $p^*,h^*$.
-That is, our prior is that $p^*(\cdot|\ell),h^*(\cdot|\ell)$ carry $\mathrm{Dirichlet}(1)$ distributions for each $\ell$.
-Following the Bayesian philosophy, we would like to incorporate any additional knowledge through conditioning.
-We have two important pieces of knowledge we would like to incorporate: that the MLM assumption holds (i.e.\ there is some $q^*$ so that $\sum_x p^*(x|\ell)q^*(y|x)=h^*(y|\ell)$) and we have observed the data $N^X,N^Y$.
-By conditioning on this knowledge, we obtain the posterior distribution carried by the variables $P,H$ described above.
-In terms of these variables, the MLM estimator $\hat q$ is simply the posterior mean for $q^*_{\mathrm{mid}}$ and the MLM credible interval $C_{x,y}$ is an interval which contains $q^*_{\mathrm{lo},x,y},q^*_{\mathrm{hi},x,y}$ with high probability.
-
 In practice, we were not able to find a way to compute the credible interval and posterior means exactly.
-It seems difficult to even sample the conditional random variables $P,H$.
+It seems difficult to even sample from the conditional distribution.
 Common approaches to this type of problem involve Markov-Chain Monte Carlo and Variational Inference, but we were unable to make these approaches work in practice.
-It seems nontrivial to work with the condition $E(p,h)=0$ that formalizes the MLM assumption.
+It seems nontrivial to work with the condition $G(p,h)=0$ that formalizes the MLM assumption.
 We instead take a somewhat na\"ive approach.
-We start with $\tilde P,\tilde H$, which carry the distribution of $p^*,h^*$ conditioned on the observed data (but not conditioned on the fact that the MLM assumption holds).
+We start by defining $\tilde P,\tilde H$ to carry the conditional distribution of $p^*,h^*$ given the observed data (but not conditioned on the fact that the MLM assumption holds).
 We then simply project $\tilde H$ to the space of parameters which are consistent with the MLM assumption.
@@
 Optimization details can be found in Appendix \ref{sec:optproblems}.
-We hope that the distribution of $\tilde P,\hat H$ will approximate the desired posterior $P,H$.
+We hope that the distribution of $\tilde P,\hat H$ will approximate the desired posterior.
 We then use samples from this distribution to compute what we need.
-The implications of this approximation are difficult to reason about; we hope further study will clarify when this is a reasonable procedure.
-In the mean time, we content ourselves that it is a practical solution to a difficult problem.
-
-A frequentist point of view might provide a valuable counterpoint to our Bayesian analysis.
-However, we were unable to produce effective confidence intervals for the link parameters.
-There are at least two difficulties: the link is unidentifiable and the true parameter values may lie on the edge of the parameter space.
-These challenges make traditional appproaches based on asymptotic normality difficult to apply.
-Perhaps future work could uncover a way to overcome these challenges.
+We asymptotically expect that $\tilde P,\tilde H$ will nearly satisfy the MLM assumption in any case, so this approximation should not make a large difference.
+However, we caution that the exact implications of the approximation are difficult to reason about.
+In any case, it is a practical solution to a difficult problem.
 
@@
 
-\section{Simulations}
+\section{Mathematical Results and Simulations}
 
@@
 
-We used simulations to test the two parts of the MLM: the point estimators $\hat q(y|x)$ and credible intervals $C_{x,y}$.  We would ordinarily hope for two things:
+There are three desiderata we might hope the Markov Link Method estimators would achieve:
 \begin{description}
     \item[Estimator convergence] As we obtain more samples, $\hat q(y|x)$ should converge to the true link $q^*(y|x)$ for each $x,y$.
+    \item[Interval concentration] As we obtain more samples, the interval $C_{x,y}$ should get smaller. Asymptotically it should include nothing but the point $q^*(y|x)$.
     \item[Conservative coverage] $q^*(y|x) \in C_{x,y}$ with high probability.
 \end{description}
-Our simulations yielded good news and bad news.
-The good news: the MLM generally achieves conservative coverage.
-In many cases, the MLM also achieves estimator convergence.
-The bad news: in many cases the MLM fails to achieve estimator convergence.
-Indeed, we can show that the identifiability issues block any possibility of success in some cases; we detail this problem in Appendix \ref{sec:casestudies}.
+However, due to certain identifiability issues, it is only actually possible to guarantee the last desideratum.
+In many cases, the MLM achieves all three.
+However, we can show that identifiability issues block any possibility of success in some cases; we detail this problem in Appendix \ref{sec:casestudies}.
+The good news is that even though the Markov Link Method credible intervals may not concentrate around the truth, they do as well as possible.
+In particular, they do appear to converge around the quantities we can estimate, namely the $q_\mathrm{lo},q_\mathrm{hi}$ bounds described in Section \ref{sec:mlm}.
+A precise  statement of what we have been able to prove may be found in Theorem \ref{thm:mainthm} of Appendix \ref{sec:casestudies}.
+
 The following two simulations give a good overview of the relevant issues:
@@
 
-The MLM appears to provide reasonable values in simulation and gives useful insight to real data.
+The MLM appears gives provable guarantees, provides reasonable values in simulation, and gives useful insight to real data.
 Applied to neuronal cell-type data, the MLM clarified how two different cell-type classification systems are related.
@@
 
-More generally, it is clear that good assumptions can help us get real insight for tough problems.
-Even if these assumptions are not sufficient to allow us to perfectly identify our object of interest, credible regions can be constructed.
-By probing these regions carefully, we can learn what the data actually has to say and what experiments will help us learn more.
+Another direction for future work would be to take a frequentist point of view.
+An obvious direction would be to use profile likelihoods as statistics to define confidence intervals for the link parameters.
+However, the distribution of these statistic is difficult to pin down.
+Traditional techniques based on asymptotic normality may fail dramatically, because the true parameters may lie at the boundary of the parameter space.
+In particular, there may be some $x,y$ for which $q^*(y|x)=0$.
+Even approximate methods are difficult to apply because the parameter space can be quite high-dimensional.
+Perhaps future work could uncover a way to overcome these challenges.
+
+It is clear that good assumptions can help us get real insight for tough problems.
+Even if these assumptions are not sufficient to allow us to perfectly identify our object of interest, we can bound our uncertainty.
+By probing this uncertainty carefully, we can learn what the data actually has to say and what experiments will help us learn more.
 
@@
 
-This theorem is certainly a step in the right direction, but we emphasize that it does require some assumptions.
-It is our opinion that the linear independence requirement is fairly mild (if it is not met then subpopulations can simply be merged together).
-However, the positivity requirement is quite troubling.
+Since it is easy to find consistent estimators for $\hat p,\hat h$, this theorem suggests we can use those estimators to get good estimates for our uncertainty about $q^*$.
+In particular, subject to the conditions of the theorem, we have that the bounds of $C_{x,y}$ converge to the bounds of what is identifiable about $q^*$.
+It is certainly a step in the right direction, but we emphasize that the theorem's conditions are nontrivial.
+It is our opinion that the linear independence condition is fairly mild (if it is not met then subpopulations can simply be merged together).
+However, the positivity condition is quite troubling.
 In many cases of interest the true link $q^*$ has genuine zeros: pairs of measurements between the two tools which are fundamentally incompatible.
@@
 A precise understanding of the discontinuity example above has remained elusive.
-Better understanding could lead to a much more accurate and practical estimates.
+Better understanding could lead to more accurate estimates.
 We leave it for future work.
+
 \end{document}

### 2018-08-10 markovlink/polytopes.py

@@ -233,7 +233,7 @@ def scipy_linprog(c,A_eq,b_eq):
     else:
         raise Exception("Linear programming failed!")    
 
-def cvxopt_linprog(c,A_eq,b_eq,initialcondition,gaptol=1e-5,verbose=True):
+def cvxopt_linprog(c,A_eq,b_eq,initialcondition,gaptol=1e-5,verbose=False):
     '''
     minimize <c, x>
     subj to Ax = b

### 2018-08-13 paper/paper.tex
rewrite qmid

@@
 It might seem sensible to start by specifying some prior beliefs about the link $q^*$.
-However, the link is in many cases `nonidentifiable';  we describe this problem in detail in Appendix \ref{sec:casestudies}.
+However, the link is in many cases `nonidentifiable,' i.e.\ even with an asymptotically infinite number of samples it is not necessarily possible to get a perfect estimate for the link.
+We describe this problem in detail in Appendix \ref{sec:casestudies}.
 One consequence of nonidentifiability is that standard bayesian inferences would have no robustness to our precise choice of prior beliefs.
@@
 So, instead of trying to estimate this inestimable quantity, we will focus on three other quantities which can be identified from data.
-Let
+For each fixed value of $p^*,h^*$, we define
     \begin{itemize}
-        \item $q_{\mathrm{lo},x,y} \triangleq \min_{q\in \Theta(p^*,h^*)} q(y|x)$
-        \item $q_{\mathrm{hi},x,y} \triangleq \max_{q\in \Theta(p^*,h^*)} q(y|x)$ 
-        \item $q_{\mathrm{mid}} \triangleq \argmin_{q\in \Theta(p^*,h^*)} \sum_{x,y} q(y|x)^2$
+        \item $q_{\mathrm{lo},x,y}(p^*,h^*) \triangleq \min_{q\in \Theta(p^*,h^*)} q(y|x)$
+        \item $q_{\mathrm{hi},x,y}(p^*,h^*) \triangleq \max_{q\in \Theta(p^*,h^*)} q(y|x)$ 
+        \item $q_{\mathrm{mid}}(p^*,h^*) \triangleq \argmin_{q\in \Theta(p^*,h^*)} \sum_{x,y} \left(q(y|x) - \mathrm{Uniform}(y)\right) ^2$
     \end{itemize}
-    Note that all three of these optimization problems have unique solutions and would be straightforward to solve if $p^*,h^*$ were known; optimization details can be found in Appendix \ref{sec:optproblems}.
-Morever, even when $q^*$ is nonidentifiable, we can still be assured that $q_{\mathrm{lo}}, q_{\mathrm{hi}}$ are identifiable and $q_{\mathrm{lo},x,y} \leq q^*(y|x) \leq q_{\mathrm{hi},x,y}$.
-We can also hope that $q_{\mathrm{mid}}$ will provide a single estimate which strikes a middle ground; in particular, $q_{\mathrm{mid}}$ is the link value which is closest to uniform among all links which are consistent with $p^*,h^*$.
+    where $\mathrm{Uniform}(y)$ indicates the uniform probability mass function for $y$.
+Note that all three of these optimization problems have unique solutions and would be straightforward to solve if $p^*,h^*$ were known.
+Indeed, the set $\Theta$ is a convex polytope, so the first two problems are linear programs and the last problem is a strictly convex quadratic program.
+Details can be found in Appendix \ref{sec:optproblems}.
+Even when $q^*$ is nonidentifiable, we can still be assured that $q_{\mathrm{lo}}, q_{\mathrm{hi}}$ are identifiable and $q_{\mathrm{lo},x,y} \leq q^*(y|x) \leq q_{\mathrm{hi},x,y}$.
+We can also hope that $q_{\mathrm{mid}}$ will provide a single estimate which strikes a middle ground; among all links which are consistent with $p^*,h^*$, $q_{\mathrm{mid}}$ is the one which is closest to uniform in a sum-of-squares sense.
 \end{itemize}  
@@
 $G(p,h)=0$ is a numerical way to indicate that the MLM assumption can hold.
-After incorporating this knowledge, we find ourselevs interested in the conditional distribution
+After incorporating this knowledge, we find ourselves interested in the conditional distribution
 \[
@@
     \[
-    \hat q \triangleq \mathbb{E}\left[q_{\mathrm{mid}} | N^X,N^Y,G(p^*,h^*)=0\right]
+    \hat q \triangleq \mathbb{E}\left[q_{\mathrm{mid}}(p^*,h^*) | N^X,N^Y,G(p^*,h^*)=0\right]
     \]
@@
     \begin{align*}
-    \mathbb{P}\left(q_{\mathrm{lo},x,y} \in C_{x,y} | N^X,N^Y,G(p^*,h^*)=0\right)\geq 97.5\% \\
-    \mathbb{P}\left(q_{\mathrm{hi},x,y} \in C_{x,y} | N^X,N^Y,G(p^*,h^*)=0\right)\geq 97.5\% 
+    \mathbb{P}\left(q_{\mathrm{lo},x,y}(p^*,h^*) \in C_{x,y} | N^X,N^Y,G(p^*,h^*)=0\right)\geq 97.5\% \\
+    \mathbb{P}\left(q_{\mathrm{hi},x,y}(p^*,h^*) \in C_{x,y} | N^X,N^Y,G(p^*,h^*)=0\right)\geq 97.5\% 
     \end{align*}
@@
 In practice, we were not able to find a way to compute the credible interval and posterior means exactly.
-It seems difficult to even sample from the conditional distribution.
+However, given samples from the posterior distribution it would be straightforward to get good estimates, using use monte-carlo approximations.
+Unfortunately, it seems difficult to obtain samples from the posterior distribution, $\mathbb{P}(p^*,h^*| N^X,N^Y,G(p^*,h^*)=0)$.
 Common approaches to this type of problem involve Markov-Chain Monte Carlo and Variational Inference, but we were unable to make these approaches work in practice.
@@
 We instead take a somewhat na\"ive approach.
-We start by defining $\tilde P,\tilde H$ to carry the conditional distribution of $p^*,h^*$ given the observed data (but not conditioned on the fact that the MLM assumption holds).
-We then simply project $\tilde H$ to the space of parameters which are consistent with the MLM assumption.
-That is, we take $\hat H$ to be the solution of the minimization problem $\min_{h'}D(\tilde H|h')$, subject to the constraint that $E(\tilde P,h')=0$.
+We start by drawing samples according to
+\[
+P,H ~ \sim \mathbb{P}(p^*,h^* | N^X,N^Y)
+\]
+This can be achieved exactly, using the the conjugacy between our prior and the Multinomial distribution.
+Notice that these samples do not incorporate our knowledge of the MLM assumption, insofar as they are not conditioned on the event $G(p^*,h^*)=0$.
+To approximately remedy this, we define $\tilde H$ as the solution of $\min_{h'}D(H|h')$, subject to the constraint that $E(P,h')=0$.
 Optimization details can be found in Appendix \ref{sec:optproblems}.
-We hope that the distribution of $\tilde P,\hat H$ will approximate the desired posterior.
-We then use samples from this distribution to compute what we need.
-We asymptotically expect that $\tilde P,\tilde H$ will nearly satisfy the MLM assumption in any case, so this approximation should not make a large difference.
+We use the pair $P,\tilde H$ as \emph{approximate} samples for the distribution $\mathbb{P}(p^*,h^*| N^X,N^Y,G(p^*,h^*)=0)$.
+We can repeat this process to produce many samples of $(P,\tilde H)$ and use those samples to produce approximate monte-carlo estimates for $\hat q(y|x),C_{x,y}$.
+We asymptotically expect that $P,H$ will nearly satisfy the MLM assumption in any case, so this approximation should not make a large difference.
 However, we caution that the exact implications of the approximation are difficult to reason about.
@@
 Each parameter has a slightly different story.
-For example, for the Standard type $x=$`Vip Rspo4' and the Patch type $y=$`Vip,' we have that $C_{x,y}=[.92,1.0]$.
-This supports the idea that the true link satisfies $q^*(y|x)\geq .92$: at least 92\% of the cells classified as type `Vip Rspo4' by the Standard method will be classified as `Vip' by the Patch method.
+For example, for the Standard type $x=$`Vip Rspo4' and the Patch type $y=$`Vip,' we have that $C_{x,y}=[.88,1.0]$.
+This supports the idea that the true link satisfies $q^*(y|x)\geq .88$: at least 88\% of the cells classified as type `Vip Rspo4' by the Standard method will be classified as `Vip' by the Patch method.
 However, for other types there is more ambiguity.

### 2018-08-15 paper/paper.tex
many notation changes

--- 2018-08-10 paper/paper.tex
@@
 \DeclareMathOperator*{\tr}{tr}
+
+\newcommand{\plx}{\ensuremath{f}}
+\newcommand{\pxy}{\ensuremath{g}}
+\newcommand{\ply}{\ensuremath{h}}
+% \newcommand{\DX}{\ensuremath{{\mathcal{D}^X}}}
+% \newcommand{\DY}{\ensuremath{{\mathcal{D}^Y}}}
+\newcommand{\DX}{\ensuremath{\mathcal{D}^X}}
+\newcommand{\DY}{\ensuremath{\mathcal{D}^Y}}
+\newcommand{\MLMH}{\ensuremath{\mathcal{A}}}
+
+\newcommand{\MLM}{MLM}
 
@@
 \title{The Markov Link Method: a nonparametric approach to combine observations from multiple experiments}
-\author{Jackson Loper, Osnat Penn, Trygve Bakken, David Blei, Liam Paninski}
+\author{Jackson Loper, Osnat Penn, Trygve Bakken, David Blei, Liam Paninski, \\Additional Authors To Be Determined}
 
@@
 
-The last century has seen a proliferation of experimental techniques.
-Different perspectives arise from different techniques (and even from different labs performing the `same' protocol).
-Ideally, this yields theories that account for all perspectives.
-Instead, sometimes, this leaves us with many perspectives and no idea how they are connected.
+Imagine we have a collection of different experiments about human cells.
+We would like to link the different perspectives that we get from each experiment.
+To make this idea rigorous, we need to define what we mean by `perspective' and what we mean by `link.'
 
-Our goal is to connect different perspectives.  To make this idea rigorous, we assume each experiment has two aspects:
+To define `perspective,' we assume each experiment has two aspects:
 %
 \begin{itemize}
-    \item a \emph{sampling procedure} determines what we study
-    \item a \emph{measurement procedure} defines the perspective we take in studying it.
+    \item A \emph{selection procedure} defines how we select cells.  For example, one experiment might select only liver cells, and another might select only cells which express the Vip protein.  
+    \item A \emph{measurement procedure} defines what we measure about each cell.
+For example, one experiment might measure the transcriptomic profile of a cell, and another experiment might measure the cell morphology.
 \end{itemize}
 %
-If two experiments differ in both aspects, it may be difficult to gain insight by combining their data.
-For example, imagine one experiment performs a transcriptomic measurement of liver cells and another applies a morphological measurement of heart cells.
-Without additional experiments or knowledge, looking at these two experiments together may not be very helpful.
-On the other hand, if two experiments have some aspect in common, there is something to be learned.
-In this sense, a collection of experiments can be understood as a network: a pair of experiments are connected if they have some aspect in common.
-In some cases, we can also connect two experiments if they differ in both aspects but we have prior knowledge (e.g.\ if the sampling procedures are different, but a reweighting can be used to correct for the difference).
-When many experiments can be joined into a common network, much can be learned.
+The selection procedure defines what the experiment studies, and the measurement procedures determines the experiment's `perspective.'
 
-We here propose the The Markov Link Method (MLM), an algorithm that uses a network of experiments to learn the relationship between two different perspectives.
-We formalize this through what we call the `measurement link' parameters, $q^*(y|x)$.
-The meaning of these parameters is defined as follows.
-Let us say we obtain measurement result $x$ from one measurement procedure.
-The number $q^*(y|x)$ indicates the probability of obtaining result $y$ from the second measurement procedure applied to measure the same specimen.
-The link tells us how the properties measured by each procedure are associated.
-For example, the link lets us answer questions such as ``I have observed the transcriptome of this cell -- what morphologies might this cell have?''
+The `link' between two measurement procedures should enable us to answer questions such as ``I have observed the transcriptome of this cell -- what morphologies might this cell have?''  We formalize this idea through what we call the `measurement link parameters,' $\pxy(y|x)$.
+Specifically, let us say we obtain measurement result $x$ from one measurement procedure.
+The number $\pxy(y|x)$ indicates the probability of obtaining result $y$ from the second measurement procedure applied to measure the same specimen.
 
-If we have a single large experiment that measures each specimen with both measurement procedures, it is straightforward to estimate the link.
-We can examine specimens through both perspectives simultaneously.
-However, in many cases we have some experiments in which specimens are measured with one procedure, other experiments using the other procedure, and few or no experiments where specimens are measured with both procedures.
-This is particularly true if the measurement techniques destroy or alter the specimen in the process of measuring it.
-
-In this paper we focus particularly on the case that we have a network of experiments based on several sampling procedures and exactly two different measurement procedures.
+We here propose the The Markov Link Method (\MLM), an algorithm that uses a collection of experiments to estimate the measurement link parameters, \pxy.
+In this paper we focus particularly on the case that we have experiments based on several selection procedures and exactly two different measurement procedures.
 We also suppose that the measurement procedures return categorical observations, i.e.\ a `measurement' assigns a specimen to one of a finite set of categories.
-The MLM could also generalize to numerical observations, but in that case it would be appropriate to incorporate some knowledge of smoothness (e.g.\ through a Gaussian assumption).
+The \MLM{} could also generalize to numerical observations, but in that case it would be appropriate to incorporate some knowledge of smoothness (e.g.\ through a Gaussian assumption).
 Categorical data requires no additional modeling assumptions, so for simplicity we will here focus on the categorical case.
-This setup, sketched in Figure \ref{fig:bigpicture}, is perhaps the simplest example that highlights all the important assumptions and properties of the MLM.
+This setup, sketched in Figure \ref{fig:bigpicture}, is perhaps the simplest example that highlights all the important assumptions and properties of the \MLM.
+It embodies a common situation in biological problems, such as is found in \cite{gouwens2018classification}.
 
@@
 \hfill{}\figgygrs{.5}{../images/bigpicture}{(a) Setup: each specimen can be \\ measured with one of two procedures, \\ but can't be measured with both.\\ How do we study the relationship \\ between the procedures?}
-\hfill{}\figgygrs{.4}{../images/plate}{(b) Key modeling assumption: \\ the link $q$ is the same for \\ every sampling procedure. }\hfill{}
+\hfill{}\figgygrs{.4}{../images/plate}{(b) Key modeling assumption: \\ the link $\pxy$ is the same for \\ every selection procedure. }\hfill{}
 
-\caption{{\bf The Markov Link Method (MLM) links different measurement procedures by using a set of different sampling procedures.}  In subfigure (a) we have neurons gathered using many different sampling procedures.
+\caption{{\bf The Markov Link Method (MLM) links different measurement procedures by using a set of different selection procedures.}  In subfigure (a) we have neurons gathered using many different selection procedures.
 We select half of the neurons and determine what type of neuron they are, using transcriptomic measurements.
@@
 In subfigure (b) we present a plate diagram articulating the statistical assumption that makes it possible (cf.\ \cite{koller2009probabilistic} for a thorough explanation of this type of diagram).
-For each neuron in each group, this diagram considers the sampling procedure ($\ell$), the transcriptomic type ($X$), and the morphological type ($Y$).
-The assumption is that the transcriptomic type is sufficiently detailed to \emph{statistically isolate} the transcriptomic type from the sampling procedure.
+For each neuron in each group, this diagram considers the selection procedure ($\ell$), the transcriptomic type ($X$), and the morphological type ($Y$).
+The assumption is that the transcriptomic type is sufficiently detailed to \emph{statistically isolate} the transcriptomic type from the selection procedure.
 This is reflected in the arrows of the diagram; all paths from $\ell$ variables to $Y$ variables pass through $X$ variables.
-The assumption allows us to estimate the link, $q$.
+The assumption allows us to estimate the link, $\pxy$.
 \label{fig:bigpicture}}
@@
 
-Imagine that we have a large collection of human cells.  For each cell we are interested in 
+The Markov Link Method is a general way to estimate measurement link parameters from a collection of experiments.
+For concreteness, we will use experiments about human cells as a running example.
+Let us say that we have gathered many human cells.
+For each cell we are interested in
 %
 \begin{enumerate}
-    \item $\ell$, the sampling procedure used to obtain the specimen.
-For example, we can select different subpopulations of cells based on proteomic markers.
-For each marker we might have a different sampling procedure designed to select cells with that marker.
-    \item $X$, the cell's `transcriptomic type,' as measured by looking at the transcriptomic activity in the cell.  
+    \item $\ell$, the selection procedure used to obtain the specimen
+    \item $X$, the cell's `transcriptomic type,' as measured through a transcriptomic analysis.  
     \item $Y$, the cell's `morphological type,' as measured by looking at an image of the cell.
@@
 %
-Our task is to reconcile the two different notions of `cell-type' indicated by $X$ and $Y$.
-We would like to know what transcriptomic types are associated with what morphological types.
-For example, we might like to say ``This cell has transcriptomic type 3, therefore it probably has morphological type F.''  Our main assumption is that these kinds of associations are unaffected by the choice of sampling procedure.
+Our task is to link the transcriptomic and morphological perspectives on `cell-type,' as manifested through the $X$ and $Y$ measurement procedures.
+For example, we might like to say ``This cell has transcriptomic type 3, therefore it probably has morphological type F.''  Our main assumption is that these kinds of associations are unaffected by the choice of selection procedure.
 In particular:
@@
 \begin{center}
-\textbf{The Markov Link Method Assumption}\\$\mathbb{P}(Y|X,\ell)=q^*(y|x)$ for each $\ell$
+\textbf{The Markov Link Method Assumption}\\$\mathbb{P}(Y|X,\ell)=\pxy(y|x)$ for each $\ell$
 \end{center}
@@
 Here $\mathbb{P}(Y|X,\ell)$ indicates the probability that the morphological type is $y$ given that the transcriptomic type is $x$ and the specimen was sampled with strategy $\ell$.
-The assumption is that this probability actually does not depend upon $\ell$.
-This can be understood intuitively through a thought experiment.
+The assumption states that that this probability actually does not depend upon $\ell$.
+This assumption can be understood intuitively through a thought experiment.
 Let us imagine we have attained perfect understanding of cellular biology.
-Someone selects a cell using one of a set of sampling procedures and measures the transcriptomic activity of the cell.
-We are then told the cell's transcriptomic activity (but we are not told which of the sampling procedured was used to gather the specimen).
-Using our perfect knowledge of the physical systems, we could then make predictions about a measurement of the cell's morphology.
-Now we are told new information: we are told which sampling procedure was used to gather the specimen.
+Someone selects a cell using one of a set of selection procedures and measures the transcriptomic activity of the cell.
+We are then told the cell's transcriptomic type (but we are not told which of the selection procedures was used to gather the specimen).
+Using our perfect knowledge of the physical systems, we could then make predictions about the cell's morphological type.
+Now we are told new information: we are told which selection procedure was used to gather the specimen.
 Would this substantially change our predictions?
@@
 
-Our task is to estimate the link, $q^*$, from a collection of experiments.
-We suppose that for each sampling procedure we have two experiments: one that measured transcriptomics and another that measured morphology.
+More formally, our task is to estimate the link parameters $\pxy$ from a collection of experiments.
+We suppose that for each selection procedure we have two experiments: one that measured transcriptomics and another that measured morphology.
 We have no experiments in which both types were measured for the same specimens.
@@
 
-In the specific case we will focus on, all of the experimental data can be summarized with two matrices, $N^X$ and $N^Y$.
-Here $N^X_{\ell x}$ indicates the number of specimens gathered with procedure $\ell$ that have transcriptomic type $x$.
-Likewise $N^Y_{\ell y}$ indicates the number of specimens gathered with procedure $\ell$ that have morphological type $y$.
+In the specific case we will focus on, all of the experimental data can be summarized with two matrices, $\DX$ and $\DY$.
+Here $\DX_{\ell x}$ indicates the number of specimens gathered with procedure $\ell$ that have transcriptomic type $x$.
+Likewise $\DY_{\ell y}$ indicates the number of specimens gathered with procedure $\ell$ that have morphological type $y$.
 For each transcriptomic type $x$ and morphological type $y$, the Markov Link Method uses the data to produce two objects:
@@
 \begin{enumerate}
-    \item $\hat q(y|x)$ -- a point estimate for the link parameter $q^*(y|x)$
-    \item $C_{x,y}$ -- a credible interval which contains the true link parameter $q^*(y|x)$ with high probability
+    \item $\hat \pxy(y|x)$ -- a point estimate for the link parameter $\pxy(y|x)$
+    \item $C_{x,y}$ -- a credible interval which contains the true link parameter $\pxy(y|x)$ with high probability
 \end{enumerate}
 %
-To produce these estimates for the link, we take a bayesian point of view.
-It might seem sensible to start by specifying some prior beliefs about the link $q^*$.
-However, the link is in many cases `nonidentifiable';  we describe this problem in detail in Appendix \ref{sec:casestudies}.
-One consequence of nonidentifiability is that standard bayesian inferences would have no robustness to our precise choice of prior beliefs.
-To avoid these problems, we focus on quantities which can be identified from the data.
-In particular,
+To estimate these, we must consider certain `nuisance parameters.'  Our observable data, $\DX,\DY$, is governed by the following parameters:
 \begin{itemize}
-    \item Let $p^*(x|\ell)$ denote the probability that a cell sampled with procedure $\ell$ will have transcriptomic type $x$.  These probabilities can be estimated from the data, $N^X$.
-    \item Let $h^*(y|\ell) \triangleq \sum_x p^*(x|\ell)q^*(y|x)$ denote the probability that a cell sampled with procedure $\ell$ will have morphological type $y$.
-These probabilities can be estimated from the data, $N^Y$.
-    \item Let $\Theta(p,h)=\{q:\ h(y|\ell) = \sum_x p(x|\ell)q(y|x)\}$ denote the set of links which are consistent with $p,h$.
-Notice that $q^*\in\Theta(p^*,h^*)$, but there may be other links $q\in\Theta(p^*,h^*)$.
-Any such link yields the same distribution on what we can observe, $N^X,N^Y$.
-Therefore there is simply no way to use data to distinguish among them.
-This is the essence of the nonidentifiability problem mentioned above.
-So, instead of trying to estimate this inestimable quantity, we will focus on three other quantities which can be identified from data.
-Let
-    \begin{itemize}
-        \item $q_{\mathrm{lo},x,y} \triangleq \min_{q\in \Theta(p^*,h^*)} q(y|x)$
-        \item $q_{\mathrm{hi},x,y} \triangleq \max_{q\in \Theta(p^*,h^*)} q(y|x)$ 
-        \item $q_{\mathrm{mid}} \triangleq \argmin_{q\in \Theta(p^*,h^*)} \sum_{x,y} q(y|x)^2$
+    \item $\plx(x|\ell)$ -- the probability that a cell sampled with procedure $\ell$ will have transcriptomic type $x$
+    \item $\ply(y|\ell)$ -- the probability that a cell sampled with procedure $\ell$ will have morphological type $y$
 \end{itemize}
-    Note that all three of these optimization problems have unique solutions and would be straightforward to solve if $p^*,h^*$ were known; optimization details can be found in Appendix \ref{sec:optproblems}.
-Morever, even when $q^*$ is nonidentifiable, we can still be assured that $q_{\mathrm{lo}}, q_{\mathrm{hi}}$ are identifiable and $q_{\mathrm{lo},x,y} \leq q^*(y|x) \leq q_{\mathrm{hi},x,y}$.
-We can also hope that $q_{\mathrm{mid}}$ will provide a single estimate which strikes a middle ground; in particular, $q_{\mathrm{mid}}$ is the link value which is closest to uniform among all links which are consistent with $p^*,h^*$.
-\end{itemize}  
-To summarize: it may not be possible to estimate $q^*$ from data, but we can always estimate $p^*,h^*$.
-We can also compute deterministic functions of $p^*,h^*$ which bound $q^*$, namely $q_{\mathrm{lo}},q_{\mathrm{hi}}$.
-We take a Bayesian point of view in estimating $p^*,h^*$.
-For prior beliefs, we assume a noninformative uniform prior $\mathbb{P}$, namely
+It is straightforward to estimate $\plx,\ply$ from $\DX,\DY$ using standard methods.
+We would like to use these parameters to tell us something about our object of interest, the link $\pxy$.
+Under the Markov Link Method assumption, this link is connected to $\plx,\ply$ through the equations
+\begin{gather}\label{eq:mlmh}
+\sum_x \plx(x|\ell)\pxy(y|x)=\ply(y|\ell) \qquad \forall y,\ell
+\end{gather}
+With these equations and knowledge of $\plx,\ply$, we can produce an estimate for the link: $\hat \pxy_{\plx,\ply}$.
+We can also produce bounds for the link.
+We define these objects precisely in Appendix \ref{sec:mlmdetails}.
+However, in practice we do not actually have exact knowledge of $\plx,\ply$.
+We only have estimates from data.
+We take a Bayesian perspective to account for our uncertainty in this estimation.
+For our prior beliefs, we assume a noninformative uniform prior $\mathbb{P}$.
+Following the Bayesian philosophy, we then incorporate new knowledge by conditioning.
+We have two important pieces of knowledge about $\plx,\ply$.
+First, we have observed the data, $\DX,\DY$.
+Second, we know from the MLM assumption that there exists \emph{some} value $\pxy$ such that Equation (\ref{eq:mlmh}) holds; let \MLMH{} denote the event that this holds.
+We then define our point estimate $\hat \pxy$ as the posterior expectation of $\hat \pxy_{\plx,\ply}$ after conditioning on this knowledge, i.e.
 \[
-\mathbb{P}(p^*,h^*) \propto 1
+\hat \pxy(y|x) = \mathbb{E}\left[\hat \pxy_{\plx,\ply} | \DX,\DY,\MLMH \right]
 \]
-Following the bayesian philosophy, we then incorporate new knowledge by conditioning.
-We have two important pieces of knowledge we would like to incorporate.
-First, we have observed the data, $N^X,N^Y$.
-Second, we know from the MLM assumption that there exists \emph{some} value $q^*$ which is consistent with $p^*,h^*$.
-To make the second idea formal, let $D(h|h')$ denote the Kullback-Leibler divergence and $G(p,h) = \min_{h':\ \Theta(p,h')\neq \emptyset} D(h|h')$.
-$G(p,h)=0$ is a numerical way to indicate that the MLM assumption can hold.
-After incorporating this knowledge, we find ourselevs interested in the conditional distribution
-\[
-\mathbb{P}\left(p^*,h^* | N^X,N^Y,G(p^*,h^*)=0\right)
-\]
-We use this posterior to produce estimates for our quantities of interest:
-\begin{enumerate}
-    \item $\hat q$ is calculated using posterior expectation:
-    \[
-    \hat q \triangleq \mathbb{E}\left[q_{\mathrm{mid}} | N^X,N^Y,G(p^*,h^*)=0\right]
-    \]
-    %
-    \item $C_{x,y}$ is calculated using credible intervals.  For each $x,y$, we define $C_{x,y}$ to be the largest interval such that 
-    \begin{align*}
-    \mathbb{P}\left(q_{\mathrm{lo},x,y} \in C_{x,y} | N^X,N^Y,G(p^*,h^*)=0\right)\geq 97.5\% \\
-    \mathbb{P}\left(q_{\mathrm{hi},x,y} \in C_{x,y} | N^X,N^Y,G(p^*,h^*)=0\right)\geq 97.5\% 
-    \end{align*}
-\end{enumerate}
-Code to compute these objects is published at https://github.com/jacksonloper/markov-link-method, including a tutorial-style ipython notebook detailing every computation made in this paper.
-
-In practice, we were not able to find a way to compute the credible interval and posterior means exactly.
-It seems difficult to even sample from the conditional distribution.
-Common approaches to this type of problem involve Markov-Chain Monte Carlo and Variational Inference, but we were unable to make these approaches work in practice.
-It seems nontrivial to work with the condition $G(p,h)=0$ that formalizes the MLM assumption.
-We instead take a somewhat na\"ive approach.
-We start by defining $\tilde P,\tilde H$ to carry the conditional distribution of $p^*,h^*$ given the observed data (but not conditioned on the fact that the MLM assumption holds).
-We then simply project $\tilde H$ to the space of parameters which are consistent with the MLM assumption.
-That is, we take $\hat H$ to be the solution of the minimization problem $\min_{h'}D(\tilde H|h')$, subject to the constraint that $E(\tilde P,h')=0$.
-Optimization details can be found in Appendix \ref{sec:optproblems}.
-We hope that the distribution of $\tilde P,\hat H$ will approximate the desired posterior.
-We then use samples from this distribution to compute what we need.
-We asymptotically expect that $\tilde P,\tilde H$ will nearly satisfy the MLM assumption in any case, so this approximation should not make a large difference.
-However, we caution that the exact implications of the approximation are difficult to reason about.
-In any case, it is a practical solution to a difficult problem.
+We can use a similar strategy to produce the credible intervals $C_{x,y}$.
+Exact details can be found in Appendix \ref{sec:mlmdetails}.
+Code to compute $\hat \pxy(y|x)$ and $C_{x,y}$ is published at https://github.com/jacksonloper/markov-link-method, including a tutorial-style ipython notebook detailing every computation made in this paper.
 
@@
 \figgy{
-\hfill{}\figgygrs{.4}{../images/simulationI}{(a) Estimation error, \\ varying number of samples \\ and number of sampling procedures.}\hfill{}
-\hfill{}\figgygrs{.5}{../images/simulationIconf}{(b) Credible intervals, \\ varying number of samples \\ and number of sampling procedures.}\hfill{}
+\hfill{}\figgygrs{.4}{../images/simulationI}{(a) Estimation error, \\ varying number of samples \\ and number of selection procedures.}\hfill{}
+\hfill{}\figgygrs{.5}{../images/simulationIconf}{(b) Credible intervals, \\ varying number of samples \\ and number of selection procedures.}\hfill{}
 
@@
 \caption{{\bf When can the MLM accurately estimate the link?} \label{fig:simI}  We use simulations to test the MLM.
-In subfigure (a) we look at the error of the MLM link estimator in many different situations, varying the numbers of samples, sampling procedures, and the link parameters.
+In subfigure (a) we look at the error of the MLM link estimator in many different situations, varying the numbers of samples, selection procedures, and the link parameters.
 In each trial our `ground truth' link parameters picked uniformly at random from the parameter space.
-The MLM estimator accurately recovers the ground truth from simulated data, as long as we have enough samples \emph{and} enough distinct sampling procedures.
+The MLM estimator accurately recovers the ground truth from simulated data, as long as we have enough samples \emph{and} enough distinct selection procedures.
 In subfigure (b) we consider the MLM's credible intervals.
-For a selection of trials we choose a random value of of $x,y$ and then compare the true link parameter $q^*(y|x)$ with the MLM interval $C_{x,y}$.
+For a selection of trials we choose a random value of of $x,y$ and then compare the true link parameter $\pxy(y|x)$ with the MLM interval $C_{x,y}$.
 We show the interval (in black) centered around the ground truth (in red).
 Since the link parameters lie in $[0,1]$, the interval is always contained within $\pm1$ of the ground truth (this larger region is shown in gray).
-The intervals mostly contain the truth, and get narrower with more samples as long as there are enough distinct sampling procedures.
-In subfigure (c) we focus on the case of exactly $4$ sampling procedures.
+The intervals mostly contain the truth, and get narrower with more samples as long as there are enough distinct selection procedures.
+In subfigure (c) we focus on the case of exactly $4$ selection procedures.
 In this case it is only possible to get very accurate estimates if the link takes on a particular form.
 If the link is independent (i.e.\ $X,Y$ are independent) or if the link is invertible and deterministic (e.g.\ $X=Y$) then we can determine what the link is.
-However, if the truth is more complex, small numbers of sampling procedures may make it impossible to determine the true link; this issue is discussed in detail in Section \ref{sec:casestudies}.}
+However, if the truth is more complex, small numbers of selection procedures may make it impossible to determine the true link; this issue is discussed in detail in Section \ref{sec:casestudies}.}
 }
@@
 \begin{description}
-    \item[Estimator convergence] As we obtain more samples, $\hat q(y|x)$ should converge to the true link $q^*(y|x)$ for each $x,y$.
-    \item[Interval concentration] As we obtain more samples, the interval $C_{x,y}$ should get smaller. Asymptotically it should include nothing but the point $q^*(y|x)$.
-    \item[Conservative coverage] $q^*(y|x) \in C_{x,y}$ with high probability.
+    \item[Estimator convergence] As we obtain more samples, $\hat \pxy(y|x)$ should converge to the true link $\pxy(y|x)$ for each $x,y$.
+    \item[Interval concentration] As we obtain more samples, the interval $C_{x,y}$ should get smaller. Asymptotically it should include nothing but the point $\pxy(y|x)$.
+    \item[Conservative coverage] $\pxy(y|x) \in C_{x,y}$ with high probability.
 \end{description}
@@
 The good news is that even though the Markov Link Method credible intervals may not concentrate around the truth, they do as well as possible.
-In particular, they do appear to converge around the quantities we can estimate, namely the $q_\mathrm{lo},q_\mathrm{hi}$ bounds described in Section \ref{sec:mlm}.
+In particular, they do appear to converge around the quantities we can estimate, namely the $q_\mathrm{lo},q_\mathrm{hi}$ bounds described in Appendix \ref{sec:mlmdetails}.
 A precise statement of what we have been able to prove may be found in Theorem \ref{thm:mainthm} of Appendix \ref{sec:casestudies}.
@@
 \begin{itemize}
-    \item How does MLM performance depend on the number of samples and the number of sampling procedures?
+    \item How does MLM performance depend on the number of samples and the number of selection procedures?
 We suppose we have two measurement procedures which return categorical measurements among six categories.
@@
 To see how the method performs in different circumstances, we will run many trials.
-In each trial we fix the number of sampling procedures and pick `ground truth' parameters uniformly at random from the parameter space.
-We then simulate a dataset from these ground truth parameters, fixing the total number of samples (spreading these samples equally among all combinations of sampling procedure and measurement procedure).
+In each trial we fix the number of selection procedures and pick `ground truth' parameters uniformly at random from the parameter space.
+We then simulate a dataset from these ground truth parameters, fixing the total number of samples (spreading these samples equally among all combinations of selection procedure and measurement procedure).
 Finally, we apply the MLM to the simulated data to get the point estimates $\tilde q(y|x)$ and the credible intervals $C_{x y}$.
@@
     \[
-    \mathrm{Error}(\tilde q,q^*)=\frac{1}{12}\sum_{x=1}^{6}\sum_{y=1}^{6} |\tilde q(y|x) - q^*(y|x)|
+    \mathrm{Error}(\tilde q,\pxy)=\frac{1}{12}\sum_{x=1}^{6}\sum_{y=1}^{6} |\tilde q(y|x) - \pxy(y|x)|
     \]
     This error ranges between zero and one.
-Zero error indicates that $\tilde q=q^*$.
+Zero error indicates that $\tilde q=\pxy$.
 An error of one indicates that the estimate has completely incorrect beliefs about the probability mass, i.e.
-$q^*(y|x)=0$ whenever $\tilde q(y|x)>0$ and vice-versa.
+$\pxy(y|x)=0$ whenever $\tilde q(y|x)>0$ and vice-versa.
 We also look at the MLM credible intervals and see how often they cover the true parameters.
@@
 In trials with more samples, the estimator generally has lower error.
-However, to make the error actually converge to zero we need at least six distinct sampling procedures.
-This figure also shows that the intervals work correctly regardless of of the number of samples or sampling procedures; they include the ground truth with high probability.
-With many samples and sampling procedures, the intervals are small and concentrated around the truth.
-With fewer samples or sampling procedures, the intervals are sometimes forced to be larger.
+However, to make the error actually converge to zero we need at least six distinct selection procedures.
+This figure also shows that the intervals work correctly regardless of of the number of samples or selection procedures; they include the ground truth with high probability.
+With many samples and selection procedures, the intervals are small and concentrated around the truth.
+With fewer samples or selection procedures, the intervals are sometimes forced to be larger.
 However, our uncertainty is not the same for every parameter.
@@
     \item How does estimator convergence depend upon the link parameters themselves?
-In each trial of this simulation we use four sampling procedures and our measurement procedures yield one of six categories.
-In the previous simulations we saw that estimator convergence was generally impossible in this situation, due to the small number of sampling procedures.
+In each trial of this simulation we use four selection procedures and our measurement procedures yield one of six categories.
+In the previous simulations we saw that estimator convergence was generally impossible in this situation, due to the small number of selection procedures.
 However, in our previous simulations we picked the link parameters uniformly from parameter space.
@@
 On one extreme, we will produce trials where the link makes $X,Y$ independent, i.e.
-$q^*(y|x)=q^*(y|x')$ for every $x,x'$.
+$\pxy(y|x)=\pxy(y|x')$ for every $x,x'$.
 On the other extreme, we will have trials where $X,Y$ are deterministically related by the equation $X=Y$.
@@
 We leave a complete mathematical understanding of this result to future work.
-For the purposes of this paper, we content ourselves by noting one interesting case: let the measurements return one of $2^k$ categories and let us use only $k+1$ carefully chosen sampling procedures.
+For the purposes of this paper, we content ourselves by noting one interesting case: let the measurements return one of $2^k$ categories and let us use only $k+1$ carefully chosen selection procedures.
 Now suppose the link defines any invertible deterministic function between $X$ and $Y$.
@@
 \hfill{}\figgygr{.95}{../images/allenNlxy}\hfill{}
-\caption{{\bf The input to the Markov Link Method: one experiment for each measurement procedure and each sampling procedure.}  Here we show a portion of the real-world data to which we applied the MLM.
-The Allen Institute gathered neurons from the visual cortex of mice, using a variety of cre/lox-based sampling procedure (cf.\ \cite{tasic2017shared}).
+\caption{{\bf The input to the Markov Link Method: one experiment for each measurement procedure and each selection procedure.}  Here we show a portion of the real-world data to which we applied the MLM.
+The Allen Institute gathered neurons from the visual cortex of mice, using a variety of cre/lox-based selection procedure (cf.\ \cite{tasic2017shared}).
 Each strategy was designed to bring out different subpopulations of cells.
@@
 
-Fortunately, there was a kind of information that seemed like it might help determine whether the two methods were properly calibrated: a variety of sampling procedures.
+Fortunately, there was a kind of information that seemed like it might help determine whether the two methods were properly calibrated: a variety of selection procedures.
 Using a cre/lox system (cf.\ \citep{tasic2017shared}) they were able to pick out various overlapping subpopulations of neurons using different procedures.
 Each procedure was expected to yield different proportions of the different cell-types.
-For each sampling procedure and each measurement procedure, many specimens were sampled and their cell-types determined.
+For each selection procedure and each measurement procedure, many specimens were sampled and their cell-types determined.
 The result of this process was two tables, parts of which are shown in Figure \ref{fig:alleninput}.
@@
 We look at a subselection of Standard types ($x$) and all of the Patch types ($y$).
-For each combination $x,y$ we draw a rectangle whose color indicates the value of the estimator $\hat q(y|x)$.
+For each combination $x,y$ we draw a rectangle whose color indicates the value of the estimator $\hat \pxy(y|x)$.
 We also show our confidence about these estimators.
@@
 For some parameters our intervals are much tighter than others.
-For example, it appears we have high confidence the Standard type `Vip Rspo4' is highly associated with the Patch type `Vip.'  By constrast, we have almost no idea what is associated with the Standard type `Meis2.'  The upper credible interval bounds suggest that $q^*(y|\mathrm{Meis2})$ could be nearly 1 for many different Patch types.
-Obviously it cannot be 1 for all of those types simultaneously, since $\sum_y q^*(y|\mathrm{Meis2})=1$, but the data simply doesn't tell us which $y$ carries the mass.
+For example, it appears we have high confidence the Standard type `Vip Rspo4' is highly associated with the Patch type `Vip.'  By constrast, we have almost no idea what is associated with the Standard type `Meis2.'  The upper credible interval bounds suggest that $\pxy(y|\mathrm{Meis2})$ could be nearly 1 for many different Patch types.
+Obviously it cannot be 1 for all of those types simultaneously, since $\sum_y \pxy(y|\mathrm{Meis2})=1$, but the data simply doesn't tell us which $y$ carries the mass.
 \label{fig:allenboot}}
@@
 
-We applied the MLM to this data to obtain point estimates $\tilde q(y|x)$ and credible intervals $C_{x y}$.
+We applied the MLM to this data to obtain point estimates $\hat \pxy(y|x)$ and credible intervals $C_{x y}$.
 In Figure \ref{fig:allenboot} we visualize these objects for selected values of $x,y$.
 Each parameter has a slightly different story.
-For example, for the Standard type $x=$`Vip Rspo4' and the Patch type $y=$`Vip,' we have that $C_{x,y}=[.92,1.0]$.
-This supports the idea that the true link satisfies $q^*(y|x)\geq .92$: at least 92\% of the cells classified as type `Vip Rspo4' by the Standard method will be classified as `Vip' by the Patch method.
+For example, for the Standard type $x=$`Vip Rspo4' and the Patch type $y=$`Vip,' we have that $C_{x,y}=[.88,1.0]$.
+This supports the idea that the true link satisfies $\pxy(y|x)\geq .88$: at least 88\% of the cells classified as type `Vip Rspo4' by the Standard method will be classified as `Vip' by the Patch method.
 However, for other types there is more ambiguity.
@@
 To get a better resolution of the calibration, we would need a sampling strategy that included `Lamp5 Egln 1' cells but represents a significantly different slice of the overall population.
-For a particular proposed experiment, simulations such as those found in Section \ref{sec:simulations} can be used to determine how many samples might be required to get an accurate estimate of the link, $q^*$.
+For a particular proposed experiment, simulations such as those found in Section \ref{sec:simulations} can be used to determine how many samples might be required to get an accurate estimate of the link, $\pxy$.
 
@@
 We discuss the identifiability issues for our problem in particular in Appendix \ref{sec:casestudies}.
-Our only assumption is that the sampling procedures can be statistically isolated from the link.
+Our only assumption is that the selection procedures can be statistically isolated from the link.
 This assumption says nothing about what the link itself is.
@@
 We formalize the `relationship' between two measurement procedures through the notion of a `measurement link.'  Given that we have measured a specimen with one procedure and obtained the outcome $x$, the link parameter $q(y|x)$ tells us the probability that we would obtain outcome $y$ if we measured the same specimen with the second procedure.
-We present the Markov Link Method assumption, which roughly states that the link can be statistically isolated from the sampling procedures.
+We present the Markov Link Method assumption, which roughly states that the link can be statistically isolated from the selection procedures.
 We provide the Markov Link Method (MLM) algorithm that uses this assumption to estimate the link and measure our uncertainty.
@@
 Traditional techniques based on asymptotic normality may fail dramatically, because the true parameters may lie at the boundary of the parameter space.
-In particular, there may be some $x,y$ for which $q^*(y|x)=0$.
+In particular, there may be some $x,y$ for which $\pxy(y|x)=0$.
 Even approximate methods are difficult to apply because the parameter space can be quite high-dimensional.
@@
 
+\section{Exact details of the Markov Link Method}
+
+\label{sec:mlmdetails}
+
+Consider experiments yielding an $\Omega_\ell\times\Omega_X$ matrix $\DX$ and an $\Omega_\ell \times \Omega_Y$ matrix $\DY$, carrying the distribution
+\begin{align*}
+(\DX_{\ell 1},\DX_{\ell 2}\cdots \DX_{\ell \Omega_X}) \sim \mathrm{Multinomial}(n_\ell,\plx(\cdot | \ell))\\
+(\DY_{\ell 1},\DY_{\ell 2}\cdots \DY_{\ell \Omega_Y}) \sim \mathrm{Multinomial}(m_\ell,\ply(\cdot | \ell))
+\end{align*}
+where $\plx(x|\ell),\ply(y|\ell)$ are conditional distributions and there is some $\pxy(y|x)$ such that $\ply(y|\ell)=\sum_x \plx(x|\ell)\pxy(y|x)$.
+This is simply a restatement of the assumptions we have made throughout this paper about how our parameters of interest $(\plx,\pxy,\ply)$ are related to the data we can observe $(\DX,\DY)$.
+
+The purpose of the MLM is to make estimates about $\pxy$ using the data $\DX,\DY$.
+Unfortunately, $\pxy$ cannot be directly determined from the data.
+Even perfect knowledge of $\plx,\ply$ may be insufficient to determine the true value of $\pxy$.
+Some examples are detailed in Appendix \ref{sec:casestudies}.
+This problem is called `nonidentifiability,' and it can have some troubling consequences.
+For example, standard Bayesian analyses applied to nonidentifiable parameters will be extremely sensitive to the precise choice of our prior beliefs.
+Even with infinite data, the prior beliefs may have a significant impact on inferences.
+To avoid these difficulties, we focus on parameters that we know we can identify from data.
+In particular, we will look at lower bounds, upper bounds, and something in-between.
+
+Let $\Theta(\plx,\ply)=\{\pxy:\ \ply(y|\ell) = \sum_x \plx(x|\ell)\pxy(y|x)\}$ denote the set of links which are consistent with $\plx,\ply$ and the Markov Link Method assumption.  We define
+\begin{itemize}
+    \item $\pxy_{\mathrm{lo},\plx,\ply}(y|x) \triangleq \min_{\pxy\in \Theta(\plx,\ply)} \pxy(y|x)$
+    \item $\pxy_{\mathrm{hi},\plx,\ply}(y|x)\triangleq \max_{\pxy\in \Theta(\plx,\ply)} \pxy(y|x)$ 
+    \item $\hat \pxy_{\plx,\ply} \triangleq \argmin_{\pxy\in \Theta(\plx,\ply)} D_{f}(\mathrm{Uniform}||\pxy)$, where $D_f$ is some $f$-divergence.\footnote{In practice, we choose a $\chi^2$ divergence because it makes the minimization problem tractable.
+See Appendix \ref{sec:optproblems} for details.}
+\end{itemize}
+Even if $\pxy$ is nonidentifiable, we can still be assured that $\pxy_{\mathrm{lo}}, \pxy_{\mathrm{hi}}$ are identifiable and $q_{\mathrm{lo}}(y|x) \leq \pxy(y|x) \leq \pxy_{\mathrm{hi}}(y|x)$.
+The estimator $\hat \pxy$ is also identifiable and we can also hope it will strike a middle ground.
+In producing this single point estimate we had to decide how to deal with the fundamental fact that actually any $\pxy \in \Theta$ might be correct.
+At a basic level, we could make two kinds of mistakes.
+We might claim a very strong association between the measurement procedures even though actually there is none.
+We might claim a very weak association even though actually there is a strong association.
+We choose to err on the side of asserting weak associations, by choosing the $\pxy$ which is as close as possible to uniform.
+We made this choice in the spirit of the Maximum Entropy Principle, i.e.\ that in the absence of other information we assume $X$ is associated with each $Y$ equally.
+This is perhaps as reasonable as any way to pick a particular $\hat \pxy$.
+However, we reiterate that $\hat \pxy$ is just one possibility among many.
+It is safest to consider the full spectrum of possibilities by looking at the extremes $\pxy_{\mathrm{lo}}, \pxy_{\mathrm{hi}}$.
+
+If we had perfect knowledge of $\plx,\ply$, the objects $\pxy_{\mathrm{lo},\plx,\ply}, \pxy_{\mathrm{hi},\plx,\ply},\hat \pxy_{\plx,\ply}$ would give us a reasonable understanding of what we can know about the link $\pxy$.
+However, in practice we do not have access to $\plx,\ply$.
+Instead, we have access to the data $\DX,\DY$ which enables us to estimate $\plx,\ply$.
+To account for our uncertainty about these estimates, we take a Bayesian perspective.
+For prior beliefs about $\plx,\ply$, we take a noninformative uniform prior $\mathbb{P}$:
+\[
+\mathbb{P}(\plx,\ply) \propto 1
+\]
+Following the Bayesian philosophy, we then incorporate new knowledge by conditioning.
+We have two important pieces of knowledge about $\plx,\ply$.
+First, we have observed the data, $\DX,\DY$.
+Second, we know from the MLM assumption that there exists \emph{some} value $\pxy$ such that Equation (\ref{eq:mlmh}) holds.
+We would like to condition on both of these facts.
+However, due to the Borel-Kolmogorov paradox, `conditioning on the MLM assumption' is not a meaningful idea.
+Instead, it is necessary to define a variable indicating how much the MLM assumption fails, and condition on this variable being zero.
+In particular, let $D(h||h')$ denote the Kullback-Leibler divergence and $\Gamma(\plx,\ply) = \min_{\ply':\ \Theta(\ply,\ply')\neq \emptyset} D(\ply||\ply')$.
+Our posterior uncertainty about $\plx,\ply$ can then be articulated through the distribution
+\[
+\mathbb{P}\left(\plx,\ply | \DX,\DY,\Gamma(\plx,\ply)=0\right)
+\]
+In terms of this posterior, we define our final point estimate $\hat \pxy$ and uncertainty bounds $C$ as follows:
+\begin{enumerate}
+    \item $\hat \pxy$ is calculated using posterior expectation:
+    \[
+    \hat \pxy \triangleq \mathbb{E}\left[\hat \pxy_{\plx,\ply} | \DX,\DY,\Gamma(\plx,\ply)=0\right]
+    \]
+    %
+    \item $C_{x,y}$ is calculated as credible intervals.  For each $x,y$, we define $C_{x,y}$ to be the largest interval such that 
+    \begin{align*}
+    \mathbb{P}\left(\pxy_{\mathrm{lo},\plx,\ply}(y|x) \in C_{x,y} | \DX,\DY,\Gamma(\plx,\ply)=0\right)\geq 97.5\% \\
+    \mathbb{P}\left(\pxy_{\mathrm{hi},\plx,\ply}(y|x) \in C_{x,y} | \DX,\DY,\Gamma(\plx,\ply)=0\right)\geq 97.5\% 
+    \end{align*}
+\end{enumerate}
+
+In practice, we were not able to find a way to compute these objects exactly.
+Given samples from the posterior distribution $\mathbb{P}(\plx,\ply| \DX,\DY,\Gamma(\plx,\ply)=0)$, it would be straightforward to get good estimates.
+As seen in Appendix \ref{sec:optproblems}, it is straightforward to compute $\pxy_{\mathrm{lo}},\pxy_{\mathrm{hi}},\hat \pxy$ from samples of $\plx,\ply$, so we could use use monte-carlo approximations for our objects of interest.
+Unfortunately, it seems difficult to obtain samples from this posterior distribution.
+Common approaches to this type of problem involve Markov-Chain Monte Carlo and Variational Inference, but we were unable to make these approaches work in practice.
+It seems nontrivial to work with the condition $\Gamma(p,h)=0$ that formalizes the \MLM{} assumption.
+We instead take a somewhat na\"ive approach.
+We start by drawing samples according to
+\[
+F,H ~ \sim \mathbb{P}(\plx,\ply | \DX,\DY)
+\]
+This can be achieved exactly, using the the conjugacy between our prior and the Multinomial distribution.
+Notice that these samples do not incorporate our knowledge of the \MLM{} assumption, insofar as they are not conditioned on the event $\Gamma(\plx,\ply)=0$.
+To approximately remedy this, we define $\tilde H$ as the solution of $\min_{h'}D(H|h')$, subject to the constraint that $\Gamma(F,h')=0$.
+Optimization details can be found in Appendix \ref{sec:optproblems}.
+We use the pair $F,\tilde H$ as \emph{approximate} samples for the distribution $\mathbb{P}(\plx,\ply| \DX,\DY,\Gamma(\plx,\ply)=0)$.
+We can repeat this process to produce many samples of $(F,\tilde H)$ and use those samples to produce approximate monte-carlo estimates for $\hat \pxy(y|x),C_{x,y}$.
+We asymptotically expect that $F,H$ will nearly satisfy the MLM assumption in any case, so this approximation should not make a large difference.
+However, we caution that the exact implications of the approximation are difficult to reason about.
+In any case, it is a practical solution to a difficult problem.
+
 \section{Numerical issues}
@@
     \item Projecting to the MLM assumption.
-Fix any parameters $P,\tilde H$.
-One step in the MLM involves projecting $\tilde H$ to the set of parameters which are consistent with $P$ and the MLM assumption.
+Fix any parameters $F,H$.
+One step in the MLM involves projecting $H$ to the set of parameters which are consistent with $F$ and the MLM assumption.
 In particular, we defined
@@
     \[
-    \min_{h} D(\tilde H|h)
+    \min_{h} D(H|h)
     \]
-    subject to the constraint that there exists some $q$ such that $h(y|\ell)=\sum_x p(x|\ell)q(y|x)$.  Parametrizing valid $h$ through $q$, we obtain the problem
+    subject to the constraint that there exists some $q$ such that $h(y|\ell)=\sum_x F(x|\ell)q(y|x)$.  Parametrizing valid $h$ through $\pxy$, we obtain the problem
     \[
-    \max_{q} \sum_{\ell,y} \tilde H(y|\ell) \log \left(\sum_x P(x|\ell) q(y|x)\right)
+    \max_{\pxy} \sum_{\ell,y} H(y|\ell) \log \left(\sum_x F(x|\ell) \pxy(y|x)\right)
     \]
-    Taking derivatives one can readily show that this problem is convex.  We solve it using an iterative method.  We initially guess that $q$ is uniform.  We then repeatedly make the updates
+    Taking derivatives one can readily show that this problem is convex.
+We solve it using exponentiated gradient ascent (cf.\ \cite{kivinen1995additive}).
+We initially guess that $\pxy$ is uniform.
+We then repeatedly make the updates
     \begin{align*}
-    q(y|x)  &\propto q(y|x)\sum_{\ell}p(x|\ell)\frac{\tilde H(y|\ell)}{\sum_x p(x|\ell)q(y|x)}\\
+    \pxy(y|x)  &\propto \pxy(y|x)\sum_{\ell}F(x|\ell)\frac{H(y|\ell)}{\sum_x F(x|\ell)\pxy(y|x)}\\
     \end{align*}
@@
     \item Linear programming.
-Fix any $P,H$.
-To deal with the identifiability issues, we defined $\Theta(p,h)=\{q:\ h(y|\ell) \triangleq \sum_x p(y|\ell)q(y|x)\}$.
+Fix any $\plx,\ply$.
+To deal with the identifiability issues, we defined $\Theta(\plx,\ply)=\{\pxy:\ \ply(y|\ell) \triangleq \sum_x \plx(x|\ell)\pxy(y|x)\}$.
 The MLM requires us to solve linear optimization problems within $\Theta$, such as
     \[
-    \min_{q\in \Theta(P,H)} q(y|x) \\
+    \min_{\pxy\in \Theta(\plx,\ply)} \pxy(y|x) \\
     \]
@@
 
-    \item Quadratic programming.  The MLM also requires us to solve quadratic optimization problems within $\Theta$:
+    \item Quadratic programming.  To obtain the minimum $\chi^2$ divergence to uniform, the MLM also requires us to solve quadratic optimization problems within $\Theta$:
     \[
-    \min_{q\in \Theta(P,H)} \sum q(y|x)^2 \\
+    \min_{\pxy \in \Theta(\plx,\ply)} \sum \pxy(y|x)^2 \\
     \]
@@
 In both cases we are able to prove something of interest -- but not quite as much as we might hope.
+Here we will use the notation introduced in Appendix \ref{sec:mlmdetails}.
 
-Consider experiments yielding an $\Omega_\ell\times\Omega_X$ matrix $N^X$ and an $\Omega_\ell \times \Omega_Y$ matrix $N^Y$, sampled according to
-\begin{align*}
-(N^X_{\ell 1},N^X_{\ell 2}\cdots N^X_{\ell \Omega_X}) \sim \mathrm{Multinomial}(n_\ell,p^*(\cdot | \ell))\\
-(N^Y_{\ell 1},N^Y_{\ell 2}\cdots N^Y_{\ell \Omega_Y}) \sim \mathrm{Multinomial}(m_\ell,h^*(\cdot | \ell))
-\end{align*}
-where $p^*(x|\ell)$ is some conditional distribution and there is some $q^*(y|x)$ such that $h^*(y|\ell)=\sum_x p^*(x|\ell)q^*(y|x)$.
-This is simply a restatement of the assumptions we have made throughout the data about how our parameters of interest $(p^*,q^*,h^*)$ are related to the data we can observe $(N^X,N^Y)$.
-
-It is well-known that we can obtain arbitrarily good estimates of $p^*,h^*$ by taking enough samples (i.e.\ taking $n_\ell,m_\ell$ sufficiently high).
-Let us therefore imagine for a moment that we in fact have \emph{perfect knowledge} of $p^*,h^*$.
-Even so, the data does not necessarily tell us the value of the link parameters $q^*$.
-There may be many possible links, $q$, which are all equally consistent with $p^*,h^*$.
-Each such link yields the exact same distribution on the data we can observe, so there can be no way to use data to distinguish among them.
-This is known as a `nonidentifiability problem.'  Even with infinite data, we simply cannot identify exactly what the value of $q^*$ might be.
+First note that we can obtain arbitrarily good estimates of $\plx,\ply$ by taking enough samples (i.e.\ taking $n_\ell,m_\ell$ sufficiently high).
+Let us therefore imagine for a moment that we in fact have \emph{perfect knowledge} of $\plx,\ply$.
+Even so, the data does not necessarily tell us the value of the link parameters $\pxy$.
+There may be many possible links, $\pxy$, which are all equally consistent with $\plx,\ply$.
+That is, we may have $\pxy_1,\pxy_2$ such that $\ply(y|\ell)=\sum_x \plx(x|\ell)\pxy_1(y|x)=\ply(y|\ell)=\sum_x \plx(x|\ell)\pxy_2(y|x)$.
+Both links yield the exact same distribution on the data we can observe, so there can be no way to use data to distinguish among them.
+This is known as a `nonidentifiability problem.'  Even with infinite data, we simply cannot identify exactly what the value of $\pxy$ might be.
 
@@
 Consider the case that $\Omega_\ell = \Omega_Y = 2$ and $\Omega_X=3$.
-That is, there are $2$ separate sampling procedures, tool I recognizes 3 categories and tool II recognizes 2 categories.
-In particular, let us imagine that $p^*(x|\ell)=A_{\ell x}$ and $h^*(y|\ell)=B_{\ell y}$ where $A,B$ are matrices given by
+That is, there are $2$ separate selection procedures, tool I recognizes 3 categories and tool II recognizes 2 categories.
+In particular, let us imagine that $\plx(x|\ell)=A_{\ell x}$ and $\ply(y|\ell)=B_{\ell y}$ where $A,B$ are matrices given by
 %
@@
 
-Rows correspond to different sampling procedures and columns correspond to a different measurement outcome.
-Now let $q^*(y|x)=C_{x y}$, another matrix.
+Rows correspond to different selection procedures and columns correspond to a different measurement outcome.
+Now let $\pxy(y|x)=C_{x y}$, another matrix.
 The Markov Link Method assumption then tells us that $A\times C=B$, where $\times$ indicates matrix multiplication.
 This corresponds to $\Omega_\ell \times \Omega_Y = 4$ equations.
-We also have a normalizing constraint that $\sum y q^*(y|x)=1$, which creates $\Omega_X=3$ additional equations.
+We also have a normalizing constraint that $\sum y \pxy(y|x)=1$, which creates $\Omega_X=3$ additional equations.
 However, these normalizing constraints actually make two of the MLM assumption constraints redundant.
@@
 However, the matrix $C$ contains six numbers.
-The result is a degree of freedom in $C$, corresponding to an aspect of $q^*$ that we simply cannot resolve.
+The result is a degree of freedom in $C$, corresponding to an aspect of $\pxy$ that we simply cannot resolve.
 For example, here are two choices of $C$ which are both consistent with the equation $A\times C=B$:
@@
 \[
-q^*(y|x)=\begin{cases}1 \quad \mathrm{if}\ x=y\\0\quad \mathrm{else}\end{cases}
+\pxy(y|x)=\begin{cases}1 \quad \mathrm{if}\ x=y\\0\quad \mathrm{else}\end{cases}
 \]
-and  $p^*(x|\ell)=A_{\ell,x}$, where this matrix is given by
+and  $\plx(x|\ell)=A_{\ell,x}$, where this matrix is given by
 \[
@@
 That is, the $x$th column of the first $k-1$ rows is the binary expansion of the number $x-1$, and the last row alternates $1$s and $0$s.
-Now let us say we have perfect knowledge of $p^*(x|\ell)$ and $h^*(y|\ell)=\sum_x p^*(y|\ell)q^*(y|x)$.
-Notice that due to the simple structure of $q^*$ we obtain $h^*(y|\ell)=A^{\ell,y}$.
-However, let us imagine we know nothing about the true value of $q^*$.
+Now let us say we have perfect knowledge of $\plx(x|\ell)$ and $\ply(y|\ell)=\sum_x \plx(y|\ell)\pxy(y|x)$.
+Notice that due to the simple structure of $\pxy$ we obtain $\ply(y|\ell)=A^{\ell,y}$.
+However, let us imagine we know nothing about the true value of $\pxy$.
 
-How much can we say about $q^*$, if we only had knowledge of $p^*$ and $h^*$?
-On the one hand, we observe that in the absence of any other constraints, the object $q^*$ has $2^{2k-3}$ degrees of freedom.
-This is because there are $2^{k-1}$ values of $\ell$ and for each subpopulation $q^*(\cdot|\ell)$ must lie in the $2^{k-2}$-dimensional simplex on $2^{k-1}$ atoms.
+How much can we say about $\pxy$, if we only had knowledge of $\plx$ and $\ply$?
+On the one hand, we observe that in the absence of any other constraints, the object $\pxy$ has $2^{2k-3}$ degrees of freedom.
+This is because there are $2^{k-1}$ values of $\ell$ and for each subpopulation $\pxy(\cdot|\ell)$ must lie in the $2^{k-2}$-dimensional simplex on $2^{k-1}$ atoms.
 On the other hand, we see that the Markov Link Assumption gives us $k\times(2^{k-1}-1)$ linear constraints on the value of $q$.
@@
 \]
-Of these $k\times2^{k-1}$ constraints, $k$ of them are redundant with the fact that $\sum_y q(y|x)=1$.
+Of these $k\times2^{k-1}$ constraints, $k$ of them are redundant with the fact that $\sum_y \pxy(y|x)=1$.
 Thus, altogether, the Markov Link Assumption together with approximate knowledge of $p,h$ gives us $k\times(2^{k-1}-1)$ linear constraints.
@@
 
-In conclusion, a simple degrees-of-freedom counting argument would suggest that there will be substantial ambiguity about what value $q$ might take on, if our only knowledge about $q$ is that it must satisfy $\sum_x p^*(y|\ell)q(y|x)=h^*(y|\ell)$.
+In conclusion, a simple degrees-of-freedom counting argument would suggest that there will be substantial ambiguity about what value $q$ might take on, if our only knowledge about $q$ is that it must satisfy $\sum_x \plx(y|\ell)q(y|x)=\ply(y|\ell)$.
 Indeed, we have \emph{exponentially many} more degrees of freedom than we have constraints.
 
-However, the reality is that $q$ is exactly determined by $p^*,h^*$.
-This is possible because there are inequality constraints which also govern $q$, namely $q(y|x)\geq 0$.
+However, the reality is that $q$ is exactly determined by $\plx,\ply$.
+This is possible because there are inequality constraints which also govern $q$, namely $\pxy(y|x)\geq 0$.
 Thus, while a simple degrees-of-freedom counting argument might suggest that we would have substantial identifiability issues in this problem, the reality is quite the opposite.
@@
 \begin{thm}\label{thm:miracle}
-Let $p^*,h^*$ be as they are defined above.  Then there is exactly one $q$ that is consistent with $p^*,h^*$ and the Markov Link assumption.  That is, $q^*$ is the only possible $q$ satisfying
+Let $\plx,\ply$ be as they are defined above.
+Then there is exactly one $\pxy$ that is consistent with $\plx,\ply$ and the Markov Link assumption.
+That is, $\pxy$ is the only possible value satisfying
 \begin{gather*}
-\sum_y q(y|x)=1 \\
-\sum_x A_{\ell,x}q(y|x)=A_{\ell,y}\\
+\sum_y \pxy(y|x)=1 \\
+\sum_x A_{\ell,x}\pxy(y|x)=A_{\ell,y}\\
 q(y|x)\geq 0
@@
 \begin{align*}
-0=&q(y|x) \qquad \forall y\leq 2^{k-2},x>2^{k-2}
+0=&\pxy(y|x) \qquad \forall y\leq 2^{k-2},x>2^{k-2}
 \end{align*}
-Indeed, for each $y\leq 2^{k-2}$ we obtain a constraint showing that $\sum_{x>2^{k-2}}q(y|x)=0$, which yields that in fact $q(y|x)=0$ for every $x>2^{k-2}$ and every $y\leq 2^{k-2}$.
+Indeed, for each $y\leq 2^{k-2}$ we obtain a constraint showing that $\sum_{x>2^{k-2}}q(y|x)=0$, which yields that in fact $\pxy(y|x)=0$ for every $x>2^{k-2}$ and every $y\leq 2^{k-2}$.
 
@@
 \[
-\sum_{x\leq2^{k-2}} A_{\ell x} q(x|y) = A_{\ell y} \qquad \forall y\leq 2^{k-2}
+\sum_{x\leq2^{k-2}} A_{\ell x} \pxy(x|y) = A_{\ell y} \qquad \forall y\leq 2^{k-2}
 \]
 This is an example of the same problem we started with -- except with $k$ one smaller.
-Applying our inductive hypothesis, we may thus obtain that $q(y|x)=q^*(y|x)$ for the first $2^{k-2}$ values of $x,y$.
-Moreover, since $\sum_{y\leq 2^{k-2}} q^*(y|x)=1$, we see that $q$ must also satisfy $q(y|x)=0$ for $y>2^{k-2}$ and $x\leq 2^{k-2}$.
-Thus we have seen that $q=q^*$ for all entries except those in which $x,y\geq 2^{k-2}$.
+Applying our inductive hypothesis, we may thus obtain that $\pxy(y|x)$ is uniquely determined for the first $2^{k-2}$ values of $x,y$.
+Moreover, since $\sum_{y\leq 2^{k-2}} \pxy(y|x)=1$, we see that $\pxy$ must also satisfy $\pxy(y|x)=0$ for $y>2^{k-2}$ and $x\leq 2^{k-2}$.
+Thus we have seen that $\pxy$ is uniquely identified for all entries except those in which $x,y\geq 2^{k-2}$.
 
 For $x,y \geq2^{k-2}$ we linearly combine equations concerning the first, last, and second to last rows of $A$ with factors of $1,1,-1$ respectively.
-We obtain constraints showing that $\sum_{x\leq2^{k-2}}q(y|x)=0$ for each $y>2^{k-2}$.
-We can then use the same reasoning to obtain that $q=q^*$ for the remaining values of $x,y$.
+We obtain constraints showing that $\sum_{x\leq2^{k-2}}\pxy(y|x)=0$ for each $y>2^{k-2}$.
+We can then use the same reasoning to obtain that $\pxy$ is uniquely identified for the remaining values of $x,y$.
 \end{proof}
 
-This result is somewhat robust to slight perturbations in $p^*,h^*$.
-In particular, if we have some $\hat p\approx p^*$ and $\hat h\approx h^*$ then at each stage of the argument we can replace statements of the form $q(y|x)=0$ with statements of the form $q(y|x)\leq \epsilon$.
-Applying this with the kinds of arguments above will show that we can be sure that $\hat q$ is arbitrarily close to $q^*$ if we know that $\hat p,\hat h$ are sufficiently close to $p^*,h^*$.
+This result is somewhat robust to slight perturbations in $\plx,\ply$.
+In particular, if we have some $\hat \plx\approx \plx$ and $\hat \ply\approx \ply$ then at each stage of the argument we can replace statements of the form $\pxy(y|x)=0$ with statements of the form $\pxy(y|x)\leq \epsilon$.
+Applying this with the kinds of arguments above will show that we can be sure that every point in $\Theta(\hat \plx,\hat \ply)$ is arbitrarily close to $\pxy$ if we know that $\hat \plx,\hat \ply$ are sufficiently close to $\plx,\ply$.
 
-However, it turns out that the relationship between $q$ and $p^*,h^*$ is not robust in every situation.  In the next section we will see that it can in fact be quite discontinuous:
+However, it turns out that the relationship between $\pxy$ and $\plx,\ply$ is not robust in every situation.  In the next section we will see that it can in fact be quite discontinuous:
 
@@
 Consider the case that $\Omega_\ell = 1$ and $\Omega_X=\Omega_Y=2$.
-That is, there is only one sampling procedure (no subpopulations) and both tool I and tool II can return one of $2$ possible values.
+That is, there is only one selection procedure (no subpopulations) and both tool I and tool II can return one of $2$ possible values.
 We will now consider two possiblities:
@@
     \begin{itemize}
-    \item $\mathbb{P}(X=1)=p^*(1)=0$
-    \item $\mathbb{P}(X=2)=p^*(2)=1$
-    \item $\mathbb{P}(Y=1)=h^*(1)=0$
-    \item $\mathbb{P}(Y=2)=h^*(2)=1$
+    \item $\mathbb{P}(X=1)=\plx(1)=0$
+    \item $\mathbb{P}(X=2)=\plx(2)=1$
+    \item $\mathbb{P}(Y=1)=\ply(1)=0$
+    \item $\mathbb{P}(Y=2)=\ply(2)=1$
     \end{itemize}
-    In this case the MLM assumption $\sum_x p^*(x)q(y|x)=h^*(y)$ can be used to prove that $q(1|2)=0,q(2|2)=1$, but we now have \emph{absolutely no} knowledge of $q(1|1),q(2|1)$.
-This is because we simply never observed the case $X=1$ (it occurs with probability zero), and so we cannot possibly have any knowledge about $q(y|x)$ for $x=1$.
+    In this case the MLM assumption $\sum_x \plx(x)\pxy(y|x)=\ply(y)$ can be used to prove that $\pxy(1|2)=0,\pxy(2|2)=1$, but we now have \emph{absolutely no} knowledge of $\pxy(1|1),\pxy(2|1)$.
+This is because we simply never observed the case $X=1$ (it occurs with probability zero), and so we cannot possibly have any knowledge about $\pxy(y|x)$ for $x=1$.
 \item Now let us take a slight variation:
     \begin{itemize}
-    \item $\mathbb{P}(X=1)=p^*(1)=0.01$
-    \item $\mathbb{P}(X=2)=p^*(2)=0.99$
-    \item $\mathbb{P}(Y=1)=h^*(1)=0$
-    \item $\mathbb{P}(Y=2)=h^*(2)=1$
+    \item $\mathbb{P}(X=1)=\plx(1)=0.01$
+    \item $\mathbb{P}(X=2)=\plx(2)=0.99$
+    \item $\mathbb{P}(Y=1)=\ply(1)=0$
+    \item $\mathbb{P}(Y=2)=\ply(2)=1$
     \end{itemize}
-    In this case we can again prove that $q(1|2)=0,q(2|2)=1$, but we can also prove that $q(1|1)=0,q(2|1)=1$.  
+    In this case we can again prove that $\pxy(1|2)=0,\pxy(2|2)=1$, but we can also prove that $\pxy(1|1)=0,\pxy(2|1)=1$.  
 \item Now we take yet another slight variation:
     \begin{itemize}
-    \item $\mathbb{P}(X=1)=p^*(1)=0.01$
-    \item $\mathbb{P}(X=2)=p^*(2)=0.99$
-    \item $\mathbb{P}(Y=1)=h^*(1)=0.01$
-    \item $\mathbb{P}(Y=2)=h^*(2)=0.99$
+    \item $\mathbb{P}(X=1)=\plx(1)=0.01$
+    \item $\mathbb{P}(X=2)=\plx(2)=0.99$
+    \item $\mathbb{P}(Y=1)=\ply(1)=0.01$
+    \item $\mathbb{P}(Y=2)=\ply(2)=0.99$
     \end{itemize}
-    In this case we can prove that $q(1|2)\leq 1/99$ and $q(2|1)\geq 1-1/99$, but we again cannot prove almost anything about $q(2|1)$.
-In particular, it is easy to produce cases in which $q(2|1)=0$ and other cases in which $q(2|1)=1$.
+    In this case we can prove that $\pxy(1|2)\leq 1/99$ and $\pxy(2|1)\geq 1-1/99$, but we again cannot prove almost anything about $\pxy(2|1)$.
+In particular, it is easy to produce cases in which $\pxy(2|1)=0$ and other cases in which $\pxy(2|1)=1$.
 
@@
 
-The disturbing thing about this example is that by making infinitesimal perturbations to $p^*$ we can pass from uncertainty to complete certainty back to uncertainty.
-It is for this reason that in this paper we refuse to ever treat $\hat p,\hat h$ as fixed and given, always considering the space of perturbations around any such values.
+The disturbing thing about this example is that by making infinitesimal perturbations to $\plx$ we can pass from uncertainty to complete certainty back to uncertainty.
+It is for this reason that in this paper we refuse to ever treat $\plx,\ply$ as fixed and given, always considering the space of perturbations around any such values.
 
-It is worth noting that these kinds of problems essentially vanish if the true $q^*$ is bounded away from zero i.e.\ $q^*(y|x)>c$ for every $x,y$ for some $c>0$.
-If this holds, together with a certain linear independence assumption, we can guarantee that the true $q^*$ is close to the set $\Theta(\hat p,\hat h)$ when $\hat p,\hat h$ are good approximations to $p^*,h^*$, where
+It is worth noting that these kinds of problems essentially vanish if the true $\pxy$ is bounded away from zero i.e.\ $\pxy(y|x)>c$ for every $x,y$ for some $c>0$.
+If this holds, together with a certain linear independence assumption, we can guarantee that the true $\pxy$ is close to the set $\Theta(\hat \plx,\hat \ply)$ when $\hat \plx,\hat \ply$ are good approximations to $\plx,\ply$, where
 \[
-\Theta(p,h) \triangleq \left\{q\in T:\ \sum_x p(x|\ell)q(y|x)= h(y|\ell) \quad \forall \ell,y \right\}
+\Theta(\plx,\ply) \triangleq \left\{\pxy\in T:\ \sum_x \plx(x|\ell)\pxy(y|x)= \ply(y|\ell) \quad \forall \ell,y \right\}
 \]
-and by $T$ we mean the transition matrix polytope, $T=\{q:\ q(y|x)\geq 0, \sum_y q(y|x)=1\}$.  In particular:\vspace{.1in}
+and by $T$ we mean the transition matrix polytope, $T=\{\pxy:\ \pxy(y|x)\geq 0, \sum_y \pxy(y|x)=1\}$.  In particular:\vspace{.1in}
 
 \begin{thm}\label{thm:mainthm}
-Fix any $q^*$ satisfying $q^*(y|x)>c$ for some $c>0$.
-Let us further assume that the matrix $B^*_{\ell,x}=p^*(x|\ell)$ has linearly independent rows.
-Fix any $p^*$ and define $h^*(y|\ell)=\sum_x p^*(x|\ell)q^*(y|x)$.
-Then by taking any $\hat p,\hat h$ sufficiently close to $p^*,h^*$ we can ensure that $q^*$ is arbitrarily close to some point in the set $\hat \Theta = \Theta(\hat p,\hat h)$.
+Fix any $\pxy^*$ satisfying $\pxy^*(y|x)>c$ for some $c>0$.
+Let us further assume that the matrix $B^*_{\ell,x}=\plx^*(x|\ell)$ has linearly independent rows.
+Fix any $\plx$ and define $\ply(y|\ell)=\sum_x \plx(x|\ell)\pxy(y|x)$.
+Then by taking any $\hat \plx,\hat \ply$ sufficiently close to $\plx,\ply$ we can ensure that $\pxy^*$ is arbitrarily close to some point in the set $\hat \Theta = \Theta(\hat \plx,\hat \ply)$.
 \end{thm}
 \begin{proof}
-Let $\hat A$ denote the affine plane $\hat A=\{q:\ \sum_x \hat p(x|\ell)q(y|x)= \hat  h(y|\ell)\}$.  Thus $\hat \Theta = T \cap \hat A$.
+Let $\hat A$ denote the affine plane $\hat A=\{\pxy:\ \sum_x \hat \plx(x|\ell)\pxy(y|x)= \hat \ply(y|\ell)\}$.  Thus $\hat \Theta = T \cap \hat A$.
 
 Now fix any $\hat p,\hat h$.
-Now let $\tilde q$ be the orthogonal Euclidean projection of $q^*$ to the affine space $A$.
-That is, $\tilde q$ is minimizes a sum-of-squares difference to $q^*$ among all the points in $A$.
-The linearly independent rows of $B^*$ allow us to bound the spectral norm of the right-pseudoinverse of $\hat B_{\ell,x}=\hat p(x|\ell)$, by taking $\hat p$ sufficiently close to $p^*$.
-If we furthermore require $\hat h$ sufficiently close to $h^*$, we can use this to ensure the projection distance is small.
-That is, we can force $\tilde q$ to be arbitrarily close to $q^*$.
-Using the fact that $q^*(y|x)>c$ we can thereby furthermore insure that $\tilde q(y|x)\geq0$.
-Finally, it is straightforward to see that the projection leaves the constraint $\sum_y q(y|x)=1$ unchanged.
-Thus $\tilde q\in \hat \Theta$ and $\tilde q$ is arbitrarily close to $q^*$.
+Now let $\tilde \pxy$ be the orthogonal Euclidean projection of $\pxy^*$ to the affine space $\hat A$.
+That is, $\tilde q$ is minimizes a sum-of-squares difference to $\pxy^*$ among all the points in $\hat A$.
+The linearly independent rows of $B^*$ allow us to bound the spectral norm of the right-pseudoinverse of $\hat B_{\ell,x}=\hat \plx(x|\ell)$, by taking $\hat \plx$ sufficiently close to $\plx^*$.
+If we furthermore require $\hat \ply$ sufficiently close to $\ply^*$, we can use this to ensure the projection distance is small.
+That is, we can force $\tilde \pxy$ to be arbitrarily close to $\pxy^*$.
+Using the fact that $\pxy^*(y|x)>c$ we can thereby furthermore insure that $\tilde \pxy(y|x)\geq0$.
+Finally, it is straightforward to see that the projection leaves the constraint $\sum_y \pxy(y|x)=1$ unchanged.
+Thus $\tilde \pxy\in \hat \Theta$ and $\tilde \pxy$ is arbitrarily close to $\pxy^*$.
 \end{proof}
 
-Since it is easy to find consistent estimators for $\hat p,\hat h$, this theorem suggests we can use those estimators to get good estimates for our uncertainty about $q^*$.
-In particular, subject to the conditions of the theorem, we have that the bounds of $C_{x,y}$ converge to the bounds of what is identifiable about $q^*$.
+Since it is easy to find consistent estimators for $\plx,\ply$, this theorem suggests we can use those estimators to get good estimates for our uncertainty about $\pxy$.
+In particular, subject to the conditions of the theorem, we have that the bounds of $C_{x,y}$ converge to the bounds of what is identifiable about $\pxy$.
 It is certainly a step in the right direction, but we emphasize that the theorem's conditions are nontrivial.
@@
 However, the positivity condition is quite troubling.
-In many cases of interest the true link $q^*$ has genuine zeros: pairs of measurements between the two tools which are fundamentally incompatible.
+In many cases of interest the true link $\pxy$ has genuine zeros: pairs of measurements between the two tools which are fundamentally incompatible.
 In this case such a theorem cannot be applied.
 
-However, it may be that the above theorem's requirement ($q(x|y)>c>0$) is much stronger than is necessary.
+However, it may be that the above theorem's requirement ($\pxy(x|y)>c>0$) is much stronger than is necessary.
 A precise understanding of the discontinuity example above has remained elusive.

### 2018-08-16 paper/paper.tex
they're not parameters anymore!

@@
 % \newcommand{\DY}{\ensuremath{{\mathcal{D}^Y}}}
-\newcommand{\DX}{\ensuremath{\mathcal{D}^X}}
-\newcommand{\DY}{\ensuremath{\mathcal{D}^Y}}
+\newcommand{\DX}{\ensuremath{{\mathcal{D}_X}}}
+\newcommand{\DY}{\ensuremath{{\mathcal{D}_Y}}}
+% \newcommand{\DX}{\ensuremath{\mathcal{D}^X}}
+% \newcommand{\DY}{\ensuremath{\mathcal{D}^Y}}
 \newcommand{\MLMH}{\ensuremath{\mathcal{A}}}
@@
 \begin{abstract}
-Using observations from multiple experiments is essential to building larger scientific theories.
-For example, we might analyze different aspects of human neurons, such as morphology, transcriptomics, or electrophysiology.
-A strong theory would account for each perspective as well as the relationship between the perspectives.
-Unfortunately, we rarely have a single experiment that simultaneously measures every viewpoint, uses a sufficient sample size, and gets sufficient measurement precision.
-In practice, we must therefore conduct different experiments and find a way to link the results.
-The Markov Link Method (MLM) offers an algorithm to reason about these `links.' The method is nonparametric: it makes no assumptions about what the link should look like.
+We propose the Markov Link Method (\MLM), a way to synthesize experiments based on several measurement procedures.
+For example, we might have different experiments measuring different aspects of human neurons, such as morphology, transcriptomics, or electrophysiology.
+The \MLM{} uses this data to link these different aspects.
+It answers the question ``We have observed the transcriptome of this cell -- what might its morphology have been?'' The method is nonparametric: it makes no assumptions about what the link looks like.
 We evaluate the MLM on a pair of single-cell RNA techniques and gain new insight about the link between the two methods.
@@
 
-Imagine we have a collection of different experiments about human cells.
-We would like to link the different perspectives that we get from each experiment.
-To make this idea rigorous, we need to define what we mean by `perspective' and what we mean by `link.'
-
-To define `perspective,' we assume each experiment has two aspects:
+We here propose the Markov Link Method (\MLM), a technique to understand how different measurement procedures are related.  For our purposes, we assume each experiment has two aspects:
 %
@@
 %
-The selection procedure defines what the experiment studies, and the measurement procedures determines the experiment's `perspective.'
-
-The `link' between two measurement procedures should enable us to answer questions such as ``I have observed the transcriptome of this cell -- what morphologies might this cell have?''  We formalize this idea through what we call the `measurement link parameters,' $\pxy(y|x)$.
+To synthesize experiments with different measurement procedures, we must understand how the measurement procedures are related.
+For example, we would like to be able to answer questions such as ``We have measured the transcriptome of this cell -- what might its morphology have been?''  We formalize this idea through what we call the `measurement link,' $\pxy(y|x)$.
 Specifically, let us say we obtain measurement result $x$ from one measurement procedure.
@@
 
-We here propose the The Markov Link Method (\MLM), an algorithm that uses a collection of experiments to estimate the measurement link parameters, \pxy.
+We here propose the The Markov Link Method (\MLM), an algorithm that uses a collection of experiments to estimate the measurement link, \pxy.
 In this paper we focus particularly on the case that we have experiments based on several selection procedures and exactly two different measurement procedures.
@@
 
-The Markov Link Method is a general way to estimate measurement link parameters from a collection of experiments.
+The Markov Link Method is a general way to estimate measurement link from a collection of experiments.
 For concreteness, we will use experiments about human cells as a running example.
@@
 
-More formally, our task is to estimate the link parameters $\pxy$ from a collection of experiments.
+More formally, our task is to estimate the link $\pxy$ from a collection of experiments.
 We suppose that for each selection procedure we have two experiments: one that measured transcriptomics and another that measured morphology.
@@
 In the specific case we will focus on, all of the experimental data can be summarized with two matrices, $\DX$ and $\DY$.
-Here $\DX_{\ell x}$ indicates the number of specimens gathered with procedure $\ell$ that have transcriptomic type $x$.
-Likewise $\DY_{\ell y}$ indicates the number of specimens gathered with procedure $\ell$ that have morphological type $y$.
-For each transcriptomic type $x$ and morphological type $y$, the Markov Link Method uses the data to produce two objects:
+The matrix $\DX$ tabulates the transcriptomic types found in specimens gathered with each selection procedure; $\DY$ does the same for morphological types.
+For each transcriptomic type $x$ and morphological type $y$, the Markov Link Method uses this data to produce two objects:
 %
 \begin{enumerate}
-    \item $\hat \pxy(y|x)$ -- a point estimate for the link parameter $\pxy(y|x)$
-    \item $C_{x,y}$ -- a credible interval which contains the true link parameter $\pxy(y|x)$ with high probability
+    \item $\hat \pxy(y|x)$ -- a point estimate for the link $\pxy(y|x)$
+    \item $C_{x,y}$ -- a credible interval which contains the true link $\pxy(y|x)$ with high probability
 \end{enumerate}
 %
-To estimate these, we must consider certain `nuisance parameters.'  Our observable data, $\DX,\DY$, is governed by the following parameters:
+To estimate these, we must consider certain `nuisance' quantities.
+The link is our primary target of study, but we cannot estimate the link without considering these other objects.
+Our observable data, $\DX,\DY$, is governed by the following conditional distributions:
 \begin{itemize}
@@
 It is straightforward to estimate $\plx,\ply$ from $\DX,\DY$ using standard methods.
-We would like to use these parameters to tell us something about our object of interest, the link $\pxy$.
+We would like to use these distributions to tell us something about our object of interest, the link $\pxy$.
 Under the Markov Link Method assumption, this link is connected to $\plx,\ply$ through the equations
@@
 With these equations and knowledge of $\plx,\ply$, we can produce an estimate for the link: $\hat \pxy_{\plx,\ply}$.
-We can also produce bounds for the link.
+If the equations are underdetermined, we can also produce bounds for the link.
 We define these objects precisely in Appendix \ref{sec:mlmdetails}.
@@
 Second, we know from the MLM assumption that there exists \emph{some} value $\pxy$ such that Equation (\ref{eq:mlmh}) holds; let \MLMH{} denote the event that this holds.
-We then define our point estimate $\hat \pxy$ as the posterior expectation of $\hat \pxy_{\plx,\ply}$ after conditioning on this knowledge, i.e.
+By conditioning on our knowledge, we obtain the posterior distribution $\mathbb{P}(d\plx,d\ply | \DX,\DY,\MLMH)$.
+Our point estimate $\hat \pxy$ is then defined as a Bayes estimator for $\hat \pxy_{\plx,\ply}$ with respect to this posterior:
 \[
-\hat \pxy(y|x) = \mathbb{E}\left[\hat \pxy_{\plx,\ply} | \DX,\DY,\MLMH \right]
+\hat \pxy(y|x) \triangleq \mathbb{E}\left[\hat \pxy_{\plx,\ply} | \DX,\DY,\MLMH \right]
+               = \int \hat \pxy_{\plx,\ply} \mathbb{P}(d\plx,d\ply | \DX,\DY,\MLMH) 
 \]
-We can use a similar strategy to produce the credible intervals $C_{x,y}$.
+This estimate minimizes the expected squared risk error under the posterior distribution.
+We can use similar strategies to produce the credible intervals $C_{x,y}$.
 Exact details can be found in Appendix \ref{sec:mlmdetails}.
@@
 \caption{{\bf When can the MLM accurately estimate the link?} \label{fig:simI}  We use simulations to test the MLM.
-In subfigure (a) we look at the error of the MLM link estimator in many different situations, varying the numbers of samples, selection procedures, and the link parameters.
-In each trial our `ground truth' link parameters picked uniformly at random from the parameter space.
+In subfigure (a) we look at the error of the MLM link estimator in many different situations, varying the numbers of samples, selection procedures, and the link itself.
+In each trial our `ground truth' link picked uniformly at random from the parameter space.
 The MLM estimator accurately recovers the ground truth from simulated data, as long as we have enough samples \emph{and} enough distinct selection procedures.
 In subfigure (b) we consider the MLM's credible intervals.
-For a selection of trials we choose a random value of of $x,y$ and then compare the true link parameter $\pxy(y|x)$ with the MLM interval $C_{x,y}$.
+For a selection of trials we choose a random value of of $x,y$ and then compare the true link $\pxy(y|x)$ with the MLM interval $C_{x,y}$.
 We show the interval (in black) centered around the ground truth (in red).
-Since the link parameters lie in $[0,1]$, the interval is always contained within $\pm1$ of the ground truth (this larger region is shown in gray).
+Since the link probabilities lie in $[0,1]$, the interval is always contained within $\pm1$ of the ground truth (this larger region is shown in gray).
 The intervals mostly contain the truth, and get narrower with more samples as long as there are enough distinct selection procedures.
@@
 \end{description}
-However, due to certain identifiability issues, it is only actually possible to guarantee the last desideratum.
-In many cases, the MLM achieves all three.
-However, we can show that identifiability issues block any possibility of success in some cases; we detail this problem in Appendix \ref{sec:casestudies}.
-The good news is that even though the Markov Link Method credible intervals may not concentrate around the truth, they do as well as possible.
-In particular, they do appear to converge around the quantities we can estimate, namely the $q_\mathrm{lo},q_\mathrm{hi}$ bounds described in Appendix \ref{sec:mlmdetails}.
-A precise statement of what we have been able to prove may be found in Theorem \ref{thm:mainthm} of Appendix \ref{sec:casestudies}.
-
+In many cases, the MLM achieves all three of these conditions.
+In other cases, we can show that an issue known as `identifiability' blocks even the possibility of estimator convergence or interval concentration.
+However, even when interval concentration is impossible, the \MLM{} appears to do as well as it possibly can: the intervals $C_{x,y}$ close around the bounds of what can possibly be learned from the data we observe.
+We have begun to develop a theory of the relevant issues, which we detail in Appendix \ref{sec:mlmdetails}.
+We give a few of the interesting corner-cases and prove a few surprising results.
+A complete story remains elusive.
 The following two simulations give a good overview of the relevant issues:
@@
 To see how the method performs in different circumstances, we will run many trials.
-In each trial we fix the number of selection procedures and pick `ground truth' parameters uniformly at random from the parameter space.
-We then simulate a dataset from these ground truth parameters, fixing the total number of samples (spreading these samples equally among all combinations of selection procedure and measurement procedure).
+In each trial we fix the number of selection procedures and pick a `ground truth' link uniformly at random from the parameter space.
+We then simulate a dataset from this ground truth link, fixing the total number of samples (spreading these samples equally among all combinations of selection procedure and measurement procedure).
 Finally, we apply the MLM to the simulated data to get the point estimates $\tilde q(y|x)$ and the credible intervals $C_{x y}$.
@@
 With fewer samples or selection procedures, the intervals are sometimes forced to be larger.
-However, our uncertainty is not the same for every parameter.
-In some cases we are able to obtain a very tight credible interval for one particular parameter even though the overall estimator error is high.
+However, our uncertainty is not the same for every aspect of the link.
+In some cases we are able to obtain a very tight credible interval for one particular value even though the overall estimator error is high.
 
-    \item How does estimator convergence depend upon the link parameters themselves?
+    \item How does estimator convergence depend upon the link itself?
 In each trial of this simulation we use four selection procedures and our measurement procedures yield one of six categories.
 In the previous simulations we saw that estimator convergence was generally impossible in this situation, due to the small number of selection procedures.
-However, in our previous simulations we picked the link parameters uniformly from parameter space.
+However, in our previous simulations we picked the link uniformly from its parameter space.
 Now we will be more choosey.
@@
 On the left we indicate the lower bounds indicated by the credible interval, on the right we indicate upper bounds.
-For some parameters our intervals are much tighter than others.
+For some aspects of the link our intervals are much tighter than others.
 For example, it appears we have high confidence the Standard type `Vip Rspo4' is highly associated with the Patch type `Vip.'  By constrast, we have almost no idea what is associated with the Standard type `Meis2.'  The upper credible interval bounds suggest that $\pxy(y|\mathrm{Meis2})$ could be nearly 1 for many different Patch types.
@@
 In Figure \ref{fig:allenboot} we visualize these objects for selected values of $x,y$.
-Each parameter has a slightly different story.
+Each part of the link has a slightly different story.
 For example, for the Standard type $x=$`Vip Rspo4' and the Patch type $y=$`Vip,' we have that $C_{x,y}=[.88,1.0]$.
@@
 In this paper we suggest one way to overcome these difficulties.
-We formalize the `relationship' between two measurement procedures through the notion of a `measurement link.'  Given that we have measured a specimen with one procedure and obtained the outcome $x$, the link parameter $q(y|x)$ tells us the probability that we would obtain outcome $y$ if we measured the same specimen with the second procedure.
+We formalize the `relationship' between two measurement procedures through the notion of a `measurement link.'  Given that we have measured a specimen with one procedure and obtained the outcome $x$, the link $q(y|x)$ tells us the probability that we would obtain outcome $y$ if we measured the same specimen with the second procedure.
 We present the Markov Link Method assumption, which roughly states that the link can be statistically isolated from the selection procedures.
@@
 Another direction for future work would be to take a frequentist point of view.
-An obvious direction would be to use profile likelihoods as statistics to define confidence intervals for the link parameters.
+An obvious direction would be to use profile likelihoods as statistics to define confidence intervals for the link.
 However, the distribution of these statistic is difficult to pin down.
-Traditional techniques based on asymptotic normality may fail dramatically, because the true parameters may lie at the boundary of the parameter space.
+Traditional techniques based on asymptotic normality may fail dramatically, because the true link may lie at the boundary of the parameter space.
 In particular, there may be some $x,y$ for which $\pxy(y|x)=0$.
@@
 where $\plx(x|\ell),\ply(y|\ell)$ are conditional distributions and there is some $\pxy(y|x)$ such that $\ply(y|\ell)=\sum_x \plx(x|\ell)\pxy(y|x)$.
-This is simply a restatement of the assumptions we have made throughout this paper about how our parameters of interest $(\plx,\pxy,\ply)$ are related to the data we can observe $(\DX,\DY)$.
+This is simply a restatement of the assumptions we have made throughout this paper about how our objects of interest $(\plx,\pxy,\ply)$ are related to the data we can observe $(\DX,\DY)$.
 
@@
 Even with infinite data, the prior beliefs may have a significant impact on inferences.
-To avoid these difficulties, we focus on parameters that we know we can identify from data.
+To avoid these difficulties, we focus on objects that we know we can identify from data.
 In particular, we will look at lower bounds, upper bounds, and something in-between.
@@
     \item $\pxy_{\mathrm{hi},\plx,\ply}(y|x)\triangleq \max_{\pxy\in \Theta(\plx,\ply)} \pxy(y|x)$ 
-    \item $\hat \pxy_{\plx,\ply} \triangleq \argmin_{\pxy\in \Theta(\plx,\ply)} D_{f}(\mathrm{Uniform}||\pxy)$, where $D_f$ is some $f$-divergence.\footnote{In practice, we choose a $\chi^2$ divergence because it makes the minimization problem tractable.
+    \item $\hat \pxy_{\plx,\ply} \triangleq \argmin_{\pxy\in \Theta(\plx,\ply)} D_{f}(\mathrm{Uniform}||\pxy)$, where $D_f$ is some $f$-divergence.\footnote{In practice, we choose a $\chi^2$ divergence because it makes the minimization problem a quadratic program; this makes it particularly easy to solve.
 See Appendix \ref{sec:optproblems} for details.}
@@
     \item Projecting to the MLM assumption.
-Fix any parameters $F,H$.
-One step in the MLM involves projecting $H$ to the set of parameters which are consistent with $F$ and the MLM assumption.
+Fix any values for $F,H$.
+One step in the MLM involves projecting $H$ to the set of distributions which are consistent with $F$ and the MLM assumption.
 In particular, we defined
@@
 Let us therefore imagine for a moment that we in fact have \emph{perfect knowledge} of $\plx,\ply$.
-Even so, the data does not necessarily tell us the value of the link parameters $\pxy$.
+Even so, the data does not necessarily tell us the value of the link $\pxy$.
 There may be many possible links, $\pxy$, which are all equally consistent with $\plx,\ply$.

### 2018-08-16 paper/paper.tex
Gave some context for size of projection distance

@@
 We asymptotically expect that $F,H$ will nearly satisfy the MLM assumption in any case, so this approximation should not make a large difference.
-However, we caution that the exact implications of the approximation are difficult to reason about.
+For example, on the Allen Institute data we found that the total variation distance between $H(\cdot|\ell)$ and $\tilde H(\cdot|\ell)$ was about 15\% (averaging over all selection procedures $\ell$ and various samples of $H$).
+For comparison, this is about three times smaller than the average total variation distance between $H(\cdot|\ell)$ and $H(\cdot|\ell')$ for a randomly selected pair of selection procedures $(\ell,\ell')$, which averages out to around 50\%.
 In any case, it is a practical solution to a difficult problem.

### 2018-08-16 markovlink/approxbayes.py

@@ -23,6 +23,25 @@ def posterior_samples(Nlx,Nly,nsamps=100,verbose=False,clipthresh=1e-6):
 
     return estimates
 
+def projection_check(Nlx,Nly,nsamps=100,verbose=False,clipthresh=1e-6):
+    estimates=[]
+    if verbose:
+        misc.pnn("\n %d samples to produce:"%nsamps)
+    for i in range(nsamps):
+        if verbose:
+            misc.pnn(i)
+        p=np.array([npr.dirichlet(x+1) for x in Nlx])
+        htilde=np.array([npr.dirichlet(x+1) for x in Nly])
+
+        qtilde=estimation.train_mlm_fixedp(htilde,p)
+
+        # get the central q, for uniqueness
+        q = polytopes.find_central(p,qtilde,clipthresh=1e-6)
+        
+        estimates.append((p,q,htilde))
+
+    return estimates
+
 
 def credible_interval(estimates,xtilde,ytilde,alpha=.05,verbose=False):
     lower,upper=get_extremes(estimates,xtilde,ytilde,verbose=verbose)

### 2018-08-17 paper/paper.tex
Dave edits

@@
 
-\title{The Markov Link Method: a nonparametric approach to combine observations from multiple experiments}
+\title{The Markov link method: a nonparametric approach to combine observations from multiple experiments}
 \author{Jackson Loper, Osnat Penn, Trygve Bakken, David Blei, Liam Paninski, \\Additional Authors To Be Determined}
@@
 \begin{abstract}
-We propose the Markov Link Method (\MLM), a way to synthesize experiments based on several measurement procedures.
+We propose the Markov link method (\MLM), a way to synthesize experiments based on several measurement procedures.
 For example, we might have different experiments measuring different aspects of human neurons, such as morphology, transcriptomics, or electrophysiology.
-The \MLM{} uses this data to link these different aspects.
-It answers the question ``We have observed the transcriptome of this cell -- what might its morphology have been?'' The method is nonparametric: it makes no assumptions about what the link looks like.
-We evaluate the MLM on a pair of single-cell RNA techniques and gain new insight about the link between the two methods.
+The \MLM{} uses this data find the  link between different aspects.
+This link answers questions such as: ``We have observed the transcriptome of this cell -- what might its morphology have been?'' The method inspects how different \emph{measurement} procedures relate by using a collection of \emph{sample selection} procedures.
+In doing so, it makes no assumptions about what the link looks like.
+Instead, the \MLM{} relies on a conditional independence assumption.
+We prove conditions under which the \MLM{} estimators are consistent, and simulations show that it gives accurate measures of uncertainty.
+We evaluate the MLM on a pair of single-cell RNA techniques and gain new insight about the link between the two methods; we see aspects in which two different notions of `cell-type' are aligned, aspects in which they are not, and aspects about which we are unsure.
+The \MLM{} suggests specific additional experiments which would help refine our understanding of the relationship between the two measurement procedures.
 \end{abstract}
 
-We here propose the Markov Link Method (\MLM), a technique to understand how different measurement procedures are related.  For our purposes, we assume each experiment has two aspects:
+It is sometimes difficult to understand how to synthesize experiments which use different measurement procedures.  We assume each experiment has two aspects:
 %
@@
 To synthesize experiments with different measurement procedures, we must understand how the measurement procedures are related.
-For example, we would like to be able to answer questions such as ``We have measured the transcriptome of this cell -- what might its morphology have been?''  We formalize this idea through what we call the `measurement link,' $\pxy(y|x)$.
+For example, we would like to be able to answer questions such as ``We have measured the transcriptome of this cell -- what might its morphology have been?''  We formalize this idea through a conditional probability distribution we call the `measurement link,' $\pxy(y|x)$.
 Specifically, let us say we obtain measurement result $x$ from one measurement procedure.
@@
 
-We here propose the The Markov Link Method (\MLM), an algorithm that uses a collection of experiments to estimate the measurement link, \pxy.
+We here propose the The Markov link method (\MLM), an algorithm that uses a collection of experiments to estimate the measurement link, \pxy.
 In this paper we focus particularly on the case that we have experiments based on several selection procedures and exactly two different measurement procedures.
@@
 
-\caption{{\bf The Markov Link Method (MLM) links different measurement procedures by using a set of different selection procedures.}  In subfigure (a) we have neurons gathered using many different selection procedures.
+\caption{{\bf The Markov link method (MLM) links different measurement procedures by using a set of different selection procedures.}  In subfigure (a) we have neurons gathered using many different selection procedures.
 We select half of the neurons and determine what type of neuron they are, using transcriptomic measurements.
@@
 
-\section{The Markov Link Method}
+\section{The Markov link method}
 
@@
 
-The Markov Link Method is a general way to estimate measurement link from a collection of experiments.
+The Markov link method is a general way to estimate measurement links using a collection of experiments.
 For concreteness, we will use experiments about human cells as a running example.
@@
 \begin{enumerate}
-    \item $\ell$, the selection procedure used to obtain the specimen
+    \item $\ell$, the selection procedure used to obtain the cell.
+For example, $\ell=1$ might correspond to a procedure that selects only cells exhibiting one proteomic marker, whereas $\ell=2$ might select cells with a different marker.
     \item $X$, the cell's `transcriptomic type,' as measured through a transcriptomic analysis.  
@@
 %
-Our task is to link the transcriptomic and morphological perspectives on `cell-type,' as manifested through the $X$ and $Y$ measurement procedures.
-For example, we might like to say ``This cell has transcriptomic type 3, therefore it probably has morphological type F.''  Our main assumption is that these kinds of associations are unaffected by the choice of selection procedure.
-In particular:
+The \MLM{} gives us a way to link the transcriptomic and morphological perspectives on `cell-type,' as manifested through the $X$ and $Y$ measurement procedures.
+For example, we might like to say ``This cell has transcriptomic type 3, therefore it probably has morphological type F.''
 
-\begin{center}
-\fbox{\parbox[c]{4in}{
-\vspace{.1in}
-\begin{center}
-\textbf{The Markov Link Method Assumption}\\$\mathbb{P}(Y|X,\ell)=\pxy(y|x)$ for each $\ell$
-\end{center}
-\vspace{.1in}
-}}
-\end{center}
-
-Here $\mathbb{P}(Y|X,\ell)$ indicates the probability that the morphological type is $y$ given that the transcriptomic type is $x$ and the specimen was sampled with strategy $\ell$.
+The \emph{Markov link method assumption} states that these kinds of associations are unaffected by the choice of selection procedure.
+That is, $\mathbb{P}(Y|X,\ell)=\pxy(y|x)$ for each $\ell$, where $\mathbb{P}(Y|X,\ell)$ indicates the probability that the morphological type is $y$ given that the transcriptomic type is $x$ and the specimen was sampled with strategy $\ell$.
 The assumption states that that this probability actually does not depend upon $\ell$.
-This assumption can be understood intuitively through a thought experiment.
+The graphical model in figure \ref{fig:bigpicture} embodies precisely this assumption.
+Intuitively, we can understand the assumption using a thought experiment.
 Let us imagine we have attained perfect understanding of cellular biology.
@@
 
-More formally, our task is to estimate the link $\pxy$ from a collection of experiments.
+The purpose of the \MLM{} is to use data and the \MLM{} assumption to estimate the link $\pxy$ from a collection of experiments.
 We suppose that for each selection procedure we have two experiments: one that measured transcriptomics and another that measured morphology.
@@
 The matrix $\DX$ tabulates the transcriptomic types found in specimens gathered with each selection procedure; $\DY$ does the same for morphological types.
-For each transcriptomic type $x$ and morphological type $y$, the Markov Link Method uses this data to produce two objects:
+For each transcriptomic type $x$ and morphological type $y$, the Markov link method uses this data to produce two objects:
 %
@@
 To estimate these, we must consider certain `nuisance' quantities.
-The link is our primary target of study, but we cannot estimate the link without considering these other objects.
-Our observable data, $\DX,\DY$, is governed by the following conditional distributions:
+The link is the primary target of study, but we cannot estimate the link without considering these other objects.
+The observable data, $\DX,\DY$, is governed by the following conditional distributions:
 \begin{itemize}
@@
 It is straightforward to estimate $\plx,\ply$ from $\DX,\DY$ using standard methods.
-We would like to use these distributions to tell us something about our object of interest, the link $\pxy$.
-Under the Markov Link Method assumption, this link is connected to $\plx,\ply$ through the equations
+We would like to use these distributions to tell us something about the the link, $\pxy$.
+Under the Markov link method assumption, this link is connected to $\plx,\ply$ through the equations
 \begin{gather}\label{eq:mlmh}
@@
 We only have estimates from data.
-We take a Bayesian perspective to account for our uncertainty in this estimation.
-For our prior beliefs, we assume a noninformative uniform prior $\mathbb{P}$.
+We take a Bayesian perspective to account for uncertainty in this estimation.
+For prior beliefs, we assume a noninformative uniform prior $\mathbb{P}$.
 Following the Bayesian philosophy, we then incorporate new knowledge by conditioning.
@@
 Second, we know from the MLM assumption that there exists \emph{some} value $\pxy$ such that Equation (\ref{eq:mlmh}) holds; let \MLMH{} denote the event that this holds.
-By conditioning on our knowledge, we obtain the posterior distribution $\mathbb{P}(d\plx,d\ply | \DX,\DY,\MLMH)$.
-Our point estimate $\hat \pxy$ is then defined as a Bayes estimator for $\hat \pxy_{\plx,\ply}$ with respect to this posterior:
+By conditioning on this knowledge, we obtain the posterior distribution $\mathbb{P}(\plx,\ply | \DX,\DY,\MLMH)$.
+The \MLM{} point estimate $\hat \pxy$ is then defined as a Bayes estimator for $\hat \pxy_{\plx,\ply}$ with respect to this posterior:
 \[
 \hat \pxy(y|x) \triangleq \mathbb{E}\left[\hat \pxy_{\plx,\ply} | \DX,\DY,\MLMH \right]
-               = \int \hat \pxy_{\plx,\ply} \mathbb{P}(d\plx,d\ply | \DX,\DY,\MLMH) 
+               = \int \hat \pxy_{\plx,\ply} \mathbb{P}(\plx,\ply | \DX,\DY,\MLMH) d\plx,d\ply
 \]
@@
 We can use similar strategies to produce the credible intervals $C_{x,y}$.
-Exact details can be found in Appendix \ref{sec:mlmdetails}.
-Code to compute $\hat \pxy(y|x)$ and $C_{x,y}$ is published at https://github.com/jacksonloper/markov-link-method, including a tutorial-style ipython notebook detailing every computation made in this paper.
+Exact details can be found in Appendix \ref{sec:mlmdetails}.\footnote{Code to compute $\hat \pxy(y|x)$ and $C_{x,y}$ is published at https://github.com/jacksonloper/markov-link-method, including a tutorial-style ipython notebook detailing every computation made in this paper.}
 
@@
 
-\section{Examples where the Markov Link Method may apply}
+\subsection{Examples}
 
@@
 
-The validity of the Markov Link Method assumption for a given situation should be closely contemplated.  Let us consider a few real-world examples where this assumption may apply.
+The \MLM{} is a general approach for synthesizing experiments, which requires that the Markov link method assumption holds.  Here are some examples where the \MLM{} could be applied:
 
@@
 These cell cultures are subject to quite different pressures, due to the fact that they survive on a plate instead of inside a human being.
-The Markov Link method can leverage the common side-information together with separate transcriptome information to understand the correspondence between in-vivo and cultured cells.
+The Markov link method can leverage the common side-information together with separate transcriptome information to understand the correspondence between in-vivo and cultured cells.
 If a particular drug is effective on a particular cultured cell-line, we can then look at the corresponding in-vivo transcriptomic profile.
@@
 Obviously the last kind is the most useful for automatic image captioning, but there is much less of it.
-The Markov Link Method suggests one way to use the more plentiful text-only and image-only data.
+The Markov link method suggests one way to use the more plentiful text-only and image-only data.
 We can first apply classic machine learning techniques to get coarse labels for both kinds of data.
@@
 In subfigure (a) we look at the error of the MLM link estimator in many different situations, varying the numbers of samples, selection procedures, and the link itself.
-In each trial our `ground truth' link picked uniformly at random from the parameter space.
+In each trial the `ground truth' link is picked uniformly at random from the parameter space.
 The MLM estimator accurately recovers the ground truth from simulated data, as long as we have enough samples \emph{and} enough distinct selection procedures.
@@
 
-There are three desiderata we might hope the Markov Link Method estimators would achieve:
+There are three desiderata we might hope the Markov link method estimators would achieve:
 \begin{description}
@@
 With fewer samples or selection procedures, the intervals are sometimes forced to be larger.
-However, our uncertainty is not the same for every aspect of the link.
+However, the uncertainty is not the same for every aspect of the link.
 In some cases we are able to obtain a very tight credible interval for one particular value even though the overall estimator error is high.
@@
     \item How does estimator convergence depend upon the link itself?
-In each trial of this simulation we use four selection procedures and our measurement procedures yield one of six categories.
+In each trial of this simulation we use four selection procedures and the measurement procedures yield one of six categories.
 In the previous simulations we saw that estimator convergence was generally impossible in this situation, due to the small number of selection procedures.
-However, in our previous simulations we picked the link uniformly from its parameter space.
+However, in previous simulations we picked the link uniformly from its parameter space.
 Now we will be more choosey.
@@
 \hfill{}\figgygr{.95}{../images/allenNlxy}\hfill{}
-\caption{{\bf The input to the Markov Link Method: one experiment for each measurement procedure and each selection procedure.}  Here we show a portion of the real-world data to which we applied the MLM.
+\caption{{\bf The input to the Markov link method: one experiment for each measurement procedure and each selection procedure.}  Here we show a portion of the real-world data to which we applied the MLM.
 The Allen Institute gathered neurons from the visual cortex of mice, using a variety of cre/lox-based selection procedure (cf.\ \cite{tasic2017shared}).
@@
 Above we show a subset of these results through heatmaps; the color of each square indicates the number of specimens found to have a particular type.
-Using this kind of data, our task is to calibrate the two classification protocols.
+Using this kind of data, the task is to calibrate the two classification protocols.
 That is, we want to be able to ask question of the following form: ``if a neuron is classified as being of type `Peri Kcnj8' by Standard, how might it have been classified by Patch?'' \label{fig:alleninput}}
@@
 For each combination $x,y$ we draw a rectangle whose color indicates the value of the estimator $\hat \pxy(y|x)$.
-We also show our confidence about these estimators.
+We also determine the \MLM{}'s confidence about these estimators.
 On the left we indicate the lower bounds indicated by the credible interval, on the right we indicate upper bounds.
-For some aspects of the link our intervals are much tighter than others.
+For some aspects of the link the intervals are much tighter than others.
 For example, it appears we have high confidence the Standard type `Vip Rspo4' is highly associated with the Patch type `Vip.'  By constrast, we have almost no idea what is associated with the Standard type `Meis2.'  The upper credible interval bounds suggest that $\pxy(y|\mathrm{Meis2})$ could be nearly 1 for many different Patch types.
@@
 The variability in the credible regions suggests what we need to do in order to more closely determine the value of the calibration.
-For example, if we could develop more unique sampling strategies which will include `Lamp5 Egln 1' cells, this would help us resolve our ambiguity about this aspect of the link.
+For example, if we could develop more unique sampling strategies which will include `Lamp5 Egln 1' cells, this would help us resolve the ambiguity about this aspect of the link.
 Indeed, going back to the original data, it is easy to see why this ambiguity appeared in the first place.
@@
 
-The main distinguishing characteristics of this paper are two-fold: we place no assumptions on the nature of the link and we take the identifiability issues seriously.
+The main distinguishing characteristics of this paper are two-fold: we take the identifiability issues seriously and place no assumptions on the nature of the link.
 We refer the reader to \cite{walter2014identifiability} for an introduction to what we mean by identifiability.
-We discuss the identifiability issues for our problem in particular in Appendix \ref{sec:casestudies}.
-Our only assumption is that the selection procedures can be statistically isolated from the link.
+We discuss the identifiability issues for this problem in particular in Appendix \ref{sec:casestudies}.
+As for assumptions, the only assumption is that the selection procedures can be statistically isolated from the link.
 This assumption says nothing about what the link itself is.
-We take a Bayesian approach, but we only use our prior beliefs to estimate quantities which actually can be estimated, carefully avoiding the identifiability issues.
+We take a Bayesian approach, but we only use prior beliefs to estimate quantities which actually can be estimated, carefully avoiding the identifiability issues.
 We use these quantities to produce intervals which account for uncertainty due both to small sample sizes and to identifiability issues.
@@
 Perhaps the closest work to this one would be \cite{makarov1982estimates}, which used two marginal distributions to get bounds on a property of the joint distribution (namely the distribution of the sum).
-We advance this approach to a more general-purpose technique, both by using many subpopulations to closely refine our estimates and by considering the entire space of possible joint distributions instead of simply a particular property of the joint.
+We advance this approach to a more general-purpose technique, both by using many subpopulations to closely refine the \MLM{} estimates and by considering the entire space of possible joint distributions instead of simply a particular property of the joint.
 
@@
 We formalize the `relationship' between two measurement procedures through the notion of a `measurement link.'  Given that we have measured a specimen with one procedure and obtained the outcome $x$, the link $q(y|x)$ tells us the probability that we would obtain outcome $y$ if we measured the same specimen with the second procedure.
-We present the Markov Link Method assumption, which roughly states that the link can be statistically isolated from the selection procedures.
-We provide the Markov Link Method (MLM) algorithm that uses this assumption to estimate the link and measure our uncertainty.
-Code is published at https://github.com/jacksonloper/markov-link-method, including a tutorial-style ipython notebook detailing every computation made in this paper.
+We present the Markov link method assumption, which roughly states that the link can be statistically isolated from the selection procedures.
+We provide the Markov link method (MLM) algorithm that uses this assumption to estimate the link and measure uncertainty.
 
@@
 In future work, it would be interesting to use additional assumptions to help us estimate calibrations.
-For example, as it stands the Markov Link Method will only yield narrow credible intervals if each measurement tool returns one of a finite number of measurement outcomes.
+For example, as it stands the Markov link method will only yield narrow credible intervals if each measurement tool returns one of a finite number of measurement outcomes.
 In some cases, we may believe that similar measurement values should have similar probabilities.
@@
 It is clear that good assumptions can help us get real insight for tough problems.
-Even if these assumptions are not sufficient to allow us to perfectly identify our object of interest, we can bound our uncertainty.
+Even if these assumptions are not sufficient to allow us to perfectly identify objects of interest, we can still get uncertainty bounds.
 By probing this uncertainty carefully, we can learn what the data actually has to say and what experiments will help us learn more.
@@
 
-\section{Exact details of the Markov Link Method}
+\section{Exact details of the Markov link method}
 
@@
 where $\plx(x|\ell),\ply(y|\ell)$ are conditional distributions and there is some $\pxy(y|x)$ such that $\ply(y|\ell)=\sum_x \plx(x|\ell)\pxy(y|x)$.
-This is simply a restatement of the assumptions we have made throughout this paper about how our objects of interest $(\plx,\pxy,\ply)$ are related to the data we can observe $(\DX,\DY)$.
+This is simply a restatement of the assumptions we have made throughout this paper about how the objects of interest $(\plx,\pxy,\ply)$ are related to the data we can observe $(\DX,\DY)$.
 
@@
 This problem is called `nonidentifiability,' and it can have some troubling consequences.
-For example, standard Bayesian analyses applied to nonidentifiable parameters will be extremely sensitive to the precise choice of our prior beliefs.
+For example, standard Bayesian analyses applied to nonidentifiable parameters will be extremely sensitive to the precise choice of prior beliefs.
 Even with infinite data, the prior beliefs may have a significant impact on inferences.
@@
 
-Let $\Theta(\plx,\ply)=\{\pxy:\ \ply(y|\ell) = \sum_x \plx(x|\ell)\pxy(y|x)\}$ denote the set of links which are consistent with $\plx,\ply$ and the Markov Link Method assumption.  We define
+Let $\Theta(\plx,\ply)=\{\pxy:\ \ply(y|\ell) = \sum_x \plx(x|\ell)\pxy(y|x)\}$ denote the set of links which are consistent with $\plx,\ply$ and the Markov link method assumption.  We define
 \begin{itemize}
@@
 Instead, we have access to the data $\DX,\DY$ which enables us to estimate $\plx,\ply$.
-To account for our uncertainty about these estimates, we take a Bayesian perspective.
+To account for uncertainty about these estimates, we take a Bayesian perspective.
 For prior beliefs about $\plx,\ply$, we take a noninformative uniform prior $\mathbb{P}$:
@@
 In particular, let $D(h||h')$ denote the Kullback-Leibler divergence and $\Gamma(\plx,\ply) = \min_{\ply':\ \Theta(\ply,\ply')\neq \emptyset} D(\ply||\ply')$.
-Our posterior uncertainty about $\plx,\ply$ can then be articulated through the distribution
+Posterior uncertainty about $\plx,\ply$ can then be articulated through the distribution
 \[
@@
 \]
-In terms of this posterior, we define our final point estimate $\hat \pxy$ and uncertainty bounds $C$ as follows:
+In terms of this posterior, we define the \MLM{} point estimate $\hat \pxy$ and uncertainty bounds $C$ as follows:
 \begin{enumerate}
@@
 \]
-This can be achieved exactly, using the the conjugacy between our prior and the Multinomial distribution.
-Notice that these samples do not incorporate our knowledge of the \MLM{} assumption, insofar as they are not conditioned on the event $\Gamma(\plx,\ply)=0$.
+This can be achieved exactly, using the the conjugacy between the prior and the Multinomial distribution.
+Notice that these samples do not incorporate knowledge of the \MLM{} assumption, insofar as they are not conditioned on the event $\Gamma(\plx,\ply)=0$.
 To approximately remedy this, we define $\tilde H$ as the solution of $\min_{h'}D(H|h')$, subject to the constraint that $\Gamma(F,h')=0$.
@@
 
-There are three numerical problems which the MLM must solve.  Here we detail our method for solving each of them.
+There are three numerical problems which the MLM must solve.  Here we detail the \MLM{} method for solving each of them.
 
@@
     \end{align*}
-    until convergence.  Our convergence criteria is that all parameters change less than $10^{-5}$ in a single iteration.
+    until convergence.  The algorithm's convergence criteria is that all parameters change less than $10^{-5}$ in a single iteration.
 
@@
 Now let $\pxy(y|x)=C_{x y}$, another matrix.
-The Markov Link Method assumption then tells us that $A\times C=B$, where $\times$ indicates matrix multiplication.
+The Markov link method assumption then tells us that $A\times C=B$, where $\times$ indicates matrix multiplication.
 This corresponds to $\Omega_\ell \times \Omega_Y = 4$ equations.
@@
 
-It follows that for $y\leq 2^{k-2}$ our original constraints may be rewritten as
+It follows that for $y\leq 2^{k-2}$ the original constraints may be rewritten as
 \[
@@
 This is an example of the same problem we started with -- except with $k$ one smaller.
-Applying our inductive hypothesis, we may thus obtain that $\pxy(y|x)$ is uniquely determined for the first $2^{k-2}$ values of $x,y$.
+Applying the inductive hypothesis, we may thus obtain that $\pxy(y|x)$ is uniquely determined for the first $2^{k-2}$ values of $x,y$.
 Moreover, since $\sum_{y\leq 2^{k-2}} \pxy(y|x)=1$, we see that $\pxy$ must also satisfy $\pxy(y|x)=0$ for $y>2^{k-2}$ and $x\leq 2^{k-2}$.
@@
 
-Since it is easy to find consistent estimators for $\plx,\ply$, this theorem suggests we can use those estimators to get good estimates for our uncertainty about $\pxy$.
+Since it is easy to find consistent estimators for $\plx,\ply$, this theorem suggests we can use those estimators to get good estimates for uncertainty about $\pxy$.
 In particular, subject to the conditions of the theorem, we have that the bounds of $C_{x,y}$ converge to the bounds of what is identifiable about $\pxy$.

### 2018-08-22 paper/paper.tex
another abstraact revision

@@
 \begin{abstract}
-We propose the Markov link method (\MLM), a way to synthesize experiments based on several measurement procedures.
+The Markov link method (\MLM) synthesizes experiments based on different measurement procedures.
 For example, we might have different experiments measuring different aspects of human neurons, such as morphology, transcriptomics, or electrophysiology.
-The \MLM{} uses this data find the  link between different aspects.
-This link answers questions such as: ``We have observed the transcriptome of this cell -- what might its morphology have been?'' The method inspects how different \emph{measurement} procedures relate by using a collection of \emph{sample selection} procedures.
-In doing so, it makes no assumptions about what the link looks like.
+To combine knowledge across these experiments, we develop the idea of a ``link'' between measurement procedures.
+This link should answer questions such as: ``We have observed the transcriptome of this cell -- what might its morphology have been?'' Without joint measurement of the transcriptome and morphology for the same neurons, it is challenging to estimate the link.
+However, it turns out the problem is not impossible.
+The \MLM{} estimates these links without joint measurement and without any assumptions about what the link looks like.
 Instead, the \MLM{} relies on a conditional independence assumption.
+The estimation problem arising from this assumption is challenging, due to nonidentifiability problems; we give a rigorous characterization of the issue and show that the \MLM{} handles it correctly.
 We prove conditions under which the \MLM{} estimators are consistent, and simulations show that it gives accurate measures of uncertainty.
-We evaluate the MLM on a pair of single-cell RNA techniques and gain new insight about the link between the two methods; we see aspects in which two different notions of `cell-type' are aligned, aspects in which they are not, and aspects about which we are unsure.
+We evaluate the \MLM{} on a pair of single-cell RNA techniques and gain new insight about the link between the two methods; we find that the two different notions of `cell-type' are generally well-aligned, with a few exceptions.
+The \MLM{} also shows that some aspects of the link simply cannot be determined from the data we have.
 The \MLM{} suggests specific additional experiments which would help refine our understanding of the relationship between the two measurement procedures.

### 2018-08-22 markovlink/polytopes.py

@@ -181,6 +181,9 @@ def find_extremal(plx,qxy,c,use_cvxopt=True,precondition=False):
     c = np.require(c,dtype=np.float).ravel()
     assert c.shape[0]==np.prod(qxy.shape),"c should be the same shape as qxy"
     
+    if qxy.shape[0] <= plx.shape[0]: # <-- trivial polytope
+        return qxy
+
     # polytope can be defined by A_eq q = b_eq, q>=0:
     A_eq,b_eq = q_with_fixedph(plx,plx@qxy,precondition=precondition)
 
